### 2022-10-24 1.Paper/Paper.Rmd
Initial

---
title: Multiple imputation of incomplete multilevel data using Heckman selection models
author:
- name: J. Munoz*
  num: a
- name: V. de Jong
  num: a
- name: T. Debray
  num: a
address:
- num: a
  org: Julius Center for Health Sciences and Primary Care, Utrecht Medical Center,Utrech, The Netherlands
corres: "* Johanna Munoz. \\email{j.munozavila@umcutrecht.nl}"
presentaddress: Julius Center for Health Sciences and Primary Care, Utrecht Medical Center, Universiteitsweg 100, 3584 CG Utrech, The Netherlands
authormark: Uthor \emph{et al}.
articletype: Research article
received: 2022-11-01
revised: 2022-12-01
accepted: 2023-01-01
abstract: "Missing data is a common problem in medical research, and is commonly addressed using multiple imputation. Although traditional imputation methods allow for valid statistical inference when data are missing at random (MAR), their implementation is not justified when observations are clustered (e.g., within studies) or when the presence of missingness depends on unobserved information. Unfortunately, this situation is increasingly common, and typically arises when individual participant data (IPD) are combined from multiple studies. While several imputation methods have been proposed for addressing individual studies where data are missing not at random (MNAR), their application and validity in large datasets with clustering remains unclear. We therefore explored the consequence of MNAR data in IPDMA in-depth, and proposed novel multilevel imputation methods for common missing patterns in clustered datasets. These methods build upon the principles of Heckman selection models, and adopt a two-stage meta-analysis approach for imputing binary and continuous variables. After evaluating the proposed imputation models in simulated scenarios, we illustrated their use in a real IPDMA based on a malaria study in five subregions in Uganda, where we focused on estimating the prevalence of parasitemia for children aged 2-10 years."
keywords: Heckman model; IPDMA; Missing not at random; Selection models;Multiple imputation;
bibliography: bibfile.bib
output: 
  rticles::sim_article:
  fig_caption: yes
  longtable: yes
  doublespace: yes

header-includes:
   - \usepackage{mathtools}
   - \usepackage{float}
   - \usepackage{multicol}
   - \usepackage{graphicx}
     \newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}
   - \usepackage{amsmath}
     \DeclareMathOperator\arctanh{arctanh}
---

# Introduction
Over the past few years, data sharing efforts have substantially increased and researchers increasingly often have access to large combined datasets derived from electronic health records or from individual participant data (IPD). For example, the clinical practice research datalink (CRPD) [@herrett2015] is an electronic health record (EHR) dataset in the UK, and has been used in a variety of medical research, such as the evaluation of health policy and drug efficacy. A recent example of an IPD-MA is the emerging risk factor collaboration [@theemergingriskfactorscollaboration2007], where data were combined from approximately 1.1 million individuals across 104 studies to investigate associations of cardiovascular diseases with several predictors. Individuals in these large datasets tend to be clustered in centres, countries or studies, where they have been subject to similar healthcare processes, and are therefore more alike than individuals from another cluster. Sometimes, clusters may also differ in participant eligibility criteria, follow-up length, predictor and outcome definitions, or in the quality of applied measurement methods. Hence, correlation is likely to be present between observations from the same cluster, which can lead to differences or ‘heterogeneity’ between clusters regarding baseline patient characteristics and subsequent outcomes.

Clustered datasets often contain  many incomplete variables. For example, in registry data it is common that test results are not available in registry data for all patients, as the decision to test may be at the discretion of the primary care physician or because the patient refuses to undergo testing . It is also possible that variables are systematically missing across clusters. For instance, in an IPD meta-analysis, it is common that studies collected information on different variables. Missing values may thus appear for all participants of a study in the combined dataset. The presence of missing data can lead to loss of statistical power, imbalance across clusters, bias in parameter estimates and therefore to erroneous conclusions as the analysis could be based on an unrepresentative sample.

To address the presence of missing data, it is important to consider  the proportion of missingness and the missing mechanism for each incomplete variable. Rubin (1976) [@rubin76] identified three missing mechanisms where the probability of missingness: 1) is constant (missing completely  at random; MCAR), 2) depends on observed data only (missing at random; MAR), or 3) depends on unobserved information even after conditioning on all observable variables (missing non-random; MNAR).  Traditional imputation methods are designed to address incomplete data sets where variables are MCAR or MAR. Their implementation is reasonable when there is no obvious mechanism of missingness, or when the observed data strongly relate to unobserved information.

Although it is possible to formally rule out MCAR [@little1988], it is impossible to assess whether data are MAR or MNAR as the observable data are not enough to test the assumptions of both mechanisms [@enders2022]. For this reason, researchers often conveniently assume that data are MAR, or present results that are based on a complete case analysis. In practice, however, unless missingness is artificially introduced by study design, e.g. when a test is only taken on patients with certain characteristics, missingness will often (partially) depend on unobserved information.

One could see, a missing variable is a mixture of MAR and MNAR, depending on the available information, making the MAR mechanism more likely as more information is recorded. For instance, healthcare professionals may decide what type of measurement on indication, and depending on their personal experience or behavior decide whether or not to record the reasons for the measurement.

Registries are notoriously prone to incomplete variables that are MNAR, due to the complex recording process <!--[@liu2021]-->, e.g. laboratory tests are taken only in certain patients based on sign and symptom information that is often incomplete or not recorded. Also IPD may suffer from MNAR, for example when study participants who experience unfavorable results drop out of the study or also due to heterogeneity of the primary objective or resources of the studies involved, which may result in variables relevant to explain the missing process not being recorded.

The MNAR mechanism is considered as non-ignorable because to deal with this type of missingness it is necessary to specify information about the missingness process in addition to assumptions about the observed data. A common approach to address MNAR is to adopt  selection models such as the one proposed by Heckman (1976) [@heckman1976]. Briefly, the Heckman selection model corrects for selection bias by estimating two linked equations, a main equation where the missing variable is associated with predictors, and a selection equation that accounts for the inclusion of observations in the sample.   An important feature of the Heckman selection model is that its implementation does not require data to be MNAR, and can also be used when data are MCAR or MAR. It therefore offers an appealing solution to incomplete data sets where the missingness mechanism is unknown.

Over the past few years, several Heckman selection models have been proposed for multiple imputation. Among them, Galimard et al.(2016) [@galimard_etal16] implemented a chained equations imputation method for continuous variables, which was extended to binary and categorical variables by employing copula estimates [@galimard_etal18].  Also, Ogundimu and Collins (2019) [@ogundimu_collins19] proposed a chained equations imputation method that is less dependent on Normality assumptions.

In clustered data sets, multilevel imputation methods are needed to properly propagate uncertainty within and across clusters [audigier_etal18]. However, to our knowledge, existing multilevel imputation methods mainly focus  on situations where data are MAR, and do not adopt Heckman selection models. Although Hammon and Zinn (2020) [@hammon_zinn20] recently proposed an extension that allows for the inclusion of random intercept effects, it can only be used for binary missing variables and assume that the effect of explanatory variables on the missingness mechanisms is common across clusters.

Therefore, the aim of this paper is to develop a multilevel imputation method for incomplete continuous and binary variables that are MNAR. In section \ref{heckman} we provide an introduction to the Heckman model and its estimation, and we extended to a hierarchical setting. Then in the section \ref{methods} we define the main steps of the proposed imputation method. Posteriorly, in the section \ref{simulation} we provide the settings and results of a Montecarlo simulation study used to evaluate the performance of the proposed impuatation method. In section \ref{example} we illustrate the method using the survey information collected in different sub-districts in Uganda to estimate the prevalence of malaria in children. 
Finally we provide a discussion  in section \ref{discussion} about  the results, limitations and propose future extension of our method.

# The Heckman model
\label{heckman}
The Heckman selection model was initially proposed as a method to correct for selection bias, in which individuals are not randomly selected from the population, leading to inconsistent estimates and erroneous conclusions [@heckman1976].

Selection bias occurs when the selection of a subject  or his observation into the sample is influenced by unobserved variables  (e.g., the respondent's level of trust toward healthcare entities may cause him or her to self-select out of the study or refuse to sign consent for a test), which in turn are correlated with unobserved variables related to a variable of interest (e.g., the result of a blood test) [@vella1998].

\begin{figure}[h!]
   \centering
   \includegraphics[scale=0.35]{DAG_heckman.png}
   \caption{DAG Heckman selection model: here the nodes (dotted = latent, continuous = observable) describe the relationship between $y^*_{ij}$ the latent response and $r^*_{ij}$ the latent  selection variables}
   \label{fig:DAG}
\end{figure}
This can be better visualized in Figure \ref{fig:DAG}, where for the $j$th individual or unit within the $i$th cluster, there is $y_{ij}^*$, a latent outcome variable, and $r_{ij}^*$, a latent selection variable, which are correlated through $u_{ij}$ an unobserved or unrecorded variable, with $i\in[1,2,...,N]$ and $j\in[1,2,...,n_i]$. Here, both latent variables are related to $x_{ij}^O$ and $x_{ij}^S$ sets of predictor covariates. From $r_{ij}^*$ one can derive $r_{ij}=I(r_{ij}^*>=0)$ a selection indicator of $y_{ij}^*$ into the sample, and with this in turn, one can define $y_{ij}=y_{ij}^*,\forall r_{ij}=1$ the observable outcome variable.

Denoting $y^*_i=(y^*_{i1},y^*_{i2},...,y^*_{in_i})^T$ and  $r^*_i=(r^*_{i1},r^*_{i2},...,r^*_{in_i})^T$ the vectors of latent outcomes and latent selections in the $i$ cluster, then Heckman's model is defined by two main equations: the outcome equation (\ref{eq:1}), which describes the relation latent outcome-exposure, and the selection equation (\ref{eq:2}) which details the likelihood that the outcome is observed in the sample.
\begin{align}
y^*_{i}&=X^O_{i}\beta^O_{i}+\epsilon^O_{i}\label{eq:1}\\
r^*_{i}&=X^S_{ij}\beta^S_{i}+\epsilon^S_{i}\label{eq:2}
\end{align}

Here $\beta^O_{i}$ and $\beta^S_{i}$ are $p\times1$ and $q\times1$ coefficient parameter vectors and $\epsilon^O_{i}=(\epsilon^O_{i1},\epsilon^O_{i2},...,\epsilon^O_{in_i})^T$ and $\epsilon^S_{i}=(\epsilon^S_{i1},\epsilon^S_{i2},...,\epsilon^S_{in_i})^T$ are the residual terms vectors for the outcome and selection equations, respectively. Generally the same variables can be used on the matrix of predictor variables $X^O_{i}$ and $X^S_{i}$. However, to avoid multicollinearity problems [@puhani1997], it is recommended to include in $X^S_{i}$ at least one variable that is not included in the outcome model [@angrist2001]. This variable is commonly known as an exclusion restriction variable, and should only be associated with the selection in the sample $r^*_{ij}$ but not with the actual observation $y^*_{ij}$.

In the presence of selection bias, aforementioned outcome equation will yield biased estimates of $\beta^O_{i}$ if no efforts are made to adjust for the non-representativeness of the observed $X^O_{i}$ and $y_{i}$ values. For this reason, the Heckman model aims to jointly estimate the outcome and selection equation by defining a relation between their respective error distributions. For instance, Heckman’s original model [@heckman1976] assumes that residual terms have a bivariate normal distribution (BVN),

\[\begin{pmatrix}
\epsilon^O_{ij} \\
\epsilon^S_{ij}
\end{pmatrix}\sim N\left(\begin{pmatrix}
0 \\
0
\end{pmatrix},\begin{pmatrix}
\sigma^2_i & \rho_i\sigma_i \\
\rho_i\sigma_i & 1
\end{pmatrix}\right)
\]

where $\sigma_i$ corresponds to the variance of the error in the outcome equation and $\rho_i$ to the correlation between the error terms of the outcome and selection equations in the $i$-th cluster. As in a probit model, this model assumes a unit variance for the error term of the selection equation. The unit variance has no consequence on the observable values of $r_{ij}=\{0,1\}$, since they only depend on the sign of $r^*_{ij}$ and not on its scale.

The interpretation of $\rho_i$ is fairly straightforward. When $\rho_i=0$, the participation does not affect the outcome model and missing data can be considered MCAR (if data are missing completely at random) or MAR (if missingness is already explained by $x_{ij}^O$). Conversely, when $\rho_i\neq 0$, this suggests that data are MNAR.

## Heckman model estimation

Under the BVN assumption, the parameters of the Heckman model coefficients can be estimated using the two-step Heckman method (H2S) [@heckman1976] or the full information maximum likelihood method (FIML) [@amemiya1984]. However, both methods can lead to inconsistent estimators when this underlying distribution cannot be assumed[@gomes_etal20].To overcome this problem, other approaches have been proposed that relax the distribution assumptions, among them copula models [@smith03]. The copula approach uses a function, known as copula, that joins the marginal distribution of the error terms of the selection and outcome equation which are specified separately.
$$ F(r^*_{i},y^*_{i}) =\Phi\left(r^*_{i}-X^S_{i}\beta^S_{i},\frac{y^*_{i}-X^O_{i}\beta^O_{i}}{\sigma_i}, \rho\right)$$
Thus, to estimate the parameters of the Heckman method, it is sufficient to specify the marginal distributions of the error terms, and link them with a suitable copula function. In our imputation method we estimate the Heckman model using the copula method, as in real life many data could not follow a BVN distribution. This facilitates the Heckman model estimation and makes it more robust to deviations from the assumptions regarding the distribution of the data.

## Hierarchical model

The Heckman model can be extended to hierarchical settings, i.e., in data where individuals or sampling units are nested within groups, as is the case in EHR or IPD. In this case, sample units from the same group are expected to share some characteristics (e.g., distribution of variables, relationships between exposure and outcomes, missing processes, missing mechanisms) given only the fact that they belong to the same group.

Since most statistical analyses assume that the sample units are independent of each other, more complex hierarchical models dealing with nested data are required.  This hierarchical complexity must not only be taken into account in the observation process, but also in the missing data process, thus requiring imputation models that are congenial to the analysis model, i.e., that have the same assumptions about the data.

Different procedures can be adopted to combine information between groups; however, in our imputation method we opted for the two-stage approach that is often used in meta-analyses. This is because such an approach is less computationally intensive and could potentially generate fewer convergence problems in the estimation of the Heckman hierarchical model compared to other approaches. Briefly, in a first stage $\theta_i$ the cluster specific parameters of the Heckman model,i.e., the parameters of the two equations in each study: the outcome model and the selection model, for each of the $N$ clusters are estimated separately and then, in a second stage, all $\theta_i$ are combined using a random effects model.

In a random effects model, $\theta_i$ parameters are assumed to be drawn independently and identically from an imaginary distribution of parameters with a $\theta_m$ population mean and a $\psi$ population variance [@higgins2009]. Thus, the  $\theta_i=\theta_m + b_i$ can be specified using $b_i \sim N(0,\psi)$ random effects to allow for between-study heterogeneity in observed data relationships, and between-study heterogeneity in missing patterns.

# Method
\label{methods}
We follow a similar approach proposed by Resche-Rigon and White (2018) [@resche-rigon2018] for multilevel imputation of data. Briefly, his method was developed to impute variables from a hierarchical structure (i.e., when there are unit samples grouped within a cluster or group).
This method involves estimating an outcome equation (describing the relationships of the observed data to the missing variable) separately in each cluster, after which the parameter estimates of that equation are pooled using random-effects meta-analysis.

With this method, values can be imputed in very common scenarios in IPD, e.g., sporadic and systematic missing patterns. In particular, when the response variable is systematically missing within a group, i.e., when $y_{ij}$ are totally missing within a group, the imputation values are drawn from a (generalized) linear model conditional on $\theta_m$ the marginal population parameters, i.e., those estimated after pooling the cluster-specific parameters $\theta_i$.On the other hand, when the variable is sporadically missing within the group, i.e., there are some observed $y_{ij}$ within the cluster, the imputation model is conditional on the shrunk-cluster parameters, i.e., those coming from the shrinkage of $\theta_i$ towards $\theta_m$.

Our imputation approach differs crucially from the previous approach as here we estimate two correlated equations (instead of a single outcome equation) in each cluster , thus obtaining $\theta_i=\{\beta_i^O,\beta_i^S,\sigma_i,\rho_i\}$ parameters from both equations which are then pooled into a $\theta_m$ parameter set at the marginal or population level. Our method is basically a univariate imputation method, but since it is implemented in a Gibbs sampling procedure, it can also be used to impute multiple incomplete variables in a data set.

## Imputation of univariate incomplete dataset
Given an outcome variable $y=(y_1,y_2,...,y_N)^T$, that consists of $y^{miss}_{ij}$ missing and $y^{obs}_{ij}$ observable values, we generate independent draws from the posterior predictive distribution for the missing data, $y_{ij}^{miss}$, given the observable data information $y_{ij}^{obs}$.
$$p(y_{ij}^{miss}|y_{ij}^{obs})=\int_{\theta} p(y_{ij}^{miss}|\theta,y_{ij}^{obs})p(\theta|y_{ij}^{obs})\mathrm{d}\theta$$
Here we implicitly assume vague prior distributions for each of the parameters included in the parameter vector $\theta$.  Because the integration can be performed computationally by sampling from the posterior predictive distribution $p(\theta|y_{ij}^{obs})$, our imputation method can be carried out in the following two steps:

1. Draw a ${\theta}$ parameter vector, ${\theta^*}$, from $p(\theta|y_{ij}^{obs})$, their posterior distribution.

2. Draw $y_{ij}^{miss}$ from $p(y_{ij}^{miss}|\theta^*)$, their predictive distribution for a given $\theta^*$ vector.

Below we describe each step in depth:

### Draw the $\theta^*$ parameter vector

#### Fit $p(y_{ij}^{obs}|\theta_i)$, the heckman selection model at group level

Initially, we estimate by the copula method the set of cluster-specific parameters, $\widehat{\theta_i}=\{\widehat{\beta^O_i}, \widehat{\beta^S_i}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, using all $j$ units with $y_{ij}^{obs}$ observable measurements within each $i$ group . 
The Heckman model is estimated with the \textbf{gjrm} function of the GJRM R package under the bivariate model with the nonrandom sample selection (BSS) specification, from which we obtain not only the parameters point estimates $\widehat{\theta_i}$, but also their corresponding $\widehat{S(\theta_i)}$ within-cluster variance-covariance matrix.

<!-- When the model does not converge in a study, we assume that data follows a MAR mechanism so we fixed $\widehat{\rho_i} = 0$ and we estimate only the outcome equation parameters from a (generalized) linear model. -->

#### Fit a meta-analysis model
In this step, we pool the parameters $\widehat{\theta_i}$ with a random-effects meta-analysis model using only the groups with observable information, i.e., those that are not systematically missing and have sufficient information to estimate the heckman model.  In particular, we pooled the $p$ coefficients of the $\beta^O$ outcome equation and estimated a multivariate random effects meta-analysis model with them, similarly we combined all $q$ coefficient parameters of the $\beta^S$ selection equation. We also performed a univariate random effects meta_analysis on $\sigma'$ the log-transformed parameter of $\sigma$ and another on $\rho'$ the fisher-transformed parameter of $\rho$.

The meta-analysis model is performed with the \textbf{mixmeta} function of the R package mixmeta, which allows the use of maximum likelihood (ML), restricted maximum likelihood (REML) and moments estimation methods. For the simulation and illustrative study, we used the restricted REML estimation method, which is recommended as it has a good balance between insensitivity and efficiency [@viechtbauer2005].

#### Draw $\theta^*_m$ the marginal parameters

From the meta-analysis model, we obtain the marginal estimates $\widehat{\theta_m}$ and the between-cluster variance $\widehat{\psi}$ with their corresponding variance-covariance matrices $\widehat{S_{\theta_m}}$ and $\widehat{S_{\psi}}$, which are used to draw the $\theta^*_m$ and $\psi^*$ parameters as follows:

\begin{align*}
\theta_m^*& \sim N(\widehat{\theta_m},\widehat{S_{\theta_m}})\\
\psi^*& \sim N(\widehat{\psi},\widehat{S_{\psi}})
\end{align*}

#### Draw $\theta^*_i$ the cluster parameters

We draw $\theta^*_i$ shrunk-cluster parameters for each group $i$ from the following posterior distribution conditional on $\theta^*_m$ and $\psi^*$.
\begin{align*}
\theta^*_i\sim N\left(\frac{ \theta^*_m/\psi^*+\widehat{\theta_i}/\widehat{S_{\theta_i}}}{1/\psi^*+1/\widehat{S_{\theta_i}}},\frac{1}{1/\psi^*+1/\widehat{S_{\theta_i}}}\right)
\end{align*}

As can be seen, the mean and variance of the posterior distribution is a combination between the estimated marginal and cluster-specific parameters. Here the weights on the $\widehat{\theta_i}$ cluster-specific and the $\theta^*_m$ marginal parameters are inversely proportional to the $\widehat{S_{\theta_i}}$ within and $\psi^*$ between clusters variances. For example, when $\widehat{S_{\theta_i}} < \psi^*$ the mean of the conditional distribution gives more weight to the estimated cluster-specific parameter.
Conversely, when $\widehat{S_{\theta_i}}>\psi^*$, more weight is given to the estimated marginal parameters.  In the case of systematic missingness, it is like considering the within-cluster variance to be infinite ($\widehat{S_{\theta_i}}\to\infty$), then letting all parameters rely only on the marginal estimates.

### Draw $y_{ij}^{miss}$ observation

Having $\theta^*_i$ the shrunk-cluster parameters vector for each group, we back transform $\sigma^*$ and $\rho^*$ to the original scale. Then $y^{miss}_{ij}$ the missing values can be drawn from $p(y^{miss}_{ij}|\theta^*_i)$ their predictive distribution given $\theta^*_i$ as follows:

#### Continuous missing variable

The imputed value of the $y_{ij}^{miss}$ missing observation can be drawn from the conditional expectation of $y_{ij}$ on unobserved measurements:
\begin{align*}
\mu&= E[y_{ij}|r_{ij}=0,{\beta^O_{i}}^*,{\beta^{S}_{i}}^*,{\rho_{i}}^*,{\sigma_i}^*]\\
\mu&= x^O_{ij}{\beta^O_{i}}^*+\rho_{i}^*\sigma_i^*\frac{-\phi(x^{S}_{ij}{\beta^{S}_{i}}^*)}{\Phi(-x^{S}_{ij}{\beta^{S}_i}^*)}\\
y_{ij}^{miss}&\sim N(\mu,{\sigma_i^*}^2)
\end{align*}

#### Binary missing variable

The missing $y_{ij}^{miss}$ is drawn from a Bernoulli distribution with $p_{ij}^*$ proportion parameter given by $P[y_{ij}=1|r_{ij}=0]$, the conditional probability that $y_{ij}=1$ given that the measure is unobservable ($r_{ij}=0$), in a bivariate probit model [@greene2018]:
\begin{align*}
p^*_{ij} &= P[y_{ij}=1|r_{ij}=0,{\beta^O_{i}}^*,{\beta^{S}_{i}}^*,{\rho_{i}}^*]\\
p^*_{ij} &=\frac{\Phi_2(x^O_{ij}{\beta^{O}_i}^*,-x^S_{ij}{\beta^{S}_i}^*,-\rho_i)}{\Phi(-x^{S}_{ij}{\beta^{S}_i}^*)}\\
y_{ij}^{miss}&\sim Ber(p^*_{ij})
\end{align*}

## Imputation of multivariate incomplete dataset

When there are simultaneous missing variables in a dataset, our imputation method can be extended in a Gibbs sampler procedure.
Particularly our imputation method has been implemented according to the structure of the MICE R package  <!--[@buuren2021]-->, that allows imputing multiple incomplete predictors and covariates in a given dataset.

The MICE package allows to specify imputation methods to each of the missing variables by setting the method and the predictive matrix for each of the missing variables. To use our method, it is necessary to specify for the MNAR missing variable the method \textbf{2l.heckman} in the mice methods vector. Furthermore, in the prediction matrix, the group or cluster variable should be specified as \textbf{'-2'}, all predictor variables belonging to the selection and outcome as \textbf{'1'}, the exclusion restrictions or predictor variables that are only included in the selection equation as \textbf{'-3'} and those that are only included in the outcome equation as \textbf{'-4'}. Please refer to the toy example in the attached github repository to better understand how to implement the imputation model.

# Simulation study
#### Aim
We design a simulation study aimed to compare the performance of imputation methods when they impute a missing variable that comes from a hierarchical dataset and follows a MNAR mechanism.

#### Data-generation mechanism

We generated the data from the Heckman selection model with bivariate normal distribution error terms. For simplicity we assume that the database collects information from $N=10$ clusters of equal number of individuals $n_i=1000$. 
For each dataset, we generated $X_{1i}$ a treatment indicator variable from a Bernoulli distribution with a probability of treatment on each cluster equal to 0.6. Next we simulated the mean of two continuous covariates from a multivariate normal distribution $\mu_h\sim N(\begin{psmallmatrix}0\\0\end{psmallmatrix},\begin{psmallmatrix}0.2 & 0.015\\0.015 & 0.2\end{psmallmatrix})$, with $h=\{2,3\}$. We then simulated for each cluster a baseline covariate $X_{2i}\sim N(\mu_2,1)$ and a exclusion restriction  $X_{3i} \sim N(\mu_3,0.5^2)$.

Here, we considered $X_{1i}$ and $X_{2i}$ as predictors in the outcome equation $X_i^O=[1,X_{1i},X_{2i}]$. For the selection equation we included both variables and the $X_{3i}$ exclusion restriction, $X_i^S=[1,X_{1i},X_{2i},X_{3i}]$. Then in case of a missing continuous variable, we calculate the latent variables  $y_{i}^*$ and $r^*_{i}$ as follows:
\begin{align*}
 y_{i}^*&= \beta_i^OX_i^O+\epsilon^O_{i} \\
 r_{i}^*&= \beta_i^SX_i^O+\epsilon^S_{i}
\end{align*}
Here we assumed that all coefficient parameters varied across studies, by including cluster-specific random effects as:
\begin{align*}
 \beta_{hi}^O&= \beta_h^O + b_{hi}^O\\
 \beta_{hi}^S&= \beta_h^S + b_{hi}^S
\end{align*}
We used these fixed coefficients $\beta_h^O=\{2,1,1\}$  and $\beta_h^S=\{0.1,1.5,-0.7,1.8\}$ in order to get around 40\% of sporadically missing values on the response $y_{ij}$ in the entire data set. Additionally, we made that the $y_{ij}^*$ observations were systematically missing in 20% of the clusters included in data set.
We assumed that random effects were independent within equations ($b_{h0}^O\indep b_{h1}^O \indep b_{h2}^O$ and $b_{h0}^S \indep b_{h1}^S \indep b_{h2}^S$), but were linked between both selection and outcome equations through a bivariate normal distributed as: 
\[\begin{pmatrix}
b^O_{h} \\
b^S_{h}
\end{pmatrix}\sim N\left(\begin{pmatrix}
0 \\
0
\end{pmatrix},\sigma_{bh}^2\begin{pmatrix}
1 & \rho*0.4 \\
\rho*0.4 &1 
\end{pmatrix}\right)
\]
with the parameters $\sigma_{b0}^2=\sigma_{b1}^2=\sigma_{b2}^2=0.5$.  We considered that the correlation parameter of the random effects between equations is a 40% of the value of the assumed correlation parameter between error terms $\rho$. In addition, we included a random effect on the exclusion restriction variable given by $b_3\sim N(0,0.3)$ assuming that the intracluster variation in the exclusion restriction effect is lower than the variation on other coefficient parameters effects. The $\rho$ parameter adopts different values depending on the simulated missing mechanism (See below additional scenarios) .

As regards the error terms, they were bivariate normal distributed as:
\[\begin{pmatrix}
\epsilon^O_{i} \\
\epsilon^S_{i}
\end{pmatrix}\sim N\left(\begin{pmatrix}
0 \\
0
\end{pmatrix},\begin{pmatrix}
\sigma^2_i & \rho\sigma_i \\
\rho\sigma_i & 1
\end{pmatrix}\right)
\]
whose $\sigma_i^2$ is variable across clusters and distributed as $log(\sigma_i) \sim N(log(0.7),0.05)$.

##### Adittional scenarios

To investigate the performance of the imputation methods under the following scenarios:

- \textbf{M(N)AR scenarios:} We assessed whether the model performed well in terms of bias and coverage when the data followed a missing MAR mechanism ($\rho=0$), and when it followed a MNAR mechanism with a low ($\rho=0.3$), intermediate ($\rho=0.6$) and strong correlation ($\rho=0.9$) between $y^*$ and $r^*$.
- \textbf{Influence of sample size}: Model sensitivity was analysed with respect to $n_i=\{50,100,1000\}$ the number of patients per cluster and $N=\{10,50,100\}$ the number of studies. We consider the 
- \textbf{Violation of distributional assumptions:} To assess how imputation models behave in the face of deviations from normality assumptions, we simulated data in which the errors followed a skewed t-distribution and also in which the missing process follows a MNAR mechanism with an explicit truncated model, i.e., the participation is directly related to the value of the outcome variable. 
- \textbf{Binary response}: We evaluated the imputation method when missing variables is binary. Therefore we simulated  $y_{i}$ binary incomplete variables, the marginal set of parameters were fixed to  $B_h^O=\{0.5,1,1\}$ and  $B_h^S=\{0.1,1.5,-0.7,1.2\}$, keeping the other parameters similar to the ones used in the simulation of missing continuous variables. The observable binary variables was defined as:
\begin{align*}
    r_{i}&=I(r*_{i}>0)\\
    y_{i}&=I(y^*_{i}>0) \forall r^*_{ij}>0
\end{align*}

#### Estimand
The estimads were the parameter coefficients of the outcome equation $\beta^O={\beta^O_0,\beta^O_1,\beta^O_2}$, with special emphasis on the treatment effect parameter $\beta^0_1$. We also report in the estimated variance of the random effects and residual errors $\sigma_{b0}^2$,$\sigma_{b1}^2$,$\sigma_{b2}^2$,$\sigma_{e}^2$.

After the imputation procedure, we estimated the following (generalized) mixed linear effect model using the lmer() function.
$y_{i}=\beta_i^OX_i^O+\epsilon^O_{i}$
In case of missing binary variable, we used the same matrix of predictors but on a binary model estimated with glmer() functions. Then, we pooled the estimates of the $\beta_i^O$ and the variance of random effect and residual errors of the multiple imputed datasets according to Rubin's rule [@rubin1987], over which we calculated the performance measures on estimands.

#### Method

For each scenario we simulated 500 datasets over which we evaluated the following imputation methods:

- \textbf{Complete case analysis (CCA):} We removed all patients with missing observations. 
- \textbf{1l.Heckman:} Multiple imputation based on the Heckman model without no study specification, following the imputation method proposed by Galimard et al.(2016) [@galimard_etal16].
- \textbf{2l.MAR}: Multiple imputation assuming MAR for hierarchical datasets, we used the multilevel imputation model proposed by Resche-Rignon and (2018) [@resche-rigon2018].
- \textbf{2l.Heckman}: The proposed imputation method based on the Heckman model for hierarchical datasets.

#### Performance measures

We calculated the following evaluation criteria according to the formulas provided in Morris et al.(2019)[@morris2019]:

- \textbf{Bias:} Bias on the coefficient and random effect parameters.
- \textbf{EmpSE:} Empirical standard errors of the estimates on the coefficient and random effect parameters.
- \textbf{ModSE:} Monte Carlo standard errors on the coefficient  parameters.
- \textbf{MSE:} Mean squared error of the coefficient  and random effect parameters.
- \textbf{Coverage(\%):} Coverage of the 95\% confidence intervals for the coefficient  parameters.
In addition we reported the average time processing and the percentage of datasets where the imputation method converged, i.e., the imputation method generated an output.

#### Software
For simulation study and illustrative examples we used  R version 4.0.4  in a linux environment.
The Heckman 2L imputation method is available  in mice R package (\textbf{mice.2l.heckman}) and also on the github repository \url{https://github.com/johamunoz/Heckman2l} where you can also find all the codes accompanying this paper and a toy example that explains how to implement the method in mice.

#### Results
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# ```{r, include=FALSE}
# library(here)
# source(here("2.Simulation", "simplot.R"),local =knitr::knit_global())
# ```

##### Descriptive results
We generated data sets of 10 groups of 1000 patients each in each scenario. For example, for the scenario in which the error terms followed a normal distribution, out of the 500 datasets generated, we obtained that on average 48.2\% of the Y response was missing, with the lowest missing percentage being 26.84\% and the maximum being 73.67\%.  At the cluster level, looking only at the sporadic missing clusters, we found an average of 35.25\% missing values, but there were clusters with no missing data at all and up to 98\% missing data in Y.

##### Results M(N)AR scenarios

# ```{r rhocont,fig.pos="h",fig.cap="\\label{fig:rhocont} Comparison of methods for continuous incomplete variable under systematical missingness by varing $\rho$, the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"}
# plot_cont_rho
# ```

Figure \ref{fig:rhocont} shows the results of simulations where the missing variable was continuous.In the MAR scenario, i.e. when $\rho=0$, all imputation methods provide similar unbiased estimates of the coefficient parameters, but as rho increases, i.e. the mechanism becomes MNAR, the estimates for the complete case analysis and the MAR-IPD imputation method turns more biased.
Overall, both Heckman-based imputation (1l. and 2l.) models provide less biased estimates of the coefficient parameters in MNAR scenarios, but the random effects estimates on the 1l.heckman are far away from the true values as no cluster information was considered at all. The precision of the Heckman IPD separated seems it is better than the Heckman IPD full version, specially in the MAR scenarios, when we checked simulation results it seems that the high variability on the full version is related to simulations in which singularity fit that it is an indication that the model is overfitted. Regarding the coverage, the Heckman-IPD imputation method provides the best coverage values across all the $\rho$ scenarios with a coverage level up to 95%.

The Heckman-IPD method, by incorporating individual and marginal level information, allows a bias-variance trade off, thus, as it was observed in Table \ref{tab:rhotable} the Heckman IPD separated resulted in better estimates in terms of RMSE than the other methods evaluated.
In particular, the method provides an advantage over the Heckman-single method in systematic missingness scenarios, since the latter does not allow to specify any cluster information implicitly, i.e. by adding a group variable in the imputation model. This can be seen in the estimates of the random effects parameters of the Heckman-single model, which are more biassed than those of the other methods in which cluster information was included.
Moreover, the method could be affected by the ratio of the pooled data (See Annex Figure\ref{fig:continuous3}). To demonstrate this, we increased the variance of the random effects in the simulated data, reducing the intra-cluster correlation (ICC) from 0.33 to 0.23. This resulted in less biassed estimates for both coefficient and random effects parameters, along with narrower confidence intervals.

##### Results sensitivity to number of clusters and sample size of clusters
We evaluated how robust our method was to variations in the sample size of the studies included in the IPD and also the number of clusters.
Our method allows for the inclusion of studies with different sample sizes, but to facilitate the evaluation of the method, in this simulation study we used 10 studies each with a fixed sample size of 1000 observations.

<!-- ```{r rhosize,fig.pos="h",fig.cap="\\label{fig:size} Comparison of methods for continuous incomplete variable under systematical missingness by varing (number of studies- sample size per study), the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"} -->
<!-- figure_final_s -->
<!-- ``` -->

When we reduced the sample size to 50 patients per study (left-most panel 10-50, Figure \ref{fig:rhosize}), we observed that the bias could be affected by some of the coefficient parameters, especially in MNAR scenarios. All simulated data were successfully imputed with the Heckman IPD methods but we observed that the precision of the parameters was drastically affected especially for the Heckman IPD full method, due to singularity fitting problems in the estimation of the mixed model, so we prefer not to show it in the graph.

On the other hand, by increasing the number of studies to 100 (rightmost panel 100-1000, Figure \ref{fig:rhosize}) we observe a higher precision of the estimates. There was an increase in the coverage of the estimates in the MAR scenario, i.e. rho=0, as the number of studies increases, which is in line with Audigier's results. However, in the case of MNAR, i.e. rho=0.6, we unfortunately observed a decrease in coverage, which could be explained by an increased complexity of the Heckman model.

Furthermore, we observed that for the full Heckman IPD method, the number of successfully imputed datasets was only 90% of the simulated datasets. This could be explained by the fact that increasing the number of studies included in the analysis might make it more difficult to estimate the marginal parameters when applying the meta-analysis model with random effects, especially when considering maximum likelihood or restrictive maximum likelihood as the estimation method. To ensure the estimation of the marginal parameters, our method evaluates the marginal parameters using the method of moments when it is not possible to estimate them with other methods chosen by the user.  We use the method of moments estimation as an alternative estimation method, as it was previously evaluated by Audigier, and was found to provide robust results when estimating the marginal parameters.

<!-- ##### Results deviations in distribution assumptions -->

<!-- ```{r dist,fig.pos="!H",fig.cap="\\label{fig:dist} Comparison of methods for continuous incomplete variable with deviations in distribution assumptions, the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"} -->
<!-- plot_cont_dist -->
<!-- ``` -->

<!-- To investigate how the imputation models behaves in settings with departures of binormal distribution, we draw error terms from a bivariate skew student-t distribution using the same location parameter and covariance matrix of the normal distributed settings,  with 8 degrees of freedom and an $\alpha=\{-5,5\}$ parameter which regulates the the slant of the density. In addition we simulated a explicit missing process, where error terms of selection and outcome equation were normally independent distributed and the inclusion depends on the value of the missing variable, as follows: -->
<!-- $ry_i=0.1y_i+\epsilon^S_{i}$ -->

<!-- These settings assure that the percentage of missingness is similar in all the evaluated scenarios around 60\%. We observed that for these simulated settings. -->

<!-- Figure \ref{fig:dist} depicts the results of this scenario. When we use the Heckman model in an explicit MNAR missing process, we observe that our model might not be completely suitable for this type of scenario, as we can observe that it affects the error variance which mainly affects the bias in the intersection coefficient parameter. and also its coverage. With respect to the t-distributed error scenario, apparently for the simulated measurements there is not much variation in the estimation of the coefficient parameters, however, we also observe that the estimation of the error variance was affected. In this scenario it is also seen that the coverage in the intercept parameter varies.  Similar variations occurred in the imputations performed with the simple Heckman imputation method, which is in line with the findings of @liu2020a. -->

<!-- ##### Bivariate incomplete variable -->
<!-- We also evaluate the performance of the imputation methods for bivariate missing variables (Figure \ref{fig:bin}). We also find that the Heckman-IPD separate method provides unbiased results on coefficient parameters and random effects with better convergence compared to the other imputation method evaluated. However, we observe that the accuracy of the random parameter effects is sensitive to $rho$ the correlation between the outcome and the selection equation, resulting in more imprecise estimators as the $rho$ parameter increases. -->

<!-- ```{r bin,fig.pos="!H",fig.cap="\\label{fig:bin} Comparison of methods for binary incomplete variable under systematical missingness, the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"} -->
<!-- plot_bin -->
<!-- ``` -->

# An ilustrative study
\label{example}
```{r, include=FALSE}
library(here)
source(here('3.Ilustrative_study','Ilustrative_study.R'),local = knitr::knit_global())
```

Malaria is a mosquito-borne disease and, especially in children and pregnant women, is the leading cause of illness and death in Africa. To prevent the spread of the disease, long-lasting nets (LLINs) and indoor residual spraying (IRS) in at-risk households are used as control measures.

Specifically, in Uganda, under the Uganda LLIN evaluation project, a LLINS distribution campaign was conducted between 2013 and 2014, and in 2017, the effect of LLIN control together with insecticides was assessed through a cross-sectional community survey in 104 health sub-districts in 48 districts located within 5 sub-regions of Uganda.

In each sub-district, a sample of households with at least one child aged 2-10 years were surveyed, where information was collected on household conditions and use of preventive measures. In addition, finger prick blood samples were taken from each child to determine the prevalence of parasitaemia and an etymological study was conducted to estimate mosquito prevalence, details of the project and survey are provided elsewhere [@staedke2019].

For this example, we used data accessed directly from CliniEpiDB [@staedke], where data were collected from 5195 households with verified consent, inhabited by 11137 residents aged 2-10 years. Blood samples were only taken from 8846 children, as 69 were excluded from the study due to lack of consent and 2222 were not present at the time of the survey. Although the original data set consists of 164 variables, here we only consider the variables described in the table \ref{tab:descriptive}, which were used as predictors of the imputation model.

```{r descriptive, results = "asis", table.pos="!H", echo=FALSE}
if (!require("xtable")) install.packages("xtable")
xt <- xtable(sum.table,
             caption = "Descriptive analysis, predictor variables", label = "tab:descriptive",align= c("p{0.0\\textwidth}|", 
                                   "p{0.11\\textwidth}|",                       
                                   "p{0.07\\textwidth}|", 
                                   "p{0.07\\textwidth}|", 
                                   "p{0.05\\textwidth}|", 
                                   "p{0.10\\textwidth}|", 
                                   "p{0.12\\textwidth}|", 
                                   "p{0.07\\textwidth}|", 
                                   "p{0.07\\textwidth}|", 
                                   "p{0.07\\textwidth}|", 
                                   "p{0.07\\textwidth}|"))
names(xt) <- c("Subregion","Districts (N)","Children (N)","Age (years)",
                "Log10 Female Anopheline","Wealth index",
                "Bednet (%)","Females (%)","Holidays (%)",
                "Missing test(%)")
print(xt, comment = FALSE, include.rownames=FALSE)
```

To illustrate our method, following the article by @rugnao2019, we estimated the prevalence of parasitemia by subregion and by age after approximately 3 years of LLIN campaigns started.  We estimated parasitemia prevalence using 3 approaches: MCAR, MAR and MNAR.

In the MCAR approach, prevalence was calculated on the basis of the actual test, i.e., we only included patients with a test result. In the case of MAR, the test values of children who were not present during the survey were imputed with the 2l.2stage.bin method of the MICEMD package, where the community was taken as the cluster and the following factors previously associated with parasitemia were used as predictors in the imputation model: sex, two-person mosquito net.  In addition, age was included as a power 3 spline function, the cluster-level Log10 mean of the number of female anopheline mosquitoes per household estimated from the etymological survey, and the household wealth index from principal components analysis calculated specifically for the surveyed households.

Under the MNAR assumption, we used the proposed 2l.Heckman method to impute missing test values.  The selection and outcome equation included the same predictor variables used under the MAR approach. 
In addition, we included a holiday indicator variable as ERV, which was calculated according to school vacation calendars and public holidays in Uganda in 2017. We examined the association of this ERV with the outcome variable (y) and with the selection indicator (ry), conditionated on the remaining imputation predictors. The model results in Table \ref{tab:exclusion} indicate that the holiday indicator could be a plausible ERV variable, as it was significantly associated with ry, but not with y.

```{r ERV, results = "asis", table.pos="!H", echo=FALSE}
if (!require("xtable")) install.packages("xtable")
xt <- xtable(tablexc,
             caption = "Evaluation of Holidays as exclusion restriction variable", label = "tab:exclusion")
names(xt) <- c("","Test result (y)"," Missing indicator (ry)")
comment<- list( pos=list(0),command=NULL)
comment$pos[[1]]<-c(nrow(xt))
comment$command<-c(paste("\\hline",
                         "***p<0.01",sep=" "))

print(xt, comment = FALSE, include.rownames=FALSE,add.to.row = comment, 
      hline.after = c(-1,0))
```
According to our imputation approach, non-participants were estimated to have a higher prevalence of malaria than participants in more than half of the districts analyzed. 
As can be seen in Figure \ref{fig:region} in terms of subregion level prevalence the estimates of the approaches do not differ significantly between methods, however except for the East-Central region, prevalence estimates under the Heckman approach are higher than those estimated using MAR or CC aproaches. 
```{r region,fig.pos="!H",fig.cap="\\label{fig:region} Estimates of malaria prevalence by Uganda subregion",echo=FALSE,fig.keep="last"}
plot_dist
```

In terms of prevalences by age, there are no significant differences between methods overall. The prevalence estimates for children aged 2 to 6 years according to the approaches evaluated in all regions are very similar (Figure \ref{figure:region_age}, which could be partly explained by the mobility of children at this age compared to that of school-age children. 
```{r region_age,fig.pos="!H",fig.cap="\\label{fig:region_age} Estimation of malaria prevalence by region and age",echo=FALSE,fig.keep="last"}
plot_sdist_age
```

However for school children, prevalences estimated with the Heckman method were found to be higher in the Mid-East and Southwest regions than those obtained with the other methods, whereas in the East-Central region the estimates with the Heckman method are lower. Possible reasons for selection bias in surveys of this type have been suggested, for example, that daytime visits might favor measurement in sick school children who stay home, leading to overestimated prevalence results as found in the East-Central region. Nevertheless, we were unable to find information that suggests or confirms the direction in which malaria prevalence is driven by selection bias in this Uganda study or in other studies similar to this one.

# Discussion
\label{discussion}
This work was done in order to extend multiple imputation for clustered datasets, in situations where some incomplete variables follow a MNAR mechanism.
For clustered datasets, only imputation methods under the MAR mechanism have been previously proposed and although imputation methods exist to handle MNAR they have only been designed for individual studies, which makes them limited in common IPD situations such as systematic missingness or when the proportion of missingness of a variable is very high in one of the included studies.   In this context, a new multiple imputation method was proposed to handle continuous and binary MNAR covariates specifically for clustered dataset, which also allows appropriate borrowing of information between the clusters to obtain more reliable imputation results at the individual cluster level.

From the results of the simulations we can observe that the imputation method we propose can be valid for the imputation of continuous and binary type missing variables that follow a MNAR mechanism according to the Heckman model and that come from multilevel data such as those used in the IPDMA studies.

Overall the method produced unbiased estimates with convergence close to 95\% for the fixed effects parameters with variation at the cluster level and also unbiased estimators for the random effects parameters.

Empirically, the method was shown to be robust to systemic and sporadic missingness in individual studies. This method, in particular, could provide more robust imputation values compared to individual-level imputation methods, as it not only allows for imputation of missing values in clusters with systematic losses, but can also shrink the values of individual clusters towards the global mean of studies, particularly advantageous in studies with extreme values or with values far away from those found on average at the global level.

The advantage of the proposed method over the previous ones is that it allows the imputation of variables from cluster level data following a MAR or MNAR mechanism according to Heckman's model. That is to say that under the specification of a validity exclusion variable the method determines by itself which is the most adjustable  correlation parameter between equations ($\rho$), or in general terms the  missingness mechanism  (MAR or MNAR), in each of the clusters evaluated.
The imputation method is built on the mice package, which allows, first of all, to be used both on the outcome and on the covariates and, in addition, offers the option of being used simultaneously with other imputation methods of the package, advantageous in databases containing missing variables with different prediction methods and models. Finally, the method can be used on systematically and sporadically missing clusters, both for continuous variables with heterogeneous error variance and for binary variables.

One of the major limitations of our method is that it needs a valid restriction variable, which in some contexts is difficult to establish at the individual study level and can be even more challenging if one tries to find a valid exclusion variable across clusters.
Also the method is sensitive to the value of the correlation between the selection equation and outcome (rho), and in general it is observed that it can lead to biassed results on fixed global parameters i.e. without variation across clusters.Similarly, the method can be sensitive to both the sample size and the number of studies included in the database. On the one hand a small sample size at the individual study level can affect not only the precision of estimates but also the convergence of the method since it requires a minimum sample size to estimate all the parameters of the Heckman model which can be at least twice the number of parameters required in a MAR prediction model. On the other hand, a high number of studies that may improve the precision of the estimators may also make the estimation of the marginal parameters more difficult and also considerably increase the processing time of our method.

The data were simulated by attributing a constant correlation across all clusters in order to evaluate the performance against M(N)AR assumptions, but in practice this parameter is variable across clusters which can considerably affect the performance of the method. So in future research the effect of this parameter could be more deeply evaluated.

On the other hand, the method can also be extended to other types of variables such as count or ordinal variables. Similarly, less restrictive Heckman based models can be considered in terms of normality distribution of errors and no specification of exclusion variables such as those proposed by Ogundimu.

### 2022-10-24 2.Simulation/simplot.R

@@ -1,0 +1,141 @@
+library(here)
+library(data.table)
+library(ggplot2)
+
+# Arrange dataset
+data<-as.data.table(readxl::read_xlsx(here('2.Simulation','Sim_results.xlsx'),sheet="Sheet1"))
+head(data)
+data<-data[,grep("Scenario|^Ns|^ni|Method|Variation|time|run|^Cov|^Qbar|^Re|^Emp|^Bias|^MSE|^MCMSE|^Mod", names(data)), with = FALSE]
+colnames(data)<-c("Scenario","Ns","ni","Variation","Method",paste0("Mean_",0:6),"time", paste0("Cov_",0:2), paste0("Bias_",0:6),
+                  paste0("MSE_",0:6),paste0("ModSE_",0:2),"run",paste0("EmpSE_",0:6),paste0("MCMSE_",0:6))
+data<- setDT(melt(data, measure.vars = c( paste0("Mean_",0:6), paste0("Cov_",0:2),paste0("Bias_",0:6),paste0("MSE_",0:6),paste0("ModSE_",0:2),paste0("EmpSE_",0:6)
+                                          ,paste0("MCMSE_",0:6)),
+                  variable.name = "Statistic", value.name = "Value"))
+data[, c("Statistic", "Parameter") := tstrsplit(Statistic, "_", fixed=TRUE)]
+data<- setDT(dcast(data, Scenario + Ns + ni+ Variation+ Method + time + run + Parameter ~ Statistic, value.var = "Value"))
+data[,SeCov:=sqrt(Cov*(1-Cov)/run)]
+data[,Coverage:=Cov]
+data[,SEBias:=EmpSE/run^0.5]
+data[,Per_run:=run/5]
+param_val <- c("0"="beta_0", "1"="beta_1", "2"="beta_2", 
+               "3"="sigma_b0", "4"="sigma_b1", "5"="sigma_b2","6"= "sigma_e")
+
+data[,Parameter := as.factor(param_val[data$Parameter])]
+
+unique(data$Scenario)
+datagraph<-data[Scenario=="Rho",c("Scenario","Variation","Method","Parameter","Bias","Cov","MSE","SEBias","SeCov","MCMSE")]
+colnames(datagraph)<-c("Scenario","Variation","Method","Parameter","p_Bias","p_Cov","p_MSE","SE_Bias","SE_Cov","SE_MSE")
+
+datagraph<-data.table::melt(setDT(datagraph), 
[…]

### 2022-10-24 3.Ilustrative_study/Additional_functions.R

@@ -1,0 +1,102 @@
+
+# 1. Functions to pool estimates first graph ----
+
+logit <- function(x) {
+  log(x/(1-x))
+}
+inv.logit <- function(x) {
+  1/(1+exp(-x))
+}
+inv.logit.SE<- function(logit.se, c) {
+  logit.se * (c*(1-c))
+}
+
+#Function for calculating absolute risk
+f.prevalence <- function(data,outcome_name) {
+  a<-summary(data[,get(outcome_name)])
+  b<-prop.test(x = a[[2]], n = sum(a), correct = TRUE)
+  prop<-b$estimate[[1]]
+  ci<-b$conf
+  return(list(prevalence= prop,ci.l=ci[1], ci.u=ci[2]))
+}
+
+
+#Function to calculate absolute prevalence of an outcome per study and per imputed dataset
+f.abs.perstudy<-function(data, outcome_name,study_name) {
+  
+  #Calculate absolute prevalence of outcome per study
+  inc.outcome<-setDT(data)[, f.prevalence(data=.SD, outcome_name=outcome_name), by = list(get(study_name),.imp)]
+  
+  #Replace 0's by 0.000001
+  inc.outcome$prevalence[inc.outcome$prevalence==0]<-0.000001
+  inc.outcome$ci.l[inc.outcome$ci.l==0]<-0.000001
+  
+  #Logit transformation
+  inc.outcome[,logit.prevalence:=logit(prevalence)]
+  inc.outcome[,logit.ci.l:=logit(ci.l)]
+  inc.outcome[,logit.ci.u:=logit(ci.u)]
+  inc.outcome[,logit.se:=(logit.ci.u-logit.ci.l)/(2*1.96)]
+  inc.outcome
+}
+
+#Function to pool the absolute prevalence per study per imputed dataset using rubins rules
+#Results in absolute risk per study
+
+f.pool<-function(data,m){
+  logit.abs <- mean(data$logit.prevalence)
+  within <- mean(data$logit.se^2)
+  between <- (1 + (1/m)) * var(data$logit.prevalence)
+  logit.var <- within + between
+  logit.se <- sqrt(logit.var)
+  logit.ci.lb <- logit.abs + qnorm(0.05/2)     * logit.se
+  logit.ci.ub <- logit.abs + qnorm(1 - 0.05/2) * logit.se
[…]

### 2022-10-24 3.Ilustrative_study/Ilustrative_study.R

@@ -1,0 +1,388 @@
+rm(list=ls())
+library(here) 
+library(data.table)
+library(micemd)
+library(mice)
+library(gamm4)
+library(tidymv)
+library(ggplot2)
+
+
+
+#0. Additional required function ----
+# 0.1 Functions to pool estimates first graph ----
+logit <- function(x) {log(x/(1-x))}
+inv.logit <- function(x) {1/(1+exp(-x))}
+inv.logit.SE<- function(logit.se, c) {logit.se * (c*(1-c))}
+
+#Function for calculating absolute risk
+f.prevalence <- function(data,outcome_name) {
+  a<-summary(data[,get(outcome_name)])
+  b<-prop.test(x = a[[2]], n = sum(a), correct = TRUE)
+  prop<-b$estimate[[1]]
+  ci<-b$conf
+  return(list(prevalence= prop,ci.l=ci[1], ci.u=ci[2]))
+}
+
+
+#Function to calculate absolute risk of an outcome per study and per imputed dataset
+f.abs.perstudy<-function(data, outcome_name,study_name) {
+  
+  #Calculate absolute risk of outcome per study
+  inc.outcome<-setDT(data)[, f.prevalence(data=.SD, outcome_name=outcome_name), by = list(get(study_name),.imp)]
+  
+  #Replace 0's by 0.000001
+  inc.outcome$prevalence[inc.outcome$prevalence==0]<-0.000001
+  inc.outcome$ci.l[inc.outcome$ci.l==0]<-0.000001
+  
+  #Logit transformation
+  inc.outcome[,logit.prevalence:=logit(prevalence)]
+  inc.outcome[,logit.ci.l:=logit(ci.l)]
+  inc.outcome[,logit.ci.u:=logit(ci.u)]
+  inc.outcome[,logit.se:=(logit.ci.u-logit.ci.l)/(2*1.96)]
+  inc.outcome
+}
+
+#Function to pool the absolute risks per study per imputed dataset using rubins rules
+#Results in absolute risk per study
+
+f.pool<-function(data,m){
+  logit.abs <- mean(data$logit.prevalence)
+  within <- mean(data$logit.se^2)
+  between <- (1 + (1/m)) * var(data$logit.prevalence)
+  logit.var <- within + between
+  logit.se <- sqrt(logit.var)
+  logit.ci.lb <- logit.abs + qnorm(0.05/2)     * logit.se
+  logit.ci.ub <- logit.abs + qnorm(1 - 0.05/2) * logit.se
[…]

### 2022-10-24 4.Codes/mice.impute.2l.heckman.R

@@ -1,0 +1,500 @@
+#' Imputation based on Heckman model for multilevel data.
+#'
+#' Imputes outcome and predictor variables that follow an MNAR mechanism
+#' according to Heckman's model and come from a multilevel database such as
+#' individual participant data.
+#' @aliases mice.impute.2l.heckman 2l.heckman
+#' @param y Vector to be imputed
+#' @param ry Logical vector of length \code{length(y)} indicating the
+#' the subset \code{y[ry]} of elements in \code{y} to which the imputation
+#' model is fitted. The \code{ry} generally distinguishes the observed
+#' (\code{TRUE}) and missing values (\code{FALSE}) in \code{y}.
+#' @param x Numeric design matrix with \code{length(y)} rows with predictors for
+#' \code{y}. Matrix \code{x} may have no missing values.
+#' @param wy Logical vector of length \code{length(y)}. A \code{TRUE} value
+#' indicates locations in \code{y} for which imputations are created.
+#' @param type type of the variable in the prediction model {0: No predictor,
+#' 1: Predictor in both the outcome and selection,-2: Cluster id (study id),
+#' -3: Predictor only in the selection model, -4: Predictor only in the outcome 
+#' model}
+#' @param pmm  predictive mean matching can be applied only for for missing continuous variables: "TRUE","FALSE"
+#' @param meta_method meta_analysis estimation method for random effects :
+#' "ml" (maximum likelihood), "reml" (restricted maximum likelihood) or "mm"
+#' method of moments.
+#' @param ... Other named arguments. Not used.
+#' @name mice.impute.2l.heckman
+#' @return Vector with imputed data, of type binary or continuous
+#' @details Imputes systematically and sporadically missing binary and continuous
+#'  univariate variables that follow a MNAR mechanism according to the Heckman 
+#'  selection model and come from a clustered dataset. 
+#'  The imputation method uses a two-stage approach in which the Heckman model 
+#'  parameters at the cluster level are estimated using the copula method.
+#'
+#' @note
+#'Missing binary variables should be included as two-level factor type variables. 
[…]

### 2022-10-24 1.Paper/Paper.Rmd
update pdf

@@
 
-Registries are notoriously prone to incomplete variables that are MNAR, due to the complex recording process <!--[@liu2021]-->, e.g.
+Registries are notoriously prone to incomplete variables that are MNAR, due to the complex recording process [@liu2021], e.g.
 laboratory tests are taken only in certain patients based on sign and symptom information that is often incomplete or not recorded.
@@
 
-In clustered data sets, multilevel imputation methods are needed to properly propagate uncertainty within and across clusters [audigier_etal18].
+In clustered data sets, multilevel imputation methods are needed to properly propagate uncertainty within and across clusters [@audigier_etal18].
 However, to our knowledge, existing multilevel imputation methods mainly focus  on situations where data are MAR, and do not adopt Heckman selection models.
@@
 # Simulation study
+\label{simulation}
+
 #### Aim
@@
 
+For coverage estimation, we obtain the confidence intervals from the Wald method.
+For the random effect variance, even when it is possible to estimate the confidence intervals through the profile or bootstrap method, we did not estimated them for computational reasons.
+
 #### Method
@@
 #### Results
+
 ```{r setup, include=FALSE} 
@@
 
-# ```{r, include=FALSE}
-# library(here)
-# source(here("2.Simulation", "simplot.R"),local =knitr::knit_global())
-# ```
+<!-- ```{r, include=FALSE} -->
+<!-- library(here) -->
+<!-- source(here("2.Simulation", "simplot.R"),local =knitr::knit_global()) -->
+<!-- ``` -->
 
@@
 
-##### Results M(N)AR scenarios
+<!-- ##### Results M(N)AR scenarios -->
 
-# ```{r rhocont,fig.pos="h",fig.cap="\\label{fig:rhocont} Comparison of methods for continuous incomplete variable under systematical missingness by varing $\rho$, the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"}
-# plot_cont_rho
-# ```
+<!-- # ```{r rhocont,fig.pos="h",fig.cap="\\label{fig:rhocont} Comparison of methods for continuous incomplete variable under systematical missingness by varing $\rho$, the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"} -->
+<!-- # plot_cont_rho -->
+<!-- # ``` -->
 
-Figure \ref{fig:rhocont} shows the results of simulations where the missing variable was continuous.In the MAR scenario, i.e.
+<!-- Figure \ref{fig:rhocont} shows the results of simulations where the missing variable was continuous.In the MAR scenario, i.e.
 when $\rho=0$, all imputation methods provide similar unbiased estimates of the coefficient parameters, but as rho increases, i.e.
 the mechanism becomes MNAR, the estimates for the complete case analysis and the MAR-IPD imputation method turns more biased.
-Overall, both Heckman-based imputation (1l.
+-->
+<!-- Overall, both Heckman-based imputation (1l.
 and 2l.) models provide less biased estimates of the coefficient parameters in MNAR scenarios, but the random effects estimates on the 1l.heckman are far away from the true values as no cluster information was considered at all.
@@
 Regarding the coverage, the Heckman-IPD imputation method provides the best coverage values across all the $\rho$ scenarios with a coverage level up to 95%.
+-->
 
-The Heckman-IPD method, by incorporating individual and marginal level information, allows a bias-variance trade off, thus, as it was observed in Table \ref{tab:rhotable} the Heckman IPD separated resulted in better estimates in terms of RMSE than the other methods evaluated.
-In particular, the method provides an advantage over the Heckman-single method in systematic missingness scenarios, since the latter does not allow to specify any cluster information implicitly, i.e.
+<!-- The Heckman-IPD method, by incorporating individual and marginal level information, allows a bias-variance trade off, thus, as it was observed in Table \ref{tab:rhotable} the Heckman IPD separated resulted in better estimates in terms of RMSE than the other methods evaluated.
+-->
+<!-- In particular, the method provides an advantage over the Heckman-single method in systematic missingness scenarios, since the latter does not allow to specify any cluster information implicitly, i.e.
 by adding a group variable in the imputation model.
 This can be seen in the estimates of the random effects parameters of the Heckman-single model, which are more biassed than those of the other methods in which cluster information was included.
-Moreover, the method could be affected by the ratio of the pooled data (See Annex Figure\ref{fig:continuous3}).
+-->
+<!-- Moreover, the method could be affected by the ratio of the pooled data (See Annex Figure\ref{fig:continuous3}).
 To demonstrate this, we increased the variance of the random effects in the simulated data, reducing the intra-cluster correlation (ICC) from 0.33 to 0.23.
 This resulted in less biassed estimates for both coefficient and random effects parameters, along with narrower confidence intervals.
+-->
 
-##### Results sensitivity to number of clusters and sample size of clusters
-We evaluated how robust our method was to variations in the sample size of the studies included in the IPD and also the number of clusters.
-Our method allows for the inclusion of studies with different sample sizes, but to facilitate the evaluation of the method, in this simulation study we used 10 studies each with a fixed sample size of 1000 observations.
+<!-- ##### Results sensitivity to number of clusters and sample size of clusters -->
+<!-- We evaluated how robust our method was to variations in the sample size of the studies included in the IPD and also the number of clusters. -->
+<!-- Our method allows for the inclusion of studies with different sample sizes, but to facilitate the evaluation of the method, in this simulation study we used 10 studies each with a fixed sample size of 1000 observations.
+-->
 
-<!-- ```{r rhosize,fig.pos="h",fig.cap="\\label{fig:size} Comparison of methods for continuous incomplete variable under systematical missingness by varing (number of studies- sample size per study), the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"} -->
-<!-- figure_final_s -->
-<!-- ``` -->
+<!-- <!-- ```{r rhosize,fig.pos="h",fig.cap="\\label{fig:size} Comparison of methods for continuous incomplete variable under systematical missingness by varing (number of studies- sample size per study), the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"} --> -->
+<!-- <!-- figure_final_s --> -->
+<!-- <!-- ``` --> -->
 
-When we reduced the sample size to 50 patients per study (left-most panel 10-50, Figure \ref{fig:rhosize}), we observed that the bias could be affected by some of the coefficient parameters, especially in MNAR scenarios.
+<!-- When we reduced the sample size to 50 patients per study (left-most panel 10-50, Figure \ref{fig:rhosize}), we observed that the bias could be affected by some of the coefficient parameters, especially in MNAR scenarios.
 All simulated data were successfully imputed with the Heckman IPD methods but we observed that the precision of the parameters was drastically affected especially for the Heckman IPD full method, due to singularity fitting problems in the estimation of the mixed model, so we prefer not to show it in the graph.
+-->
 
-On the other hand, by increasing the number of studies to 100 (rightmost panel 100-1000, Figure \ref{fig:rhosize}) we observe a higher precision of the estimates.
+<!-- On the other hand, by increasing the number of studies to 100 (rightmost panel 100-1000, Figure \ref{fig:rhosize}) we observe a higher precision of the estimates.
 There was an increase in the coverage of the estimates in the MAR scenario, i.e.
@@
 rho=0.6, we unfortunately observed a decrease in coverage, which could be explained by an increased complexity of the Heckman model.
+-->
 
-Furthermore, we observed that for the full Heckman IPD method, the number of successfully imputed datasets was only 90% of the simulated datasets.
+<!-- Furthermore, we observed that for the full Heckman IPD method, the number of successfully imputed datasets was only 90% of the simulated datasets.
 This could be explained by the fact that increasing the number of studies included in the analysis might make it more difficult to estimate the marginal parameters when applying the meta-analysis model with random effects, especially when considering maximum likelihood or restrictive maximum likelihood as the estimation method.
@@
 We use the method of moments estimation as an alternative estimation method, as it was previously evaluated by Audigier, and was found to provide robust results when estimating the marginal parameters.
+-->
 
@@
 However for school children, prevalences estimated with the Heckman method were found to be higher in the Mid-East and Southwest regions than those obtained with the other methods, whereas in the East-Central region the estimates with the Heckman method are lower.
-Possible reasons for selection bias in surveys of this type have been suggested, for example, that daytime visits might favor measurement in sick school children who stay home, leading to overestimated prevalence results as found in the East-Central region.
+Possible reasons for selection bias in surveys of this type have been suggested [@thedhsprogram2020], for example, that daytime visits might favor measurement in sick school children who stay home, leading to overestimated prevalence results as found in the East-Central region.
 Nevertheless, we were unable to find information that suggests or confirms the direction in which malaria prevalence is driven by selection bias in this Uganda study or in other studies similar to this one.

### 2022-10-24 1.Paper/Paper.Rmd
update text

@@
 These methods build upon the principles of Heckman selection models, and adopt a two-stage meta-analysis approach for imputing binary and continuous variables.
-After evaluating the proposed imputation models in simulated scenarios, we illustrated their use in a real IPDMA based on a malaria study in five subregions in Uganda, where we focused on estimating the prevalence of parasitemia for children aged 2-10 years."
+After evaluating the proposed imputation models in simulated scenarios, we illustrated their use in a malaria cross-sectional community survey to estimate the prevalence of parasitemia for children aged 2-10 years in five subregions in Uganda."
 keywords: Heckman model; IPDMA; Missing not at random; Selection models;Multiple imputation;
@@
 The copula approach uses a function, known as copula, that joins the marginal distribution of the error terms of the selection and outcome equation which are specified separately.
-$$ F(r^*_{i},y^*_{i}) =\Phi\left(r^*_{i}-X^S_{i}\beta^S_{i},\frac{y^*_{i}-X^O_{i}\beta^O_{i}}{\sigma_i}, \rho\right)$$
+$$ F(r^*_{i},y^*_{i}) =\Phi\left(r^*_{i}-X^S_{i}\beta^S_{i},\frac{y^*_{i}-X^O_{i}\beta^O_{i}}{\sigma_i}, \rho_i\right)$$
 Thus, to estimate the parameters of the Heckman method, it is sufficient to specify the marginal distributions of the error terms, and link them with a suitable copula function.
@@
 
-For coverage estimation, we obtain the confidence intervals from the Wald method.
-For the random effect variance, even when it is possible to estimate the confidence intervals through the profile or bootstrap method, we did not estimated them for computational reasons.
+To calculate the coverage of the parameter coefficients, we estimate confidence intervals (CI) with the Wald method.
+Although it is possible to obtain CI through the profile or bootstrap method for the variances of the random effects, we prefer not to estimate them (and hence the coverage) for these parameters due to computational time.
 

### 2022-10-25 1.Paper/Paper.Rmd
update

--- 2022-10-24 1.Paper/Paper.Rmd
@@
 
+To calculate the coverage of the parameter coefficients, we estimate confidence intervals (CI) with the Wald method.
+Although it is possible to obtain CI through the profile or bootstrap method for the variances of the random effects, we prefer not to estimate them (and hence the coverage) for these parameters due to computational time.
+
 #### Method
@@
 
-# ```{r, include=FALSE}
-# library(here)
-# source(here("2.Simulation", "simplot.R"),local =knitr::knit_global())
-# ```
+```{r, include=FALSE}
+library(here)
+source(here("2.Simulation","simplot.R"),local =knitr::knit_global())
+```
 
@@
 
-# ```{r rhocont,fig.pos="h",fig.cap="\\label{fig:rhocont} Comparison of methods for continuous incomplete variable under systematical missingness by varing $\rho$, the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"}
-# plot_cont_rho
-# ```
+ ```{r rhocont,fig.pos="h",fig.cap="\\label{fig:rhocont} Comparison of methods for continuous incomplete variable under systematical missingness by varing $\rho$, the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"}
+plot_rho
+ ```
 
@@
 
-<!-- ```{r rhosize,fig.pos="h",fig.cap="\\label{fig:size} Comparison of methods for continuous incomplete variable under systematical missingness by varing (number of studies- sample size per study), the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"} -->
-<!-- figure_final_s -->
-<!-- ``` -->
+```{r rhosize,fig.pos="h",fig.cap="\\label{fig:size} Comparison of methods for continuous incomplete variable under systematical missingness by varing (number of studies- sample size per study), the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"}
+plot_N
+```
 
@@
 
-<!-- ##### Results deviations in distribution assumptions -->
+##### Results deviations in distribution assumptions
 
-<!-- ```{r dist,fig.pos="!H",fig.cap="\\label{fig:dist} Comparison of methods for continuous incomplete variable with deviations in distribution assumptions, the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"} -->
-<!-- plot_cont_dist -->
-<!-- ``` -->
+```{r dist,fig.pos="!H",fig.cap="\\label{fig:dist} Comparison of methods for continuous incomplete variable with deviations in distribution assumptions, the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"}
+plot_S
+```
 
-<!-- To investigate how the imputation models behaves in settings with departures of binormal distribution, we draw error terms from a bivariate skew student-t distribution using the same location parameter and covariance matrix of the normal distributed settings,  with 8 degrees of freedom and an $\alpha=\{-5,5\}$ parameter which regulates the the slant of the density.
-In addition we simulated a explicit missing process, where error terms of selection and outcome equation were normally independent distributed and the inclusion depends on the value of the missing variable, as follows: -->
-<!-- $ry_i=0.1y_i+\epsilon^S_{i}$ -->
+To investigate how the imputation models behaves in settings with departures of binormal distribution, we draw error terms from a bivariate skew student-t distribution using the same location parameter and covariance matrix of the normal distributed settings,  with 8 degrees of freedom and an $\alpha=\{-5,5\}$ parameter which regulates the the slant of the density.
+In addition we simulated a explicit missing process, where error terms of selection and outcome equation were normally independent distributed and the inclusion depends on the value of the missing variable, as follows:
+$ry_i=0.1y_i+\epsilon^S_{i}$
 
-<!-- These settings assure that the percentage of missingness is similar in all the evaluated scenarios around 60\%. We observed that for these simulated settings. -->
+These settings assure that the percentage of missingness is similar in all the evaluated scenarios around 60\%. We observed that for these simulated settings.
 
-<!-- Figure \ref{fig:dist} depicts the results of this scenario.
+Figure \ref{fig:dist} depicts the results of this scenario.
 When we use the Heckman model in an explicit MNAR missing process, we observe that our model might not be completely suitable for this type of scenario, as we can observe that it affects the error variance which mainly affects the bias in the intersection coefficient parameter.
@@
 Similar variations occurred in the imputations performed with the simple Heckman imputation method, which is in line with the findings of @liu2020a.
--->
 
-<!-- ##### Bivariate incomplete variable -->
-<!-- We also evaluate the performance of the imputation methods for bivariate missing variables (Figure \ref{fig:bin}).
+##### Bivariate incomplete variable
+We also evaluate the performance of the imputation methods for bivariate missing variables (Figure \ref{fig:bin}).
 We also find that the Heckman-IPD separate method provides unbiased results on coefficient parameters and random effects with better convergence compared to the other imputation method evaluated.
 However, we observe that the accuracy of the random parameter effects is sensitive to $rho$ the correlation between the outcome and the selection equation, resulting in more imprecise estimators as the $rho$ parameter increases.
--->
 
-<!-- ```{r bin,fig.pos="!H",fig.cap="\\label{fig:bin} Comparison of methods for binary incomplete variable under systematical missingness, the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"} -->
-<!-- plot_bin -->
-<!-- ``` -->
+```{r bin,fig.pos="!H",fig.cap="\\label{fig:bin} Comparison of methods for binary incomplete variable under systematical missingness, the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"}
+plot_bin
+```
 
@@
                                    "p{0.07\\textwidth}|"))
-names(xt) <- c("Subregion","Districts (N)","Children (N)","Age (years)",
+names(xt) <- c("Subregion","Districts (N)","Children (N)","Age mean (years)",
                 "Log10 Female Anopheline","Wealth index",

### 2022-10-25 2.Simulation/simplot.R

@@ -3,13 +3,12 @@ library(data.table)
 library(ggplot2)
 
 # Arrange dataset
-data<-as.data.table(readxl::read_xlsx(here('2.Simulation','Sim_results.xlsx'),sheet="Sheet1"))
-head(data)
-data<-data[,grep("Scenario|^Ns|^ni|Method|Variation|time|run|^Cov|^Qbar|^Re|^Emp|^Bias|^MSE|^MCMSE|^Mod", names(data)), with = FALSE]
+data<-as.data.table(readxl::read_xlsx(here('2.Simulation','Sim_results.xlsx'),sheet="Sheet3"))
+data<-data[,grep("Scenario|^Ns|^ni|Method|Variation|time|run|^Cov|^Qbar|^Re|^Emp|^Bias|^MSE|^MCMSE|^Mod|^SE", names(data)), with = FALSE]
 colnames(data)<-c("Scenario","Ns","ni","Variation","Method",paste0("Mean_",0:6),"time", paste0("Cov_",0:2), paste0("Bias_",0:6),
-                  paste0("MSE_",0:6),paste0("ModSE_",0:2),"run",paste0("EmpSE_",0:6),paste0("MCMSE_",0:6))
+                  paste0("MSE_",0:6),paste0("ModSE_",0:2),"run",paste0("EmpSE_",0:6),paste0("MCMSE_",0:6),paste0("SEEmp_",0:6))
 data<- setDT(melt(data, measure.vars = c( paste0("Mean_",0:6), paste0("Cov_",0:2),paste0("Bias_",0:6),paste0("MSE_",0:6),paste0("ModSE_",0:2),paste0("EmpSE_",0:6)
-                                          ,paste0("MCMSE_",0:6)),
+                                          ,paste0("MCMSE_",0:6),paste0("SEEmp_",0:6)),
                   variable.name = "Statistic", value.name = "Value"))
 data[, c("Statistic", "Parameter") := tstrsplit(Statistic, "_", fixed=TRUE)]
 data<- setDT(dcast(data, Scenario + Ns + ni+ Variation+ Method + time + run + Parameter ~ Statistic, value.var = "Value"))
@@ -19,123 +18,55 @@ data[,SEBias:=EmpSE/run^0.5]
 data[,Per_run:=run/5]
 param_val <- c("0"="beta_0", "1"="beta_1", "2"="beta_2", 
[…]

### 2022-10-26 1.Paper/Paper.Rmd
graph update

@@
 We generated data sets of 10 groups of 1000 patients each in each scenario.
-For example, for the scenario in which the error terms followed a normal distribution, out of the 500 datasets generated, we obtained that on average 48.2\% of the Y response was missing, with the lowest missing percentage being 26.84\% and the maximum being 73.67\%.
+For example, for the scenario in which the error terms followed a normal distribution, out of the 500 datasets generated, we obtained that on average 60.8\% of the Y response was missing, with the lowest missing percentage being 26.84\% and the maximum being 73.67\%.
 At the cluster level, looking only at the sporadic missing clusters, we found an average of 35.25\% missing values, but there were clusters with no missing data at all and up to 98\% missing data in Y.
+Regarding the binary response scenarios, the proportion mean of treatment is around 60% CI[20%,90%].
+Processing times differ in the imputation methods evaluated, for example, for 10 clusters of 1000 units, the 1l.
+Heckman takes an average of 2 minutes, the 2l.MAR takes 10 minutes and the 2l.
+Heckman takes about 19 minutes.
+It should be noted that the time of the last two approaches depends strongly on the number of clusters included and the sample size in the data set.
 
@@
 and 2l.) models provide less biased estimates of the coefficient parameters in MNAR scenarios, but the random effects estimates on the 1l.heckman are far away from the true values as no cluster information was considered at all.
+ Regarding the coverage, the Heckman-IPD imputation method provides the best coverage values across all the $\rho$ scenarios with a coverage level close to the nominal 95\% interval.
+Even 2l.Heckman is not properly a randomization-valid (Rubin 1987b) method across all the $\rho$ values it leads to better results in terms of bias and coverage compared to the evalueated methods.
+
 The precision of the Heckman IPD separated seems it is better than the Heckman IPD full version, specially in the MAR scenarios, when we checked simulation results it seems that the high variability on the full version is related to simulations in which singularity fit that it is an indication that the model is overfitted.
-Regarding the coverage, the Heckman-IPD imputation method provides the best coverage values across all the $\rho$ scenarios with a coverage level up to 95%.
 

### 2022-10-26 2.Simulation/simplot.R

@@ -3,19 +3,16 @@ library(data.table)
 library(ggplot2)
 
 # Arrange dataset
-data<-as.data.table(readxl::read_xlsx(here('2.Simulation','Sim_results.xlsx'),sheet="Sheet3"))
-data<-data[,grep("Scenario|^Ns|^ni|Method|Variation|time|run|^Cov|^Qbar|^Re|^Emp|^Bias|^MSE|^MCMSE|^Mod|^SE", names(data)), with = FALSE]
-colnames(data)<-c("Scenario","Ns","ni","Variation","Method",paste0("Mean_",0:6),"time", paste0("Cov_",0:2), paste0("Bias_",0:6),
-                  paste0("MSE_",0:6),paste0("ModSE_",0:2),"run",paste0("EmpSE_",0:6),paste0("MCMSE_",0:6),paste0("SEEmp_",0:6))
-data<- setDT(melt(data, measure.vars = c( paste0("Mean_",0:6), paste0("Cov_",0:2),paste0("Bias_",0:6),paste0("MSE_",0:6),paste0("ModSE_",0:2),paste0("EmpSE_",0:6)
-                                          ,paste0("MCMSE_",0:6),paste0("SEEmp_",0:6)),
-                  variable.name = "Statistic", value.name = "Value"))
-data[, c("Statistic", "Parameter") := tstrsplit(Statistic, "_", fixed=TRUE)]
-data<- setDT(dcast(data, Scenario + Ns + ni+ Variation+ Method + time + run + Parameter ~ Statistic, value.var = "Value"))
-data[,SeCov:=sqrt(Cov*(1-Cov)/run)]
-data[,Coverage:=Cov]
-data[,SEBias:=EmpSE/run^0.5]
-data[,Per_run:=run/5]
+data<-as.data.table(readxl::read_xlsx(here('2.Simulation','Sim_results.xlsx'),sheet="Sheet4"))
+data<-data[,grep("Scenario|Method|Variation|^Bias|^EmpSE|^RMSE|^CIW|^Cov|^SE", names(data)), with = FALSE]
+colnames(data)<-c("Scenario","Variation","Method", paste0("p_Coverage_",0:2), paste0("p_Bias_",0:6),
+                  paste0("p_RMSE_",0:6),paste0("p_Width_",0:2),paste0("p_EmpSE_",0:6),paste0("SE_Bias_",0:6),
[…]

### 2022-10-27 2.Simulation/simplot.R
update plots

@@ -4,13 +4,18 @@ library(ggplot2)
 
 # Arrange dataset
 data<-as.data.table(readxl::read_xlsx(here('2.Simulation','Sim_results.xlsx'),sheet="Sheet4"))
-data<-data[,grep("Scenario|Method|Variation|^Bias|^EmpSE|^RMSE|^CIW|^Cov|^SE", names(data)), with = FALSE]
+#data<-data[,grep("Scenario|Method|Variation|^Bias|^EmpSE|^RMSE|^CIW|^Cov|^SE", names(data)), with = FALSE]
+#colnames(data)<-c("Scenario","Variation","Method", paste0("p_Coverage_",0:2), paste0("p_Bias_",0:6),paste0("p_RMSE_",0:6),paste0("p_Width_",0:2),paste0("p_EmpSE_",0:6),paste0("SE_Bias_",0:6),paste0("SE_EmpSE_",0:6),paste0("SE_RMSE_",0:6),paste0("SE_Width_",0:2), paste0("SE_Coverage_",0:2))
+#data<- setDT(melt(data, measure.vars = c( paste0("p_Bias_",0:6), paste0("p_Coverage_",0:2), paste0("p_EmpSE_",0:6), paste0("p_RMSE_",0:6),paste0("p_Width_",0:2),paste0("SE_Bias_",0:6), paste0("SE_Coverage_",0:2),paste0("SE_EmpSE_",0:6),paste0("SE_RMSE_",0:6),paste0("SE_Width_",0:2)),variable.name = "Statistic", value.name = "Value"))
+data<-data[Variation!="10-50",]
+data<-data[,grep("Scenario|Method|Variation|^Bias|^RMSE|^CIW|^Cov|^SE_B|^SE_C|^SE_W|^SE_RM", names(data)), with = FALSE]
 colnames(data)<-c("Scenario","Variation","Method", paste0("p_Coverage_",0:2), paste0("p_Bias_",0:6),
-                  paste0("p_RMSE_",0:6),paste0("p_Width_",0:2),paste0("p_EmpSE_",0:6),paste0("SE_Bias_",0:6),
-                  paste0("SE_EmpSE_",0:6),paste0("SE_RMSE_",0:6),paste0("SE_Width_",0:2), paste0("SE_Coverage_",0:2))
[…]

### 2022-10-27 1.Paper/Paper.Rmd
update codes

@@
 Next we simulated the mean of two continuous covariates from a multivariate normal distribution $\mu_h\sim N(\begin{psmallmatrix}0\\0\end{psmallmatrix},\begin{psmallmatrix}0.2 & 0.015\\0.015 & 0.2\end{psmallmatrix})$, with $h=\{2,3\}$.
-We then simulated for each cluster a baseline covariate $X_{2i}\sim N(\mu_2,1)$ and a exclusion restriction  $X_{3i} \sim N(\mu_3,0.5^2)$.
+We then simulated for each cluster a baseline covariate $X_{2i}\sim N(\mu_2,1)$ and a exclusion restriction  $X_{3i} \sim N(\mu_3,0.5)$.
 
@@
 \end{align*}
-We used these fixed coefficients $\beta_h^O=\{2,1,1\}$  and $\beta_h^S=\{0.1,1.5,-0.7,1.8\}$ in order to get around 40\% of sporadically missing values on the response $y_{ij}$ in the entire data set.
+We used these fixed coefficients $\beta_h^O=\{0.3,1,1\}$  and $\beta_h^S=\{-0.8,1.3,-0.7,1.2\}$ in order to get around 40\% of sporadically missing values on the response $y_{ij}$ in the entire data set.
 Additionally, we made that the $y_{ij}^*$ observations were systematically missing in 20% of the clusters included in data set.
@@
 \]
-with the parameters $\sigma_{b0}^2=\sigma_{b1}^2=\sigma_{b2}^2=0.5$.
+with the parameters $\sigma_{b0}^2=\sigma_{b1}^2=\sigma_{b2}^2=0.4$.
 We considered that the correlation parameter of the random effects between equations is a 40% of the value of the assumed correlation parameter between error terms $\rho$.
-In addition, we included a random effect on the exclusion restriction variable given by $b_3\sim N(0,0.3)$ assuming that the intracluster variation in the exclusion restriction effect is lower than the variation on other coefficient parameters effects.
+In addition, we included a random effect on the exclusion restriction variable given by $b_3\sim N(0,0.2)$ assuming that the intracluster variation in the exclusion restriction effect is lower than the variation on other coefficient parameters effects.
 The $\rho$ parameter adopts different values depending on the simulated missing mechanism (See below additional scenarios) .
@@
 \]
-whose $\sigma_i^2$ is variable across clusters and distributed as $log(\sigma_i) \sim N(log(0.7),0.05)$.
+whose $\sigma_i^2$ is variable across clusters and distributed as $log(\sigma_i) \sim N(0,0.05)$.
 
@@
 - \textbf{Binary response}: We evaluated the imputation method when missing variables is binary.
-Therefore we simulated  $y_{i}$ binary incomplete variables, the marginal set of parameters were fixed to  $B_h^O=\{0.5,1,1\}$ and  $B_h^S=\{0.1,1.5,-0.7,1.2\}$, keeping the other parameters similar to the ones used in the simulation of missing continuous variables.
-The observable binary variables was defined as:
+Therefore we simulated  $y_{i}$ binary incomplete variables, we keep the parameters similar to the ones used in the simulation of missing continuous variables,but the observable binary variables was defined as:
 \begin{align*}
@@
 To calculate the coverage of the parameter coefficients, we estimate confidence intervals (CI) with the Wald method.
-Although it is possible to obtain CI through the profile or bootstrap method for the variances of the random effects, we prefer not to estimate them (and hence the coverage) for these parameters due to computational time.
+Although it is possible to obtain CI through the profile or bootstrap method for the variances of the random effects, we prefer not to estimate them (and hence the coverage) for the random effects parameters due to computational time.
 
@@
 
-We calculated the following evaluation criteria according to the formulas provided in Morris et al.(2019)[@morris2019]:
+We calculated the following evaluation criteria [@VanBureen] according to the formulas provided in Morris et al.(2019)[@morris2019]:
 
 - \textbf{Bias:} Bias on the coefficient and random effect parameters.
-- \textbf{EmpSE:} Empirical standard errors of the estimates on the coefficient and random effect parameters.
-- \textbf{ModSE:} Monte Carlo standard errors on the coefficient  parameters.
-- \textbf{MSE:} Mean squared error of the coefficient  and random effect parameters.
 - \textbf{Coverage(\%):} Coverage of the 95\% confidence intervals for the coefficient  parameters.
-In addition we reported the average time processing and the percentage of datasets where the imputation method converged, i.e., the imputation method generated an output.
+-\testbf{Width}: Width of the confidence interval on the coefficient parameters.
+- \textbf{RMSE:} Root mean squared error of the coefficient  and random effect parameters.
+
+In addition, in the appendix table we reported the empirical standard errors (EmpSE), Monte Carlo standard errors (ModSE) on the coefficient  parameters,average time processing (time in seconds) and the percentage of datasets where the imputation method converged (run), i.e., the imputation method generated an output.
 

### 2022-10-27 2.Simulation/simplot.R

@@ -3,7 +3,7 @@ library(data.table)
 library(ggplot2)
 
 # Arrange dataset
-data<-as.data.table(readxl::read_xlsx(here('2.Simulation','Sim_results.xlsx'),sheet="Sheet4"))
+data<-as.data.table(readxl::read_xlsx(here('2.Simulation','Sim_results.xlsx'),sheet="Sheet1"))
 #data<-data[,grep("Scenario|Method|Variation|^Bias|^EmpSE|^RMSE|^CIW|^Cov|^SE", names(data)), with = FALSE]
 #colnames(data)<-c("Scenario","Variation","Method", paste0("p_Coverage_",0:2), paste0("p_Bias_",0:6),paste0("p_RMSE_",0:6),paste0("p_Width_",0:2),paste0("p_EmpSE_",0:6),paste0("SE_Bias_",0:6),paste0("SE_EmpSE_",0:6),paste0("SE_RMSE_",0:6),paste0("SE_Width_",0:2), paste0("SE_Coverage_",0:2))
 #data<- setDT(melt(data, measure.vars = c( paste0("p_Bias_",0:6), paste0("p_Coverage_",0:2), paste0("p_EmpSE_",0:6), paste0("p_RMSE_",0:6),paste0("p_Width_",0:2),paste0("SE_Bias_",0:6), paste0("SE_Coverage_",0:2),paste0("SE_EmpSE_",0:6),paste0("SE_RMSE_",0:6),paste0("SE_Width_",0:2)),variable.name = "Statistic", value.name = "Value"))
@@ -61,7 +61,7 @@ return(plot)
 }
 
 
-plot_rho<-plotsim(Scenariov="Rho3",levelsvar=c("0","0.3","0.6","0.9"))
-plot_bin<-plotsim(Scenariov="Bin3",levelsvar=c("0","0.3","0.6","0.9"))
-plot_N<-plotsim(Scenariov="N3",levelsvar=c("10-50","10-100","10-1000","50-1000","100-1000"))
-plot_S<-plotsim(Scenariov="S3",levelsvar=c("e~BVGamma","e~BVN","R=f(Y)" ))
+plot_rho<-plotsim(Scenariov="Rho",levelsvar=c("0","0.3","0.6","0.9"))
+plot_bin<-plotsim(Scenariov="Bin",levelsvar=c("0","0.3","0.6","0.9"))
[…]

### 2022-10-27 4.Codes/Functions_HPC.R

@@ -1,0 +1,919 @@
+rm(list = ls())
+library(lme4)
+library(data.table)
+library(mvtnorm)
+library(fMultivar)
+library(data.table)
+library(mgcv)
+library(GJRM)
+library(Matrix)
+library(mixmeta)
+library(mvtnorm)
+library(pbivnorm)
+library(mice)
+library(micemd)
+library(miceMNAR)
+library(glmmTMB)
+
+
+
+# Define list of parameters----
+par_data <- list ( N = 10, # Number of studies
+                   ni = 1000, # Number of patients per study
+                   sigmav = 1, # Mean of sigma_e parameter (error terms)
+                   sdsigma = 0.05, # Variance of the sigma_e parameter (error terms)
+                   het = 1, # Homogeneous(=0) or heterogeneous (=1) variation on error terms
+                   rho = 0.6, # Correlation of errors of the outcome and selection equation
+                   vrho = 0.0, # Variance of rho parameter
+                   vint =  0.4, # Variance of random intercept ( correlation o-s 0.3*rho)
+                   vslp1 = 0.4, # Variance of random slope X1 ( correlation o-s 0.5*rho)
+                   vslp2 = 0.4, # Variance of random slope X2 ( correlation o-s 0.5*rho)
+                   vslp3 = 0.2, # Variance of random slope X3 ( correlation o-s 0.5*rho)
+                   skew = 0, # 1= if error terms are skewed t distributed, 0= Normal distributed
+                   tau = 0.4, # correlation of random effects of outcome and selection (same for random intercept and slope)
+                   syst = 1, # 1= 20% of data are systematically missing otherwise all studies are sporadically missing
+                   typey = "normal", # Type continuous variable"probit"
+                   betao = c(0.3,1,1), # Beta output model(continuous)
+                   betas = c(-0.8,1.3,-0.7,1.2),
+                   typem = 0, # Type missing generation 0 = Heckman type 1= truncated
+                   valadd=0
+)
+
+
[…]

### 2022-10-27 1.Paper/Paper.Rmd
update

@@
 
-We calculated the following evaluation criteria [@VanBureen] according to the formulas provided in Morris et al.(2019)[@morris2019]:
+We calculated the following evaluation criteria [@buuren18c] according to the formulas provided in Morris et al.(2019)[@morris2019]:
 
 - \textbf{Bias:} Bias on the coefficient and random effect parameters.
-- \textbf{Coverage(\%):} Coverage of the 95\% confidence intervals for the coefficient  parameters.
--\testbf{Width}: Width of the confidence interval on the coefficient parameters.
+- \textbf{Coverage:} Coverage of the 95\% confidence intervals for the coefficient  parameters.
+- \textbf{Width}: Width of the confidence interval on the coefficient parameters.
 - \textbf{RMSE:} Root mean squared error of the coefficient  and random effect parameters.
@@
 Processing times differ in the imputation methods evaluated, for example, for 10 clusters of 1000 units, the 1l.
-Heckman takes an average of 2 minutes, the 2l.MAR takes 10 minutes and the 2l.
-Heckman takes about 19 minutes.
+Heckman takes an average of 10 seconds, the 2l.MAR takes 1.5 and the 2l.
+Heckman takes about 18 seconds.
 It should be noted that the time of the last two approaches depends strongly on the number of clusters included and the sample size in the data set.
@@
 
- ```{r rhocont,fig.pos="h",fig.cap="\\label{fig:rhocont} Comparison of methods for continuous incomplete variable under systematical missingness by varing $\rho$, the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"}
+ ```{r rhocont,fig.pos="h",fig.cap="\\label{fig:rhocont} Comparison of methods for continuous incomplete variable under systematical missingness by varing $\\rho$, the dashed line depicts the target performance criteria value",echo=FALSE,fig.keep="last"}
 plot_rho
@@
 the mechanism becomes MNAR, the estimates for the complete case analysis and the MAR-IPD imputation method turns more biased.
+
 Overall, both Heckman-based imputation (1l.
 and 2l.) models provide less biased estimates of the coefficient parameters in MNAR scenarios, but the random effects estimates on the 1l.heckman are far away from the true values as no cluster information was considered at all.
- Regarding the coverage, the Heckman-IPD imputation method provides the best coverage values across all the $\rho$ scenarios with a coverage level close to the nominal 95\% interval.
-Even 2l.Heckman is not properly a randomization-valid (Rubin 1987b) method across all the $\rho$ values it leads to better results in terms of bias and coverage compared to the evalueated methods.
 
-The precision of the Heckman IPD separated seems it is better than the Heckman IPD full version, specially in the MAR scenarios, when we checked simulation results it seems that the high variability on the full version is related to simulations in which singularity fit that it is an indication that the model is overfitted.
+ Regarding the coverage, the 2l.Heckman imputation method provides the best coverage values across all the $\rho$ scenarios with a coverage level close to the nominal 95\% interval.
+Even 2l.Heckman is not properly a randomization-valid [@buuren18c] method across all the $\rho$ values it leads to better results in terms of bias and coverage compared to the evalueated methods.
 
-The Heckman-IPD method, by incorporating individual and marginal level information, allows a bias-variance trade off, thus, as it was observed in Table \ref{tab:rhotable} the Heckman IPD separated resulted in better estimates in terms of RMSE than the other methods evaluated.
-In particular, the method provides an advantage over the Heckman-single method in systematic missingness scenarios, since the latter does not allow to specify any cluster information implicitly, i.e.
+The 2l.Heckman method, by incorporating individual and marginal level information, allows a bias-variance trade off, thus, as it was observed the Heckman IPD separated resulted in better estimates in terms of RMSE than the other methods evaluated.
+In particular, the method provides an advantage over the 1l.Heckman method in systematic missingness scenarios, since the latter does not allow to specify any cluster information implicitly, i.e.
 by adding a group variable in the imputation model.
-This can be seen in the estimates of the random effects parameters of the Heckman-single model, which are more biassed than those of the other methods in which cluster information was included.
-Moreover, the method could be affected by the ratio of the pooled data (See Annex Figure\ref{fig:continuous3}).
-To demonstrate this, we increased the variance of the random effects in the simulated data, reducing the intra-cluster correlation (ICC) from 0.33 to 0.23.
-This resulted in less biassed estimates for both coefficient and random effects parameters, along with narrower confidence intervals.
+This can be seen in the estimates of the random effects parameters of the 1l.Heckman method, which are more biased than those of the other methods in which cluster information was included.
+
+However, the accuracy (Width) of the 2l.Heckman model is higher than that of the other evaluated methods, which generally increases as $\rho$ moves away from zero.
+In particular, we noticed in $\beta_2$ that the width was larger in the MAR scenario, when we checked the simulation results, it seems that the high variability came from simulations in which the linear model estimation had singularity problems.
 
 ##### Results sensitivity to number of clusters and sample size of clusters
-We evaluated how robust our method was to variations in the sample size of the studies included in the IPD and also the number of clusters.
-Our method allows for the inclusion of studies with different sample sizes, but to facilitate the evaluation of the method, in this simulation study we used 10 studies each with a fixed sample size of 1000 observations.
+We evaluated how robust our method was to variations in the number of clusters and also in the sample size of cluster.
 
-```{r rhosize,fig.pos="h",fig.cap="\\label{fig:size} Comparison of methods for continuous incomplete variable under systematical missingness by varing (number of studies- sample size per study), the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"}
+```{r rhosize,fig.pos="h",fig.cap="\\label{fig:size} Comparison of methods for continuous incomplete variable under systematical missingness by varing (N=number of clusters- $n_i$=sample size per cluster), the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"}
 plot_N
@@
 
-When we reduced the sample size to 50 patients per study (left-most panel 10-50, Figure \ref{fig:rhosize}), we observed that the bias could be affected by some of the coefficient parameters, especially in MNAR scenarios.
-All simulated data were successfully imputed with the Heckman IPD methods but we observed that the precision of the parameters was drastically affected especially for the Heckman IPD full method, due to singularity fitting problems in the estimation of the mixed model, so we prefer not to show it in the graph.
+By increasing N, the number of clusters from 10 to 100 ( Figure \ref{fig:rhosize}), we observe that the bias was not affected but the average width of the estimates decreases (larger precision) and hence the RMSE.
+On the other hand, by decresing the number of units per cluster (from $n_i$=1000 to $n_i$=100) the precision decreased for all coefficients, the bias on coefficient estimates were not drastically affected but the variances of random effects do.
 
-On the other hand, by increasing the number of studies to 100 (rightmost panel 100-1000, Figure \ref{fig:rhosize}) we observe a higher precision of the estimates.
-There was an increase in the coverage of the estimates in the MAR scenario, i.e.
-rho=0, as the number of studies increases, which is in line with Audigier's results.
-However, in the case of MNAR, i.e.
-rho=0.6, we unfortunately observed a decrease in coverage, which could be explained by an increased complexity of the Heckman model.
-
-Furthermore, we observed that for the full Heckman IPD method, the number of successfully imputed datasets was only 90% of the simulated datasets.
-This could be explained by the fact that increasing the number of studies included in the analysis might make it more difficult to estimate the marginal parameters when applying the meta-analysis model with random effects, especially when considering maximum likelihood or restrictive maximum likelihood as the estimation method.
-To ensure the estimation of the marginal parameters, our method evaluates the marginal parameters using the method of moments when it is not possible to estimate them with other methods chosen by the user.
-We use the method of moments estimation as an alternative estimation method, as it was previously evaluated by Audigier, and was found to provide robust results when estimating the marginal parameters.
+When we reduced the sample size to 50 patients per study (not shown here but available in the appendix), the bias and RMSE of the $\sigma_b2$ was drastically affected, with large CI on these criteria.
+This could be explained in part for the scarce information on certain cluster that affects directly the estimation of the Heckman model on those clusters.
 
@@
 
-To investigate how the imputation models behaves in settings with departures of binormal distribution, we draw error terms from a bivariate skew student-t distribution using the same location parameter and covariance matrix of the normal distributed settings,  with 8 degrees of freedom and an $\alpha=\{-5,5\}$ parameter which regulates the the slant of the density.
-In addition we simulated a explicit missing process, where error terms of selection and outcome equation were normally independent distributed and the inclusion depends on the value of the missing variable, as follows:
-$ry_i=0.1y_i+\epsilon^S_{i}$
-
-These settings assure that the percentage of missingness is similar in all the evaluated scenarios around 60\%. We observed that for these simulated settings.
+To investigate how the imputation models behaves in settings with departures of bivariate normal distribution, we draw error terms from a bivariate skew student-t distribution using the same location parameter and covariance matrix of the normal distributed settings,  with 4 degrees of freedom and an $\alpha=\{-2,6\}$ parameter which regulates the the slant of the density.
+In addition we simulated a explicit missing process, where error terms of selection and outcome equation were normally independent distributed and the selection of observations depends on the value of the outcome variable, as $ry^*_i=0.3y^*_i+\epsilon^S_{i}$.These settings assure that the percentage of missingness on the outcome variable is around 60\% in all the evaluated scenarios.
 
 Figure \ref{fig:dist} depicts the results of this scenario.
-When we use the Heckman model in an explicit MNAR missing process, we observe that our model might not be completely suitable for this type of scenario, as we can observe that it affects the error variance which mainly affects the bias in the intersection coefficient parameter.
-and also its coverage.
-With respect to the t-distributed error scenario, apparently for the simulated measurements there is not much variation in the estimation of the coefficient parameters, however, we also observe that the estimation of the error variance was affected.
-In this scenario it is also seen that the coverage in the intercept parameter varies.
-Similar variations occurred in the imputations performed with the simple Heckman imputation method, which is in line with the findings of @liu2020a.
+When we use the Heckman model in an explicit MNAR process ($ry^*=f(y^*)$), we observe that our model might not be completely suitable for this type of scenario, as we can observe that it highly affects the bias of the intercept parameter ($\beta_0$) and the width of all the coefficient parmeters, Also the bias of the random effects parameters ($\sigma_0$ and $\sigma_2$) was affected.
+With respect to the t-distributed error scenario, apparently only the bias of $\beta_0$ was affected.
+For both scenarios, it is also seen that the coverage in the $\beta_0$ was highly affected.
 
 ##### Bivariate incomplete variable
-We also evaluate the performance of the imputation methods for bivariate missing variables (Figure \ref{fig:bin}).
-We also find that the Heckman-IPD separate method provides unbiased results on coefficient parameters and random effects with better convergence compared to the other imputation method evaluated.
-However, we observe that the accuracy of the random parameter effects is sensitive to $rho$ the correlation between the outcome and the selection equation, resulting in more imprecise estimators as the $rho$ parameter increases.
+We also find (Figure \ref{fig:bin}) that the 2l.Heckman method provides unbiased results on coefficient parameters and random effects when the missing variable is binary.
+However, we observe more variability in the estimates of random effects, reflected on longer confidence intervals on the Bias and RMSE criteria.
 
@@
 As can be seen in Figure \ref{fig:region} in terms of subregion level prevalence the estimates of the approaches do not differ significantly between methods, however except for the East-Central region, prevalence estimates under the Heckman approach are higher than those estimated using MAR or CC aproaches.
-```{r region,fig.pos="!H",fig.cap="\\label{fig:region} Estimates of malaria prevalence by Uganda subregion",echo=FALSE,fig.keep="last"}
+```{r region,fig.width = 5, fig.asp =.67, fig.pos="!H", fig.cap="\\label{fig:region} Estimates of malaria prevalence by Uganda subregion",echo=FALSE,fig.keep="last"}
 plot_dist
@@
 The prevalence estimates for children aged 2 to 6 years according to the approaches evaluated in all regions are very similar (Figure \ref{figure:region_age}, which could be partly explained by the mobility of children at this age compared to that of school-age children.
-```{r region_age,fig.pos="!H",fig.cap="\\label{fig:region_age} Estimation of malaria prevalence by region and age",echo=FALSE,fig.keep="last"}
+```{r region_age, fig.pos="!H", fig.cap="\\label{fig:region_age} Estimation of malaria prevalence by region and age",echo=FALSE,fig.keep="last"}
 plot_sdist_age

### 2022-10-27 2.Simulation/simplot.R

@@ -64,4 +64,4 @@ return(plot)
 plot_rho<-plotsim(Scenariov="Rho",levelsvar=c("0","0.3","0.6","0.9"))
 plot_bin<-plotsim(Scenariov="Bin",levelsvar=c("0","0.3","0.6","0.9"))
 plot_N<-plotsim(Scenariov="N",levelsvar=c("10-50","10-100","10-1000","50-1000","100-1000"))
-plot_S<-plotsim(Scenariov="S",levelsvar=c("e~BVGamma","e~BVN","R=f(Y)" ))
+plot_S<-plotsim(Scenariov="S",levelsvar=c("Skewed-t","Normal","ry*=f(y*)"))

### 2022-10-28 1.Paper/Paper.Rmd
fix some typos

--- 2022-10-27 1.Paper/Paper.Rmd
@@
 A recent example of an IPD-MA is the emerging risk factor collaboration [@theemergingriskfactorscollaboration2007], where data were combined from approximately 1.1 million individuals across 104 studies to investigate associations of cardiovascular diseases with several predictors.
-Individuals in these large datasets tend to be clustered in centres, countries or studies, where they have been subject to similar healthcare processes, and are therefore more alike than individuals from another cluster.
+Individuals in these large datasets tend to be clustered in centers, countries or studies, where they have been subject to similar healthcare processes, and are therefore more alike than individuals from another cluster.
 Sometimes, clusters may also differ in participant eligibility criteria, follow-up length, predictor and outcome definitions, or in the quality of applied measurement methods.
@@
 Then in the section \ref{methods} we define the main steps of the proposed imputation method.
-Posteriorly, in the section \ref{simulation} we provide the settings and results of a Montecarlo simulation study used to evaluate the performance of the proposed impuatation method.
+Posteriorly, in the section \ref{simulation} we provide the settings and results of a Monte Carlo simulation study used to evaluate the performance of the proposed imputation method.
 In section \ref{example} we illustrate the method using the survey information collected in different sub-districts in Uganda to estimate the prevalence of malaria in children.
@@
 #### Fit a meta-analysis model
-In this step, we pool the parameters $\widehat{\theta_i}$ with a random-effects meta-analysis model using only the groups with observable information, i.e., those that are not systematically missing and have sufficient information to estimate the heckman model.
+In this step, we pool the parameters $\widehat{\theta_i}$ with a random-effects meta-analysis model using only the groups with observable information, i.e., those that are not systematically missing and have sufficient information to estimate the Heckman model.
 In particular, we pooled the $p$ coefficients of the $\beta^O$ outcome equation and estimated a multivariate random effects meta-analysis model with them, similarly we combined all $q$ coefficient parameters of the $\beta^S$ selection equation.
@@
 We considered that the correlation parameter of the random effects between equations is a 40% of the value of the assumed correlation parameter between error terms $\rho$.
-In addition, we included a random effect on the exclusion restriction variable given by $b_3\sim N(0,0.2)$ assuming that the intracluster variation in the exclusion restriction effect is lower than the variation on other coefficient parameters effects.
+In addition, we included a random effect on the exclusion restriction variable given by $b_3\sim N(0,0.2)$ assuming that the intra-cluster variation in the exclusion restriction effect is lower than the variation on other coefficient parameters effects.
 The $\rho$ parameter adopts different values depending on the simulated missing mechanism (See below additional scenarios) .
@@
 #### Estimand
-The estimads were the parameter coefficients of the outcome equation $\beta^O={\beta^O_0,\beta^O_1,\beta^O_2}$, with special emphasis on the treatment effect parameter $\beta^0_1$.
+The estimands were the parameter coefficients of the outcome equation $\beta^O={\beta^O_0,\beta^O_1,\beta^O_2}$, with special emphasis on the treatment effect parameter $\beta^0_1$.
 We also report in the estimated variance of the random effects and residual errors $\sigma_{b0}^2$,$\sigma_{b1}^2$,$\sigma_{b2}^2$,$\sigma_{e}^2$.
@@
 Overall, both Heckman-based imputation (1l.
-and 2l.) models provide less biased estimates of the coefficient parameters in MNAR scenarios, but the random effects estimates on the 1l.heckman are far away from the true values as no cluster information was considered at all.
+and 2l.) models provide less biased estimates of the coefficient parameters in MNAR scenarios, but the random effects estimates on the 1l.Heckman are far away from the true values as no cluster information was considered at all.
  Regarding the coverage, the Heckman-IPD imputation method provides the best coverage values across all the $\rho$ scenarios with a coverage level close to the nominal 95\% interval.
-Even 2l.Heckman is not properly a randomization-valid (Rubin 1987b) method across all the $\rho$ values it leads to better results in terms of bias and coverage compared to the evalueated methods.
+Even 2l.Heckman is not properly a randomization-valid (Rubin 1987b) method across all the $\rho$ values it leads to better results in terms of bias and coverage compared to the evaluated methods.
 
-The precision of the Heckman IPD separated seems it is better than the Heckman IPD full version, specially in the MAR scenarios, when we checked simulation results it seems that the high variability on the full version is related to simulations in which singularity fit that it is an indication that the model is overfitted.
+The precision of the Heckman IPD separated seems it is better than the Heckman IPD full version, specially in the MAR scenarios, when we checked simulation results it seems that the high variability on the full version is related to simulations in which singularity fit that it is an indication that the model is over-fitted.
 
@@
 by adding a group variable in the imputation model.
-This can be seen in the estimates of the random effects parameters of the Heckman-single model, which are more biassed than those of the other methods in which cluster information was included.
+This can be seen in the estimates of the random effects parameters of the Heckman-single model, which are more biased than those of the other methods in which cluster information was included.
 Moreover, the method could be affected by the ratio of the pooled data (See Annex Figure\ref{fig:continuous3}).
 To demonstrate this, we increased the variance of the random effects in the simulated data, reducing the intra-cluster correlation (ICC) from 0.33 to 0.23.
-This resulted in less biassed estimates for both coefficient and random effects parameters, along with narrower confidence intervals.
+This resulted in less biased estimates for both coefficient and random effects parameters, along with narrower confidence intervals.
 
@@
 
-To illustrate our method, following the article by @rugnao2019, we estimated the prevalence of parasitemia by subregion and by age after approximately 3 years of LLIN campaigns started.
+To illustrate our method, following the article by @rugnao2019, we estimated the prevalence of parasitemia by sub-region and by age after approximately 3 years of LLIN campaigns started.
 We estimated parasitemia prevalence using 3 approaches: MCAR, MAR and MNAR.
@@
 In addition, we included a holiday indicator variable as ERV, which was calculated according to school vacation calendars and public holidays in Uganda in 2017.
-We examined the association of this ERV with the outcome variable (y) and with the selection indicator (ry), conditionated on the remaining imputation predictors.
+We examined the association of this ERV with the outcome variable (y) and with the selection indicator (ry), conditioned on the remaining imputation predictors.
 The model results in Table \ref{tab:exclusion} indicate that the holiday indicator could be a plausible ERV variable, as it was significantly associated with ry, but not with y.
@@
 According to our imputation approach, non-participants were estimated to have a higher prevalence of malaria than participants in more than half of the districts analyzed. 
-As can be seen in Figure \ref{fig:region} in terms of subregion level prevalence the estimates of the approaches do not differ significantly between methods, however except for the East-Central region, prevalence estimates under the Heckman approach are higher than those estimated using MAR or CC aproaches.
+As can be seen in Figure \ref{fig:region} in terms of sub-region level prevalence the estimates of the approaches do not differ significantly between methods, however except for the East-Central region, prevalence estimates under the Heckman approach are higher than those estimated using MAR or CC approaches.
 ```{r region,fig.pos="!H",fig.cap="\\label{fig:region} Estimates of malaria prevalence by Uganda subregion",echo=FALSE,fig.keep="last"}
@@
 One of the major limitations of our method is that it needs a valid restriction variable, which in some contexts is difficult to establish at the individual study level and can be even more challenging if one tries to find a valid exclusion variable across clusters.
-Also the method is sensitive to the value of the correlation between the selection equation and outcome (rho), and in general it is observed that it can lead to biassed results on fixed global parameters i.e.
+Also the method is sensitive to the value of the correlation between the selection equation and outcome (rho), and in general it is observed that it can lead to biased results on fixed global parameters i.e.
 without variation across clusters.Similarly, the method can be sensitive to both the sample size and the number of studies included in the database.

### 2022-10-31 1.Paper/Paper.Rmd

--- 2022-10-27 1.Paper/Paper.Rmd
@@
 Overall, both Heckman-based imputation (1l.
-and 2l.) models provide less biased estimates of the coefficient parameters in MNAR scenarios, but the random effects estimates on the 1l.heckman are far away from the true values as no cluster information was considered at all.
+and 2l.) models provide less biased estimates of the coefficient parameters in MNAR scenarios, but the random effects estimates on the 1l.Heckman are far away from the true values as no cluster information was considered at all.
+Regarding the coverage, the Heckman-IPD imputation method provides the best coverage values across all the $\rho$ scenarios with a coverage level close to the nominal 95\% interval.
+Even 2l.Heckman is not properly a randomization-valid (Rubin 1987b) method across all the $\rho$ values it leads to better results in terms of bias and coverage compared to the evaluated methods.
 
- Regarding the coverage, the 2l.Heckman imputation method provides the best coverage values across all the $\rho$ scenarios with a coverage level close to the nominal 95\% interval.
-Even 2l.Heckman is not properly a randomization-valid [@buuren18c] method across all the $\rho$ values it leads to better results in terms of bias and coverage compared to the evalueated methods.
+The precision of the Heckman IPD separated seems it is better than the Heckman IPD full version, specially in the MAR scenarios, when we checked simulation results it seems that the high variability on the full version is related to simulations in which singularity fit that it is an indication that the model is over-fitted.
 
-The 2l.Heckman method, by incorporating individual and marginal level information, allows a bias-variance trade off, thus, as it was observed the Heckman IPD separated resulted in better estimates in terms of RMSE than the other methods evaluated.
+The 2l.Heckman method, by incorporating individual and marginal level information, allows a bias-variance trade off, thus, as it was observed the 2l.Heckman separated resulted in better estimates in terms of RMSE than the other methods evaluated.
 In particular, the method provides an advantage over the 1l.Heckman method in systematic missingness scenarios, since the latter does not allow to specify any cluster information implicitly, i.e.
@@
 This can be seen in the estimates of the random effects parameters of the 1l.Heckman method, which are more biased than those of the other methods in which cluster information was included.
-
 However, the accuracy (Width) of the 2l.Heckman model is higher than that of the other evaluated methods, which generally increases as $\rho$ moves away from zero.
@@
 According to our imputation approach, non-participants were estimated to have a higher prevalence of malaria than participants in more than half of the districts analyzed.
-As can be seen in Figure \ref{fig:region} in terms of subregion level prevalence the estimates of the approaches do not differ significantly between methods, however except for the East-Central region, prevalence estimates under the Heckman approach are higher than those estimated using MAR or CC aproaches.
-```{r region,fig.width = 5, fig.asp =.67, fig.pos="!H", fig.cap="\\label{fig:region} Estimates of malaria prevalence by Uganda subregion",echo=FALSE,fig.keep="last"}
+
+As can be seen in Figure \ref{fig:region} in terms of sub-region level prevalence the estimates of the approaches do not differ significantly between methods, however except for the East-Central region, prevalence estimates under the Heckman approach are higher than those estimated using MAR or CC approaches.
+```{r region,fig.pos="!H",fig.cap="\\label{fig:region} Estimates of malaria prevalence by Uganda subregion",echo=FALSE,fig.keep="last"}
 plot_dist

### 2022-10-31 1.Paper/Paper.Rmd
Update Paper.Rmd

@@
 and 2l.) models provide less biased estimates of the coefficient parameters in MNAR scenarios, but the random effects estimates on the 1l.Heckman are far away from the true values as no cluster information was considered at all.
-Regarding the coverage, the Heckman-IPD imputation method provides the best coverage values across all the $\rho$ scenarios with a coverage level close to the nominal 95\% interval.
+Regarding the coverage, the 2l.Heckman imputation method provides the best coverage values across all the $\rho$ scenarios with a coverage level close to the nominal 95\% interval.
 Even 2l.Heckman is not properly a randomization-valid (Rubin 1987b) method across all the $\rho$ values it leads to better results in terms of bias and coverage compared to the evaluated methods.
-
-The precision of the Heckman IPD separated seems it is better than the Heckman IPD full version, specially in the MAR scenarios, when we checked simulation results it seems that the high variability on the full version is related to simulations in which singularity fit that it is an indication that the model is over-fitted.
 
@@
 One of the major limitations of our method is that it needs a valid restriction variable, which in some contexts is difficult to establish at the individual study level and can be even more challenging if one tries to find a valid exclusion variable across clusters.
+
 Also the method is sensitive to the value of the correlation between the selection equation and outcome (rho), and in general it is observed that it can lead to biased results on fixed global parameters i.e.

### 2022-10-31 1.Paper/Paper.Rmd
Update figures

@@
 
-```{r rhosize,fig.pos="h",fig.cap="\\label{fig:size} Comparison of methods for continuous incomplete variable under systematical missingness by varing (N=number of clusters- $n_i$=sample size per cluster), the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"}
+```{r rhosize,fig.pos="h",fig.cap="\\label{fig:size} Comparison of methods for continuous incomplete variable under systematical missingness by varing (N=number of clusters;$n_i$=sample size per cluster), the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"}
 plot_N

### 2022-10-31 2.Simulation/simplot.R

@@ -7,7 +7,7 @@ data<-as.data.table(readxl::read_xlsx(here('2.Simulation','Sim_results.xlsx'),sh
 #data<-data[,grep("Scenario|Method|Variation|^Bias|^EmpSE|^RMSE|^CIW|^Cov|^SE", names(data)), with = FALSE]
 #colnames(data)<-c("Scenario","Variation","Method", paste0("p_Coverage_",0:2), paste0("p_Bias_",0:6),paste0("p_RMSE_",0:6),paste0("p_Width_",0:2),paste0("p_EmpSE_",0:6),paste0("SE_Bias_",0:6),paste0("SE_EmpSE_",0:6),paste0("SE_RMSE_",0:6),paste0("SE_Width_",0:2), paste0("SE_Coverage_",0:2))
 #data<- setDT(melt(data, measure.vars = c( paste0("p_Bias_",0:6), paste0("p_Coverage_",0:2), paste0("p_EmpSE_",0:6), paste0("p_RMSE_",0:6),paste0("p_Width_",0:2),paste0("SE_Bias_",0:6), paste0("SE_Coverage_",0:2),paste0("SE_EmpSE_",0:6),paste0("SE_RMSE_",0:6),paste0("SE_Width_",0:2)),variable.name = "Statistic", value.name = "Value"))
-data<-data[Variation!="10-50",]
+data<-data[Variation!="10;50",]
 data<-data[,grep("Scenario|Method|Variation|^Bias|^RMSE|^CIW|^Cov|^SE_B|^SE_C|^SE_W|^SE_RM", names(data)), with = FALSE]
 colnames(data)<-c("Scenario","Variation","Method", paste0("p_Coverage_",0:2), paste0("p_Bias_",0:6),
                   paste0("p_RMSE_",0:6),paste0("p_Width_",0:2),paste0("SE_Bias_",0:6),paste0("SE_RMSE_",0:6),paste0("SE_Width_",0:2), paste0("SE_Coverage_",0:2))
@@ -29,7 +29,7 @@ data[,Parameter := factor(Parameter, levels = c("beta_0","beta_1","beta_2","sigm
                                                    expression(sigma[e])))]
 
 
-plotsim<-function(Scenariov,levelsvar){
+plotsim<-function(Scenariov,levelsvar,yaxis){
 datagraph<-data[Scenario==Scenariov,]
[…]

### 2022-10-31 3.Ilustrative_study/Ilustrative_study.R

@@ -1,4 +1,3 @@
-rm(list=ls())
 library(here) 
 library(data.table)
 library(micemd)
@@ -100,7 +99,7 @@ pred_model<-function(model,source){
 
 # 1. Subset original dataset ----
 
-data_original <- read.delim(here('3.Ilustrative_study','ISASimple_Gates_LLINE-UP_rct_RSRC.txt'))
+data_original <- read.delim(here('3.Ilustrative_study','ISASimple_Gates_LLINE-UP_rct_RSRC.txt')) # Dataset retrieved from https://clinepidb.org/ce/app/workspace/analyses/DS_7c4cd6bba9/new/details#AccessRequest
 dist <- readxl::read_xlsx(here('3.Ilustrative_study','Districts.xlsx')) #district specification
 source(here('3.Ilustrative_study','Additional_functions.R'))
 source(here('4.Codes','mice.impute.2l.heckman.R'))
@@ -293,7 +292,7 @@ pd <- position_dodge(width = 0.4)
 plot_dist<-ggplot(data_wide, aes(x = subreg_id, y = p, colour = Parameter)) +
                   geom_pointrange(aes(ymax = UC, ymin = LC),position = pd) + 
                   geom_point(position = pd)+ theme_light()+
-                  ylab("Parasite prevalence (%)")+xlab("Sub region")+
+                  ylab("Parasite prevalence (%)")+xlab("Sub-region")+
                   scale_color_brewer(palette="Dark2")
                   
 

### 2022-11-02 1.Paper/Paper.Rmd
Comments and suggestions for the entire paper. Due to a missing file I commented out the case study.

--- 2022-10-31 1.Paper/Paper.Rmd
@@
   num: a
-- name: V. de Jong
-  num: a
-- name: T. Debray
+- name: Valentijn M. T. de Jong
+  num: a, b
+- name: Thomas. P. A. Debray
   num: a
@@
 - num: a
-  org: Julius Center for Health Sciences and Primary Care, Utrecht Medical Center,Utrech, The Netherlands
+  org: Julius Center for Health Sciences and Primary Care, University Medical Center Utrecht, Utrecht University, Utrecht, The Netherlands
+- num: b
+  org: Data Analytics and Methods Task Force, European Medicines Agency, Amsterdam, the Netherlands
 corres: "* Johanna Munoz. \\email{j.munozavila@umcutrecht.nl}"
-presentaddress: Julius Center for Health Sciences and Primary Care, Utrecht Medical Center, Universiteitsweg 100, 3584 CG Utrech, The Netherlands
+presentaddress: Julius Center for Health Sciences and Primary Care, University Medical Center Utrecht, Universiteitsweg 100, 3584 CG Utrecht, the Netherlands
 authormark: Uthor \emph{et al}.
@@
 Although traditional imputation methods allow for valid statistical inference when data are missing at random (MAR), their implementation is not justified when observations are clustered (e.g., within studies) or when the presence of missingness depends on unobserved information.
-Unfortunately, this situation is increasingly common, and typically arises when individual participant data (IPD) are combined from multiple studies.
+Unfortunately, this situation is increasingly common, and typically arises when individual participant data (IPD) from multiple studies are combined.
 While several imputation methods have been proposed for addressing individual studies where data are missing not at random (MNAR), their application and validity in large datasets with clustering remains unclear.
-We therefore explored the consequence of MNAR data in IPDMA in-depth, and proposed novel multilevel imputation methods for common missing patterns in clustered datasets.
+We therefore explored the consequence of MNAR data in IPD meta-analysis in-depth, and proposed novel multilevel imputation methods for common missing patterns in clustered datasets.
 These methods build upon the principles of Heckman selection models, and adopt a two-stage meta-analysis approach for imputing binary and continuous variables.
@@
 ---
+<!-- I tried to add a second affiliation for me, but I get 1,1 behind my name instead of 1,2. Do you happen to know how to fix this? -->
+<!-- I find the last sentence of the abstract a bit hard to read. Does the following rephrasing make sense? -->
+<!-- "..., we illustrated their use in a cross-sectional community survey to estimate the prevalence of malaria parasitemia in children aged 2-10 years in five subregions in Uganda."  -->
 
 # Introduction
-Over the past few years, data sharing efforts have substantially increased and researchers increasingly often have access to large combined datasets derived from electronic health records or from individual participant data (IPD).
-For example, the clinical practice research datalink (CRPD) [@herrett2015] is an electronic health record (EHR) dataset in the UK, and has been used in a variety of medical research, such as the evaluation of health policy and drug efficacy.
+Over the past few years, data sharing efforts have substantially increased and researchers increasingly often have access to large combined datasets derived from electronic health records (EHR) or from individual participant data (IPD).
+For example, the clinical practice research datalink (CRPD) [@herrett2015] is an electronic health record dataset in the UK, which has been used in a variety of medical research, such as the evaluation of health policy and drug efficacy.
 A recent example of an IPD-MA is the emerging risk factor collaboration [@theemergingriskfactorscollaboration2007], where data were combined from approximately 1.1 million individuals across 104 studies to investigate associations of cardiovascular diseases with several predictors.
-Individuals in these large datasets tend to be clustered in centers, countries or studies, where they have been subject to similar healthcare processes, and are therefore more alike than individuals from another cluster.
+Individuals in these large datasets tend to be clustered in centres, countries or studies, where they have been subject to similar healthcare processes, and are therefore more alike than individuals from another cluster.
 Sometimes, clusters may also differ in participant eligibility criteria, follow-up length, predictor and outcome definitions, or in the quality of applied measurement methods.
@@
 Clustered datasets often contain  many incomplete variables.
-For example, in registry data it is common that test results are not available in registry data for all patients, as the decision to test may be at the discretion of the primary care physician or because the patient refuses to undergo testing .
+For example, in registry data it is common that test results are not available in registry data for all patients, as the decision to test may be at the discretion of the primary care physician or because the patient refuses to undergo testing.
 It is also possible that variables are systematically missing across clusters.
@@
 To address the presence of missing data, it is important to consider  the proportion of missingness and the missing mechanism for each incomplete variable.
-Rubin (1976) [@rubin76] identified three missing mechanisms where the probability of missingness: 1) is constant (missing completely  at random; MCAR), 2) depends on observed data only (missing at random; MAR), or 3) depends on unobserved information even after conditioning on all observable variables (missing non-random; MNAR).
+Rubin (1976) [@rubin76] identified three missing mechanisms where the probability of missingness: 1) is constant (missing completely  at random; MCAR), 2) depends on observed data only (missing at random; MAR), or 3) depends on unobserved information even after conditioning on all observable variables (missing not at random; MNAR).
 Traditional imputation methods are designed to address incomplete data sets where variables are MCAR or MAR.
@@
 
-Although it is possible to formally rule out MCAR [@little1988], it is impossible to assess whether data are MAR or MNAR as the observable data are not enough to test the assumptions of both mechanisms [@enders2022].
+Although it is possible to formally rule out MCAR [@little1988], 
+<!-- Stat med uses the AMA citation style.
+citations at the end of sentence go after the period: "out MCAR.
+[@little1988]" , and those at the end of a clause go after the period: "out MCAR, [@little1988]".
+Inline citations like "the clinical practice research datalink (CRPD) [@herrett2015]" are fine though.
+See an example here:
+https://onlinelibrary.wiley.com/doi/epdf/10.1002/sim.9221 -->
+it is impossible to assess whether data are MAR or MNAR as the observable data are not enough to test the assumptions of both mechanisms [@enders2022].
 For this reason, researchers often conveniently assume that data are MAR, or present results that are based on a complete case analysis.
@@
 
-One could see, a missing variable is a mixture of MAR and MNAR, depending on the available information, making the MAR mechanism more likely as more information is recorded.
+The missingness mechanism of a variable may be a mixture of MAR and MNAR, depending on the availability of data that may predict the missing data.
+Therefore, the MAR mechanism may become more influential than the MNAR mechanism as more information is recorded.
+<!-- MAR does not really become more likely.
+It is rather that the MAR mechanism becomes stronger than the MNAR mechanism.
+I have tried to capture this with the alternative text -->
 For instance, healthcare professionals may decide what type of measurement on indication, and depending on their personal experience or behavior decide whether or not to record the reasons for the measurement.
 
-Registries are notoriously prone to incomplete variables that are MNAR, due to the complex recording process [@liu2021], e.g.
+Registries are notoriously prone to incomplete variables that are MNAR, due to the complex recording process.
+[@liu2021] e.g.
 laboratory tests are taken only in certain patients based on sign and symptom information that is often incomplete or not recorded.
-Also IPD may suffer from MNAR, for example when study participants who experience unfavorable results drop out of the study or also due to heterogeneity of the primary objective or resources of the studies involved, which may result in variables relevant to explain the missing process not being recorded.
+Also IPD may suffer from MNAR, for example when study participants who experience unfavorable results drop out of the study.
+Also, heterogeneity of the primary objective or resources of the studies involved may result in variables relevant to explain the missing process not being recorded.
 
-The MNAR mechanism is considered as non-ignorable because to deal with this type of missingness it is necessary to specify information about the missingness process in addition to assumptions about the observed data.
+The MNAR mechanism is considered as non-ignorable, because to deal with this type of missingness it is necessary to specify information about the missingness process in addition to assumptions about the observed data.
 A common approach to address MNAR is to adopt  selection models such as the one proposed by Heckman (1976) [@heckman1976].
@@
 An important feature of the Heckman selection model is that its implementation does not require data to be MNAR, and can also be used when data are MCAR or MAR.
-It therefore offers an appealing solution to incomplete data sets where the missingness mechanism is unknown.
+It therefore offers an appealing solution to incomplete data sets where the missingness mechanism is not precisely known.
 
-Over the past few years, several Heckman selection models have been proposed for multiple imputation.
+Over the past few years, several extensions and adaptations to the Heckman selection model have been proposed for multiple imputation.
 Among them, Galimard et al.(2016) [@galimard_etal16] implemented a chained equations imputation method for continuous variables, which was extended to binary and categorical variables by employing copula estimates [@galimard_etal18].
@@
 In section \ref{heckman} we provide an introduction to the Heckman model and its estimation, and we extended to a hierarchical setting.
-Then in the section \ref{methods} we define the main steps of the proposed imputation method.
-Posteriorly, in the section \ref{simulation} we provide the settings and results of a Monte Carlo simulation study used to evaluate the performance of the proposed imputation method.
+Then in section \ref{methods} we define the main steps of the proposed imputation method.
+In section \ref{simulation} we provide the settings and results of a Monte Carlo simulation study used to evaluate the performance of the proposed imputation method.
 In section \ref{example} we illustrate the method using the survey information collected in different sub-districts in Uganda to estimate the prevalence of malaria in children.
@@
 
-Selection bias occurs when the selection of a subject  or his observation into the sample is influenced by unobserved variables  (e.g., the respondent's level of trust toward healthcare entities may cause him or her to self-select out of the study or refuse to sign consent for a test), which in turn are correlated with unobserved variables related to a variable of interest (e.g., the result of a blood test) [@vella1998].
+Selection bias occurs when the selection of a subject  or their observation into the sample is influenced by unobserved variables  (e.g., the respondent's level of trust toward healthcare entities may cause them to self-select out of the study or refuse to sign consent for a test), which in turn are correlated with unobserved variables related to a variable of interest (e.g., the result of a blood test) [@vella1998].
 
@@
 \end{figure}
-This can be better visualized in Figure \ref{fig:DAG}, where for the $j$th individual or unit within the $i$th cluster, there is $y_{ij}^*$, a latent outcome variable, and $r_{ij}^*$, a latent selection variable, which are correlated through $u_{ij}$ an unobserved or unrecorded variable, with $i\in[1,2,...,N]$ and $j\in[1,2,...,n_i]$.
-Here, both latent variables are related to $x_{ij}^O$ and $x_{ij}^S$ sets of predictor covariates.
+This is visualized in Figure \ref{fig:DAG}, where for the $j$th individual or unit within the $i$th cluster, there is $y_{ij}^*$, a latent outcome variable, and $r_{ij}^*$, a latent selection variable, which are correlated through $u_{ij}$ an unobserved or unrecorded variable, with $i\in[1,2,...,N]$ and $j\in[1,2,...,n_i]$.
+Here, both latent variables are related to the sets of predictor covariates $x_{ij}^O$ and $x_{ij}^S$.
 From $r_{ij}^*$ one can derive $r_{ij}=I(r_{ij}^*>=0)$ a selection indicator of $y_{ij}^*$ into the sample, and with this in turn, one can define $y_{ij}=y_{ij}^*,\forall r_{ij}=1$ the observable outcome variable.
 
-Denoting $y^*_i=(y^*_{i1},y^*_{i2},...,y^*_{in_i})^T$ and  $r^*_i=(r^*_{i1},r^*_{i2},...,r^*_{in_i})^T$ the vectors of latent outcomes and latent selections in the $i$ cluster, then Heckman's model is defined by two main equations: the outcome equation (\ref{eq:1}), which describes the relation latent outcome-exposure, and the selection equation (\ref{eq:2}) which details the likelihood that the outcome is observed in the sample.
+Denoting $y^*_i=(y^*_{i1},y^*_{i2},...,y^*_{in_i})^T$ and  $r^*_i=(r^*_{i1},r^*_{i2},...,r^*_{in_i})^T$ the vectors of latent outcomes and latent selections in the cluster $i$, then Heckman's model is defined by two main equations: the outcome equation (\ref{eq:1}), which describes the relation latent outcome-exposure association, and the selection equation (\ref{eq:2}) which details the likelihood that the outcome is observed in the sample.
 \begin{align}
@@
 This variable is commonly known as an exclusion restriction variable, and should only be associated with the selection in the sample $r^*_{ij}$ but not with the actual observation $y^*_{ij}$.
+<!-- Can you give some explanation (1-3 sentences?) on the exclusion and restriction assumptions here? -->
 
@@
 Under the BVN assumption, the parameters of the Heckman model coefficients can be estimated using the two-step Heckman method (H2S) [@heckman1976] or the full information maximum likelihood method (FIML) [@amemiya1984].
-However, both methods can lead to inconsistent estimators when this underlying distribution cannot be assumed[@gomes_etal20].To overcome this problem, other approaches have been proposed that relax the distribution assumptions, among them copula models [@smith03].
-The copula approach uses a function, known as copula, that joins the marginal distribution of the error terms of the selection and outcome equation which are specified separately.
+However, both methods can lead to inconsistent estimators when this underlying distribution cannot be assumed[@gomes_etal20].
+To overcome this problem, other approaches have been proposed that relax the distribution assumptions, among them copula models [@smith03].
+The copula approach uses a function, known as a copula, that joins the marginal distribution of the error terms of the selection and outcome equation which are specified separately.
 $$ F(r^*_{i},y^*_{i}) =\Phi\left(r^*_{i}-X^S_{i}\beta^S_{i},\frac{y^*_{i}-X^O_{i}\beta^O_{i}}{\sigma_i}, \rho_i\right)$$
@@
 This is because such an approach is less computationally intensive and could potentially generate fewer convergence problems in the estimation of the Heckman hierarchical model compared to other approaches.
-Briefly, in a first stage $\theta_i$ the cluster specific parameters of the Heckman model,i.e., the parameters of the two equations in each study: the outcome model and the selection model, for each of the $N$ clusters are estimated separately and then, in a second stage, all $\theta_i$ are combined using a random effects model.
 
-In a random effects model, $\theta_i$ parameters are assumed to be drawn independently and identically from an imaginary distribution of parameters with a $\theta_m$ population mean and a $\psi$ population variance [@higgins2009].
+<!-- This sentence was too long and has the verbs all the way at the end, making it hard to read. I suggest to rephrase to three shorter sentences: -->
+Briefly, in a first stage $\theta_i$ the cluster specific parameters of the Heckman model are estimated.
+That is, the parameters of the two equations in each study, the outcome model and the selection model, are estimated separately for each of the $N$ clusters.
+In the second stage, all $\theta_i$ are combined using a random effects meta-analysis model.
+
+<!-- I wouldn't call the distribution imaginary. It is assumed to be real but latent. Some words we could use: conjectured, latent, presupposed, presumed  -->
+In a random effects model, $\theta_i$ parameters are assumed to be drawn independently and identically from a latent distribution of parameters with a  population mean $\theta_m$ and a population variance $\psi$ [@higgins2009].
 Thus, the  $\theta_i=\theta_m + b_i$ can be specified using $b_i \sim N(0,\psi)$ random effects to allow for between-study heterogeneity in observed data relationships, and between-study heterogeneity in missing patterns.
@@
 We follow a similar approach proposed by Resche-Rigon and White (2018) [@resche-rigon2018] for multilevel imputation of data.
-Briefly, his method was developed to impute variables from a hierarchical structure (i.e., when there are unit samples grouped within a cluster or group).
-This method involves estimating an outcome equation (describing the relationships of the observed data to the missing variable) separately in each cluster, after which the parameter estimates of that equation are pooled using random-effects meta-analysis.
+Briefly, their method was developed to impute variables from a hierarchical structure (i.e., when there are samples unit grouped within a cluster or group).
+This method involves estimating an outcome equation (describing the relationships of the observed data to the missing variable) separately in each cluster, after which the parameter estimates of that equation are pooled using random effects meta-analysis.
 
 With this method, values can be imputed in very common scenarios in IPD, e.g., sporadic and systematic missing patterns.
-In particular, when the response variable is systematically missing within a group, i.e., when $y_{ij}$ are totally missing within a group, the imputation values are drawn from a (generalized) linear model conditional on $\theta_m$ the marginal population parameters, i.e., those estimated after pooling the cluster-specific parameters $\theta_i$.On the other hand, when the variable is sporadically missing within the group, i.e., there are some observed $y_{ij}$ within the cluster, the imputation model is conditional on the shrunk-cluster parameters, i.e., those coming from the shrinkage of $\theta_i$ towards $\theta_m$.
+In particular, when the response variable is systematically missing within a group, i.e., when $y_{ij}$ are totally missing within a group, the imputation values are drawn from a (generalized) linear model conditional on $\theta_m$ the marginal population parameters, i.e., those estimated after pooling the cluster-specific parameters $\theta_i$.
+On the other hand, when the variable is sporadically missing within the group, i.e., there are some observed $y_{ij}$ within the cluster, the imputation model is conditional on the shrunk-cluster parameters, i.e., those coming from the shrinkage of $\theta_i$ towards $\theta_m$.
 
-Our imputation approach differs crucially from the previous approach as here we estimate two correlated equations (instead of a single outcome equation) in each cluster , thus obtaining $\theta_i=\{\beta_i^O,\beta_i^S,\sigma_i,\rho_i\}$ parameters from both equations which are then pooled into a $\theta_m$ parameter set at the marginal or population level.
+Our imputation approach differs crucially from the previous approach as here we estimate two correlated equations (instead of a single outcome equation) in each cluster, thus obtaining $\theta_i=\{\beta_i^O,\beta_i^S,\sigma_i,\rho_i\}$ parameters from both equations which are then pooled into a $\theta_m$ parameter set at the marginal or population level.
 Our method is basically a univariate imputation method, but since it is implemented in a Gibbs sampling procedure, it can also be used to impute multiple incomplete variables in a data set.
+<!-- I don't understand what you mean here: it is univariate and it can be used to impute multiple variables? I would say these two statements contradict each other. -->
 
@@
 
-Initially, we estimate by the copula method the set of cluster-specific parameters, $\widehat{\theta_i}=\{\widehat{\beta^O_i}, \widehat{\beta^S_i}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, using all $j$ units with $y_{ij}^{obs}$ observable measurements within each $i$ group .
-The Heckman model is estimated with the \textbf{gjrm} function of the GJRM R package under the bivariate model with the nonrandom sample selection (BSS) specification, from which we obtain not only the parameters point estimates $\widehat{\theta_i}$, but also their corresponding $\widehat{S(\theta_i)}$ within-cluster variance-covariance matrix.
+Initially, we use the copulat method to estimate the set of cluster-specific parameters, $\widehat{\theta_i}=\{\widehat{\beta^O_i}, \widehat{\beta^S_i}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, using all $j$ units with observable measurements $y_{ij}^{obs}$ within each group $i$.
+The Heckman model is estimated with the \textbf{gjrm} function of the GJRM R package under the bivariate model with the nonrandom sample selection (BSS) specification, from which we obtain not only the parameters' point estimates $\widehat{\theta_i}$, but also their corresponding $\widehat{S(\theta_i)}$ within-cluster variance-covariance matrix.
+
+<!-- Can you add a reference to the R package here? -->
 
@@
 #### Fit a meta-analysis model
-In this step, we pool the parameters $\widehat{\theta_i}$ with a random-effects meta-analysis model using only the groups with observable information, i.e., those that are not systematically missing and have sufficient information to estimate the Heckman model.
+<!-- This paragraph starts of in the the present tense " we pool ", and later it's past tense " we pooled" etc.
+Stick to one tense.
+In applied papers, it is normal to use only past tense.
+In stats papers it is common to use present tense when describing a (new) method in the methods section.
+-->
+In this step, we pool the parameters $\widehat{\theta_i}$ with a random effects meta-analysis model using only the groups with observable information, i.e., those that are not systematically missing and have sufficient information to estimate the heckman model.
 In particular, we pooled the $p$ coefficients of the $\beta^O$ outcome equation and estimated a multivariate random effects meta-analysis model with them, similarly we combined all $q$ coefficient parameters of the $\beta^S$ selection equation.
-We also performed a univariate random effects meta_analysis on $\sigma'$ the log-transformed parameter of $\sigma$ and another on $\rho'$ the fisher-transformed parameter of $\rho$.
+<!-- The phrase " the log-transformed parameter of $\sigma$" is a description of $\sigma'$, so it is what is known as a commenting or parenthetical clause.
+That means you need to put a comma before and after this clause: "...
+on $\sigma'$, the log-transformed parameter of $\sigma$, ...".
+Although this article is not for the BMJ, I can recommend reading their house-style for writing: https://www.bmj.com/about-bmj/resources-authors/house-style -->
+<!-- Read more here if you like: http://content.principia.edu/teaching-excellence/commenting-with-commas-when-and-where-to-use-them/ -->
+We also performed a univariate random effects meta_analysis on $\sigma'$, the log-transformed parameter of $\sigma$, and another on $\rho'$, the fisher-transformed parameter of $\rho$.
 
-The meta-analysis model is performed with the \textbf{mixmeta} function of the R package mixmeta, which allows the use of maximum likelihood (ML), restricted maximum likelihood (REML) and moments estimation methods.
+The meta-analysis model is performed with the \textbf{mixmeta} function of the R package *mixmeta*, which allows the use of maximum likelihood (ML), restricted maximum likelihood (REML), and moments estimation methods.
 For the simulation and illustrative study, we used the restricted REML estimation method, which is recommended as it has a good balance between insensitivity and efficiency [@viechtbauer2005].
+<!-- Insensitivity is not generally a good thing.
+I think you should either explain why it is a good thing here: "...
+it is insensitive to the true value of the mean of the distribution of random effects and it is efficient" (or something similar, feel free to rephrase) , or simplify it to "...
+a good balance between unbiasedness and efficiency."  -->
 
@@
 
-#### Draw $\theta^*_i$ the cluster parameters
-
-We draw $\theta^*_i$ shrunk-cluster parameters for each group $i$ from the following posterior distribution conditional on $\theta^*_m$ and $\psi^*$.
+#### Draw the cluster parameters $\theta^*_i$
+<!-- Throughout the text, consider the order in which you put the symbol and the text: -->
+<!-- Put the symbol after the text if the text is a description of the symbol: "shrunk-cluster parameters $\theta^*_i$" . -->
+<!-- Put the text after the symbol if the symbol quantifies the text: "$n$ patients" , or " $N$ clusters" . -->
+We draw the shrunk-cluster parameters $\theta^*_i$ for each group $i$ from the following posterior distribution conditional on $\theta^*_m$ and $\psi^*$.
 \begin{align*}
@@
 As can be seen, the mean and variance of the posterior distribution is a combination between the estimated marginal and cluster-specific parameters.
-Here the weights on the $\widehat{\theta_i}$ cluster-specific and the $\theta^*_m$ marginal parameters are inversely proportional to the $\widehat{S_{\theta_i}}$ within and $\psi^*$ between clusters variances.
+Here the weights on the cluster-specific parameters $\widehat{\theta_i}$ and the marginal parameters $\theta^*_m$ are inversely proportional to the within cluster variance $\widehat{S_{\theta_i}}$ and between clusters variance $\psi^*$.
 For example, when $\widehat{S_{\theta_i}} < \psi^*$ the mean of the conditional distribution gives more weight to the estimated cluster-specific parameter.
@@
 In the case of systematic missingness, it is like considering the within-cluster variance to be infinite ($\widehat{S_{\theta_i}}\to\infty$), then letting all parameters rely only on the marginal estimates.
+<!-- What does then mean in this last part? should it be " and lettting all parameters rely ..." ? -->
 
@@
 
-Having $\theta^*_i$ the shrunk-cluster parameters vector for each group, we back transform $\sigma^*$ and $\rho^*$ to the original scale.
+Having $\theta^*_i$ the shrunk-cluster parameters vector for each group, we back-transform $\sigma^*$ and $\rho^*$ to the original scale.
 Then $y^{miss}_{ij}$ the missing values can be drawn from $p(y^{miss}_{ij}|\theta^*_i)$ their predictive distribution given $\theta^*_i$ as follows:
@@
 When there are simultaneous missing variables in a dataset, our imputation method can be extended in a Gibbs sampler procedure.
-Particularly our imputation method has been implemented according to the structure of the MICE R package  <!--[@buuren2021]-->, that allows imputing multiple incomplete predictors and covariates in a given dataset.
+Particularly, our imputation method has been implemented according to the structure of the MICE R package  <!--[@buuren2021]-->, that allows imputing multiple incomplete predictors and covariates in a given dataset.
 
-The MICE package allows to specify imputation methods to each of the missing variables by setting the method and the predictive matrix for each of the missing variables.
-To use our method, it is necessary to specify for the MNAR missing variable the method \textbf{2l.heckman} in the mice methods vector.
+<!-- The next paragraph belongs in the manual for the package, or a vignette. I don't think it is necessary to put it here. -->
+The MICE package allows one to specify imputation methods to each of the missing variables by setting the method and the predictive matrix for each of the missing variables.
+To use our proposed method, it is necessary to specify for the MNAR missing variable the method \textbf{2l.heckman} in the mice methods vector.
 Furthermore, in the prediction matrix, the group or cluster variable should be specified as \textbf{'-2'}, all predictor variables belonging to the selection and outcome as \textbf{'1'}, the exclusion restrictions or predictor variables that are only included in the selection equation as \textbf{'-3'} and those that are only included in the outcome equation as \textbf{'-4'}.
@@
 #### Aim
-We design a simulation study aimed to compare the performance of imputation methods when they impute a missing variable that comes from a hierarchical dataset and follows a MNAR mechanism.
+We designed a simulation study aimed to compare the performance of imputation methods for imputing a missing variable in a hierarchical dataset, where the missingness follows a MNAR mechanism.
 
@@
 For simplicity we assume that the database collects information from $N=10$ clusters of equal number of individuals $n_i=1000$.
+<!-- But you also have scenarios with different numbers for N and n_i.
+You could write here that for the "basic scenario" we have N =  and n_i = , and that we alter this in other scenarios, or in " senstivity analyses" .
+-->
 For each dataset, we generated $X_{1i}$ a treatment indicator variable from a Bernoulli distribution with a probability of treatment on each cluster equal to 0.6.
-Next we simulated the mean of two continuous covariates from a multivariate normal distribution $\mu_h\sim N(\begin{psmallmatrix}0\\0\end{psmallmatrix},\begin{psmallmatrix}0.2 & 0.015\\0.015 & 0.2\end{psmallmatrix})$, with $h=\{2,3\}$.
+Next, we simulated the mean of two continuous covariates from a multivariate normal distribution $\mu_h\sim N(\begin{psmallmatrix}0\\0\end{psmallmatrix},\begin{psmallmatrix}0.2 & 0.015\\0.015 & 0.2\end{psmallmatrix})$, with $h=\{2,3\}$.
 We then simulated for each cluster a baseline covariate $X_{2i}\sim N(\mu_2,1)$ and a exclusion restriction  $X_{3i} \sim N(\mu_3,0.5)$.
@@
 \end{align*}
-We used these fixed coefficients $\beta_h^O=\{0.3,1,1\}$  and $\beta_h^S=\{-0.8,1.3,-0.7,1.2\}$ in order to get around 40\% of sporadically missing values on the response $y_{ij}$ in the entire data set.
-Additionally, we made that the $y_{ij}^*$ observations were systematically missing in 20% of the clusters included in data set.
+We fixed coefficients $\beta_h^O=\{0.3,1,1\}$  and $\beta_h^S=\{-0.8,1.3,-0.7,1.2\}$ in order to get around 40\% of sporadically missing values on the response $y_{ij}$ in the entire data set.
+Additionally, we ensured that the $y_{ij}^*$ observations were systematically missing in 20% of the clusters included in the data set.
+<!-- How did we ensure that these were systematically missing in 20% of the clusters?
+Does that follow from the fixing of the betas?
+Or did you remove the y values in 20% of the clusters?
+Please be specific.
+-->
 We assumed that random effects were independent within equations ($b_{h0}^O\indep b_{h1}^O \indep b_{h2}^O$ and $b_{h0}^S \indep b_{h1}^S \indep b_{h2}^S$), but were linked between both selection and outcome equations through a bivariate normal distributed as:
@@
 with the parameters $\sigma_{b0}^2=\sigma_{b1}^2=\sigma_{b2}^2=0.4$.
-We considered that the correlation parameter of the random effects between equations is a 40% of the value of the assumed correlation parameter between error terms $\rho$.
-In addition, we included a random effect on the exclusion restriction variable given by $b_3\sim N(0,0.2)$ assuming that the intra-cluster variation in the exclusion restriction effect is lower than the variation on other coefficient parameters effects.
-The $\rho$ parameter adopts different values depending on the simulated missing mechanism (See below additional scenarios) .
+We considered that the correlation parameter of the random effects between equations is 40% of the value of the assumed correlation parameter between error terms $\rho$.
+In addition, we included a random effect on the exclusion restriction variable given by $b_3\sim N(0,0.2)$ assuming that the intracluster variation in the exclusion restriction effect is lower than the variation on other coefficient parameters effects.
+The $\rho$ parameter was given different values depending on the simulated missing mechanism (See below additional scenarios) .
 
@@
 We consider the
-- \textbf{Violation of distributional assumptions:} To assess how imputation models behave in the face of deviations from normality assumptions, we simulated data in which the errors followed a skewed t-distribution and also in which the missing process follows a MNAR mechanism with an explicit truncated model, i.e., the participation is directly related to the value of the outcome variable.
-- \textbf{Binary response}: We evaluated the imputation method when missing variables is binary.
-Therefore we simulated  $y_{i}$ binary incomplete variables, we keep the parameters similar to the ones used in the simulation of missing continuous variables,but the observable binary variables was defined as:
+<!-- The sentence above stops mid-sentence. -->
+- \textbf{Violation of distributional assumptions:} To assess how the imputation models behave in the face of deviations from normality assumptions, we simulated data in which the errors followed a skewed t-distribution and also in which the missing process follows a MNAR mechanism with an explicit truncated model, i.e., the participation is directly related to the value of the outcome variable.
+- \textbf{Binary response}: We evaluated the imputation method when the missing variables are binary.
+Therefore we simulated  $y_{i}$ binary incomplete variables, we keep the parameters similar to the ones used in the simulation of missing continuous variables, but the observable binary variables were defined as:
 \begin{align*}
@@
 \end{align*}
+<!-- One binary variable or multiple binary variables? -->
 
@@
 The estimands were the parameter coefficients of the outcome equation $\beta^O={\beta^O_0,\beta^O_1,\beta^O_2}$, with special emphasis on the treatment effect parameter $\beta^0_1$.
-We also report in the estimated variance of the random effects and residual errors $\sigma_{b0}^2$,$\sigma_{b1}^2$,$\sigma_{b2}^2$,$\sigma_{e}^2$.
+We also report the estimated variance of the random effects and residual errors $\sigma_{b0}^2$,$\sigma_{b1}^2$,$\sigma_{b2}^2$,$\sigma_{e}^2$.
 
-After the imputation procedure, we estimated the following (generalized) mixed linear effect model using the lmer() function.
+After the imputation procedure, we estimated the following (generalized) mixed linear effect model using the *lmer()* function from the *lme4* R package.
+<!-- Add ref to lme4 package -->
 $y_{i}=\beta_i^OX_i^O+\epsilon^O_{i}$
-In case of missing binary variable, we used the same matrix of predictors but on a binary model estimated with glmer() functions.
-Then, we pooled the estimates of the $\beta_i^O$ and the variance of random effect and residual errors of the multiple imputed datasets according to Rubin's rule [@rubin1987], over which we calculated the performance measures on estimands.
+In case of missing binary variable, we used the same matrix of predictors but on a binary model estimated with the *glmer()* function from the *lme4* R package.
+Then, we pooled the estimates of the $\beta_i^O$ and the variance of the random effect and residual errors of the multiple imputed datasets according to Rubin's rule [@rubin1987], over which we calculated the performance measures on the estimands.
 
-To calculate the coverage of the parameter coefficients, we estimate confidence intervals (CI) with the Wald method.
+To calculate the coverage of the parameter coefficients' \95% confidence intervals (CI), we estimate CI with the Wald method.
 Although it is possible to obtain CI through the profile or bootstrap method for the variances of the random effects, we prefer not to estimate them (and hence the coverage) for the random effects parameters due to computational time.
@@
 
-In addition, in the appendix table we reported the empirical standard errors (EmpSE), Monte Carlo standard errors (ModSE) on the coefficient  parameters,average time processing (time in seconds) and the percentage of datasets where the imputation method converged (run), i.e., the imputation method generated an output.
+In addition, in the appendix table we reported the empirical standard errors (EmpSE), Monte Carlo standard errors (ModSE) on the coefficient  parameters, average processing time (time in seconds) and the percentage of datasets where the imputation method converged (run), i.e., the imputation method generated an output.
 
 #### Software
-For simulation study and illustrative examples we used  R version 4.0.4  in a linux environment.
-The Heckman 2L imputation method is available  in mice R package (\textbf{mice.2l.heckman}) and also on the github repository \url{https://github.com/johamunoz/Heckman2l} where you can also find all the codes accompanying this paper and a toy example that explains how to implement the method in mice.
+For the simulation study and illustrative examples we used R version 4.0.4 in a linux environment. 
+<!-- Add a reference to R if you haven't done so yet. Run citation() in R to get a bibtex entry. -->
+The Heckman 2L imputation method is available in the mice R package (as \textbf{mice.2l.heckman()}) and also on the github repository \url{https://github.com/johamunoz/Heckman2l} where you can also find all the codes accompanying this paper and a toy example that explains how to implement the method in mice.
 
@@
 At the cluster level, looking only at the sporadic missing clusters, we found an average of 35.25\% missing values, but there were clusters with no missing data at all and up to 98\% missing data in Y.
-Regarding the binary response scenarios, the proportion mean of treatment is around 60% CI[20%,90%].
+Regarding the binary response scenarios, the proportion of treated  patients is around 60% CI[20%,90%].
 Processing times differ in the imputation methods evaluated, for example, for 10 clusters of 1000 units, the 1l.
@@
 
-Figure \ref{fig:rhocont} shows the results of simulations where the missing variable was continuous.In the MAR scenario, i.e.
+Figure \ref{fig:rhocont} shows the results of simulations where the missing variable was continuous.
+In the MAR scenario, i.e.
 when $\rho=0$, all imputation methods provide similar unbiased estimates of the coefficient parameters, but as rho increases, i.e.
-the mechanism becomes MNAR, the estimates for the complete case analysis and the MAR-IPD imputation method turns more biased.
+the mechanism becomes MNAR, the estimates for the complete case analysis and the MAR-IPD imputation method become biased.
 
 Overall, both Heckman-based imputation (1l.
-and 2l.) models provide less biased estimates of the coefficient parameters in MNAR scenarios, but the random effects estimates on the 1l.Heckman are far away from the true values as no cluster information was considered at all.
+and 2l.) models provide less biased estimates of the coefficient parameters in MNAR scenarios, but the random effects estimates on the 1l.heckman are far away from the true values as no cluster information was considered at all.
+
  Regarding the coverage, the 2l.Heckman imputation method provides the best coverage values across all the $\rho$ scenarios with a coverage level close to the nominal 95\% interval.
-Even 2l.Heckman is not properly a randomization-valid (Rubin 1987b) method across all the $\rho$ values it leads to better results in terms of bias and coverage compared to the evaluated methods.
+Even 2l.Heckman is not properly a randomization-valid [@buuren18c] method across all the $\rho$ values it leads to better results in terms of bias and coverage compared to the evaluated methods.
+ <!-- What do you mean with "Even" here? Do you mean to say "Even though" ? Can you add one sentence on what you mean that it is not randomization valid? -->
 
-The 2l.Heckman method, by incorporating individual and marginal level information, allows a bias-variance trade off, thus, as it was observed the 2l.Heckman separated resulted in better estimates in terms of RMSE than the other methods evaluated.
-In particular, the method provides an advantage over the 1l.Heckman method in systematic missingness scenarios, since the latter does not allow to specify any cluster information implicitly, i.e.
+The 2l.Heckman method, by incorporating individual and marginal level information, allows a bias-variance trade-off.
+Thus, we observed that the Heckman IPD separated resulted in better estimates in terms of RMSE than the other methods evaluated.
+<!-- What is "Heckman IPD separated" Is that the 2l.Heckman method? -->
+In particular, the method provides an advantage over the 1l.Heckman method in systematic missingness scenarios, since the latter does not allow the specification of any cluster information implicitly, i.e.
 by adding a group variable in the imputation model.
 This can be seen in the estimates of the random effects parameters of the 1l.Heckman method, which are more biased than those of the other methods in which cluster information was included.
-However, the accuracy (Width) of the 2l.Heckman model is higher than that of the other evaluated methods, which generally increases as $\rho$ moves away from zero.
-In particular, we noticed in $\beta_2$ that the width was larger in the MAR scenario, when we checked the simulation results, it seems that the high variability came from simulations in which the linear model estimation had singularity problems.
+
+<!-- I would not equate accuracy with CI width -->
+However, the 95\% CI width of the 2l.Heckman model is greater than that of the other evaluated methods, which generally increases as $\rho$ moves away from zero.
+In particular, we noticed in $\beta_2$ that the width was larger in the MAR scenario.
+It seems that the high variability came from simulations in which the linear model estimation had singularity problems (not shown).
 
 ##### Results sensitivity to number of clusters and sample size of clusters
-We evaluated how robust our method was to variations in the number of clusters and also in the sample size of cluster.
+We evaluated how robust our method was to variations in the number of clusters and also in the sample size of cluster (Figure \ref{fig:rhosize}).
 
@@
 
-By increasing N, the number of clusters from 10 to 100 ( Figure \ref{fig:rhosize}), we observe that the bias was not affected but the average width of the estimates decreases (larger precision) and hence the RMSE.
-On the other hand, by decresing the number of units per cluster (from $n_i$=1000 to $n_i$=100) the precision decreased for all coefficients, the bias on coefficient estimates were not drastically affected but the variances of random effects do.
+<!-- Again, try to use only one tense: past or present tense. This paragraph mixed both within a single sentence. It is standard to write the results and discussion sections in past tense. -->
+<!-- What do you mean with "the average width of the estimates"?
+Are you referring to the CI width for the individual estimates?
+Or are you referring to the width of the estimates of the bias as resulting from monte carlo error?
+-->
+By increasing N, the number of clusters, from 10 to 100, we observe that the bias was not affected but the average width of the estimates decreases (larger precision) and hence the RMSE decreased.
+On the other hand, by decreasing the number of units per cluster (from $n_i$=1000 to $n_i$=100) the precision decreased for all coefficients, the bias on coefficient estimates were not drastically affected but the variances of random effects did.
 
-When we reduced the sample size to 50 patients per study (not shown here but available in the appendix), the bias and RMSE of the $\sigma_b2$ was drastically affected, with large CI on these criteria.
-This could be explained in part for the scarce information on certain cluster that affects directly the estimation of the Heckman model on those clusters.
-
-##### Results deviations in distribution assumptions
-
-```{r dist,fig.pos="!H",fig.cap="\\label{fig:dist} Comparison of methods for continuous incomplete variable with deviations in distribution assumptions, the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"}
-plot_S
-```
-
-To investigate how the imputation models behaves in settings with departures of bivariate normal distribution, we draw error terms from a bivariate skew student-t distribution using the same location parameter and covariance matrix of the normal distributed settings,  with 4 degrees of freedom and an $\alpha=\{-2,6\}$ parameter which regulates the the slant of the density.
-In addition we simulated a explicit missing process, where error terms of selection and outcome equation were normally independent distributed and the selection of observations depends on the value of the outcome variable, as $ry^*_i=0.3y^*_i+\epsilon^S_{i}$.These settings assure that the percentage of missingness on the outcome variable is around 60\% in all the evaluated scenarios.
-
-Figure \ref{fig:dist} depicts the results of this scenario.
-When we use the Heckman model in an explicit MNAR process ($ry^*=f(y^*)$), we observe that our model might not be completely suitable for this type of scenario, as we can observe that it highly affects the bias of the intercept parameter ($\beta_0$) and the width of all the coefficient parmeters, Also the bias of the random effects parameters ($\sigma_0$ and $\sigma_2$) was affected.
-With respect to the t-distributed error scenario, apparently only the bias of $\beta_0$ was affected.
-For both scenarios, it is also seen that the coverage in the $\beta_0$ was highly affected.
+When we reduced the sample size to 50 patients per study (Appendix), the bias and RMSE of the $\sigma_b2$ were drastically affected, with variability on these criteria.
+This could be explained in part due to the scarce information on certain clusters, which affects directly the estimation of the Heckman model on those clusters.
 
 ##### Bivariate incomplete variable
+This is part of the main analyses is in't it? I moved it up accordingly.
 We also find (Figure \ref{fig:bin}) that the 2l.Heckman method provides unbiased results on coefficient parameters and random effects when the missing variable is binary.
@@
 
-# An ilustrative study
+### Sensitivity analysis: distributional assumptions
+In this sensitivity analysis, we aimed to investigate the importance of distributional assumptions regarding the error terms.
+
+```{r dist,fig.pos="!H",fig.cap="\\label{fig:dist} Comparison of methods for continuous incomplete variable with deviations in distribution assumptions, where the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"}
+plot_S
+```
+<!-- Ideally, the phrase " where the dashed line depicts the true simulated parameter" belongs in a note under the caption. -->
+
+<!-- This next paragraph belongs in the methods section.
+Alternatively, you could make a (sub) section for a sensitivity analysis, where you could describe both the methods and results of this sensitivity analysis (which I did here.
+-->
+
+#### Methods
+To investigate how the imputation models behave in settings with departures from the bivariate normal distribution, provide two sensitivity analyses.
+1.
+\textbf{Skewed-t:} We drew error terms from a bivariate skewed student-t distribution using the same location parameter and covariance matrix of the normal distributed settings,  with 4 degrees of freedom and an $\alpha=\{-2,6\}$ parameter which regulates the the slant of the density.
+2.
+\textbf{Name 2 :} In addition we simulated an explicit missingness process, where error terms of the selection and outcome equations were independently normal distributed and the selection of observations depended on the value of the outcome variable, as $ry^*_i=0.3y^*_i+\epsilon^S_{i}$.
+These settings assure that the percentage of missingness on the outcome variable is around 60\% in all the evaluated scenarios.
+3. \textbf{Normal :} As a reference, we provide the results from the basic scenario of the main simulation study, where the error terms were Normally distributed.  <!-- Multivariate normal? -->
+
+<!-- Replace "Name 2" with a short name for this scenario, preferably not an equation. Then use the same name in the Figure. -->
+
+#### Results
+
+When we use the Heckman model in an explicit MNAR process ($ry^*=f(y^*)$), we observe that our model might not be completely suitable for this type of scenario, as it highly affected the bias of the intercept parameter ($\beta_0$) and the width of the 95\% CI of all the coefficient parameters \ref{fig:dist}.
+Also the bias of the random effects parameters ($\sigma_0$ and $\sigma_2$) was affected.
+
+With respect to the t-distributed error scenario, only the bias of $\beta_0$ was affected. For both scenarios, it is also seen that the coverage in the $\beta_0$ was highly affected.
+<!-- I don't see $\sigma_0$ and $\sigma_2$ in the figure. Are they called $\sigma_{b0}$ and $\sigma_{b2}$ there? -->
+<!-- Can you add some more description of the results of the sensitivity analyses?
+For instance, that in case of the skewed distribution the other methods are also biased and also have poor coverage?
+For the explicit MNAR process, the other methods are also biased.
+-->
+
+# An illustrative study
 \label{example}
@@
 library(here)
-source(here('3.Ilustrative_study','Ilustrative_study.R'),local = knitr::knit_global())
+# source(here('3.Ilustrative_study','Ilustrative_study.R'),local = knitr::knit_global())
 ```
@@
 
-Specifically, in Uganda, under the Uganda LLIN evaluation project, a LLINS distribution campaign was conducted between 2013 and 2014, and in 2017, the effect of LLIN control together with insecticides was assessed through a cross-sectional community survey in 104 health sub-districts in 48 districts located within 5 sub-regions of Uganda.
+<!-- Can you clarify this? Do you mean that both the campaign and the survey took place in 2013, 2014 and 2017? In that case i suggest rephrasing -->
+<!-- else split it up into two sentences to make it easier to read (see below). -->
+Specifically, in Uganda, under the Uganda LLIN evaluation project, a LLINS distribution campaign was conducted between 2013 and 2014.
+In 2017, the effect of LLIN control together with insecticides was assessed through a cross-sectional community survey in 104 health sub-districts in 48 districts located within 5 sub-regions of Uganda.
 
-In each sub-district, a sample of households with at least one child aged 2-10 years were surveyed, where information was collected on household conditions and use of preventive measures.
-In addition, finger prick blood samples were taken from each child to determine the prevalence of parasitaemia and an etymological study was conducted to estimate mosquito prevalence, details of the project and survey are provided elsewhere [@staedke2019].
+In each sub-district, a sample of households with at least one child aged 2-10 years was surveyed, where information was collected on household conditions and use of preventive measures.
+In addition, finger prick blood samples were taken from each child to determine the prevalence of parasitaemia and an etymological study was conducted to estimate mosquito prevalence.
+Details of the project and survey are provided elsewhere [@staedke2019].
 
@@
 Blood samples were only taken from 8846 children, as 69 were excluded from the study due to lack of consent and 2222 were not present at the time of the survey.
-Although the original data set consists of 164 variables, here we only consider the variables described in the table \ref{tab:descriptive}, which were used as predictors of the imputation model.
+Although the original data set consists of 164 variables, here we only consider the variables described in Table \ref{tab:descriptive}, which were used as predictors in the imputation model.
 
 ```{r descriptive, results = "asis", table.pos="!H", echo=FALSE}
-if (!require("xtable")) install.packages("xtable")
-xt <- xtable(sum.table,
-             caption = "Descriptive analysis, predictor variables", label = "tab:descriptive",align= c("p{0.0\\textwidth}|", 
-                                   "p{0.11\\textwidth}|",                       
-                                   "p{0.07\\textwidth}|", 
-                                   "p{0.07\\textwidth}|", 
-                                   "p{0.05\\textwidth}|", 
-                                   "p{0.10\\textwidth}|", 
-                                   "p{0.12\\textwidth}|", 
-                                   "p{0.07\\textwidth}|", 
-                                   "p{0.07\\textwidth}|", 
-                                   "p{0.07\\textwidth}|", 
-                                   "p{0.07\\textwidth}|"))
-names(xt) <- c("Subregion","Districts (N)","Children (N)","Age mean (years)",
-                "Log10 Female Anopheline","Wealth index",
-                "Bednet (%)","Females (%)","Holidays (%)",
-                "Missing test(%)")
-print(xt, comment = FALSE, include.rownames=FALSE)
+# if (!require("xtable")) install.packages("xtable")
+# xt <- xtable(sum.table,
+#              caption = "Descriptive analysis, predictor variables", label = "tab:descriptive",align= c("p{0.0\\textwidth}|",
+#                                    "p{0.11\\textwidth}|",
+#                                    "p{0.07\\textwidth}|",
+#                                    "p{0.07\\textwidth}|",
+#                                    "p{0.05\\textwidth}|",
+#                                    "p{0.10\\textwidth}|",
+#                                    "p{0.12\\textwidth}|",
+#                                    "p{0.07\\textwidth}|",
+#                                    "p{0.07\\textwidth}|",
+#                                    "p{0.07\\textwidth}|",
+#                                    "p{0.07\\textwidth}|"))
+# names(xt) <- c("Subregion","Districts (N)","Children (N)","Age mean (years)",
+#                 "Log10 Female Anopheline","Wealth index",
+#                 "Bednet (%)","Females (%)","Holidays (%)",
+#                 "Missing test(%)")
+# print(xt, comment = FALSE, include.rownames=FALSE)
 ```
 
-To illustrate our method, following the article by @rugnao2019, we estimated the prevalence of parasitemia by sub-region and by age after approximately 3 years of LLIN campaigns started.
-We estimated parasitemia prevalence using 3 approaches: MCAR, MAR and MNAR.
+<!-- I see you wrote both parasitemia and parasitaemia. Choose one -->
+To illustrate our proposed method, following the article by @rugnao2019, we estimated the prevalence of parasitemia by subregion and by age after approximately 3 years of LLIN campaigns started.
+We estimated parasitemia prevalence using 3 approaches that made different assumptions on the missingness mechanism: MCAR, MAR and MNAR.
 
-In the MCAR approach, prevalence was calculated on the basis of the actual test, i.e., we only included patients with a test result.
-In the case of MAR, the test values of children who were not present during the survey were imputed with the 2l.2stage.bin method of the MICEMD package, where the community was taken as the cluster and the following factors previously associated with parasitemia were used as predictors in the imputation model: sex, two-person mosquito net.
-In addition, age was included as a power 3 spline function, the cluster-level Log10 mean of the number of female anopheline mosquitoes per household estimated from the etymological survey, and the household wealth index from principal components analysis calculated specifically for the surveyed households.
+Under the MCAR assumption, prevalence was calculated on the basis of the recorded tests, i.e., we only included patients with a test result.
+Under the MAR assumption, the test values of children who were not present during the survey were imputed with the 2l.2stage.bin method of the MICEMD package, where the community was taken as the cluster and the following factors previously associated with parasitemia were used as predictors in the imputation model: sex, two-person mosquito net.
+In addition, we included age as a power 3 spline function, the cluster-level Log10 mean of the number of female anopheline mosquitoes per household estimated from the etymological survey, and the household wealth index from principal components analysis calculated specifically for the surveyed households.
 
-Under the MNAR assumption, we used the proposed 2l.Heckman method to impute missing test values.  The selection and outcome equation included the same predictor variables used under the MAR approach. 
+Under the MNAR assumption, we used the proposed 2l.Heckman method to impute missing test values.
+The selection and outcome equation included the same predictor variables as used under the MAR approach.
 In addition, we included a holiday indicator variable as ERV, which was calculated according to school vacation calendars and public holidays in Uganda in 2017.
@@
 ```{r ERV, results = "asis", table.pos="!H", echo=FALSE}
-if (!require("xtable")) install.packages("xtable")
-xt <- xtable(tablexc,
-             caption = "Evaluation of Holidays as exclusion restriction variable", label = "tab:exclusion")
-names(xt) <- c("","Test result (y)"," Missing indicator (ry)")
-comment<- list( pos=list(0),command=NULL)
-comment$pos[[1]]<-c(nrow(xt))
-comment$command<-c(paste("\\hline",
-                         "***p<0.01",sep=" "))
-
-print(xt, comment = FALSE, include.rownames=FALSE,add.to.row = comment, 
-      hline.after = c(-1,0))
+# if (!require("xtable")) install.packages("xtable")
+# xt <- xtable(tablexc,
+#              caption = "Evaluation of Holidays as exclusion restriction variable", label = "tab:exclusion")
+# names(xt) <- c("","Test result (y)"," Missing indicator (ry)")
+# comment<- list( pos=list(0),command=NULL)
+# comment$pos[[1]]<-c(nrow(xt))
+# comment$command<-c(paste("\\hline",
+#                          "***p<0.01",sep=" "))
+# 
+# print(xt, comment = FALSE, include.rownames=FALSE,add.to.row = comment,
+#       hline.after = c(-1,0))
 ```
+<!-- What are non-participants? Children with missing values? -->
 According to our imputation approach, non-participants were estimated to have a higher prevalence of malaria than participants in more than half of the districts analyzed. 
-
-As can be seen in Figure \ref{fig:region} in terms of sub-region level prevalence the estimates of the approaches do not differ significantly between methods, however except for the East-Central region, prevalence estimates under the Heckman approach are higher than those estimated using MAR or CC approaches.
-```{r region,fig.pos="!H",fig.cap="\\label{fig:region} Estimates of malaria prevalence by Uganda subregion",echo=FALSE,fig.keep="last"}
-plot_dist
+As can be seen in Figure \ref{fig:region}, for each subregion the prevalence estimates of the approaches do not differ significantly between methods.
+However, prevalence estimates under the MNAR assumption (i.e.
+2 level Heckman)  are higher than those estimated under the MAR or MCAR aproaches, except for the East-Central region.
+```{r region,fig.width = 5, fig.asp =.67, fig.pos="!H", fig.cap="\\label{fig:region} Estimates of malaria prevalence by Uganda subregion",echo=FALSE,fig.keep="last"}
+# plot_dist
 ```
 
-In terms of prevalences by age, there are no significant differences between methods overall.
-The prevalence estimates for children aged 2 to 6 years according to the approaches evaluated in all regions are very similar (Figure \ref{figure:region_age}, which could be partly explained by the mobility of children at this age compared to that of school-age children.
+<!-- The age at which children start going to school is not universal, so you need to mention this. -->
+In terms of prevalence by age, there are no significant differences between methods (Figure \ref{figure:region_age}).
+The prevalence estimates for children aged 2 to 6 years are very similar in all regions under the different assumptions.
+This could be partly explained by the mobility of children at this age compared to that of school-age children.
+<!-- Can you explain in one sentence how mobility of children is related to prevalence here? -->
+
 ```{r region_age, fig.pos="!H", fig.cap="\\label{fig:region_age} Estimation of malaria prevalence by region and age",echo=FALSE,fig.keep="last"}
-plot_sdist_age
+# plot_sdist_age
 ```
@@
 However for school children, prevalences estimated with the Heckman method were found to be higher in the Mid-East and Southwest regions than those obtained with the other methods, whereas in the East-Central region the estimates with the Heckman method are lower.
-Possible reasons for selection bias in surveys of this type have been suggested [@thedhsprogram2020], for example, that daytime visits might favor measurement in sick school children who stay home, leading to overestimated prevalence results as found in the East-Central region.
-Nevertheless, we were unable to find information that suggests or confirms the direction in which malaria prevalence is driven by selection bias in this Uganda study or in other studies similar to this one.
+A possible reasons for selection bias in surveys of this type is, for example, that daytime visits might favor measurement in sick school children who stay home, leading to overestimated prevalence results as found in the East-Central region.
+[@thedhsprogram2020] Nevertheless, we were unable to find information that suggests or confirms the direction in which malaria prevalence is driven by selection bias in this Uganda study or in other studies similar to this one.
 
@@
 \label{discussion}
-This work was done in order to extend multiple imputation for clustered datasets, in situations where some incomplete variables follow a MNAR mechanism.
-For clustered datasets, only imputation methods under the MAR mechanism have been previously proposed and although imputation methods exist to handle MNAR they have only been designed for individual studies, which makes them limited in common IPD situations such as systematic missingness or when the proportion of missingness of a variable is very high in one of the included studies.
-In this context, a new multiple imputation method was proposed to handle continuous and binary MNAR covariates specifically for clustered dataset, which also allows appropriate borrowing of information between the clusters to obtain more reliable imputation results at the individual cluster level.
+We have extended and evaluated methods for multiple imputation of clustered datasets, in situations where some incomplete variables follow a MNAR mechanism.
+For clustered datasets, only imputation methods under the MAR mechanism had previously been proposed.
+Although imputation methods exist to handle MNAR they have only been designed for individual studies.
+This makes them limited in common IPD situations such as systematic missingness or when the proportion of missingness of a variable is very high in one of the included studies.
+In this context, we proposed a new multiple imputation method to handle continuous and binary MNAR covariates specifically for a clustered dataset, which also allows appropriate borrowing of information between the clusters to obtain more reliable imputation results at the individual cluster level.
+<!-- If you say " a new multiple imputation method was proposed", it is unclear who proposed it. It is fine to say " We proposed a new multiple imputation method" (which is also shorter). -->
 
@@
 Overall the method produced unbiased estimates with convergence close to 95\% for the fixed effects parameters with variation at the cluster level and also unbiased estimators for the random effects parameters.
+<!-- What are "the fixed effects parameters with variation at the cluster level"? Isn't that a contradiction ? -->
 
-Empirically, the method was shown to be robust to systemic and sporadic missingness in individual studies.
-This method, in particular, could provide more robust imputation values compared to individual-level imputation methods, as it not only allows for imputation of missing values in clusters with systematic losses, but can also shrink the values of individual clusters towards the global mean of studies, particularly advantageous in studies with extreme values or with values far away from those found on average at the global level.
+Empirically, we showed that the proposed method was robust to systemic and sporadic missingness in individual studies.
+This method, in particular, could provide more robust imputation values compared to individual-level imputation methods, as it not only allows for imputation of missing values in clusters with systematic losses, but can also shrink the values of individual clusters towards the global mean of studies.
+This is particularly advantageous in studies with extreme values or with values far away from those found on average at the global level.
+<!-- But why is it advantageous for these studies?
+If the prevalence truly is large in these studies, they should not be shrunk.
+It may be advantageous (I guess) because these studies are too small to estimate the prevalence reliably.
+I suggest something like this:
 
-The advantage of the proposed method over the previous ones is that it allows the imputation of variables from cluster level data following a MAR or MNAR mechanism according to Heckman's model.
-That is to say that under the specification of a validity exclusion variable the method determines by itself which is the most adjustable  correlation parameter between equations ($\rho$), or in general terms the  missingness mechanism  (MAR or MNAR), in each of the clusters evaluated.
-The imputation method is built on the mice package, which allows, first of all, to be used both on the outcome and on the covariates and, in addition, offers the option of being used simultaneously with other imputation methods of the package, advantageous in databases containing missing variables with different prediction methods and models.
+"This is particularly advantageous in studies with small sample sizes, where an analysis approach that ignores the data from other studies may lead to extreme estimates of prevalence.
+When using our proposed method, these would be shrunk to the overall mean."
+-->
+
+The advantage of the proposed method over methods that assume MAR is that it allows the imputation of variables from cluster level data following a MAR or MNAR mechanism according to Heckman's model.
+That is to say that under the specification of a valid exclusion variable the method determines by itself which is the most adjustable  correlation parameter between equations ($\rho$), or in general terms the  missingness mechanism  (MAR or MNAR), in each of the clusters evaluated.
+Our implementation of the imputation method is built on the mice R package, which allows, first of all, to be used both on the outcome and on the covariates.
+In addition, it offers the option of being used simultaneously with other imputation methods implemented the package, which is advantageous in databases containing missing variables with different prediction methods and models.
 Finally, the method can be used on systematically and sporadically missing clusters, both for continuous variables with heterogeneous error variance and for binary variables.
 
-One of the major limitations of our method is that it needs a valid restriction variable, which in some contexts is difficult to establish at the individual study level and can be even more challenging if one tries to find a valid exclusion variable across clusters.
+## Limitations and future directions
 
-Also the method is sensitive to the value of the correlation between the selection equation and outcome (rho), and in general it is observed that it can lead to biased results on fixed global parameters i.e.
-without variation across clusters.Similarly, the method can be sensitive to both the sample size and the number of studies included in the database.
-On the one hand a small sample size at the individual study level can affect not only the precision of estimates but also the convergence of the method since it requires a minimum sample size to estimate all the parameters of the Heckman model which can be at least twice the number of parameters required in a MAR prediction model.
+A major limitation of our method is that it needs a valid restriction variable, which in some contexts is difficult to establish at the individual study level and can be even more challenging if one tries to find a valid exclusion variable across clusters.
+Also, the method is sensitive to the value of the correlation between the selection equation and outcome (rho), and in general it is observed that it can lead to biased results on fixed global parameters i.e.
+without variation across clusters.
+Similarly, the method can be sensitive to both the sample size and the number of studies included in the database.
+On the one hand, a small sample size at the individual study level can affect not only the precision of estimates but also the convergence of the method since it requires a minimum sample size to estimate all the parameters of the Heckman model which can be at least twice the number of parameters required in an imputation model that assumes MAR.
 On the other hand, a high number of studies that may improve the precision of the estimators may also make the estimation of the marginal parameters more difficult and also considerably increase the processing time of our method.
@@
 The data were simulated by attributing a constant correlation across all clusters in order to evaluate the performance against M(N)AR assumptions, but in practice this parameter is variable across clusters which can considerably affect the performance of the method.
-So in future research the effect of this parameter could be more deeply evaluated.
+Therefore, in future research the effect of this parameter could be more deeply evaluated.
+One might also consider relaxing this assumption of constant correlation to allow for a random effects distribution for the correlation parameter.
+<!-- Does my suggestion make sense? -->
 
-On the other hand, the method can also be extended to other types of variables such as count or ordinal variables.
+Further, the method can also be extended to other types of variables such as count or ordinal variables.
 Similarly, less restrictive Heckman based models can be considered in terms of normality distribution of errors and no specification of exclusion variables such as those proposed by Ogundimu.
+<!-- Add reference to Ogundimu -->
+
+### Conclusion
+<!-- Always add a conclusion, also in stats papers. I'm thinking it should mention the following. I am not so sure on especially the last sentence, so feel free to rephrase -->
+We have <!--(well actually you) --> proposed an extension to the Heckman model that can account for MNAR, MAR or MCAR of a continuous or binary variable in clustered data sets.
+Our simulations showed that it has favorable statistical properties, when assumptions were met, and provided that the sample size was sufficiently large.
+Regarding deviations from distributional assumptions of the error terms, the coefficient parameters were fairly robust in terms of bias, but the intercept was not.
+
+<!-- Can you add this disclaimer for me at the bottom? -->
+<!-- Also, I added recodid. If Harlan or Paul joins as co-author, don't forget to add the Canadian line as well -->
+<!-- If the editors ask you to make a separate Data Availability section, you can add that below here has well -->
+
+# Footnotes {.unnumbered}
+### Disclaimer {.unnumbered}
+The views expressed in this paper are the personal views of the authors and may not be understood or quoted as being made on behalf of or reflecting the position of the regulatory agency/agencies or organizations with which the authors are employed/affiliated.
+
+### Acknowledgements {.unnumbered}
+<!-- I usually put recodid as follows in latex: -->
+\begin{minipage}{.08\textwidth}
+\includegraphics[width=\textwidth]{flag_EU}
+\end{minipage}
+\begin{minipage}{.90\textwidth}
+This project has received funding from the European Union's Horizon 2020 research and innovation programme under ReCoDID grant agreement No 825746.
+\end{minipage}

### 2022-11-04 1.Paper/Paper.Rmd
update paper

@@
 author:
-- name: J. Munoz*
+- name: Johanna Munoz*
   num: a
 - name: Valentijn M. T. de Jong
-  num: a, b
+  num: a,b
 - name: Thomas. P. A. Debray
-  num: a
+  num: a,c
 address:
@@
 - num: b
-  org: Data Analytics and Methods Task Force, European Medicines Agency, Amsterdam, the Netherlands
+  org: Data Analytics and Methods Task Force, European Medicines Agency, Amsterdam, The Netherlands
+- num: c
+  org: Smart Data Analysis and Statistics, Utrecht, The Netherlands
+
 corres: "* Johanna Munoz. \\email{j.munozavila@umcutrecht.nl}"
@@
 These methods build upon the principles of Heckman selection models, and adopt a two-stage meta-analysis approach for imputing binary and continuous variables.
-After evaluating the proposed imputation models in simulated scenarios, we illustrated their use in a malaria cross-sectional community survey to estimate the prevalence of parasitemia for children aged 2-10 years in five subregions in Uganda."
+After evaluating the proposed imputation models in simulated scenarios, we illustrated their use in a cross-sectional community survey to estimate the prevalence of malaria parasitemia in children aged 2-10 years in five subregions in Uganda."
+
 keywords: Heckman model; IPDMA; Missing not at random; Selection models;Multiple imputation;
@@
 ---
-<!-- I tried to add a second affiliation for me, but I get 1,1 behind my name instead of 1,2. Do you happen to know how to fix this? -->
-<!-- I find the last sentence of the abstract a bit hard to read. Does the following rephrasing make sense? -->
-<!-- "..., we illustrated their use in a cross-sectional community survey to estimate the prevalence of malaria parasitemia in children aged 2-10 years in five subregions in Uganda."  -->
+<!-- I tried to add a second affiliation for me, but I get 1,1 behind my name instead of 1,2. Do you happen to know how to fix this? Aparently we should not add space between comas :S -->
+<!-- I find the last sentence of the abstract a bit hard to read. Does the following rephrasing make sense? yes!! thanks i just add yours!! -->
 
@@
 For example, the clinical practice research datalink (CRPD) [@herrett2015] is an electronic health record dataset in the UK, which has been used in a variety of medical research, such as the evaluation of health policy and drug efficacy.
-A recent example of an IPD-MA is the emerging risk factor collaboration [@theemergingriskfactorscollaboration2007], where data were combined from approximately 1.1 million individuals across 104 studies to investigate associations of cardiovascular diseases with several predictors.
+A recent example of an IPD-MA is the emerging risk factor collaboration, [@theemergingriskfactorscollaboration2007] where data were combined from approximately 1.1 million individuals across 104 studies to investigate associations of cardiovascular diseases with several predictors.
 Individuals in these large datasets tend to be clustered in centres, countries or studies, where they have been subject to similar healthcare processes, and are therefore more alike than individuals from another cluster.
@@
 
-Although it is possible to formally rule out MCAR [@little1988], 
+Although it is possible to formally rule out MCAR, [@little1988]  
 <!-- Stat med uses the AMA citation style.
@@
 See an example here:
-https://onlinelibrary.wiley.com/doi/epdf/10.1002/sim.9221 -->
-it is impossible to assess whether data are MAR or MNAR as the observable data are not enough to test the assumptions of both mechanisms [@enders2022].
-For this reason, researchers often conveniently assume that data are MAR, or present results that are based on a complete case analysis.
+https://onlinelibrary.wiley.com/doi/epdf/10.1002/sim.9221 Cool thanks a lot!! -->
+it is impossible to assess whether data are MAR or MNAR as the observable data are not enough to test the assumptions of both mechanisms.[@enders2022]  For this reason, researchers often conveniently assume that data are MAR, or present results that are based on a complete case analysis.
 In practice, however, unless missingness is artificially introduced by study design, e.g.
@@
 It is rather that the MAR mechanism becomes stronger than the MNAR mechanism.
-I have tried to capture this with the alternative text -->
+I have tried to capture this with the alternative text !!
+perfect thanks :D-->
 For instance, healthcare professionals may decide what type of measurement on indication, and depending on their personal experience or behavior decide whether or not to record the reasons for the measurement.
@@
 The MNAR mechanism is considered as non-ignorable, because to deal with this type of missingness it is necessary to specify information about the missingness process in addition to assumptions about the observed data.
-A common approach to address MNAR is to adopt  selection models such as the one proposed by Heckman (1976) [@heckman1976].
-Briefly, the Heckman selection model corrects for selection bias by estimating two linked equations, a main equation where the missing variable is associated with predictors, and a selection equation that accounts for the inclusion of observations in the sample.
+A common approach to address MNAR is to adopt  selection models such as the one proposed by Heckman (1976).
+[@heckman1976] Briefly, the Heckman selection model corrects for selection bias by estimating two linked equations, a main equation where the missing variable is associated with predictors, and a selection equation that accounts for the inclusion of observations in the sample.
 An important feature of the Heckman selection model is that its implementation does not require data to be MNAR, and can also be used when data are MCAR or MAR.
@@
 Over the past few years, several extensions and adaptations to the Heckman selection model have been proposed for multiple imputation.
-Among them, Galimard et al.(2016) [@galimard_etal16] implemented a chained equations imputation method for continuous variables, which was extended to binary and categorical variables by employing copula estimates [@galimard_etal18].
-Also, Ogundimu and Collins (2019) [@ogundimu_collins19] proposed a chained equations imputation method that is less dependent on Normality assumptions.
+Among them, Galimard et al.(2016) [@galimard_etal16] implemented a chained equations imputation method for continuous variables, which was extended to binary and categorical variables by employing copula estimates.
+[@galimard_etal18]  Also, Ogundimu and Collins (2019) [@ogundimu_collins19] proposed a chained equations imputation method that is less dependent on Normality assumptions.
 
-In clustered data sets, multilevel imputation methods are needed to properly propagate uncertainty within and across clusters [@audigier_etal18].
-However, to our knowledge, existing multilevel imputation methods mainly focus  on situations where data are MAR, and do not adopt Heckman selection models.
+In clustered data sets, multilevel imputation methods are needed to properly propagate uncertainty within and across clusters.
+[@audigier_etal18]  However, to our knowledge, existing multilevel imputation methods mainly focus  on situations where data are MAR, and do not adopt Heckman selection models.
 Although Hammon and Zinn (2020) [@hammon_zinn20] recently proposed an extension that allows for the inclusion of random intercept effects, it can only be used for binary missing variables and assume that the effect of explanatory variables on the missingness mechanisms is common across clusters.
@@
 \label{heckman}
-The Heckman selection model was initially proposed as a method to correct for selection bias, in which individuals are not randomly selected from the population, leading to inconsistent estimates and erroneous conclusions [@heckman1976].
+The Heckman selection model was initially proposed as a method to correct for selection bias, in which individuals are not randomly selected from the population, leading to inconsistent estimates and erroneous conclusions.
+[@heckman1976]
 
-Selection bias occurs when the selection of a subject  or their observation into the sample is influenced by unobserved variables  (e.g., the respondent's level of trust toward healthcare entities may cause them to self-select out of the study or refuse to sign consent for a test), which in turn are correlated with unobserved variables related to a variable of interest (e.g., the result of a blood test) [@vella1998].
+Selection bias occurs when the selection of a subject  or their observation into the sample is influenced by unobserved variables  (e.g., the respondent's level of trust toward healthcare entities may cause them to self-select out of the study or refuse to sign consent for a test), which in turn are correlated with unobserved variables related to a variable of interest (e.g., the result of a blood test).
+[@vella1998]
 
@@
 Generally the same variables can be used on the matrix of predictor variables $X^O_{i}$ and $X^S_{i}$.
-However, to avoid multicollinearity problems [@puhani1997], it is recommended to include in $X^S_{i}$ at least one variable that is not included in the outcome model [@angrist2001].
-This variable is commonly known as an exclusion restriction variable, and should only be associated with the selection in the sample $r^*_{ij}$ but not with the actual observation $y^*_{ij}$.
+However, to avoid multicollinearity problems, [@puhani1997] it is recommended to include in $X^S_{i}$ at least one variable that is not included in the outcome model.
+[@angrist2001] This variable is commonly known as an exclusion restriction variable, and should only be associated with the selection in the sample $r^*_{ij}$ but not with the actual observation $y^*_{ij}$.
 <!-- Can you give some explanation (1-3 sentences?) on the exclusion and restriction assumptions here? -->
@@
 
-Under the BVN assumption, the parameters of the Heckman model coefficients can be estimated using the two-step Heckman method (H2S) [@heckman1976] or the full information maximum likelihood method (FIML) [@amemiya1984].
-However, both methods can lead to inconsistent estimators when this underlying distribution cannot be assumed[@gomes_etal20].
-To overcome this problem, other approaches have been proposed that relax the distribution assumptions, among them copula models [@smith03].
-The copula approach uses a function, known as a copula, that joins the marginal distribution of the error terms of the selection and outcome equation which are specified separately.
+Under the BVN assumption, the parameters of the Heckman model coefficients can be estimated using the two-step Heckman method (H2S) [@heckman1976] or the full information maximum likelihood method (FIML).[@amemiya1984] However, both methods can lead to inconsistent estimators when this underlying distribution cannot be assumed.
+[@gomes_etal20] To overcome this problem, other approaches have been proposed that relax the distribution assumptions, among them copula models.
+[@smith03] The copula approach uses a function, known as a copula, that joins the marginal distribution of the error terms of the selection and outcome equation which are specified separately.
 $$ F(r^*_{i},y^*_{i}) =\Phi\left(r^*_{i}-X^S_{i}\beta^S_{i},\frac{y^*_{i}-X^O_{i}\beta^O_{i}}{\sigma_i}, \rho_i\right)$$
@@
 <!-- I wouldn't call the distribution imaginary. It is assumed to be real but latent. Some words we could use: conjectured, latent, presupposed, presumed  -->
-In a random effects model, $\theta_i$ parameters are assumed to be drawn independently and identically from a latent distribution of parameters with a  population mean $\theta_m$ and a population variance $\psi$ [@higgins2009].
-Thus, the  $\theta_i=\theta_m + b_i$ can be specified using $b_i \sim N(0,\psi)$ random effects to allow for between-study heterogeneity in observed data relationships, and between-study heterogeneity in missing patterns.
+In a random effects model, $\theta_i$ parameters are assumed to be drawn independently and identically from a latent distribution of parameters with a  population mean $\theta_m$ and a population variance $\psi$.
+[@higgins2009] Thus, the  $\theta_i=\theta_m + b_i$ can be specified using $b_i \sim N(0,\psi)$ random effects to allow for between-study heterogeneity in observed data relationships, and between-study heterogeneity in missing patterns.
 
@@
 Initially, we use the copulat method to estimate the set of cluster-specific parameters, $\widehat{\theta_i}=\{\widehat{\beta^O_i}, \widehat{\beta^S_i}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, using all $j$ units with observable measurements $y_{ij}^{obs}$ within each group $i$.
-The Heckman model is estimated with the \textbf{gjrm} function of the GJRM R package under the bivariate model with the nonrandom sample selection (BSS) specification, from which we obtain not only the parameters' point estimates $\widehat{\theta_i}$, but also their corresponding $\widehat{S(\theta_i)}$ within-cluster variance-covariance matrix.
+The Heckman model is estimated with the \textbf{gjrm} function of the GJRM R package, [@radiceGJRMGeneralisedJoint2021] under the bivariate model with the nonrandom sample selection (BSS) specification, from which we obtain not only the parameters' point estimates $\widehat{\theta_i}$, but also their corresponding $\widehat{S(\theta_i)}$ within-cluster variance-covariance matrix.
 
@@
 The meta-analysis model is performed with the \textbf{mixmeta} function of the R package *mixmeta*, which allows the use of maximum likelihood (ML), restricted maximum likelihood (REML), and moments estimation methods.
-For the simulation and illustrative study, we used the restricted REML estimation method, which is recommended as it has a good balance between insensitivity and efficiency [@viechtbauer2005].
+For the simulation and illustrative study, we used the restricted REML estimation method, which is recommended as it has a good balance between insensitivity and efficiency.
+[@viechtbauer2005]
 <!-- Insensitivity is not generally a good thing.
@@
 
-The missing $y_{ij}^{miss}$ is drawn from a Bernoulli distribution with $p_{ij}^*$ proportion parameter given by $P[y_{ij}=1|r_{ij}=0]$, the conditional probability that $y_{ij}=1$ given that the measure is unobservable ($r_{ij}=0$), in a bivariate probit model [@greene2018]:
+The missing $y_{ij}^{miss}$ is drawn from a Bernoulli distribution with $p_{ij}^*$ proportion parameter given by $P[y_{ij}=1|r_{ij}=0]$, the conditional probability that $y_{ij}=1$ given that the measure is unobservable ($r_{ij}=0$), in a bivariate probit model: [@greene2018]
 \begin{align*}
@@
 When there are simultaneous missing variables in a dataset, our imputation method can be extended in a Gibbs sampler procedure.
-Particularly, our imputation method has been implemented according to the structure of the MICE R package  <!--[@buuren2021]-->, that allows imputing multiple incomplete predictors and covariates in a given dataset.
+Particularly, our imputation method has been implemented according to the structure of the MICE R package,[@buuren2021] that allows imputing multiple incomplete predictors and covariates in a given dataset.
 
-<!-- The next paragraph belongs in the manual for the package, or a vignette. I don't think it is necessary to put it here. -->
+<!-- The next paragraph belongs in the manual for the package, or a vignette. I don't think it is necessary to put it here. 
 The MICE package allows one to specify imputation methods to each of the missing variables by setting the method and the predictive matrix for each of the missing variables.
@@
 Please refer to the toy example in the attached github repository to better understand how to implement the imputation model.
+-->
 
@@
 
-After the imputation procedure, we estimated the following (generalized) mixed linear effect model using the *lmer()* function from the *lme4* R package.
+After the imputation procedure, we estimated the following (generalized) mixed linear effect model using the *lmer()* function from the *lme4* R package.[@bates2022]
 <!-- Add ref to lme4 package -->
@@
 In case of missing binary variable, we used the same matrix of predictors but on a binary model estimated with the *glmer()* function from the *lme4* R package.
-Then, we pooled the estimates of the $\beta_i^O$ and the variance of the random effect and residual errors of the multiple imputed datasets according to Rubin's rule [@rubin1987], over which we calculated the performance measures on the estimands.
+Then, we pooled the estimates of the $\beta_i^O$ and the variance of the random effect and residual errors of the multiple imputed datasets according to Rubin's rule, [@rubin1987] over which we calculated the performance measures on the estimands.
 
@@
 - \textbf{Complete case analysis (CCA):} We removed all patients with missing observations. 
-- \textbf{1l.Heckman:} Multiple imputation based on the Heckman model without no study specification, following the imputation method proposed by Galimard et al.(2016) [@galimard_etal16].
-- \textbf{2l.MAR}: Multiple imputation assuming MAR for hierarchical datasets, we used the multilevel imputation model proposed by Resche-Rignon and (2018) [@resche-rigon2018].
+- \textbf{1l.Heckman:} Multiple imputation based on the Heckman model without no study specification, following the imputation method proposed by Galimard et al.(2016). [@galimard_etal16]
+- \textbf{2l.MAR}: Multiple imputation assuming MAR for hierarchical datasets, we used the multilevel imputation model ( 2l.2stage.norm and 2l.2stage.bin) from the micemd R package, [@audigier2021] which are described in Audigier et al.
+(2018) paper.
+[@audigier_etal18]
+
 - \textbf{2l.Heckman}: The proposed imputation method based on the Heckman model for hierarchical datasets.
@@
 
-We calculated the following evaluation criteria [@buuren18c] according to the formulas provided in Morris et al.(2019)[@morris2019]:
+We calculated the following evaluation criteria, usually employed to evaluate imputation methods, [@buuren18c] according to the formulas provided in Morris et al.(2019): [@morris2019]
 
@@
  Regarding the coverage, the 2l.Heckman imputation method provides the best coverage values across all the $\rho$ scenarios with a coverage level close to the nominal 95\% interval.
-Even 2l.Heckman is not properly a randomization-valid [@buuren18c] method across all the $\rho$ values it leads to better results in terms of bias and coverage compared to the evaluated methods.
+Even though 2l.Heckman is not properly a randomization-valid method [@buuren18c], as it is not unbiased and with a coverage above 95\% across all the $\rho$ values, it leads to better results in terms of bias and coverage compared to the evaluated methods.
  <!-- What do you mean with "Even" here? Do you mean to say "Even though" ? Can you add one sentence on what you mean that it is not randomization valid? -->
@@
 The 2l.Heckman method, by incorporating individual and marginal level information, allows a bias-variance trade-off.
-Thus, we observed that the Heckman IPD separated resulted in better estimates in terms of RMSE than the other methods evaluated.
+Thus, we observed that the 2l.Heckman resulted in better estimates in terms of RMSE than the other methods evaluated.
 <!-- What is "Heckman IPD separated" Is that the 2l.Heckman method? -->
@@
 In addition, finger prick blood samples were taken from each child to determine the prevalence of parasitaemia and an etymological study was conducted to estimate mosquito prevalence.
-Details of the project and survey are provided elsewhere [@staedke2019].
+Details of the project and survey are provided elsewhere.
+[@staedke2019]
 
-For this example, we used data accessed directly from CliniEpiDB [@staedke], where data were collected from 5195 households with verified consent, inhabited by 11137 residents aged 2-10 years.
+For this example, we used data accessed directly from CliniEpiDB, [@staedke] where data were collected from 5195 households with verified consent, inhabited by 11137 residents aged 2-10 years.
 Blood samples were only taken from 8846 children, as 69 were excluded from the study due to lack of consent and 2222 were not present at the time of the survey.

### 2022-11-07 4.Codes/Functions_HPC.R
Update Functions_HPC.R

@@ -783,7 +783,7 @@ get_estimates <- function( data = data, # data with missing values
     m <- 1
     
   }else {
-    if(ImpID != 3){ #HeckmanIPD or Riche-Rignon
+    if(ImpID != 3){ #HeckmanIPD or 2l.2stage.norm
       
       name <- par_imp$dm[[ImpID,"name"]]
       methv <- ifelse(par_imp$dm[[ImpID,"meth"]] == "R"&family == "probit",

### 2022-11-10 1.Paper/Paper.Rmd
Update paper

@@
 
-Although it is possible to formally rule out MCAR, [@little1988]  
 <!-- Stat med uses the AMA citation style.
@@
 https://onlinelibrary.wiley.com/doi/epdf/10.1002/sim.9221 Cool thanks a lot!! -->
-it is impossible to assess whether data are MAR or MNAR as the observable data are not enough to test the assumptions of both mechanisms.[@enders2022]  For this reason, researchers often conveniently assume that data are MAR, or present results that are based on a complete case analysis.
+Although it is possible to formally rule out MCAR, [@little1988]  it is impossible to assess whether data are MAR or MNAR as the observable data are not enough to test the assumptions of both mechanisms.[@enders2022]  For this reason, researchers often conveniently assume that data are MAR, or present results that are based on a complete case analysis.
 In practice, however, unless missingness is artificially introduced by study design, e.g.
@@
 I have tried to capture this with the alternative text !!
-perfect thanks :D-->
+perfect thanks, your point is correct and now sound so much better :D-->
 For instance, healthcare professionals may decide what type of measurement on indication, and depending on their personal experience or behavior decide whether or not to record the reasons for the measurement.
@@
 In section \ref{heckman} we provide an introduction to the Heckman model and its estimation, and we extended to a hierarchical setting.
-Then in section \ref{methods} we define the main steps of the proposed imputation method.
+Then in section \ref{methodsn} we define the main steps of the proposed imputation method.
 In section \ref{simulation} we provide the settings and results of a Monte Carlo simulation study used to evaluate the performance of the proposed imputation method.
@@
 Here $\beta^O_{i}$ and $\beta^S_{i}$ are $p\times1$ and $q\times1$ coefficient parameter vectors and $\epsilon^O_{i}=(\epsilon^O_{i1},\epsilon^O_{i2},...,\epsilon^O_{in_i})^T$ and $\epsilon^S_{i}=(\epsilon^S_{i1},\epsilon^S_{i2},...,\epsilon^S_{in_i})^T$ are the residual terms vectors for the outcome and selection equations, respectively.
+
 Generally the same variables can be used on the matrix of predictor variables $X^O_{i}$ and $X^S_{i}$.
 However, to avoid multicollinearity problems, [@puhani1997] it is recommended to include in $X^S_{i}$ at least one variable that is not included in the outcome model.
-[@angrist2001] This variable is commonly known as an exclusion restriction variable, and should only be associated with the selection in the sample $r^*_{ij}$ but not with the actual observation $y^*_{ij}$.
+[@angrist2001] This variable is commonly known as an exclusion restriction variable (ERV), and should only be associated with the selection in the sample $r^*_{ij}$ (relevance condition) but not with the actual observation $y^*_{ij}$ (exclusion condition).[@gomes_etal20] The ERV should meet the relevance and exclusion conditions in order to provide independent information about the selection processing in order to facilitate the estimation of the Heckman model specially when dataset does not follow the BVN assumption.
 <!-- Can you give some explanation (1-3 sentences?) on the exclusion and restriction assumptions here? -->
@@
 
-<!-- This sentence was too long and has the verbs all the way at the end, making it hard to read. I suggest to rephrase to three shorter sentences: -->
-Briefly, in a first stage $\theta_i$ the cluster specific parameters of the Heckman model are estimated.
-That is, the parameters of the two equations in each study, the outcome model and the selection model, are estimated separately for each of the $N$ clusters.
-In the second stage, all $\theta_i$ are combined using a random effects meta-analysis model.
+<!-- This sentence was too long and has the verbs all the way at the end, making it hard to read. I suggest to rephrase to three shorter sentences: OK!!-->
+Briefly, in a first stage  we estimate $\theta_i$ the cluster specific parameters of the Heckman model.
+That is, we estimate separately for each of the $N$ clusters,the parameters of the two equations in each study, the outcome model and the selection model.
+In the second stage, we combine all $\theta_i$  by using a random effects meta-analysis model.
 
-<!-- I wouldn't call the distribution imaginary. It is assumed to be real but latent. Some words we could use: conjectured, latent, presupposed, presumed  -->
+<!-- I wouldn't call the distribution imaginary. It is assumed to be real but latent. Some words we could use: conjectured, latent, presupposed, presumed, Thanks :D -->
 In a random effects model, $\theta_i$ parameters are assumed to be drawn independently and identically from a latent distribution of parameters with a  population mean $\theta_m$ and a population variance $\psi$.
@@
 # Method
-\label{methods}
+\label{methodsn}
 We follow a similar approach proposed by Resche-Rigon and White (2018) [@resche-rigon2018] for multilevel imputation of data.
@@
 Our imputation approach differs crucially from the previous approach as here we estimate two correlated equations (instead of a single outcome equation) in each cluster, thus obtaining $\theta_i=\{\beta_i^O,\beta_i^S,\sigma_i,\rho_i\}$ parameters from both equations which are then pooled into a $\theta_m$ parameter set at the marginal or population level.
-Our method is basically a univariate imputation method, but since it is implemented in a Gibbs sampling procedure, it can also be used to impute multiple incomplete variables in a data set.
-<!-- I don't understand what you mean here: it is univariate and it can be used to impute multiple variables? I would say these two statements contradict each other. -->
+Our imputation method was created for impute univariate incomplete datasets, but since it is implemented in a Gibbs sampling procedure, it can also be used to impute multiple incomplete variables in a data set.
+<!-- I don't understand what you mean here: it is univariate and it can be used to impute multiple variables?
+I would say these two statements contradict each other.
+yes you are right i tried to say is a mehtod for univariate incomplete dataset -->
 
@@
 
-<!-- Can you add a reference to the R package here? -->
+<!-- Can you add a reference to the R package here? OK -->
 
@@
 In this step, we pool the parameters $\widehat{\theta_i}$ with a random effects meta-analysis model using only the groups with observable information, i.e., those that are not systematically missing and have sufficient information to estimate the heckman model.
-In particular, we pooled the $p$ coefficients of the $\beta^O$ outcome equation and estimated a multivariate random effects meta-analysis model with them, similarly we combined all $q$ coefficient parameters of the $\beta^S$ selection equation.
+In particular, we pool the $p$ coefficients of the $\beta^O$ outcome equation and estimated a multivariate random effects meta-analysis model with them, similarly we combine all $q$ coefficient parameters of the $\beta^S$ selection equation.
 <!-- The phrase " the log-transformed parameter of $\sigma$" is a description of $\sigma'$, so it is what is known as a commenting or parenthetical clause.
@@
 Although this article is not for the BMJ, I can recommend reading their house-style for writing: https://www.bmj.com/about-bmj/resources-authors/house-style -->
-<!-- Read more here if you like: http://content.principia.edu/teaching-excellence/commenting-with-commas-when-and-where-to-use-them/ -->
+<!-- Read more here if you like: http://content.principia.edu/teaching-excellence/commenting-with-commas-when-and-where-to-use-them/ Thanks a lot you are a walking library :D, very appreciated all your lessons!!
+-->
 We also performed a univariate random effects meta_analysis on $\sigma'$, the log-transformed parameter of $\sigma$, and another on $\rho'$, the fisher-transformed parameter of $\rho$.
@@
 The meta-analysis model is performed with the \textbf{mixmeta} function of the R package *mixmeta*, which allows the use of maximum likelihood (ML), restricted maximum likelihood (REML), and moments estimation methods.
-For the simulation and illustrative study, we used the restricted REML estimation method, which is recommended as it has a good balance between insensitivity and efficiency.
+For the simulation and illustrative study, we used the restricted REML estimation method, which is recommended as it has a good balance between unbiasedness and efficiency.
 [@viechtbauer2005]
@@
 it is insensitive to the true value of the mean of the distribution of random effects and it is efficient" (or something similar, feel free to rephrase) , or simplify it to "...
-a good balance between unbiasedness and efficiency."  -->
+a good balance between unbiasedness and efficiency.
+jajaj I think at certain point i used the translator to correct my writing and i did not notice that change unbiasedness by insensitive..But thanks God you have eagle eyes @@) .."  -->
 
@@
 <!-- Put the symbol after the text if the text is a description of the symbol: "shrunk-cluster parameters $\theta^*_i$" . -->
-<!-- Put the text after the symbol if the symbol quantifies the text: "$n$ patients" , or " $N$ clusters" . -->
+<!-- Put the text after the symbol if the symbol quantifies the text: "$n$ patients" , or " $N$ clusters" . Whoahh you are a gramatical Lord thanks a lot!! -->
 We draw the shrunk-cluster parameters $\theta^*_i$ for each group $i$ from the following posterior distribution conditional on $\theta^*_m$ and $\psi^*$.
@@
 Conversely, when $\widehat{S_{\theta_i}}>\psi^*$, more weight is given to the estimated marginal parameters.
-In the case of systematic missingness, it is like considering the within-cluster variance to be infinite ($\widehat{S_{\theta_i}}\to\infty$), then letting all parameters rely only on the marginal estimates.
-<!-- What does then mean in this last part? should it be " and lettting all parameters rely ..." ? -->
+Therefore, in case of a cluster with systematic missingness, it is as if the within-cluster variance is infinite ($\widehat{S_{\theta_i}}\to\infty$), so all the weight is assigned to the parameters estimated at the marginal level.
+That is, when there is no information in a cluster, we rely entirely on the marginal information.
+
+<!-- What does then mean in this last part? should it be " and lettting all parameters rely ..." ? I think now is more clear, thanks!! -->
 
@@
 
-Having $\theta^*_i$ the shrunk-cluster parameters vector for each group, we back-transform $\sigma^*$ and $\rho^*$ to the original scale.
-Then $y^{miss}_{ij}$ the missing values can be drawn from $p(y^{miss}_{ij}|\theta^*_i)$ their predictive distribution given $\theta^*_i$ as follows:
+Having $\theta^*_i$, the shrunk-cluster parameters vector for each group, we back-transform $\sigma^*$ and $\rho^*$ to the original scale.
+Then $y^{miss}_{ij}$, the missing values, can be drawn from $p(y^{miss}_{ij}|\theta^*_i)$, their predictive distribution given $\theta^*_i$, as follows:
 
@@
 
-The imputed value of the $y_{ij}^{miss}$ missing observation can be drawn from the conditional expectation of $y_{ij}$ on unobserved measurements:
+The imputed value of $y_{ij}^{miss}$ can be drawn from the conditional expectation of $y_{ij}$ on unobserved measurements:
 \begin{align*}
@@
 
-The missing $y_{ij}^{miss}$ is drawn from a Bernoulli distribution with $p_{ij}^*$ proportion parameter given by $P[y_{ij}=1|r_{ij}=0]$, the conditional probability that $y_{ij}=1$ given that the measure is unobservable ($r_{ij}=0$), in a bivariate probit model: [@greene2018]
+When $y_{ij}^{miss}$ is a binary variable, the imputed value is drawn from a Bernoulli distribution with a proportion parameter $p_{ij}^*$  given by $P[y_{ij}=1|r_{ij}=0]$, that is the conditional probability that $y_{ij}=1$ given that the measure is unobservable ($r_{ij}=0$).
+The $p_{ij}^*$ is obtained from a bivariate probit model, as follows: [@greene2018]
 \begin{align*}
@@
 
-#### Aim
+## Aim
 We designed a simulation study aimed to compare the performance of imputation methods for imputing a missing variable in a hierarchical dataset, where the missingness follows a MNAR mechanism.
 
-#### Data-generation mechanism
+## Data-generation mechanism
 
 We generated the data from the Heckman selection model with bivariate normal distribution error terms.
-For simplicity we assume that the database collects information from $N=10$ clusters of equal number of individuals $n_i=1000$.
+For simplicity we assume a "basic scenario" where the database collects information from $N=10$ clusters of equal number of individuals $n_i=1000$.
+However, we altered both $N$ and $n_i$ in the sensitivity analysis (see below).
 <!-- But you also have scenarios with different numbers for N and n_i.
-You could write here that for the "basic scenario" we have N =  and n_i = , and that we alter this in other scenarios, or in " senstivity analyses" .
+You could write here that for the "basic scenario" we have N =  and n_i = , and that we alter this in other scenarios, or in " senstivity analyses" ok .
 -->
@@
 We fixed coefficients $\beta_h^O=\{0.3,1,1\}$  and $\beta_h^S=\{-0.8,1.3,-0.7,1.2\}$ in order to get around 40\% of sporadically missing values on the response $y_{ij}$ in the entire data set.
-Additionally, we ensured that the $y_{ij}^*$ observations were systematically missing in 20% of the clusters included in the data set.
+Additionally, we ensured that the $y_{ij}^*$ observations were systematically missing in 20% of the clusters included in the data set, by removing the outcome values in the 20% of the clusters.
 <!-- How did we ensure that these were systematically missing in 20% of the clusters?
@@
 Please be specific.
--->
+OK-->
 We assumed that random effects were independent within equations ($b_{h0}^O\indep b_{h1}^O \indep b_{h2}^O$ and $b_{h0}^S \indep b_{h1}^S \indep b_{h2}^S$), but were linked between both selection and outcome equations through a bivariate normal distributed as:
@@
 
-##### Adittional scenarios
+## Adittional scenarios
 
@@
 
-- \textbf{M(N)AR scenarios:} We assessed whether the model performed well in terms of bias and coverage when the data followed a missing MAR mechanism ($\rho=0$), and when it followed a MNAR mechanism with a low ($\rho=0.3$), intermediate ($\rho=0.6$) and strong correlation ($\rho=0.9$) between $y^*$ and $r^*$.
-- \textbf{Influence of sample size}: Model sensitivity was analysed with respect to $n_i=\{50,100,1000\}$ the number of patients per cluster and $N=\{10,50,100\}$ the number of studies.
-We consider the
-<!-- The sentence above stops mid-sentence. -->
-- \textbf{Violation of distributional assumptions:} To assess how the imputation models behave in the face of deviations from normality assumptions, we simulated data in which the errors followed a skewed t-distribution and also in which the missing process follows a MNAR mechanism with an explicit truncated model, i.e., the participation is directly related to the value of the outcome variable.
-- \textbf{Binary response}: We evaluated the imputation method when the missing variables are binary.
-Therefore we simulated  $y_{i}$ binary incomplete variables, we keep the parameters similar to the ones used in the simulation of missing continuous variables, but the observable binary variables were defined as:
+- \textbf{M(N)AR scenarios:} We assessed whether the model performed well in terms of bias and coverage when the data followed a missing MAR mechanism ($\rho=0$), and when data followed a MNAR mechanism with a low ($\rho=0.3$), intermediate ($\rho=0.6$) and strong correlation ($\rho=0.9$) between $y^*$ and $r^*$.
+We evaluated also the imputation method when the missing variable is binary.
+Therefore we simulated a $y_{i}$ binary incomplete variable, by keeping similar parameters to the ones used in the simulation of a missing continuous variable, but here we defined the observable binary variable as:
 \begin{align*}
@@
 \end{align*}
-<!-- One binary variable or multiple binary variables? -->
 
-#### Estimand
+- \textbf{Influence of sample size}: Model sensitivity was analysed with respect to the the number of patients per cluster $n_i=\{50,100,1000\}$  and the number of studies $N=\{10,50,100\}$.
+
+- \textbf{Violation of distributional assumptions:} To assess how the imputation models behave in the face of deviations from normality assumptions, we simulated data in which the errors followed a skewed t-distribution and also in which the missing process follows a MNAR mechanism with an explicit truncated model, i.e., the participation is directly related to the value of the outcome variable.
+
+## Estimand
 The estimands were the parameter coefficients of the outcome equation $\beta^O={\beta^O_0,\beta^O_1,\beta^O_2}$, with special emphasis on the treatment effect parameter $\beta^0_1$.
@@
 After the imputation procedure, we estimated the following (generalized) mixed linear effect model using the *lmer()* function from the *lme4* R package.[@bates2022]
-<!-- Add ref to lme4 package -->
 $y_{i}=\beta_i^OX_i^O+\epsilon^O_{i}$
@@
 
-#### Method
+## Method
 
@@
 
-#### Performance measures
+## Performance measures
 
@@
 - \textbf{Coverage:} Coverage of the 95\% confidence intervals for the coefficient  parameters.
-- \textbf{Width}: Width of the confidence interval on the coefficient parameters.
+- \textbf{Width}: Average width of the confidence interval on the coefficient parameters.
 - \textbf{RMSE:} Root mean squared error of the coefficient  and random effect parameters.
@@
 
-#### Software
-For the simulation study and illustrative examples we used R version 4.0.4 in a linux environment. 
-<!-- Add a reference to R if you haven't done so yet. Run citation() in R to get a bibtex entry. -->
-The Heckman 2L imputation method is available in the mice R package (as \textbf{mice.2l.heckman()}) and also on the github repository \url{https://github.com/johamunoz/Heckman2l} where you can also find all the codes accompanying this paper and a toy example that explains how to implement the method in mice.
+## Software
+For the simulation study and illustrative examples we used R version 4.0.4 in a linux environment. [@Rpackage]
+<!-- Add a reference to R if you haven't done so yet. Run citation() in R to get a bibtex entry. Mago!!! -->
+The Heckman 2L imputation method is available in the mice R package (as \textbf{mice.2l.heckman()}) and also on the github repository \url{https://github.com/johamunoz/Statsmed_Heckman} where you can also find all the codes accompanying this paper and a toy example that explains how to implement the method in mice.
 
-#### Results
+## Results
 
@@
 
-##### Descriptive results
+### Descriptive results
 We generated data sets of 10 groups of 1000 patients each in each scenario.
@@
 
-##### Results M(N)AR scenarios
+### Results M(N)AR scenarios
 
- ```{r rhocont,fig.pos="h",fig.cap="\\label{fig:rhocont} Comparison of methods for continuous incomplete variable under systematical missingness by varing $\\rho$, the dashed line depicts the target performance criteria value",echo=FALSE,fig.keep="last"}
+#### Continuous incomplete variable
+ ```{r rhocont,fig.pos="!htb",fig.cap="\\label{fig:rhocont} Continuous incomplete variable by varing $\\rho$, the dashed line depicts the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
 plot_rho
@@
 
+ <!-- What do you mean with "Even" here? Do you mean to say "Even though" ? Can you add one sentence on what you mean that it is not randomization valid? OK -->
+<!-- What is "Heckman IPD separated" Is that the 2l.Heckman method? Yes jaja-->
+<!-- I would not equate accuracy with CI width -->
 Figure \ref{fig:rhocont} shows the results of simulations where the missing variable was continuous.
 In the MAR scenario, i.e.
-when $\rho=0$, all imputation methods provide similar unbiased estimates of the coefficient parameters, but as rho increases, i.e.
-the mechanism becomes MNAR, the estimates for the complete case analysis and the MAR-IPD imputation method become biased.
+when $\rho=0$, all imputation methods provided similar unbiased estimates of the coefficient parameters, but as rho increased, i.e.
+the mechanism becomed MNAR, the estimates for the complete case analysis and the MAR-IPD imputation method became biased.
 
 Overall, both Heckman-based imputation (1l.
-and 2l.) models provide less biased estimates of the coefficient parameters in MNAR scenarios, but the random effects estimates on the 1l.heckman are far away from the true values as no cluster information was considered at all.
+and 2l.) models provided less biased estimates of the coefficient parameters in MNAR scenarios, but the random effects estimates on the 1l.heckman were far away from the true values as no cluster information was considered at all.
 
- Regarding the coverage, the 2l.Heckman imputation method provides the best coverage values across all the $\rho$ scenarios with a coverage level close to the nominal 95\% interval.
-Even though 2l.Heckman is not properly a randomization-valid method [@buuren18c], as it is not unbiased and with a coverage above 95\% across all the $\rho$ values, it leads to better results in terms of bias and coverage compared to the evaluated methods.
- <!-- What do you mean with "Even" here? Do you mean to say "Even though" ? Can you add one sentence on what you mean that it is not randomization valid? -->
+Regarding the coverage, the 2l.Heckman imputation method provided the best coverage values across all the $\rho$ scenarios with a coverage level close to the nominal 95\% interval.
+Even though 2l.Heckman was not properly a randomization-valid method [@buuren18c], as it was not unbiased and had a coverage above 95\% across all the $\rho$ values, it leaded to better results in terms of bias and coverage compared to the evaluated methods.
 
-The 2l.Heckman method, by incorporating individual and marginal level information, allows a bias-variance trade-off.
+The 2l.Heckman method, by incorporating individual and marginal level information, allowed a bias-variance trade-off.
 Thus, we observed that the 2l.Heckman resulted in better estimates in terms of RMSE than the other methods evaluated.
-<!-- What is "Heckman IPD separated" Is that the 2l.Heckman method? -->
-In particular, the method provides an advantage over the 1l.Heckman method in systematic missingness scenarios, since the latter does not allow the specification of any cluster information implicitly, i.e.
+
+In particular, the method provided an advantage over the 1l.Heckman method in systematic missingness scenarios, since the latter does not allow the specification of any cluster information implicitly, i.e.
 by adding a group variable in the imputation model.
-This can be seen in the estimates of the random effects parameters of the 1l.Heckman method, which are more biased than those of the other methods in which cluster information was included.
+This can be seen in the estimates of the random effects parameters of the 1l.Heckman method, which were more biased than those of the other methods in which cluster information was included.
 
-<!-- I would not equate accuracy with CI width -->
-However, the 95\% CI width of the 2l.Heckman model is greater than that of the other evaluated methods, which generally increases as $\rho$ moves away from zero.
-In particular, we noticed in $\beta_2$ that the width was larger in the MAR scenario.
-It seems that the high variability came from simulations in which the linear model estimation had singularity problems (not shown).
+Regarding the CI length of the 95th percentile of the estimated parameters, we observed that using the 2l.Heckman model we obtained larger CIs than when using other evaluated methods, which generally increased as $\rho$ moved away from zero.
+In particular, we observed in $\beta_2$ that the CI length was larger in the MAR scenario.
+By checking the simulation results further, we noted that the high variability came from one simulation in which the linear model estimation had singularity problems (not shown here).
 
-##### Results sensitivity to number of clusters and sample size of clusters
-We evaluated how robust our method was to variations in the number of clusters and also in the sample size of cluster (Figure \ref{fig:rhosize}).
+#### Bivariate incomplete variable
 
-```{r rhosize,fig.pos="h",fig.cap="\\label{fig:size} Comparison of methods for continuous incomplete variable under systematical missingness by varing (N=number of clusters- $n_i$=sample size per cluster), the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"}
+In case of a bivariate missing variable, we also find (Figure \ref{fig:bin}) that the 2l.Heckman method provides unbiased results on coefficient parameters and random effects when the missing variable is binary.
+However, we observe more variability in the estimates of random effects, reflected on longer confidence intervals on the Bias and RMSE criteria.
+
+ ```{r rhobin,fig.pos="!htb",fig.cap="\\label{fig:bin} Binary incomplete variable by varing $\\rho$, the dashed line depicts the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
+plot_bin
+```
+
+### Sensitivity analysis: number of clusters and sample size of clusters
+We evaluated how robust our method was to variations in the number of clusters and also in the sample size of cluster (Figure \ref{fig:size}).
+
+```{r size,fig.pos="h",fig.cap="\\label{fig:size} Continuous incomplete variable under systematical missingness by varing: number of clusters (N); sample size per cluster ($n_i$), the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
 plot_N
@@
 -->
-By increasing N, the number of clusters, from 10 to 100, we observe that the bias was not affected but the average width of the estimates decreases (larger precision) and hence the RMSE decreased.
-On the other hand, by decreasing the number of units per cluster (from $n_i$=1000 to $n_i$=100) the precision decreased for all coefficients, the bias on coefficient estimates were not drastically affected but the variances of random effects did.
+By increasing N, the number of clusters, from 10 to 100, we observed that the bias was not affected but the CI width of the estimates decreased (larger precision) and hence the RMSE decreased.
+On the other hand, by reducing the number of units per cluster (from $n_i$=1000 to $n_i$=100) the precision decreased for all coefficients, the bias on coefficient estimates were not drastically affected but the variances of random effects did.
 
-When we reduced the sample size to 50 patients per study (Appendix), the bias and RMSE of the $\sigma_b2$ were drastically affected, with variability on these criteria.
+When we reduced the sample size to 50 patients per study (Appendix), the bias and RMSE of the $\sigma_b2$ were drastically affected compared to the previous $N; n_i$ settings.
 This could be explained in part due to the scarce information on certain clusters, which affects directly the estimation of the Heckman model on those clusters.
 
-##### Bivariate incomplete variable
-This is part of the main analyses is in't it? I moved it up accordingly.
-We also find (Figure \ref{fig:bin}) that the 2l.Heckman method provides unbiased results on coefficient parameters and random effects when the missing variable is binary.
-However, we observe more variability in the estimates of random effects, reflected on longer confidence intervals on the Bias and RMSE criteria.
+### Sensitivity analysis: distributional assumptions
 
-```{r bin,fig.pos="!H",fig.cap="\\label{fig:bin} Comparison of methods for binary incomplete variable under systematical missingness, the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"}
-plot_bin
-```
-
-### Sensitivity analysis: distributional assumptions
-In this sensitivity analysis, we aimed to investigate the importance of distributional assumptions regarding the error terms.
-
-```{r dist,fig.pos="!H",fig.cap="\\label{fig:dist} Comparison of methods for continuous incomplete variable with deviations in distribution assumptions, where the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last"}
+```{r dist,fig.pos="!H",fig.cap="\\label{fig:dist} Comparison of methods for continuous incomplete variable with deviations in distribution assumptions, where the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
 plot_S
@@
 
-#### Methods
-To investigate how the imputation models behave in settings with departures from the bivariate normal distribution, provide two sensitivity analyses.
+Here, we aimed to investigate how the imputation models behave in settings with departures from the bivariate normal distribution, provide two sensitivity analyses.
+
  1.
@@
  2.
-\textbf{Name 2 :} In addition we simulated an explicit missingness process, where error terms of the selection and outcome equations were independently normal distributed and the selection of observations depended on the value of the outcome variable, as $ry^*_i=0.3y^*_i+\epsilon^S_{i}$.
+\textbf{Explicit:} In addition we simulated an explicit missingness process, where error terms of the selection and outcome equations were independently normal distributed and the selection of observations depended on the value of the outcome variable, as $ry^*_i=0.3y^*_i+\epsilon^S_{i}$.
 These settings assure that the percentage of missingness on the outcome variable is around 60\% in all the evaluated scenarios.
-3. \textbf{Normal :} As a reference, we provide the results from the basic scenario of the main simulation study, where the error terms were Normally distributed.  <!-- Multivariate normal? -->
+ 3. \textbf{Normal:} As a reference, we provide the results from the basic scenario of the main simulation study, where the error terms were Bivariate normal distributed.
 
@@
 
-#### Results
-
-When we use the Heckman model in an explicit MNAR process ($ry^*=f(y^*)$), we observe that our model might not be completely suitable for this type of scenario, as it highly affected the bias of the intercept parameter ($\beta_0$) and the width of the 95\% CI of all the coefficient parameters \ref{fig:dist}.
+When we used the Heckman model in an explicit MNAR process, we observed that our model might not be completely suitable for this type of scenario, as it highly affected the bias of the intercept parameter ($\beta_0$) and the width of the 95\% CI of all the coefficient parameters \ref{fig:dist}.
 Also the bias of the random effects parameters ($\sigma_0$ and $\sigma_2$) was affected.
 
-With respect to the t-distributed error scenario, only the bias of $\beta_0$ was affected. For both scenarios, it is also seen that the coverage in the $\beta_0$ was highly affected.
+With respect to the Skewed-t scenario, only the bias of $\beta_0$ was affected. For both scenarios, it is also seen that the coverage in the $\beta_0$ was highly affected.
+
 <!-- I don't see $\sigma_0$ and $\sigma_2$ in the figure. Are they called $\sigma_{b0}$ and $\sigma_{b2}$ there? -->

### 2022-11-10 2.Simulation/simplot.R

@@ -54,11 +54,15 @@ plot<-ggplot(datagraph[Parameter%in%c("beta[0]","beta[1]","beta[2]","sigma[b * 0
   scale_color_brewer(palette="Dark2")+
   guides(color=guide_legend(title="Imputation method"))+
   theme_light()+  
-  theme(strip.background =element_rect(fill="white"),legend.position="bottom")+
+  theme(strip.background =element_rect(fill="white"),legend.position="bottom",legend.margin=margin(t=-12))+
   theme(strip.text = element_text(colour = 'black'))+
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.15,size=7))+
+  theme(axis.text.x = element_text(angle = 45, vjust = 0.15,size=7),
+        axis.title=element_text(size=8))+
   theme(axis.text.y = element_text(size=7))+
-  xlab("Value")+ylab(yaxis)
+  xlab("Performance measure")+ylab(yaxis)+
+  theme(legend.text=element_text(size=6))+
+  theme(legend.title=element_text(size=7))
+
 return(plot)
 }
 

### 2022-11-10 1.Paper/Paper.Rmd
Update cache

@@
 #### Continuous incomplete variable
- ```{r rhocont,fig.pos="!htb",fig.cap="\\label{fig:rhocont} Continuous incomplete variable by varing $\\rho$, the dashed line depicts the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
+ ```{r rhocont,fig.align='center',fig.pos="!htb",fig.cap="\\label{fig:rhocont} Continuous incomplete variable by varing $\\rho$, the dashed line depicts the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
 plot_rho
@@
 
- ```{r rhobin,fig.pos="!htb",fig.cap="\\label{fig:bin} Binary incomplete variable by varing $\\rho$, the dashed line depicts the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
+ ```{r rhobin,fig.align='center',fig.pos="!htb",fig.cap="\\label{fig:bin} Binary incomplete variable by varing $\\rho$, the dashed line depicts the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
 plot_bin
@@
 
-```{r size,fig.pos="h",fig.cap="\\label{fig:size} Continuous incomplete variable under systematical missingness by varing: number of clusters (N); sample size per cluster ($n_i$), the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
+```{r size,fig.align='center',fig.pos="!htb",fig.cap="\\label{fig:size} Continuous incomplete variable under systematical missingness by varing: number of clusters (N); sample size per cluster ($n_i$), the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
 plot_N
@@
 
-```{r dist,fig.pos="!H",fig.cap="\\label{fig:dist} Comparison of methods for continuous incomplete variable with deviations in distribution assumptions, where the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
+```{r dist,fig.align='center',fig.pos="!htb",fig.cap="\\label{fig:dist} Comparison of methods for continuous incomplete variable with deviations in distribution assumptions, where the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
 plot_S
@@
 \label{example}
-```{r, include=FALSE}
+```{r, include=FALSE,cache = TRUE}
 library(here)
-# source(here('3.Ilustrative_study','Ilustrative_study.R'),local = knitr::knit_global())
+ source(here('3.Ilustrative_study','Ilustrative_study.R'),local = knitr::knit_global())
 ```
@@
 
-```{r descriptive, results = "asis", table.pos="!H", echo=FALSE}
-# if (!require("xtable")) install.packages("xtable")
-# xt <- xtable(sum.table,
-#              caption = "Descriptive analysis, predictor variables", label = "tab:descriptive",align= c("p{0.0\\textwidth}|",
-#                                    "p{0.11\\textwidth}|",
-#                                    "p{0.07\\textwidth}|",
-#                                    "p{0.07\\textwidth}|",
-#                                    "p{0.05\\textwidth}|",
-#                                    "p{0.10\\textwidth}|",
-#                                    "p{0.12\\textwidth}|",
-#                                    "p{0.07\\textwidth}|",
-#                                    "p{0.07\\textwidth}|",
-#                                    "p{0.07\\textwidth}|",
-#                                    "p{0.07\\textwidth}|"))
-# names(xt) <- c("Subregion","Districts (N)","Children (N)","Age mean (years)",
-#                 "Log10 Female Anopheline","Wealth index",
-#                 "Bednet (%)","Females (%)","Holidays (%)",
-#                 "Missing test(%)")
-# print(xt, comment = FALSE, include.rownames=FALSE)
+```{r descriptive,fig.align='center',cache = TRUE, results = "asis", table.pos="!htb", echo=FALSE}
+if (!require("xtable")) install.packages("xtable")
+xt <- xtable(sum.table,
+             caption = "Descriptive analysis, predictor variables", label = "tab:descriptive",align= c("p{0.0\\textwidth}|",
+                                   "p{0.11\\textwidth}|",
+                                   "p{0.07\\textwidth}|",
+                                   "p{0.07\\textwidth}|",
+                                   "p{0.05\\textwidth}|",
+                                   "p{0.10\\textwidth}|",
+                                   "p{0.12\\textwidth}|",
+                                   "p{0.07\\textwidth}|",
+                                   "p{0.07\\textwidth}|",
+                                   "p{0.07\\textwidth}|",
+                                   "p{0.07\\textwidth}|"))
+names(xt) <- c("Subregion","Districts (N)","Children (N)","Age mean (years)",
+                "Log10 Female Anopheline","Wealth index",
+                "Bednet (%)","Females (%)","Holidays (%)",
+                "Missing test(%)")
+print(xt, comment = FALSE, include.rownames=FALSE)
 ```
@@
 
-```{r ERV, results = "asis", table.pos="!H", echo=FALSE}
-# if (!require("xtable")) install.packages("xtable")
-# xt <- xtable(tablexc,
-#              caption = "Evaluation of Holidays as exclusion restriction variable", label = "tab:exclusion")
-# names(xt) <- c("","Test result (y)"," Missing indicator (ry)")
-# comment<- list( pos=list(0),command=NULL)
-# comment$pos[[1]]<-c(nrow(xt))
-# comment$command<-c(paste("\\hline",
-#                          "***p<0.01",sep=" "))
-# 
-# print(xt, comment = FALSE, include.rownames=FALSE,add.to.row = comment,
-#       hline.after = c(-1,0))
+```{r ERV,fig.align='center', cache = TRUE,results = "asis", table.pos="!htb", echo=FALSE}
+if (!require("xtable")) install.packages("xtable")
+xt <- xtable(tablexc,
+             caption = "Evaluation of Holidays as exclusion restriction variable", label = "tab:exclusion")
+names(xt) <- c("","Test result (y)"," Missing indicator (ry)")
+comment<- list( pos=list(0),command=NULL)
+comment$pos[[1]]<-c(nrow(xt))
+comment$command<-c(paste("\\hline",
+                         "***p<0.01",sep=" "))
+
+print(xt, comment = FALSE, include.rownames=FALSE,add.to.row = comment,
+      hline.after = c(-1,0))
 ```
@@
 2 level Heckman)  are higher than those estimated under the MAR or MCAR aproaches, except for the East-Central region.
-```{r region,fig.width = 5, fig.asp =.67, fig.pos="!H", fig.cap="\\label{fig:region} Estimates of malaria prevalence by Uganda subregion",echo=FALSE,fig.keep="last"}
-# plot_dist
+```{r region,fig.align='center',fig.width = 5, fig.asp =.67, fig.pos="!htb", fig.cap="\\label{fig:region} Estimates of malaria prevalence by Uganda subregion",echo=FALSE,fig.keep="last",cache = TRUE}
+ plot_dist
 ```
@@
 
-```{r region_age, fig.pos="!H", fig.cap="\\label{fig:region_age} Estimation of malaria prevalence by region and age",echo=FALSE,fig.keep="last"}
-# plot_sdist_age
+```{r region_age,fig.align='center', fig.pos="!htb", fig.cap="\\label{fig:region_age} Estimation of malaria prevalence by region and age",echo=FALSE,fig.keep="last",cache = TRUE}
+plot_sdist_age
 ```
@@
 
-### Conclusion
+## Conclusion
 <!-- Always add a conclusion, also in stats papers. I'm thinking it should mention the following. I am not so sure on especially the last sentence, so feel free to rephrase -->

### 2022-11-10 2.Simulation/simplot.R

@@ -70,4 +70,4 @@ return(plot)
 plot_rho<-plotsim(Scenariov="Rho",levelsvar=c("0","0.3","0.6","0.9"),yaxis=expression(rho))
 plot_bin<-plotsim(Scenariov="Bin",levelsvar=c("0","0.3","0.6","0.9"),yaxis=expression(rho))
 plot_N<-plotsim(Scenariov="N",levelsvar=c("10;50","10;100","10;1000","50;1000","100;1000"),yaxis="Number of clusters; sample size per cluster")
-plot_S<-plotsim(Scenariov="S",levelsvar=c("Skewed-t","Normal","ry*=f(y*)"),yaxis="Missing process")
+plot_S<-plotsim(Scenariov="S",levelsvar=c("Skewed-t","Normal","Explicit"),yaxis="Missing process")

### 2022-11-10 3.Ilustrative_study/Ilustrative_study.R

@@ -225,30 +225,30 @@ pred["test","sage2"]<- 1
 meth<-ini$method
 meth[c("test")]<-"4l.heckman"
 
-# Full correlation of parameters
-# data_heck <- mice(hdataF0f, # dataset with missing values
-#                   m = 20,   # number of imputations
-#                   seed = 1234, #seed attached to the dataID
-#                   meth = meth, #imputation method vector
-#                   pred = pred, #imputation predictors matrix
-#                   maxit=1,
-#                   meta_method="reml",
-#                   pmm=FALSE)
-# save(data_heck,file=(here('3.Ilustrative_study','MalariaMNAR.RData')))
-load(file=here('3.Ilustrative_study','MalariaMNAR.RData'))
+#Full correlation of parameters
+data_heck <- mice(hdataF0f, # dataset with missing values
+                  m = 20,   # number of imputations
+                  seed = 1234, #seed attached to the dataID
+                  meth = meth, #imputation method vector
+                  pred = pred, #imputation predictors matrix
+                  maxit=1,
+                  meta_method="reml",
+                  pmm=FALSE)
+#save(data_heck,file=(here('3.Ilustrative_study','MalariaMNAR.RData')))
+#load(file=here('3.Ilustrative_study','MalariaMNAR.RData'))
 
 #3.4. Multiple imputation with  2l.2stage.bin method----
 
-# meth[c("test")]<-"2l.2stage.bin" 
-# pred["test","fundays"] <- 0
-# data_mar <- mice( hdataF0f, # dataset with missing values
-#                        m = 20,   # number of imputations
-#                        seed = 1234, #seed attached to the dataID
-#                        meth = meth, #imputation method vector
-#                        pred = pred, #imputation predictors matrix
-#                        maxit=1)
-# save(data_mar,file=(here('3.Ilustrative_study','MalariaMAR.RData')))
-load(file=here('3.Ilustrative_study','MalariaMAR.RData'))
+meth[c("test")]<-"2l.2stage.bin"
+pred["test","fundays"] <- 0
+data_mar <- mice( hdataF0f, # dataset with missing values
+                       m = 20,   # number of imputations
+                       seed = 1234, #seed attached to the dataID
+                       meth = meth, #imputation method vector
+                       pred = pred, #imputation predictors matrix
+                       maxit=1)
+#save(data_mar,file=(here('3.Ilustrative_study','MalariaMAR.RData')))
[…]

### 2022-11-10 1.Paper/Paper.Rmd
Update figures

@@
 #### Continuous incomplete variable
- ```{r rhocont,fig.align='center',fig.pos="!htb",fig.cap="\\label{fig:rhocont} Continuous incomplete variable by varing $\\rho$, the dashed line depicts the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
+ ```{r rhocont,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable by varing $\\rho$, the dashed line depicts the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
 plot_rho
@@
 
-In case of a bivariate missing variable, we also find (Figure \ref{fig:bin}) that the 2l.Heckman method provides unbiased results on coefficient parameters and random effects when the missing variable is binary.
+In case of a bivariate missing variable, we also find (Figure \ref{fig:rhobin}) that the 2l.Heckman method provides unbiased results on coefficient parameters and random effects when the missing variable is binary.
 However, we observe more variability in the estimates of random effects, reflected on longer confidence intervals on the Bias and RMSE criteria.
 
- ```{r rhobin,fig.align='center',fig.pos="!htb",fig.cap="\\label{fig:bin} Binary incomplete variable by varing $\\rho$, the dashed line depicts the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
+ ```{r rhobin,fig.align='center',fig.pos="!htb",fig.cap="Binary incomplete variable by varing $\\rho$, the dashed line depicts the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
 plot_bin
@@
 
-```{r size,fig.align='center',fig.pos="!htb",fig.cap="\\label{fig:size} Continuous incomplete variable under systematical missingness by varing: number of clusters (N); sample size per cluster ($n_i$), the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
+```{r size,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable under systematical missingness by varing: number of clusters (N); sample size per cluster ($n_i$), the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
 plot_N
@@
 
-```{r dist,fig.align='center',fig.pos="!htb",fig.cap="\\label{fig:dist} Comparison of methods for continuous incomplete variable with deviations in distribution assumptions, where the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
+```{r dist,fig.align='center',fig.pos="!htb",fig.cap="Comparison of methods for continuous incomplete variable with deviations in distribution assumptions, where the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
 plot_S
@@
 In each sub-district, a sample of households with at least one child aged 2-10 years was surveyed, where information was collected on household conditions and use of preventive measures.
-In addition, finger prick blood samples were taken from each child to determine the prevalence of parasitaemia and an etymological study was conducted to estimate mosquito prevalence.
+In addition, finger prick blood samples were taken from each child to determine the prevalence of parasitemia and an etymological study was conducted to estimate mosquito prevalence.
 Details of the project and survey are provided elsewhere.
@@
 2 level Heckman)  are higher than those estimated under the MAR or MCAR aproaches, except for the East-Central region.
-```{r region,fig.align='center',fig.width = 5, fig.asp =.67, fig.pos="!htb", fig.cap="\\label{fig:region} Estimates of malaria prevalence by Uganda subregion",echo=FALSE,fig.keep="last",cache = TRUE}
+```{r region,fig.align='center',fig.width = 5, fig.asp =.67, fig.pos="!htb", fig.cap="Estimates of malaria prevalence by Uganda subregion",echo=FALSE,fig.keep="last",cache=TRUE}
  plot_dist
@@
 
-```{r region_age,fig.align='center', fig.pos="!htb", fig.cap="\\label{fig:region_age} Estimation of malaria prevalence by region and age",echo=FALSE,fig.keep="last",cache = TRUE}
+```{r region_age,fig.align='center', fig.pos="!htb", fig.cap="Estimation of malaria prevalence by region and age",echo=FALSE,fig.keep="last",cache = TRUE}
 plot_sdist_age

### 2022-11-10 3.Ilustrative_study/Ilustrative_study.R

@@ -287,13 +287,21 @@ data_long[, c("Statistic", "Parameter") := tstrsplit(Statistic, "_", fixed=TRUE)
 data_wide<- setDT(dcast(data_long, subreg_id + count+ Miss+Parameter ~ Statistic, value.var = "Value"))
 data_wide[,Parameter:=factor(Parameter,levels=c("MCAR","MAR","MNAR"))]
 levels(data_wide$Parameter)<-c("CC","2l.MAR","2l.Heckman")
-
+levels(data_wide$subreg_id)<-c("East\nCentral","Mid\nEastern","Mid\nWestern","North\nEast","South\nWestern")
 pd <- position_dodge(width = 0.4)
 plot_dist<-ggplot(data_wide, aes(x = subreg_id, y = p, colour = Parameter)) +
                   geom_pointrange(aes(ymax = UC, ymin = LC),position = pd) + 
                   geom_point(position = pd)+ theme_light()+
-                  ylab("Parasite prevalence (%)")+xlab("Sub-region")+
-                  scale_color_brewer(palette="Dark2")
+                  ylab("Parasitemia prevalence (%)")+xlab("Sub-region")+
+                  scale_color_brewer(palette="Dark2")+labs(color='Method')+ 
+                  theme(strip.background =element_rect(fill="white"),legend.position="bottom",legend.margin=margin(t=-12))+
+                  theme(axis.text.x = element_text(size=7),
+                        axis.text.y = element_text(size=7),
+                        axis.title=element_text(size=8),
+                        legend.text=element_text(size=6),
+                        legend.title=element_text(size=7))
+
+  
                   
 
 
@@ -323,7 +331,7 @@ mod2<-pred_model(model=mod.mar,source="2l.MAR")
 mod<-rbind(mod0,mod1,mod2)
 setDT(mod)[,source:=as.factor(source)]
 mod[,source:=factor(source,levels=c("CC","2l.MAR","2l.Heckman"))]
-
+levels(mod$subreg_id)<-c("East Central","Mid Eastern","Mid Western","North East","South Western")
 #Prevalence plot
 plot_sdist_age <- ggplot(mod, aes(x = age, y = pred, group = source)) + 
   geom_line(aes(colour=source))+ 
@@ -338,8 +346,13 @@ plot_sdist_age <- ggplot(mod, aes(x = age, y = pred, group = source)) +
   theme(panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank(), 
         legend.position = "bottom",
[…]

### 2022-11-10 1.Paper/Paper.Rmd
tables update

@@
   doublespace: yes
-
+longtable: true
 header-includes:
@@
      \DeclareMathOperator\arctanh{arctanh}
+   - \usepackage{booktabs}
+   - \usepackage{longtable}
+   - \usepackage{array}
+   - \usepackage{multirow}
+   - \usepackage{wrapfig}
+   - \usepackage{float}
+   - \usepackage{colortbl}
+   - \usepackage{pdflscape}
+   - \usepackage{tabu}
+   - \usepackage{threeparttable}
+   - \usepackage{makecell}
+   - \usepackage{xcolor}
+   - \usepackage{array}
+     \newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
+
 ---
@@
 [@vella1998]
-
 \begin{figure}[h!]
+\vspace{-0.7cm}
    \centering
@@
 #### Continuous incomplete variable
- ```{r rhocont,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable by varing $\\rho$, the dashed line depicts the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
+ ```{r rhocont,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable by varing $\\rho$, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
 plot_rho
@@
 
- ```{r rhobin,fig.align='center',fig.pos="!htb",fig.cap="Binary incomplete variable by varing $\\rho$, the dashed line depicts the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
+```{r rhobin,fig.align='center',fig.pos="!htb",fig.cap="Binary incomplete variable by varing $\\rho$, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
 plot_bin
@@
 
-```{r size,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable under systematical missingness by varing: number of clusters (N); sample size per cluster ($n_i$), the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
+```{r size,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable under systematical missingness by varing: number of clusters (N); sample size per cluster ($n_i$), where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
 plot_N
@@
 
-```{r dist,fig.align='center',fig.pos="!htb",fig.cap="Comparison of methods for continuous incomplete variable with deviations in distribution assumptions, where the dashed line depicts the true simulated parameter value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
+```{r dist,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable with deviations in distribution assumptions, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
 plot_S
@@
 
-<!-- Replace "Name 2" with a short name for this scenario, preferably not an equation. Then use the same name in the Figure. -->
+<!-- Replace "Name 2" with a short name for this scenario, preferably not an equation. Then use the same name in the Figure. Thanks a lot it is awesome :D!!-->
 
@@
 library(here)
+
  source(here('3.Ilustrative_study','Ilustrative_study.R'),local = knitr::knit_global())
@@
 Although the original data set consists of 164 variables, here we only consider the variables described in Table \ref{tab:descriptive}, which were used as predictors in the imputation model.
-
-```{r descriptive,fig.align='center',cache = TRUE, results = "asis", table.pos="!htb", echo=FALSE}
-if (!require("xtable")) install.packages("xtable")
-xt <- xtable(sum.table,
-             caption = "Descriptive analysis, predictor variables", label = "tab:descriptive",align= c("p{0.0\\textwidth}|",
-                                   "p{0.11\\textwidth}|",
-                                   "p{0.07\\textwidth}|",
-                                   "p{0.07\\textwidth}|",
-                                   "p{0.05\\textwidth}|",
-                                   "p{0.10\\textwidth}|",
-                                   "p{0.12\\textwidth}|",
-                                   "p{0.07\\textwidth}|",
-                                   "p{0.07\\textwidth}|",
-                                   "p{0.07\\textwidth}|",
-                                   "p{0.07\\textwidth}|"))
-names(xt) <- c("Subregion","Districts (N)","Children (N)","Age mean (years)",
-                "Log10 Female Anopheline","Wealth index",
-                "Bednet (%)","Females (%)","Holidays (%)",
-                "Missing test(%)")
-print(xt, comment = FALSE, include.rownames=FALSE)
+```{r descriptive, echo = FALSE, warning = FALSE,cache = TRUE}
+options(kableExtra.latex.load_packages = FALSE) 
+library(kableExtra)
+kbl(sum.table, booktabs = T, align="r",
+    caption = "Descriptive analysis, predictor variables",
+    col.names = c("Sub-region","District (N)","Children (N)",
+                  "Age mean (years)","Log10 Female Anopheline","Wealth index",
+                "Bednet (%)","Girls (%)","Holiday (%)",
+                "No test (%)"),
+    row.names=FALSE)%>%
+  column_spec(c(1,5,6), width = "2,2cm")%>%
+  column_spec(c(2:3,7:10), width = "1cm")%>%
+  column_spec(c(4), width = "1,5cm")%>%
+  kable_styling(latex_options = c("striped", "scale_down"))
 ```
@@
 The model results in Table \ref{tab:exclusion} indicate that the holiday indicator could be a plausible ERV variable, as it was significantly associated with ry, but not with y.
+```{r exclusion, echo = FALSE, warning = FALSE,cache = TRUE}
+options(kableExtra.latex.load_packages = FALSE) 
+library(kableExtra)
+kbl(tablexc, booktabs = T, align="l",
+    caption = "Evaluation of Holidays as exclusion restriction variable",
+    col.names = c("Predictors/response","Test result (y)","Test taken (ry)"),
+    row.names=FALSE)%>%
+  column_spec(c(1,2,3), width = "2,5cm")%>%
+  kable_styling(position = "center",latex_options = c("striped"))%>%
+  footnote(general="***p<0.01",footnote_as_chunk = T)
+```
 
-```{r ERV,fig.align='center', cache = TRUE,results = "asis", table.pos="!htb", echo=FALSE}
-if (!require("xtable")) install.packages("xtable")
-xt <- xtable(tablexc,
-             caption = "Evaluation of Holidays as exclusion restriction variable", label = "tab:exclusion")
-names(xt) <- c("","Test result (y)"," Missing indicator (ry)")
-comment<- list( pos=list(0),command=NULL)
-comment$pos[[1]]<-c(nrow(xt))
-comment$command<-c(paste("\\hline",
-                         "***p<0.01",sep=" "))
-
-print(xt, comment = FALSE, include.rownames=FALSE,add.to.row = comment,
-      hline.after = c(-1,0))
-```
 <!-- What are non-participants? Children with missing values? -->
-According to our imputation approach, non-participants were estimated to have a higher prevalence of malaria than participants in more than half of the districts analyzed. 
+According to our imputation approach, non-tested children were estimated to have a higher prevalence of malaria than participants in more than half of the districts analyzed. 
 As can be seen in Figure \ref{fig:region}, for each subregion the prevalence estimates of the approaches do not differ significantly between methods.
@@
 2 level Heckman)  are higher than those estimated under the MAR or MCAR aproaches, except for the East-Central region.
-```{r region,fig.align='center',fig.width = 5, fig.asp =.67, fig.pos="!htb", fig.cap="Estimates of malaria prevalence by Uganda subregion",echo=FALSE,fig.keep="last",cache=TRUE}
+```{r region,fig.align='center',fig.width = 5, fig.asp =.67, fig.pos="!htb", fig.cap="Estimates of malaria prevalence by sub-region",echo=FALSE,fig.keep="last",cache=TRUE}
  plot_dist
@@
 
-<!-- The age at which children start going to school is not universal, so you need to mention this. -->
+<!-- The age at which children start going to school is not universal, so you need to mention this. done!! thanks -->
 In terms of prevalence by age, there are no significant differences between methods (Figure \ref{figure:region_age}).
 The prevalence estimates for children aged 2 to 6 years are very similar in all regions under the different assumptions.
-This could be partly explained by the mobility of children at this age compared to that of school-age children.
-<!-- Can you explain in one sentence how mobility of children is related to prevalence here? -->
+Assuming that children start going to school after the age of 6, the results could be partly explained by the mobility of 2-6 year olds compared to school-age children, i.e.
+school-age children spend more time outdoors and travel more than younger children.
+<!-- Can you explain in one sentence how mobility of children is related to prevalence here? OK-->
 
-```{r region_age,fig.align='center', fig.pos="!htb", fig.cap="Estimation of malaria prevalence by region and age",echo=FALSE,fig.keep="last",cache = TRUE}
+```{r region_age,fig.align='center', fig.pos="!htb", fig.cap="Estimates of malaria prevalence by sub-region and age",echo=FALSE,fig.keep="last",cache = TRUE}
 plot_sdist_age
@@
 In this context, we proposed a new multiple imputation method to handle continuous and binary MNAR covariates specifically for a clustered dataset, which also allows appropriate borrowing of information between the clusters to obtain more reliable imputation results at the individual cluster level.
-<!-- If you say " a new multiple imputation method was proposed", it is unclear who proposed it. It is fine to say " We proposed a new multiple imputation method" (which is also shorter). -->
+<!-- If you say " a new multiple imputation method was proposed", it is unclear who proposed it. It is fine to say " We proposed a new multiple imputation method" (which is also shorter). thanks :D-->
 
@@
 
-Overall the method produced unbiased estimates with convergence close to 95\% for the fixed effects parameters with variation at the cluster level and also unbiased estimators for the random effects parameters.
-<!-- What are "the fixed effects parameters with variation at the cluster level"? Isn't that a contradiction ? -->
+Overall, the method produced unbiased estimates with convergence close to 95\% for the coefficient parameters and for the random effects parameters resulted in less biased estimates compared to the other methods evaluated.
 
-Empirically, we showed that the proposed method was robust to systemic and sporadic missingness in individual studies.
-This method, in particular, could provide more robust imputation values compared to individual-level imputation methods, as it not only allows for imputation of missing values in clusters with systematic losses, but can also shrink the values of individual clusters towards the global mean of studies.
-This is particularly advantageous in studies with extreme values or with values far away from those found on average at the global level.
+Empirically, we showed that the proposed method was robust when there are systematic missigness at the cluster level.
+This method, in particular, could provide more robust imputation values compared with individual-level imputation methods, as it not only allows imputation of missing values in clusters with systematic missigness, but can also reduce the values of individual clusters toward the overall study mean.
+This is especially advantageous in studies with small sample sizes, where an analysis approach that ignores data from other studies may lead to extreme estimates of prevalence.
+By using our proposed method, these would be reduced to the overall mean.
+
 <!-- But why is it advantageous for these studies?
@@
 "This is particularly advantageous in studies with small sample sizes, where an analysis approach that ignores the data from other studies may lead to extreme estimates of prevalence.
-When using our proposed method, these would be shrunk to the overall mean."
+When using our proposed method, these would be shrunk to the overall mean.
+Sounds perfect!!
+thanks for ur suggestion."
 -->
 
-The advantage of the proposed method over methods that assume MAR is that it allows the imputation of variables from cluster level data following a MAR or MNAR mechanism according to Heckman's model.
+The advantage of the proposed method over methods that assume MAR is that it allows the imputation of variables from cluster level data following a MAR or MNAR mechanism according to the Heckman model.
 That is to say that under the specification of a valid exclusion variable the method determines by itself which is the most adjustable  correlation parameter between equations ($\rho$), or in general terms the  missingness mechanism  (MAR or MNAR), in each of the clusters evaluated.
-Our implementation of the imputation method is built on the mice R package, which allows, first of all, to be used both on the outcome and on the covariates.
-In addition, it offers the option of being used simultaneously with other imputation methods implemented the package, which is advantageous in databases containing missing variables with different prediction methods and models.
-Finally, the method can be used on systematically and sporadically missing clusters, both for continuous variables with heterogeneous error variance and for binary variables.
+Our implementation of the imputation method was built according to the specifications of the mice R package, which allows, first of all, to be used both on the outcome and on the covariates.
+In addition, it offers the option of being used simultaneously with other imputation methods implemented in this package, which is advantageous in databases containing missing variables with different prediction methods and models.
+Finally, the method can be used on systematically and sporadically missing clusters, both for continuous and binary missing variables with heterogeneous effects and error variances.
 
@@
 A major limitation of our method is that it needs a valid restriction variable, which in some contexts is difficult to establish at the individual study level and can be even more challenging if one tries to find a valid exclusion variable across clusters.
-Also, the method is sensitive to the value of the correlation between the selection equation and outcome (rho), and in general it is observed that it can lead to biased results on fixed global parameters i.e.
-without variation across clusters.
-Similarly, the method can be sensitive to both the sample size and the number of studies included in the database.
+Also, the method can be sensitive to both the sample size and the number of studies included in the database.
 On the one hand, a small sample size at the individual study level can affect not only the precision of estimates but also the convergence of the method since it requires a minimum sample size to estimate all the parameters of the Heckman model which can be at least twice the number of parameters required in an imputation model that assumes MAR.
@@
 One might also consider relaxing this assumption of constant correlation to allow for a random effects distribution for the correlation parameter.
-<!-- Does my suggestion make sense? -->
+
+<!-- Does my suggestion make sense? yes!!! thanks a lot-->
 
 Further, the method can also be extended to other types of variables such as count or ordinal variables.
-Similarly, less restrictive Heckman based models can be considered in terms of normality distribution of errors and no specification of exclusion variables such as those proposed by Ogundimu.
+Similarly, less restrictive Heckman based models can be considered in terms of normality distribution of errors and no specification of exclusion variables such as those proposed by @ogundimu_collins19.
 <!-- Add reference to Ogundimu -->
@@
 <!-- Always add a conclusion, also in stats papers. I'm thinking it should mention the following. I am not so sure on especially the last sentence, so feel free to rephrase -->
-We have <!--(well actually you) --> proposed an extension to the Heckman model that can account for MNAR, MAR or MCAR of a continuous or binary variable in clustered data sets.
+We have <!--(well actually you) no actually Thomas :D --> proposed an extension to the Heckman model that can account for MNAR, MAR or MCAR of a continuous or binary variable in clustered data sets.
 Our simulations showed that it has favorable statistical properties, when assumptions were met, and provided that the sample size was sufficiently large.

### 2022-11-10 Untitled/Untitled.Rmd
tables update

---
title: A demonstration of the \LaTeX class file for Statistics in Medicine with Rmarkdown
author:
- name: A. Uthor*
  num: a,b
- name: O. Tro
  num: b
- name: O. Vriga
  num: c
address:
- num: a
  org: Department of Incredible Research, University A, City A, Country A
- num: b
  org: Department of Applied Things, University B, City B, Country B
- num: c
  org: Very Important Stuff Committee, Institute C, City C, Country C
corres: "*Corresponding author name, This is sample corresponding address. \\email{authorone@gmail.com}"
presentaddress: This is sample for present address text this is sample for present address text
authormark: Uthor \emph{et al}.
articletype: Research article
received: 2017-01-01
revised: 2017-02-01
accepted: 2017-03-01
abstract: "Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aenean ut elit odio. Donec fermentum tellus neque, vitae fringilla orci pretium vitae. Fusce maximus finibus facilisis. Donec ut ullamcorper turpis. Donec ut porta ipsum. Nullam cursus mauris a sapien ornare pulvinar. Aenean malesuada molestie erat quis mattis. Praesent scelerisque posuere faucibus. Praesent nunc nulla, ullamcorper ut ullamcorper sed, molestie ut est. Donec consequat libero nisi, non semper velit vulputate et. Quisque eleifend tincidunt ligula, bibendum finibus massa cursus eget. Curabitur aliquet vehicula quam non pulvinar. Aliquam facilisis tortor nec purus finibus, sit amet elementum eros sodales. Ut porta porttitor vestibulum. Integer molestie, leo ut maximus aliquam, velit dui iaculis nibh, eget hendrerit purus risus sit amet dolor. Sed sed tincidunt ex. Curabitur imperdiet egestas tellus in iaculis. Maecenas ante neque, pretium vel nisl at, lobortis lacinia neque. In gravida elit vel volutpat imperdiet. Sed ut nulla arcu. Proin blandit interdum ex sit amet laoreet. Phasellus efficitur, sem hendrerit mattis dapibus, nunc tellus ornare nisi, nec eleifend enim nibh ac ipsum. Aenean tincidunt nisl sit amet facilisis faucibus. Donec odio erat, bibendum eu imperdiet sed, gravida luctus turpis."
keywords: Class file; \LaTeX; Statist. Med.; Rmarkdown;
bibliography: bibfile.bib
output: rticles::sim_article
longtable: true
header-includes: 
- \usepackage{longtable}
---

# Introduction

A minimal working example of markdown tables with the Statistics in Medicine R Markdown template

# Table

Table produced by plain markdown:

First Header | Second Header
------------ | -------------
Content cell | Content cell
Content cell | Content cell

`corres`: Author and address for correspondence

* `authormark`: Short author list for header

* `received`, `revised`, `accepted`: dates of submission, revision, and acceptance of the manuscript

* `abstract`: Limited to 250 words

* `keywords`: Up to 6 keywords

* `bibliography`: BibTeX `.bib` file

* `classoption`: options of the `WileyNJD-v2` class

* `longtable`: set to `true` to include the `longtable` package, used by default from `pandoc` to convert markdown to \LaTeX code

## Remarks

1. In `authormark` use _et al._ if there are three or more authors.

2. Note the use of `num` to link names and addresses.

3. For submitting a double-spaced manuscript, add `doublespace` as an option to a `classoption` line in the YAML header: `classoption: doublespace`.

4. Keywords are separated by semicolons.

# The Body of the Article

## Mathematics

Use mathematics in Rmarkdown as usual.

## Figures and Tables

Figures are supported from R code:

```{r plot-ref, fig.cap = "Fancy Caption\\label{fig:plot}"}
x = rnorm(10)
y = rnorm(10)
plot(x, y)
```

...and can be referenced (Figure \ref{fig:plot}) by including the `\\label{}` tag in the `fig.cap` attribute of the R chunk: `fig.cap = "Fancy Caption\\label{fig:plot}"`. It is a quirky hack at the moment, see [here](https://github.com/yihui/knitr/issues/323).

Analogously, use Rmarkdown to produce tables as usual:

```{r, results = "asis"}
if (!require("xtable")) install.packages("xtable")
xt <- xtable(head(cars), caption = "A table", label = "tab:table")
print(xt, comment = FALSE)
```

Referenced via \ref{tab:table}. You can also use the YAML option `header-includes` to includes custom \LaTeX packages for tables (keep in mind that `pandoc` uses `longtables` by default, and it is hardcoded; some things may require including the package `longtable`). E.g., using `ctable`:
```
header-includes:
- \usepackage{ctable}
```
Then, just write straight-up \LaTeX code and reference is as usual (`\ref{tab:ctable}`):
```
\ctable[cap = {Short caption},
        caption = {A long, long, long, long, long caption for this table.},
        label={tab:ctable},]
        {cc}
        {
        \tnote[$\ast$]{Footnote 1}
        \tnote[$\dagger$]{Other footnote}
        \tnote[b]{Mistakes are possible.}
        }{
        \FL
        COL 1\tmark[a] & COL 2\tmark[$\ast$]
        \ML
        6.92\tmark[$\dagger$] & 0.09781 \\
        6.93\tmark[$\dagger$] & 0.09901 \\
        97 & 2000
        \LL
}
```

It is also possible to set the `YAML` option `longtable: true` and use markdown tables (or the `knitr::kable` function): `knitr::kable(head(cars))` produces the same table as the `xtable` example presented before.

```{r dos, results = "asis"}
if (!require("xtable")) install.packages("xtable")
knitr::kable(head(cars))
```

## Cross-referencing

The use of the Rmarkdown equivalent of the \LaTeX cross-reference system
for figures, tables, equations, etc., is encouraged (using `[@<name>]`, equivalent of `\ref{<name>}` and `\label{<name>}`). That works well for citations in Rmarkdown, not so well for figures and tables. In that case, it is possible to revert to standard \LaTeX syntax.

## Double Spacing

If you need to double space your document for submission please use the `doublespace` option in the header.

# Bibliography

Link a `.bib` document via the YAML header, and bibliography will be printed at the very end (as usual). The default bibliography style is provided by Wiley as in `WileyNJD-AMA.bst`, do not delete that file.

Use the Rmarkdown equivalent of the \LaTeX citation system using `[@<name>]`. Example: [@Taylor1937], [@Knupp1999; @Kamm2000].

To include all citation from the `.bib` file, add `\nocite{*}` before the end of the document.

# Further information

All \LaTeX enviroments supported by the main template are supported here as well; see the `.tex` sample file [here](http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1097-0258/homepage/la_tex_class_file.htm) for more details and example.

### 2022-11-10 3.Ilustrative_study/Ilustrative_study.R

@@ -156,23 +156,26 @@ hdataF0$dis_id<-as.integer(as.factor(hdataF0$dis_id)) #convert as numeric for im
 
 sum.table<-setDT(hdataF0)[,.(did=(length(unique(dis_name))),
                              N=.N,
-                             age=round(mean(age),2),
+                             age=format(round(mean(age),2),nsmall = 2),
                              p_anopheles= round(mean(exp(p_anopheles)),2),
                              LIp_an=round(exp(quantile(p_anopheles,probs=0.025,type=1)),1),
                              UIp_an=round(exp(quantile(p_anopheles,probs=0.975,type=1)),1),
-                             wealthin_hou=round(mean(wealthin_hou),2),
-                             LIw=round(quantile(wealthin_hou,probs=0.025,type=1),1),
-                             UIw=round(quantile(wealthin_hou,probs=0.975,type=1),1),
+                             wealthin_hou=format(round(mean(wealthin_hou),2), nsmall = 2),
+                             LIw=format(round(quantile(wealthin_hou,probs=0.025,type=1),1), nsmall = 1),
+                             UIw=format(round(quantile(wealthin_hou,probs=0.975,type=1),1), nsmall = 1),
                              bednets_pp=round(sum(ifelse(bednets_pp=="Yes",1,0))/.N*100,1),
-                             sex=round(sum(ifelse(sex=="Male",1,0))/.N*100,1),
+                             sex=format(round(sum(ifelse(sex=="Male",1,0))/.N*100,1),nsmall = 1),
                              fundays=round(sum(ifelse(fundays=="yes",1,0))/.N*100,1),
-                             ry=round((.N-sum(ry))/.N*100,1)),by=list(subreg_id)]
+                             ry=format(round((.N-sum(ry))/.N*100,1),nsmall = 1)),by=list(subreg_id)]
 
 
 sum.table[,p_anopheles:=paste0(p_anopheles,"[",LIp_an,",",UIp_an,"]")]
 sum.table[,wealthin_hou:=paste0(wealthin_hou,"[",LIw,",",UIw,"]")]
-
+sum.table$subreg_id<- c("North East","Mid Eastern","South Western","Mid Western","East Central")
[…]

### 2022-11-11 1.Paper/Paper.Rmd
update references

@@
 Therefore, the aim of this paper is to develop a multilevel imputation method for incomplete continuous and binary variables that are MNAR.
-In section \ref{heckman} we provide an introduction to the Heckman model and its estimation, and we extended to a hierarchical setting.
-Then in section \ref{methodsn} we define the main steps of the proposed imputation method.
-In section \ref{simulation} we provide the settings and results of a Monte Carlo simulation study used to evaluate the performance of the proposed imputation method.
-In section \ref{example} we illustrate the method using the survey information collected in different sub-districts in Uganda to estimate the prevalence of malaria in children.
-Finally we provide a discussion  in section \ref{discussion} about  the results, limitations and propose future extension of our method.
+In section \ref{sec:heckman} we provide an introduction to the Heckman model and its estimation, and we extended to a hierarchical setting.
+Then in section \ref{sec:methods} we define the main steps of the proposed imputation method.
+In section \ref{sec:simulation} we provide the settings and results of a Monte Carlo simulation study used to evaluate the performance of the proposed imputation method.
+In section \ref{sec:illustration} we illustrate the method using the survey information collected in different sub-districts in Uganda to estimate the prevalence of malaria in children.
+Finally we provide a discussion in section \ref{sec:discussion} about  the results, limitations and propose future extension of our method.
 
 # The Heckman model
-\label{heckman}
+\label{sec:heckman}
 The Heckman selection model was initially proposed as a method to correct for selection bias, in which individuals are not randomly selected from the population, leading to inconsistent estimates and erroneous conclusions.
@@
 # Method
-\label{methodsn}
+\label{sec:methods}
 We follow a similar approach proposed by Resche-Rigon and White (2018) [@resche-rigon2018] for multilevel imputation of data.
@@
 
-Initially, we use the copulat method to estimate the set of cluster-specific parameters, $\widehat{\theta_i}=\{\widehat{\beta^O_i}, \widehat{\beta^S_i}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, using all $j$ units with observable measurements $y_{ij}^{obs}$ within each group $i$.
+Initially, we use the copula method to estimate the set of cluster-specific parameters, $\widehat{\theta_i}=\{\widehat{\beta^O_i}, \widehat{\beta^S_i}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, using all $j$ units with observable measurements $y_{ij}^{obs}$ within each group $i$.
 The Heckman model is estimated with the \textbf{gjrm} function of the GJRM R package, [@radiceGJRMGeneralisedJoint2021] under the bivariate model with the nonrandom sample selection (BSS) specification, from which we obtain not only the parameters' point estimates $\widehat{\theta_i}$, but also their corresponding $\widehat{S(\theta_i)}$ within-cluster variance-covariance matrix.
@@
 
-The meta-analysis model is performed with the \textbf{mixmeta} function of the R package *mixmeta*, which allows the use of maximum likelihood (ML), restricted maximum likelihood (REML), and moments estimation methods.
+The meta-analysis model is performed with the \textbf{mixmeta} function of the R package mixmeta[@gasparrini2021], which allows the use of maximum likelihood (ML), restricted maximum likelihood (REML), and moments estimation methods.
 For the simulation and illustrative study, we used the restricted REML estimation method, which is recommended as it has a good balance between unbiasedness and efficiency.
@@
 # Simulation study
-\label{simulation}
+\label{sec:simulation}
 
@@
 # An illustrative study
-\label{example}
+\label{sec:illustration}
 ```{r, include=FALSE,cache = TRUE}
@@
 <!-- I see you wrote both parasitemia and parasitaemia. Choose one -->
-To illustrate our proposed method, following the article by @rugnao2019, we estimated the prevalence of parasitemia by subregion and by age after approximately 3 years of LLIN campaigns started.
+To illustrate our proposed method, following the article by Rugnao et al.
+(2019),[@rugnao2019] we estimated the prevalence of parasitemia by subregion and by age after approximately 3 years of LLIN campaigns started.
 We estimated parasitemia prevalence using 3 approaches that made different assumptions on the missingness mechanism: MCAR, MAR and MNAR.
@@
 <!-- The age at which children start going to school is not universal, so you need to mention this. done!! thanks -->
-In terms of prevalence by age, there are no significant differences between methods (Figure \ref{figure:region_age}).
+In terms of prevalence by age, there are no significant differences between methods (Figure \ref{fig:region_age}).
 The prevalence estimates for children aged 2 to 6 years are very similar in all regions under the different assumptions.
@@
 # Discussion
-\label{discussion}
+\label{sec:discussion}
 We have extended and evaluated methods for multiple imputation of clustered datasets, in situations where some incomplete variables follow a MNAR mechanism.
@@
 Further, the method can also be extended to other types of variables such as count or ordinal variables.
-Similarly, less restrictive Heckman based models can be considered in terms of normality distribution of errors and no specification of exclusion variables such as those proposed by @ogundimu_collins19.
+Similarly, less restrictive Heckman based models can be considered in terms of normality distribution of errors and no specification of exclusion variables such as those proposed by Ogundimu & Collins (2019).[@ogundimu_collins19].
 <!-- Add reference to Ogundimu -->

### 2022-11-15 1.Paper/Paper.Rmd
update

@@
   num: a
+- name: Matthias Egger 
+  num: d  
+- name: Orestis Efthimiou
+  num: d
 - name: Valentijn M. T. de Jong
-  num: a,b
+  num: a,b,e
 - name: Thomas. P. A. Debray
-  num: a,c
+  num: a,c,e
 address:
@@
   org: Smart Data Analysis and Statistics, Utrecht, The Netherlands 
+- num: d
+  org: University of Bern, Switzerland
+- num: e
+  org: These authors contributed equally
 

### 2022-11-22 1.Paper/Paper.Rmd
update

@@
 abstract: "Missing data is a common problem in medical research, and is commonly addressed using multiple imputation.
-Although traditional imputation methods allow for valid statistical inference when data are missing at random (MAR), their implementation is not justified when observations are clustered (e.g., within studies) or when the presence of missingness depends on unobserved information.
+Although traditional imputation methods allow for valid statistical inference when data are missing at random (MAR), their implementation is not justified when observations are clustered (e.g., within studies) or when the presence of missingness depends on unobserved information, i.e.
+the data are missing not at random (MNAR).
 Unfortunately, this situation is increasingly common, and typically arises when individual participant data (IPD) from multiple studies are combined.
-While several imputation methods have been proposed for addressing individual studies where data are missing not at random (MNAR), their application and validity in large datasets with clustering remains unclear.
-We therefore explored the consequence of MNAR data in IPD meta-analysis in-depth, and proposed novel multilevel imputation methods for common missing patterns in clustered datasets.
-These methods build upon the principles of Heckman selection models, and adopt a two-stage meta-analysis approach for imputing binary and continuous variables.
-After evaluating the proposed imputation models in simulated scenarios, we illustrated their use in a cross-sectional community survey to estimate the prevalence of malaria parasitemia in children aged 2-10 years in five subregions in Uganda."
+While several imputation methods have been proposed for addressing individual studies when data are MNAR, their application and validity in large datasets with clustering remains unclear.
+We therefore explored the consequence of MNAR data in IPD meta-analysis in-depth, and proposed a novel multilevel imputation method for common missing patterns in clustered datasets.
+This method is based on the principles of Heckman selection models and adopts a two-stage meta-analysis approach to impute binary and continuous variables that may be outcomes or predictors and that are systematically or sporadically missing.
+After evaluating the proposed imputation model in simulated scenarios, we illustrate it use in a cross-sectional community survey to estimate the prevalence of malaria parasitemia in children aged 2-10 years in five subregions in Uganda."
 
@@
 ---
-<!-- I tried to add a second affiliation for me, but I get 1,1 behind my name instead of 1,2. Do you happen to know how to fix this? Aparently we should not add space between comas :S -->
-<!-- I find the last sentence of the abstract a bit hard to read. Does the following rephrasing make sense? yes!! thanks i just add yours!! -->
 
 # Introduction
-Over the past few years, data sharing efforts have substantially increased and researchers increasingly often have access to large combined datasets derived from electronic health records (EHR) or from individual participant data (IPD).
+Over the past few years, data sharing efforts have substantially increased, and researchers increasingly often have access to IPD from large combined datasets derived from electronic health records (EHR) or from multiple randomized or observable trials (i.e.
+in IPD meta-analysis, IPD-MA).
 For example, the clinical practice research datalink (CRPD) [@herrett2015] is an electronic health record dataset in the UK, which has been used in a variety of medical research, such as the evaluation of health policy and drug efficacy.
-A recent example of an IPD-MA is the emerging risk factor collaboration, [@theemergingriskfactorscollaboration2007] where data were combined from approximately 1.1 million individuals across 104 studies to investigate associations of cardiovascular diseases with several predictors.
-Individuals in these large datasets tend to be clustered in centres, countries or studies, where they have been subject to similar healthcare processes, and are therefore more alike than individuals from another cluster.
-Sometimes, clusters may also differ in participant eligibility criteria, follow-up length, predictor and outcome definitions, or in the quality of applied measurement methods.
-Hence, correlation is likely to be present between observations from the same cluster, which can lead to differences or ‘heterogeneity’ between clusters regarding baseline patient characteristics and subsequent outcomes.
+A recent example of an IPD-MA is the emerging risk factor collaboration, [@theemergingriskfactorscollaboration2007] where data were combined from approximately 1.1 million individuals across 104 observational studies to investigate associations of cardiovascular diseases with several predictors.
+Individuals in these large datasets tend to be clustered in centres, countries or studies, where they have been subject to similar healthcare processes.
+Moreover, clusters may also differ in participant eligibility criteria, follow-up length, predictor and outcome definitions, or in the quality of applied measurement methods.
+Hence, differences (i.e.,heterogeneity) between clusters with respect to baseline covariates and outcomes, while the structure of the correlations between these variables is likely to be different across different clusters.
 
-Clustered datasets often contain  many incomplete variables.
-For example, in registry data it is common that test results are not available in registry data for all patients, as the decision to test may be at the discretion of the primary care physician or because the patient refuses to undergo testing.
+A usual problem is that such clustered datasets may contain  many incomplete variables.
+For example, in registry data it is common that test results are not available for all patients, as the decision to test may be at the discretion of the primary care physician or because the patient refuses to undergo testing.
 It is also possible that variables are systematically missing across clusters.
-For instance, in an IPD meta-analysis, it is common that studies collected information on different variables.
+For instance, in an IPD-MA, studies may have collected information on different variables.
 Missing values may thus appear for all participants of a study in the combined dataset.
-The presence of missing data can lead to loss of statistical power, imbalance across clusters, bias in parameter estimates and therefore to erroneous conclusions as the analysis could be based on an unrepresentative sample.
+The presence of missing data can lead to loss of statistical power, imbalance in cluster size, bias in parameter estimates and therefore to erroneous conclusions as the analysis could be based on an unrepresentative sample.
 
@@
 Traditional imputation methods are designed to address incomplete data sets where variables are MCAR or MAR.
-Their implementation is reasonable when there is no obvious mechanism of missingness, or when the observed data strongly relate to unobserved information.
+Their implementation is justified when there is not systematic difference between units with missing or complete data or when the missingness of variable is strongly related to events or variables measured in the study.
 
-<!-- Stat med uses the AMA citation style.
-citations at the end of sentence go after the period: "out MCAR.
-[@little1988]" , and those at the end of a clause go after the period: "out MCAR, [@little1988]".
-Inline citations like "the clinical practice research datalink (CRPD) [@herrett2015]" are fine though.
-See an example here:
-https://onlinelibrary.wiley.com/doi/epdf/10.1002/sim.9221 Cool thanks a lot!! -->
-Although it is possible to formally rule out MCAR, [@little1988]  it is impossible to assess whether data are MAR or MNAR as the observable data are not enough to test the assumptions of both mechanisms.[@enders2022]  For this reason, researchers often conveniently assume that data are MAR, or present results that are based on a complete case analysis.
+Although it is possible to formally rule out MCAR, [@little1988]  it is impossible to assess whether data are MAR or MNAR as the observable data are not enough to test the assumptions of both mechanisms.[@enders2022]  For this reason, researchers often conveniently assume that data are MAR, or present results of a complete case analysis.
 In practice, however, unless missingness is artificially introduced by study design, e.g.
-when a test is only taken on patients with certain characteristics, missingness will often (partially) depend on unobserved information.
+when a test is only taken on patients with certain characteristics, missingness will often (at least partially) depend on unobserved information.
 
-The missingness mechanism of a variable may be a mixture of MAR and MNAR, depending on the availability of data that may predict the missing data.
-Therefore, the MAR mechanism may become more influential than the MNAR mechanism as more information is recorded.
-<!-- MAR does not really become more likely.
-It is rather that the MAR mechanism becomes stronger than the MNAR mechanism.
-I have tried to capture this with the alternative text !!
-perfect thanks, your point is correct and now sound so much better :D-->
-For instance, healthcare professionals may decide what type of measurement on indication, and depending on their personal experience or behavior decide whether or not to record the reasons for the measurement.
+The missingness mechanism of a variable may be a mixture of MAR and MNAR, depending on the availability of data that may predict the missingness patterns.
+Therefore, the MAR mechanism may become more influential than MNAR when more information is recorded.
+For instance, healthcare professionals may decide which vital measures to take at the time of patient admission.
+If low blood pressure (BP) measurements are more likely to be unrecorded, this variable could be assumed to be MNAR, since its missingness depends on its own value, however BP could be considered MAR as more information explaining its missingness is recorded such as the severity of a patient's admission.
 
-Registries are notoriously prone to incomplete variables that are MNAR, due to the complex recording process.
-[@liu2021] e.g.
-laboratory tests are taken only in certain patients based on sign and symptom information that is often incomplete or not recorded.
-Also IPD may suffer from MNAR, for example when study participants who experience unfavorable results drop out of the study.
-Also, heterogeneity of the primary objective or resources of the studies involved may result in variables relevant to explain the missing process not being recorded.
+Registries are notoriously prone to incomplete variables that are MNAR, due complex recording processes.
+[@liu2021] For example, laboratory tests are taken only in certain patients based on symptoms that are often incompletely recorded.
+Data from randomized trials may also suffer from MNAR, for example when study participants that experience unfavorable results drop out of the study.
+Also, heterogeneity of the primary objective or resources of the studies may result in variables relevant to explain the missing process not being recorded at all in some of the studies.
 
-The MNAR mechanism is considered as non-ignorable, because to deal with this type of missingness it is necessary to specify information about the missingness process in addition to assumptions about the observed data.
-A common approach to address MNAR is to adopt  selection models such as the one proposed by Heckman (1976).
-[@heckman1976] Briefly, the Heckman selection model corrects for selection bias by estimating two linked equations, a main equation where the missing variable is associated with predictors, and a selection equation that accounts for the inclusion of observations in the sample.
-An important feature of the Heckman selection model is that its implementation does not require data to be MNAR, and can also be used when data are MCAR or MAR.
-It therefore offers an appealing solution to incomplete data sets where the missingness mechanism is not precisely known.
+The MNAR mechanism is considered as non-ignorable, in this case it is necessary to specify information about the missingness process in addition to assumptions about the observed data.
+A popular approach to address MNAR is selection models such as the one proposed by Heckman (1976).
+[@heckman1976] Briefly, the Heckman selection model corrects for selection bias by estimating two linked equations: a outcome equation, where the missing variable is associated with predictors, and a selection equation, that accounts for the inclusion of observations in the sample.
+An important feature of the Heckman selection model is that it does not assume data to be MNAR, so that it can also be used when data are MCAR or MAR.
+It therefore offers an appealing solution to incomplete data sets when the missingness mechanism is not precisely known.
 
@@
 Among them, Galimard et al.(2016) [@galimard_etal16] implemented a chained equations imputation method for continuous variables, which was extended to binary and categorical variables by employing copula estimates.
-[@galimard_etal18]  Also, Ogundimu and Collins (2019) [@ogundimu_collins19] proposed a chained equations imputation method that is less dependent on Normality assumptions.
+[@galimard_etal18]  Also, Ogundimu and Collins (2019) [@ogundimu_collins19] proposed a chained equations imputation method that is less dependent on normality assumptions.
 
-In clustered data sets, multilevel imputation methods are needed to properly propagate uncertainty within and across clusters.
+In clustered data sets, multilevel imputation methods are required to properly propagate uncertainty within and across clusters.
 [@audigier_etal18]  However, to our knowledge, existing multilevel imputation methods mainly focus  on situations where data are MAR, and do not adopt Heckman selection models.
-Although Hammon and Zinn (2020) [@hammon_zinn20] recently proposed an extension that allows for the inclusion of random intercept effects, it can only be used for binary missing variables and assume that the effect of explanatory variables on the missingness mechanisms is common across clusters.
+Although Hammon and Zinn (2020) [@hammon_zinn20] recently proposed an extension that allows for the inclusion of random intercept effects, it can only be used for binary missing variables and assumes that the effect of explanatory variables on the missingness mechanisms is common across clusters.
 
-Therefore, the aim of this paper is to develop a multilevel imputation method for incomplete continuous and binary variables that are MNAR.
-In section \ref{sec:heckman} we provide an introduction to the Heckman model and its estimation, and we extended to a hierarchical setting.
-Then in section \ref{sec:methods} we define the main steps of the proposed imputation method.
-In section \ref{sec:simulation} we provide the settings and results of a Monte Carlo simulation study used to evaluate the performance of the proposed imputation method.
+Therefore, the aim of this work is to develop a multilevel imputation method for continuous and binary variables that are both sporadically and systematically MNAR, and that this can be applied for an incomplete outcome or multiple incomplete predictors in the data sets.
+
+In section \ref{sec:heckman} we provide an introduction to the Heckman model and its estimation, and we extend it to a hierarchical setting.
+In section \ref{sec:methods} we define the main steps of the proposed imputation method.
+In section \ref{sec:simulation} we provide the settings and results of a simulation study to evaluate the performance of the proposed imputation method.
 In section \ref{sec:illustration} we illustrate the method using the survey information collected in different sub-districts in Uganda to estimate the prevalence of malaria in children.
-Finally we provide a discussion in section \ref{sec:discussion} about  the results, limitations and propose future extension of our method.
+Finally in section \ref{sec:discussion} we summarize the results, outline limitations and propose future extension of our method.
 
@@
 
-Selection bias occurs when the selection of a subject  or their observation into the sample is influenced by unobserved variables  (e.g., the respondent's level of trust toward healthcare entities may cause them to self-select out of the study or refuse to sign consent for a test), which in turn are correlated with unobserved variables related to a variable of interest (e.g., the result of a blood test).
+Selection bias occurs when the inclusion of a subject or an observation into the sample is influenced by unobserved variables  (e.g., the respondent's level of trust toward healthcare entities may cause them to self-select out of the study or refuse to sign consent for a test), which in turn are correlated with unobserved variables related to an outcome of interest (e.g., the result of a blood test).
 [@vella1998]
@@
    \includegraphics[scale=0.35]{DAG_heckman.png}
-   \caption{DAG Heckman selection model: here the nodes (dotted = latent, continuous = observable) describe the relationship between $y^*_{ij}$ the latent response and $r^*_{ij}$ the latent  selection variables}
+   \caption{DAG Heckman selection model: here the nodes (dotted lines = latent variables, continuous lines = observed variables) describe the relationship between $y^*_{ij}$ the latent response and $r^*_{ij}$ the latent selection variables of a $j$th unit in a $i$th cluster,that are dependable on $x_{ij}^O$ and $x_{ij}^S$ set of predictors and are correlated through $u_{ij}$ unobservable variables.}
    \label{fig:DAG}
@@
 
-Denoting $y^*_i=(y^*_{i1},y^*_{i2},...,y^*_{in_i})^T$ and  $r^*_i=(r^*_{i1},r^*_{i2},...,r^*_{in_i})^T$ the vectors of latent outcomes and latent selections in the cluster $i$, then Heckman's model is defined by two main equations: the outcome equation (\ref{eq:1}), which describes the relation latent outcome-exposure association, and the selection equation (\ref{eq:2}) which details the likelihood that the outcome is observed in the sample.
+Denoting $\boldsymbol{y^*_i}=(y^*_{i1},y^*_{i2},...,y^*_{in_i})^T$ and  $\boldsymbol{r^*_i}=(r^*_{i1},r^*_{i2},...,r^*_{in_i})^T$ the vectors of latent outcomes and latent selections in the cluster $i$, then Heckman's model is defined by two main equations: the outcome equation (\ref{eq:1}), which describes the relation between the latent outcome ($y_{ij}^*$) and a set of covariates ($\boldsymbol{X_{i}^O}=(\boldsymbol{x^{O}_{i1}},\boldsymbol{x^{O}_{i2}},...,\boldsymbol{x^{O}_{in_i}})^T$), and the selection equation (\ref{eq:2}) which models the likelihood that the outcome is observed in the sample as a function of another set of covariates ($\boldsymbol{X_{i}^S}=(\boldsymbol{x^{S}_{i1}},\boldsymbol{x^{S}_{i2}},...,\boldsymbol{x^{S}_{in_i}})^T$).
 \begin{align}
-y^*_{i}&=X^O_{i}\beta^O_{i}+\epsilon^O_{i}\label{eq:1}\\
-r^*_{i}&=X^S_{ij}\beta^S_{i}+\epsilon^S_{i}\label{eq:2}
+\boldsymbol{y^*_{i}}&=\boldsymbol{X^O_{i}}\boldsymbol{\beta^O_{i}}+\boldsymbol{\epsilon^O_{i}}\label{eq:1}\\
+\boldsymbol{r^*_{i}}&=\boldsymbol{X^S_{i}}\boldsymbol{\beta^S_{i}}+\boldsymbol{\epsilon^S_{i}}\label{eq:2}
 \end{align}
 
-Here $\beta^O_{i}$ and $\beta^S_{i}$ are $p\times1$ and $q\times1$ coefficient parameter vectors and $\epsilon^O_{i}=(\epsilon^O_{i1},\epsilon^O_{i2},...,\epsilon^O_{in_i})^T$ and $\epsilon^S_{i}=(\epsilon^S_{i1},\epsilon^S_{i2},...,\epsilon^S_{in_i})^T$ are the residual terms vectors for the outcome and selection equations, respectively.
+Here $\boldsymbol{\beta^O_{i}}$ and $\boldsymbol{\beta^S_{i}}$ are $p\times1$ and $q\times1$ coefficient parameter vectors and $\boldsymbol{\epsilon^O_{i}}=(\epsilon^O_{i1},\epsilon^O_{i2},...,\epsilon^O_{in_i})^T$ and $\boldsymbol{\epsilon^S_{i}}=(\epsilon^S_{i1},\epsilon^S_{i2},...,\epsilon^S_{in_i})^T$ are the residual terms vectors for the outcome and selection equations, respectively.
 

### 2022-11-22 3.Ilustrative_study/Ilustrative_study.R

@@ -228,7 +228,7 @@ pred["test","sage2"]<- 1
 
 #3.3. Multiple imputation with heckman model----
 meth<-ini$method
-meth[c("test")]<-"4l.heckman"
+meth[c("test")]<-"2l.heckman"
 
 #Full correlation of parameters
 data_heck <- mice(hdataF0f, # dataset with missing values

### 2022-11-22 4.Codes/mice.impute.2l.heckman.R

@@ -59,7 +59,7 @@
 #' @export
 #'
 
-mice.impute.4l.heckman <-function(y,ry,x,wy = NULL, type, pmm = FALSE, meta_method ="reml",...) {
+mice.impute.2l.heckman <-function(y,ry,x,wy = NULL, type, pmm = FALSE, meta_method ="reml",...) {
   
   
   # 1. Define variables and dataset----

### 2022-11-22 1.Paper/Paper.Rmd
update text

@@
 abstract: "Missing data is a common problem in medical research, and is commonly addressed using multiple imputation.
-Although traditional imputation methods allow for valid statistical inference when data are missing at random (MAR), their implementation is not justified when observations are clustered (e.g., within studies) or when the presence of missingness depends on unobserved information, i.e.
+Although traditional imputation methods allow for valid statistical inference when data are missing at random (MAR), their implementation is problematic when the presence of missingness depends on unobserved information, i.e.
 the data are missing not at random (MNAR).
-Unfortunately, this situation is increasingly common, and typically arises when individual participant data (IPD) from multiple studies are combined.
+Unfortunately, this situation is rather common, in observational studies, registries and other sources of real-world data.
 While several imputation methods have been proposed for addressing individual studies when data are MNAR, their application and validity in large datasets with clustering remains unclear.
-We therefore explored the consequence of MNAR data in IPD meta-analysis in-depth, and proposed a novel multilevel imputation method for common missing patterns in clustered datasets.
+We therefore explored the consequence of MNAR data in hierarchical data in-depth, and proposed a novel multilevel imputation method for common missing patterns in clustered datasets.
 This method is based on the principles of Heckman selection models and adopts a two-stage meta-analysis approach to impute binary and continuous variables that may be outcomes or predictors and that are systematically or sporadically missing.
@@
    \includegraphics[scale=0.35]{DAG_heckman.png}
-   \caption{DAG Heckman selection model: here the nodes (dotted lines = latent variables, continuous lines = observed variables) describe the relationship between $y^*_{ij}$ the latent response and $r^*_{ij}$ the latent selection variables of a $j$th unit in a $i$th cluster,that are dependable on $x_{ij}^O$ and $x_{ij}^S$ set of predictors and are correlated through $u_{ij}$ unobservable variables.}
+   \caption{DAG Heckman selection model: here the nodes (dotted lines = latent variables, continuous lines = observed variables) describe the relationship between $y^*_{ij}$ the latent response and $r^*_{ij}$ the latent selection variables of a $j$th unit in a $i$th cluster,that are dependable on $ \boldsymbol{x_{ij}^O}$ and $ \boldsymbol{x_{ij}^S}$ set of predictors and are correlated through $ \boldsymbol{u_{ij}}$ unobservable variables.}
    \label{fig:DAG}
 \end{figure}
-This is visualized in Figure \ref{fig:DAG}, where for the $j$th individual or unit within the $i$th cluster, there is $y_{ij}^*$, a latent outcome variable, and $r_{ij}^*$, a latent selection variable, which are correlated through $u_{ij}$ an unobserved or unrecorded variable, with $i\in[1,2,...,N]$ and $j\in[1,2,...,n_i]$.
-Here, both latent variables are related to the sets of predictor covariates $x_{ij}^O$ and $x_{ij}^S$.
+This is visualized in Figure \ref{fig:DAG}, where for the $j$th individual or unit within the $i$th cluster, there is $y_{ij}^*$, a latent outcome variable, and $r_{ij}^*$, a latent selection variable, which are correlated through $\boldsymbol{u_{ij}}$ an unobserved or unrecorded variable, with $i\in[1,2,...,N]$ and $j\in[1,2,...,n_i]$.
+Here, both latent variables are related to the sets of predictor covariates $\boldsymbol{x_{ij}^O}$ and $\boldsymbol{x_{ij}^S}$.
 From $r_{ij}^*$ one can derive $r_{ij}=I(r_{ij}^*>=0)$ a selection indicator of $y_{ij}^*$ into the sample, and with this in turn, one can define $y_{ij}=y_{ij}^*,\forall r_{ij}=1$ the observable outcome variable.
@@
 
-Generally the same variables can be used on the matrix of predictor variables $X^O_{i}$ and $X^S_{i}$.
-However, to avoid multicollinearity problems, [@puhani1997] it is recommended to include in $X^S_{i}$ at least one variable that is not included in the outcome model.
-[@angrist2001] This variable is commonly known as an exclusion restriction variable (ERV), and should only be associated with the selection in the sample $r^*_{ij}$ (relevance condition) but not with the actual observation $y^*_{ij}$ (exclusion condition).[@gomes_etal20] The ERV should meet the relevance and exclusion conditions in order to provide independent information about the selection processing in order to facilitate the estimation of the Heckman model specially when dataset does not follow the BVN assumption.
-<!-- Can you give some explanation (1-3 sentences?) on the exclusion and restriction assumptions here? -->
+Generally the same variables can be used on the matrix of predictor variables $\boldsymbol{X^O_{i}}$ and $\boldsymbol{X^S_{i}}$.
+However, to avoid multicollinearity problems, [@puhani1997] it is recommended to include in $\boldsymbol{X^S_{i}}$ at least one variable that is not included in the outcome model.
+[@angrist2001] This variable is commonly known as an exclusion restriction variable (ERV), and should only be associated with the selection in the sample $r^*_{ij}$ (relevance condition) but not with the actual observation $y^*_{ij}$ (exclusion condition).[@gomes_etal20] The ERV should meet the relevance and exclusion conditions in order to provide independent information about the selection process that facilitates the estimation of the Heckman model.
 
-In the presence of selection bias, aforementioned outcome equation will yield biased estimates of $\beta^O_{i}$ if no efforts are made to adjust for the non-representativeness of the observed $X^O_{i}$ and $y_{i}$ values.
+In the presence of selection bias,the aforementioned outcome equation will yield biased estimates of $\boldsymbol{\beta^O_{i}}$ if no efforts are made to adjust for the non-representativeness of the observed $\boldsymbol{X^O_{i}}$ and $\boldsymbol{y_{i}}$ values.
 For this reason, the Heckman model aims to jointly estimate the outcome and selection equation by defining a relation between their respective error distributions.
@@
 The interpretation of $\rho_i$ is fairly straightforward.
-When $\rho_i=0$, the participation does not affect the outcome model and missing data can be considered MCAR (if data are missing completely at random) or MAR (if missingness is already explained by $x_{ij}^O$).
+When $\rho_i=0$, the participation does not affect the outcome model and missing data can be considered MCAR (if data are missing completely at random) or MAR (if missingness is already explained by $\boldsymbol{x_{ij}^O}$).
 Conversely, when $\rho_i\neq 0$, this suggests that data are MNAR.
@@
 
-Under the BVN assumption, the parameters of the Heckman model coefficients can be estimated using the two-step Heckman method (H2S) [@heckman1976] or the full information maximum likelihood method (FIML).[@amemiya1984] However, both methods can lead to inconsistent estimators when this underlying distribution cannot be assumed.
+Under the BVN assumption, the parameters of the Heckman model coefficients can be estimated using the two-step Heckman method (H2S) [@heckman1976] or the full information maximum likelihood method (FIML).[@amemiya1984] However, both methods may lead to inconsistent estimators when the true underlying distribution is not the BVN.
 [@gomes_etal20] To overcome this problem, other approaches have been proposed that relax the distribution assumptions, among them copula models.
 [@smith03] The copula approach uses a function, known as a copula, that joins the marginal distribution of the error terms of the selection and outcome equation which are specified separately.
-$$ F(r^*_{i},y^*_{i}) =\Phi\left(r^*_{i}-X^S_{i}\beta^S_{i},\frac{y^*_{i}-X^O_{i}\beta^O_{i}}{\sigma_i}, \rho_i\right)$$
+The Heckman model can be expressed with the following bivariate normal copula:
+$$ F(\boldsymbol{r^*_{i}},\boldsymbol{y^*_{i}}) =\Phi\left(\boldsymbol{r^*_{i}}-\boldsymbol{X^S_{i}}\boldsymbol{\beta^S_{i}},\frac{\boldsymbol{y^*_{i}}-\boldsymbol{X^O_{i}}\boldsymbol{\beta^O_{i}}}{\sigma_i}, \rho_i\right),$$
+where $\Phi$ is the bivariate standard normal cumulative distribution function.
 Thus, to estimate the parameters of the Heckman method, it is sufficient to specify the marginal distributions of the error terms, and link them with a suitable copula function.
-In our imputation method we estimate the Heckman model using the copula method, as in real life many data could not follow a BVN distribution.
-This facilitates the Heckman model estimation and makes it more robust to deviations from the assumptions regarding the distribution of the data.
+In our imputation method we estimate the Heckman model using the copula method, as in real life data generating mechanism may deviate a BVN distribution.
+This facilitates the Heckman model estimation and makes it more robust to deviations of distributional assumptions.
 
@@
 
-The Heckman model can be extended to hierarchical settings, i.e., in data where individuals or sampling units are nested within groups, as is the case in EHR or IPD.
-In this case, sample units from the same group are expected to share some characteristics (e.g., distribution of variables, relationships between exposure and outcomes, missing processes, missing mechanisms) given only the fact that they belong to the same group.
+The Heckman model can be extended to hierarchical settings, i.e., in cases when individuals or sampling units are nested within clusters, as is the case in EHR or IPD.
+In this case, sample units from the same cluster are expected to share some characteristics (e.g., the distribution of variables and correlations between them, relationships between exposure and outcomes, missing processes).
 
 Since most statistical analyses assume that the sample units are independent of each other, more complex hierarchical models dealing with nested data are required.
-This hierarchical complexity must not only be taken into account in the observation process, but also in the missing data process, thus requiring imputation models that are congenial to the analysis model, i.e., that have the same assumptions about the data.
+This hierarchical complexity must not only be taken into account in the observation process,i.e.
+model that connects outcome and predictors, but also in the missing process,i.e related to the distribution and reasons for missingness, thus requiring imputation models that are congenial to the analysis model, i.e., that make the same assumptions about the data.
 
-Different procedures can be adopted to combine information between groups; however, in our imputation method we opted for the two-stage approach that is often used in meta-analyses.
-This is because such an approach is less computationally intensive and could potentially generate fewer convergence problems in the estimation of the Heckman hierarchical model compared to other approaches.
+Different procedures can be adopted to combine information between clusters; however, in our imputation method we opted for the two-stage approach that is often used in meta-analyses.[@simmonds2005] This is because such an approach is computationally less intensive and could potentially generate fewer convergence problems in the estimation of the Heckman hierarchical model compared to other approaches.
 
-<!-- This sentence was too long and has the verbs all the way at the end, making it hard to read. I suggest to rephrase to three shorter sentences: OK!!-->
-Briefly, in a first stage  we estimate $\theta_i$ the cluster specific parameters of the Heckman model.
+Briefly, in a first stage  we estimate $\boldsymbol{\theta_i}$ the cluster specific parameters of the Heckman model.
 That is, we estimate separately for each of the $N$ clusters,the parameters of the two equations in each study, the outcome model and the selection model.
-In the second stage, we combine all $\theta_i$  by using a random effects meta-analysis model.
+In the second stage, we combine all $\boldsymbol{\theta_i}$  by using a random effects multivariate meta-analysis model.
 
-<!-- I wouldn't call the distribution imaginary. It is assumed to be real but latent. Some words we could use: conjectured, latent, presupposed, presumed, Thanks :D -->
-In a random effects model, $\theta_i$ parameters are assumed to be drawn independently and identically from a latent distribution of parameters with a  population mean $\theta_m$ and a population variance $\psi$.
-[@higgins2009] Thus, the  $\theta_i=\theta_m + b_i$ can be specified using $b_i \sim N(0,\psi)$ random effects to allow for between-study heterogeneity in observed data relationships, and between-study heterogeneity in missing patterns.
+In a random effects model, $\boldsymbol{\theta_i}$ parameters are assumed to be drawn independently and identically from a latent multivariate normal distribution of parameters with a  population mean $\boldsymbol{\theta_m}$ and a population variance $\boldsymbol{\psi}$.
+[@higgins2009] Thus, we can write $\boldsymbol{\theta_i}=\boldsymbol{\theta_m} + \boldsymbol{b_i}$  using $\boldsymbol{b_i} \sim N(0,\boldsymbol{\psi})$ random effects to allow for between-study heterogeneity in observed data relationships, and between-study heterogeneity in missing patterns.
 
-# Method
+# Using the Heckman model to impute missing data
 \label{sec:methods}
 We follow a similar approach proposed by Resche-Rigon and White (2018) [@resche-rigon2018] for multilevel imputation of data.
-Briefly, their method was developed to impute variables from a hierarchical structure (i.e., when there are samples unit grouped within a cluster or group).
+Briefly, their method was developed to impute variables from a hierarchical structure (i.e., when there are units grouped within a cluster or group).
 This method involves estimating an outcome equation (describing the relationships of the observed data to the missing variable) separately in each cluster, after which the parameter estimates of that equation are pooled using random effects meta-analysis.
@@
 With this method, values can be imputed in very common scenarios in IPD, e.g., sporadic and systematic missing patterns.
-In particular, when the response variable is systematically missing within a group, i.e., when $y_{ij}$ are totally missing within a group, the imputation values are drawn from a (generalized) linear model conditional on $\theta_m$ the marginal population parameters, i.e., those estimated after pooling the cluster-specific parameters $\theta_i$.
-On the other hand, when the variable is sporadically missing within the group, i.e., there are some observed $y_{ij}$ within the cluster, the imputation model is conditional on the shrunk-cluster parameters, i.e., those coming from the shrinkage of $\theta_i$ towards $\theta_m$.
+In particular, when the response variable is systematically missing within a group, i.e., when $y_{ij}$ are totally missing within a cluster, the imputation values are drawn from a (generalized) linear model conditional on $\boldsymbol{\theta_m}$ the marginal population parameters, i.e., those estimated after pooling the cluster-specific parameters $\boldsymbol{\theta_i}$.
+On the other hand, when the variable is sporadically missing within the cluster, i.e., there are some observed $y_{ij}$ within the cluster, the imputation model is conditional on the shrunk-cluster parameters, i.e., those coming from the shrinkage of $\boldsymbol{\theta_i}$ towards $\boldsymbol{\theta_m}$.
 
-Our imputation approach differs crucially from the previous approach as here we estimate two correlated equations (instead of a single outcome equation) in each cluster, thus obtaining $\theta_i=\{\beta_i^O,\beta_i^S,\sigma_i,\rho_i\}$ parameters from both equations which are then pooled into a $\theta_m$ parameter set at the marginal or population level.
-Our imputation method was created for impute univariate incomplete datasets, but since it is implemented in a Gibbs sampling procedure, it can also be used to impute multiple incomplete variables in a data set.
-<!-- I don't understand what you mean here: it is univariate and it can be used to impute multiple variables?
-I would say these two statements contradict each other.
-yes you are right i tried to say is a mehtod for univariate incomplete dataset -->
+Our imputation approach differs crucially from the previous approach as here we estimate two correlated equations (instead of a single outcome equation) in each cluster, thus obtaining $\boldsymbol{\theta_i}=\{\boldsymbol{\beta_i^O},\boldsymbol{\beta_i^S},\sigma_i,\rho_i\}$ parameters from both equations which are then pooled into a $\boldsymbol{\theta_m}$ parameter set at the marginal or population level.
+Our imputation method was created for imputing datasets with a single missing outcome or covariate variable, but it can also be used to impute multiple incomplete variables in a data set.
 
-## Imputation of univariate incomplete dataset
-Given an outcome variable $y=(y_1,y_2,...,y_N)^T$, that consists of $y^{miss}_{ij}$ missing and $y^{obs}_{ij}$ observable values, we generate independent draws from the posterior predictive distribution for the missing data, $y_{ij}^{miss}$, given the observable data information $y_{ij}^{obs}$.
-$$p(y_{ij}^{miss}|y_{ij}^{obs})=\int_{\theta} p(y_{ij}^{miss}|\theta,y_{ij}^{obs})p(\theta|y_{ij}^{obs})\mathrm{d}\theta$$
-Here we implicitly assume vague prior distributions for each of the parameters included in the parameter vector $\theta$.
-Because the integration can be performed computationally by sampling from the posterior predictive distribution $p(\theta|y_{ij}^{obs})$, our imputation method can be carried out in the following two steps:
+In the following subsections, we explain the method when the incomplete variable is the outcome variable but this method can also be used to impute any incomplete MNAR predictor variable in the dataset.
+In the latter case, in the outcome equation (\ref{eq:1}), the incomplete predictor variable must be specified as the dependent variable and $\boldsymbol{X^O_{i}}$ as the set of covariates associated with it.
+On the other hand, in the selection model(\ref{eq:2}), $\boldsymbol{r^*_{i}}$ corresponds to the latent variable of selection of the incomplete predictor variable together with its $\boldsymbol{X^S_{i}}$ associated set of predictors.
 
-1. Draw a ${\theta}$ parameter vector, ${\theta^*}$, from $p(\theta|y_{ij}^{obs})$, their posterior distribution.
+## Univariate imputation
+Given an outcome variable $\boldsymbol{y}=(y_1,y_2,...,y_N)^T$, that consists of $y^{miss}_{ij}$ missing and $y^{obs}_{ij}$ observable values, we generate independent draws from the posterior predictive distribution for the missing data, $y_{ij}^{miss}$, given the observable data information $y_{ij}^{obs}$.
+$$p(y_{ij}^{miss}|y_{ij}^{obs})=\int_{\boldsymbol{\theta}} p(y_{ij}^{miss}|\boldsymbol{\theta},y_{ij}^{obs})p(\boldsymbol{\theta}|y_{ij}^{obs})\mathrm{d}\boldsymbol{\theta}$$
+Here we implicitly assume vague prior distributions for each of the parameters included in the parameter vector $\boldsymbol{\theta}$.
+Because the integration can be performed computationally by sampling from the posterior predictive distribution $p(\boldsymbol{\theta}|y_{ij}^{obs})$, our imputation method can be carried out in the following two steps:
 
-2. Draw $y_{ij}^{miss}$ from $p(y_{ij}^{miss}|\theta^*)$, their predictive distribution for a given $\theta^*$ vector.
+1. Draw a $\boldsymbol{\theta}$ parameter vector, $\boldsymbol{\theta^*}$, from $p(\boldsymbol{\theta}|y_{ij}^{obs})$, their posterior distribution.
+
+2. Draw $y_{ij}^{miss}$ from $p(y_{ij}^{miss}|\boldsymbol{\theta^*})$, their predictive distribution for a given $\boldsymbol{\theta^*}$ vector.
 
@@
 
-### Draw the $\theta^*$ parameter vector
+### Draw the $\boldsymbol{\theta^*}$ parameter vector
 
-#### Fit $p(y_{ij}^{obs}|\theta_i)$, the heckman selection model at group level
+#### Fit $p(y_{ij}^{obs}|\boldsymbol{\theta_i})$, the heckman selection model at cluster level
 
-Initially, we use the copula method to estimate the set of cluster-specific parameters, $\widehat{\theta_i}=\{\widehat{\beta^O_i}, \widehat{\beta^S_i}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, using all $j$ units with observable measurements $y_{ij}^{obs}$ within each group $i$.
-The Heckman model is estimated with the \textbf{gjrm} function of the GJRM R package, [@radiceGJRMGeneralisedJoint2021] under the bivariate model with the nonrandom sample selection (BSS) specification, from which we obtain not only the parameters' point estimates $\widehat{\theta_i}$, but also their corresponding $\widehat{S(\theta_i)}$ within-cluster variance-covariance matrix.
-
-<!-- Can you add a reference to the R package here? OK -->
-
-<!-- When the model does not converge in a study, we assume that data follows a MAR mechanism so we fixed $\widehat{\rho_i} = 0$ and we estimate only the outcome equation parameters from a (generalized) linear model.
--->
+Initially, we use the copula method to estimate the set of cluster-specific parameters, $\boldsymbol{\widehat{\theta_i}}=\boldsymbol{\{\widehat{\beta^O_i}}, \boldsymbol{\widehat{\beta^S_i}}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, using all $j$ units with observable measurements $y_{ij}^{obs}$ within each cluster $i$.
+The Heckman model is estimated with the \textbf{gjrm} function of the GJRM R package, [@radiceGJRMGeneralisedJoint2021] under the bivariate model with the nonrandom sample selection (BSS) specification, from which we obtain not only the parameters' point estimates $\boldsymbol{\widehat{\theta_i}}$, but also their corresponding $\boldsymbol{\widehat{S(\theta_i)}}$ within-cluster variance-covariance matrix.
 
 #### Fit a meta-analysis model
-<!-- This paragraph starts of in the the present tense " we pool ", and later it's past tense " we pooled" etc.
-Stick to one tense.
-In applied papers, it is normal to use only past tense.
-In stats papers it is common to use present tense when describing a (new) method in the methods section.
--->
-In this step, we pool the parameters $\widehat{\theta_i}$ with a random effects meta-analysis model using only the groups with observable information, i.e., those that are not systematically missing and have sufficient information to estimate the heckman model.
-In particular, we pool the $p$ coefficients of the $\beta^O$ outcome equation and estimated a multivariate random effects meta-analysis model with them, similarly we combine all $q$ coefficient parameters of the $\beta^S$ selection equation.
-<!-- The phrase " the log-transformed parameter of $\sigma$" is a description of $\sigma'$, so it is what is known as a commenting or parenthetical clause.
-That means you need to put a comma before and after this clause: "...
-on $\sigma'$, the log-transformed parameter of $\sigma$, ...".
-Although this article is not for the BMJ, I can recommend reading their house-style for writing: https://www.bmj.com/about-bmj/resources-authors/house-style -->
-<!-- Read more here if you like: http://content.principia.edu/teaching-excellence/commenting-with-commas-when-and-where-to-use-them/ Thanks a lot you are a walking library :D, very appreciated all your lessons!!
--->
-We also performed a univariate random effects meta_analysis on $\sigma'$, the log-transformed parameter of $\sigma$, and another on $\rho'$, the fisher-transformed parameter of $\rho$.
+
+In this step, we pool the parameters $\boldsymbol{\widehat{\theta_i}}$ with a random effects meta-analysis model using only the clusters with observable information, i.e., those with no systematically missing outcome.
+In particular, we pool the $p$ coefficients of the $\boldsymbol{\beta^O}$ in the outcome equation and estimate a multivariate random effects meta-analysis model with them.
+Similarly we combine all $q$ coefficient parameters of the $\boldsymbol{\beta^S}$ in the selection equation.
+
+We also perform a univariate random effects meta_analysis on $\sigma'$, the log-transformed parameter of $\sigma$, and another on $\rho'$, the fisher-transformed parameter of $\rho$.
 
 The meta-analysis model is performed with the \textbf{mixmeta} function of the R package mixmeta[@gasparrini2021], which allows the use of maximum likelihood (ML), restricted maximum likelihood (REML), and moments estimation methods.
-For the simulation and illustrative study, we used the restricted REML estimation method, which is recommended as it has a good balance between unbiasedness and efficiency.
+For the simulation and illustrative study, we use the restricted REML estimation method, which is recommended as it has a good balance between unbiasedness and efficiency.
 [@viechtbauer2005]
-<!-- Insensitivity is not generally a good thing.
-I think you should either explain why it is a good thing here: "...
-it is insensitive to the true value of the mean of the distribution of random effects and it is efficient" (or something similar, feel free to rephrase) , or simplify it to "...
-a good balance between unbiasedness and efficiency.
-jajaj I think at certain point i used the translator to correct my writing and i did not notice that change unbiasedness by insensitive..But thanks God you have eagle eyes @@) .."  -->
 
-#### Draw $\theta^*_m$ the marginal parameters
+#### Draw $\boldsymbol{\theta^*_m}$ the marginal parameters
 
-From the meta-analysis model, we obtain the marginal estimates $\widehat{\theta_m}$ and the between-cluster variance $\widehat{\psi}$ with their corresponding variance-covariance matrices $\widehat{S_{\theta_m}}$ and $\widehat{S_{\psi}}$, which are used to draw the $\theta^*_m$ and $\psi^*$ parameters as follows:
+From the meta-analysis model, we obtain the marginal estimates $\boldsymbol{\widehat{\theta_m}}$ and the between-cluster variance matrix $\boldsymbol{\widehat{\psi}}$,i.e.
+variance-covariance matrix of the random effects, with their corresponding variance-covariance matrices $\boldsymbol{\widehat{S_{\theta_m}}}$ and $\boldsymbol{\widehat{S_{\psi}}}$, which are used to draw the $\boldsymbol{\theta^*_m}$ and $\boldsymbol{\psi^*}$ parameters, from their posterior distribution as follows: [@resche-rigon2013]
 
 \begin{align*}
-\theta_m^*& \sim N(\widehat{\theta_m},\widehat{S_{\theta_m}})\\
-\psi^*& \sim N(\widehat{\psi},\widehat{S_{\psi}})
+\boldsymbol{\theta_m^*}& \sim N(\boldsymbol{\widehat{\theta_m}},\boldsymbol{\widehat{S_{\theta_m}}})\\
+\boldsymbol{\psi^*}& \sim N(\boldsymbol{\widehat{\psi}},\boldsymbol{\widehat{S_{\psi}}})
 \end{align*}
 
-#### Draw the cluster parameters $\theta^*_i$
-<!-- Throughout the text, consider the order in which you put the symbol and the text: -->
-<!-- Put the symbol after the text if the text is a description of the symbol: "shrunk-cluster parameters $\theta^*_i$" . -->
-<!-- Put the text after the symbol if the symbol quantifies the text: "$n$ patients" , or " $N$ clusters" . Whoahh you are a gramatical Lord thanks a lot!! -->
-We draw the shrunk-cluster parameters $\theta^*_i$ for each group $i$ from the following posterior distribution conditional on $\theta^*_m$ and $\psi^*$.
+#### Draw the cluster parameters $\boldsymbol{\theta^*_i}$
+
+We draw the shrunked-cluster-parameters $\boldsymbol{\theta^*_i}$ for each cluster $i$ from the following posterior distribution conditional on $\boldsymbol{\theta^*_m}$ and $\boldsymbol{\psi^*}$.
+
 \begin{align*}
-\theta^*_i\sim N\left(\frac{ \theta^*_m/\psi^*+\widehat{\theta_i}/\widehat{S_{\theta_i}}}{1/\psi^*+1/\widehat{S_{\theta_i}}},\frac{1}{1/\psi^*+1/\widehat{S_{\theta_i}}}\right)
+\theta^*_i\sim N\left(\left({\boldsymbol{\psi^*}}^{-1}+{\boldsymbol{\widehat{S_{\theta_i}}}}^{-1}\right)^{-1}\left(\boldsymbol{\psi^*}^{-1}\boldsymbol{\theta^*_m}+\boldsymbol{\widehat{S_{\theta_i}}}^{-1}\boldsymbol{\widehat{\theta_i}} \right),\left({\boldsymbol{\psi^*}}^{-1}+{\boldsymbol{\widehat{S_{\theta_i}}}}^{-1}\right)^{-1} \right)
 \end{align*}
@@
 As can be seen, the mean and variance of the posterior distribution is a combination between the estimated marginal and cluster-specific parameters.
-Here the weights on the cluster-specific parameters $\widehat{\theta_i}$ and the marginal parameters $\theta^*_m$ are inversely proportional to the within cluster variance $\widehat{S_{\theta_i}}$ and between clusters variance $\psi^*$.
-For example, when $\widehat{S_{\theta_i}} < \psi^*$ the mean of the conditional distribution gives more weight to the estimated cluster-specific parameter.
-Conversely, when $\widehat{S_{\theta_i}}>\psi^*$, more weight is given to the estimated marginal parameters.
-Therefore, in case of a cluster with systematic missingness, it is as if the within-cluster variance is infinite ($\widehat{S_{\theta_i}}\to\infty$), so all the weight is assigned to the parameters estimated at the marginal level.
+Here the weights on the cluster-specific parameters $\boldsymbol{\widehat{\theta_i}}$ and the marginal parameters $\boldsymbol{\theta^*_m}$ are inversely proportional to the within cluster variance $\boldsymbol{\widehat{S_{\theta_i}}}$ and between clusters variance $\boldsymbol{\psi^*}$.
+For example, when $\boldsymbol{\widehat{S_{\theta_i}}} < \boldsymbol{\psi^*}$ the mean of the conditional distribution gives more weight to the estimated cluster-specific parameter.
+Conversely, when $\boldsymbol{\widehat{S_{\theta_i}}}>\boldsymbol{\psi^*}$, more weight is given to the estimated marginal parameters.
+Therefore, in case of a cluster with systematic missingness, it is as if the within-cluster variance is infinite ($\boldsymbol{\widehat{S_{\theta_i}}}\to\infty$), so all the weight is assigned to the parameters estimated at the marginal level.
 That is, when there is no information in a cluster, we rely entirely on the marginal information.
-
-<!-- What does then mean in this last part? should it be " and lettting all parameters rely ..." ? I think now is more clear, thanks!! -->
 
@@
 
-Having $\theta^*_i$, the shrunk-cluster parameters vector for each group, we back-transform $\sigma^*$ and $\rho^*$ to the original scale.
-Then $y^{miss}_{ij}$, the missing values, can be drawn from $p(y^{miss}_{ij}|\theta^*_i)$, their predictive distribution given $\theta^*_i$, as follows:
+Having $\boldsymbol{\theta^*_i}$, the shrunk-cluster parameters vector for each cluster, we back-transform $\sigma^*$ and $\rho^*$ to the original scale.
+Then $y^{miss}_{ij}$, the missing values, can be drawn from $p(y^{miss}_{ij}|\boldsymbol{\theta^*_i})$, their predictive distribution given $\boldsymbol{\theta^*_i}$, as follows:
 
@@
 
-The imputed value of $y_{ij}^{miss}$ can be drawn from the conditional expectation of $y_{ij}$ on unobserved measurements:
+The imputed value of $y_{ij}^{miss}$ can be drawn from the conditional expectation of $y_{ij}$ on unobserved measurements: [@greene2018]
 \begin{align*}
-\mu&= E[y_{ij}|r_{ij}=0,{\beta^O_{i}}^*,{\beta^{S}_{i}}^*,{\rho_{i}}^*,{\sigma_i}^*]\\
-\mu&= x^O_{ij}{\beta^O_{i}}^*+\rho_{i}^*\sigma_i^*\frac{-\phi(x^{S}_{ij}{\beta^{S}_{i}}^*)}{\Phi(-x^{S}_{ij}{\beta^{S}_i}^*)}\\
+\mu&= E[y_{ij}|r_{ij}=0,{\boldsymbol{\beta^O_{i}}^*},\boldsymbol{{\beta^{S}_{i}}^*},{\rho_{i}}^*,{\sigma_i}^*]\\
+\mu&= \boldsymbol{x^O_{ij}}\boldsymbol{{\beta^O_{i}}^*}+\rho_{i}^*\sigma_i^*\frac{-\phi(\boldsymbol{x^{S}_{ij}}\boldsymbol{{\beta^{S}_{i}}^*})}{\Phi(-\boldsymbol{x^{S}_{ij}}{\boldsymbol{\beta^{S}_i}^*})}\\
 y_{ij}^{miss}&\sim N(\mu,{\sigma_i^*}^2)
 \end{align*}
+
+where $\phi(.)$ is the probability density function and $\Phi(.)$ is the cumulative distribution function of the standard normal distribution.
 
@@
 \begin{align*}
-p^*_{ij} &= P[y_{ij}=1|r_{ij}=0,{\beta^O_{i}}^*,{\beta^{S}_{i}}^*,{\rho_{i}}^*]\\
-p^*_{ij} &=\frac{\Phi_2(x^O_{ij}{\beta^{O}_i}^*,-x^S_{ij}{\beta^{S}_i}^*,-\rho_i)}{\Phi(-x^{S}_{ij}{\beta^{S}_i}^*)}\\
+p^*_{ij} &= P[y_{ij}=1|r_{ij}=0,{\boldsymbol{\beta^O_{i}}^*},{\boldsymbol{\beta^{S}_{i}}^*},{\rho_{i}}^*]\\
+p^*_{ij} &=\frac{\Phi_2(\boldsymbol{x^O_{ij}}\boldsymbol{{\beta^{O}_i}^*},-\boldsymbol{x^S_{ij}}\boldsymbol{{\beta^{S}_i}^*},-\rho_i)}{\Phi(-\boldsymbol{x^{S}_{ij}}{\boldsymbol{\beta^{S}_i}^*})}\\
 y_{ij}^{miss}&\sim Ber(p^*_{ij})
 \end{align*}
+where  $\Phi_2(.)$ corresponds to the bivariate normal cumulative distribution function.
 
-## Imputation of multivariate incomplete dataset
+## Multivariate imputation
 
@@
 Particularly, our imputation method has been implemented according to the structure of the MICE R package,[@buuren2021] that allows imputing multiple incomplete predictors and covariates in a given dataset.
-
-<!-- The next paragraph belongs in the manual for the package, or a vignette. I don't think it is necessary to put it here. 
-The MICE package allows one to specify imputation methods to each of the missing variables by setting the method and the predictive matrix for each of the missing variables.
-To use our proposed method, it is necessary to specify for the MNAR missing variable the method \textbf{2l.heckman} in the mice methods vector.
-Furthermore, in the prediction matrix, the group or cluster variable should be specified as \textbf{'-2'}, all predictor variables belonging to the selection and outcome as \textbf{'1'}, the exclusion restrictions or predictor variables that are only included in the selection equation as \textbf{'-3'} and those that are only included in the outcome equation as \textbf{'-4'}.
-Please refer to the toy example in the attached github repository to better understand how to implement the imputation model.
--->
 
@@
 ## Aim
-We designed a simulation study aimed to compare the performance of imputation methods for imputing a missing variable in a hierarchical dataset, where the missingness follows a MNAR mechanism.
+We designed a simulation study aimed to compare the performance of alternative methods for imputing a single missing outcome variable in a hierarchical dataset, where the missingness follows a MNAR mechanism.
+In our scenarios we considered systematically missingness,i.e.
+the case some of the studies did not report hte variable at all.
 
@@
 
-We generated the data from the Heckman selection model with bivariate normal distribution error terms.
-For simplicity we assume a "basic scenario" where the database collects information from $N=10$ clusters of equal number of individuals $n_i=1000$.
-However, we altered both $N$ and $n_i$ in the sensitivity analysis (see below).
-<!-- But you also have scenarios with different numbers for N and n_i.
-You could write here that for the "basic scenario" we have N =  and n_i = , and that we alter this in other scenarios, or in " senstivity analyses" ok .
--->
+We generated the data from a Heckman selection model with bivariate normal distribution error terms.
+For simplicity we started from "basic scenario" where the database collected information from $N=10$ clusters of $n_i=1000$  individuals.
+However, we altered both $N$ and $n_i$ in sensitivity analyses (see below).
+
 For each dataset, we generated $X_{1i}$ a treatment indicator variable from a Bernoulli distribution with a probability of treatment on each cluster equal to 0.6.
 Next, we simulated the mean of two continuous covariates from a multivariate normal distribution $\mu_h\sim N(\begin{psmallmatrix}0\\0\end{psmallmatrix},\begin{psmallmatrix}0.2 & 0.015\\0.015 & 0.2\end{psmallmatrix})$, with $h=\{2,3\}$.
-We then simulated for each cluster a baseline covariate $X_{2i}\sim N(\mu_2,1)$ and a exclusion restriction  $X_{3i} \sim N(\mu_3,0.5)$.
+We then simulated for each cluster a baseline covariate $X_{2i}\sim N(\mu_2,1)$ and a exclusion restriction variable  $X_{3i} \sim N(\mu_3,0.5)$.
 
-Here, we considered $X_{1i}$ and $X_{2i}$ as predictors in the outcome equation $X_i^O=[1,X_{1i},X_{2i}]$.
-For the selection equation we included both variables and the $X_{3i}$ exclusion restriction, $X_i^S=[1,X_{1i},X_{2i},X_{3i}]$.
-Then in case of a missing continuous variable, we calculate the latent variables  $y_{i}^*$ and $r^*_{i}$ as follows:
+Here, we considered $X_{1i}$ and $X_{2i}$ as predictors in the outcome equation, i.e., $\boldsymbol{X_i}^O=[1,\boldsymbol{X_{1i}},\boldsymbol{X_{2i}}]$.
+For the selection equation we included both variables and the $X_{3i}$ exclusion restriction variable, $\boldsymbol{X_i^S}=[1,\boldsymbol{X_{1i}},\boldsymbol{X_{2i}},\boldsymbol{X_{3i}}]$.
+Then in case of a missing continuous variable, we calculate the latent variables  $\boldsymbol{y_{i}^*}$ and $\boldsymbol{r^*_{i}}$ as follows:
 \begin{align*}
- y_{i}^*&= \beta_i^OX_i^O+\epsilon^O_{i} \\
- r_{i}^*&= \beta_i^SX_i^O+\epsilon^S_{i}
+ \boldsymbol{y_{i}^*}&= \boldsymbol{\beta_i^O}\boldsymbol{X_i^O}+\boldsymbol{\epsilon^O_{i}} \\
+ \boldsymbol{r_{i}^*}&= \boldsymbol{\beta_i^S}\boldsymbol{X_i^O}+\boldsymbol{\epsilon^S_{i}}
 \end{align*}
@@
 Additionally, we ensured that the $y_{ij}^*$ observations were systematically missing in 20% of the clusters included in the data set, by removing the outcome values in the 20% of the clusters.
-<!-- How did we ensure that these were systematically missing in 20% of the clusters?
-Does that follow from the fixing of the betas?
-Or did you remove the y values in 20% of the clusters?
-Please be specific.
-OK-->
+
 We assumed that random effects were independent within equations ($b_{h0}^O\indep b_{h1}^O \indep b_{h2}^O$ and $b_{h0}^S \indep b_{h1}^S \indep b_{h2}^S$), but were linked between both selection and outcome equations through a bivariate normal distributed as:
@@
 
-To investigate the performance of the imputation methods under the following scenarios:
+In addition to the basic scenario described above, we explored additional data generating mechanisms.
+Specifically we investigated the performance of the imputation methods under the following scenarios:
 
-- \textbf{M(N)AR scenarios:} We assessed whether the model performed well in terms of bias and coverage when the data followed a missing MAR mechanism ($\rho=0$), and when data followed a MNAR mechanism with a low ($\rho=0.3$), intermediate ($\rho=0.6$) and strong correlation ($\rho=0.9$) between $y^*$ and $r^*$.
-We evaluated also the imputation method when the missing variable is binary.
+- \textbf{M(N)AR scenarios:} We explored a missing MAR mechanism ($\rho=0$), a scenario when data followed a MNAR mechanism with a low ($\rho=0.3$), intermediate ($\rho=0.6$) and strong correlation ($\rho=0.9$) between $y^*$ and $r^*$.
+We also explored a scenario when the missing variable is binary.
 Therefore we simulated a $y_{i}$ binary incomplete variable, by keeping similar parameters to the ones used in the simulation of a missing continuous variable, but here we defined the observable binary variable as:
@@
 
-- \textbf{Influence of sample size}: Model sensitivity was analysed with respect to the the number of patients per cluster $n_i=\{50,100,1000\}$  and the number of studies $N=\{10,50,100\}$.
+- \textbf{Influence of sample size}: we explored different configurations regarding the number of patients per cluster $n_i=\{50,100,1000\}$  and the number of studies $N=\{10,50,100\}$.
 
-- \textbf{Violation of distributional assumptions:} To assess how the imputation models behave in the face of deviations from normality assumptions, we simulated data in which the errors followed a skewed t-distribution and also in which the missing process follows a MNAR mechanism with an explicit truncated model, i.e., the participation is directly related to the value of the outcome variable.
+- \textbf{Violation of distributional assumptions:} Here, we aimed to investigate how the imputation models behave in settings with departures from the bivariate normal distribution, performed two sensitivity analyses.
 
-## Estimand
+ 1.
+\textbf{Skewed-t:} We drew error terms from a bivariate skewed student-t distribution using the same location parameter and covariance matrix of the normal distributed settings,  with 4 degrees of freedom and an $\alpha=\{-2,6\}$ parameter which regulates the the slant of the density.
+ 2.
+\textbf{Explicit:} In addition we simulated an explicit missingness process, where error terms of the selection and outcome equations were independently normal distributed and the selection of observations depended on the value of the outcome variable, as $ry^*_i=0.3y^*_i+\epsilon^S_{i}$.
+These settings assure that the percentage of missingness on the outcome variable is around 60\% in all the evaluated scenarios.
+ 3. \textbf{Normal:} As a reference, we provide the results from the basic scenario of the main simulation study, where the error terms were Bivariate normal distributed.
+
+## Estimands
 The estimands were the parameter coefficients of the outcome equation $\beta^O={\beta^O_0,\beta^O_1,\beta^O_2}$, with special emphasis on the treatment effect parameter $\beta^0_1$.
@@
 
-After the imputation procedure, we estimated the following (generalized) mixed linear effect model using the *lmer()* function from the *lme4* R package.[@bates2022]
-$y_{i}=\beta_i^OX_i^O+\epsilon^O_{i}$
-In case of missing binary variable, we used the same matrix of predictors but on a binary model estimated with the *glmer()* function from the *lme4* R package.
-Then, we pooled the estimates of the $\beta_i^O$ and the variance of the random effect and residual errors of the multiple imputed datasets according to Rubin's rule, [@rubin1987] over which we calculated the performance measures on the estimands.
-
-To calculate the coverage of the parameter coefficients' \95% confidence intervals (CI), we estimate CI with the Wald method.
-Although it is possible to obtain CI through the profile or bootstrap method for the variances of the random effects, we prefer not to estimate them (and hence the coverage) for the random effects parameters due to computational time.
-
-## Method
+## Analysis Methods
 
@@
 
+## Estimating procedures
+
+After the imputation procedure, we estimated the following (generalized) mixed linear effect model using the *lmer()* function from the *lme4* R package.[@bates2022]
+$y_{i}=\beta_i^OX_i^O+\epsilon^O_{i}$
+In case of missing binary variable, we used the same matrix of predictors but on a binary model estimated with the *glmer()* function from the *lme4* R package.
+Then, we pooled the estimates of the $\beta_i^O$ and the variance of the random effect and residual errors of the multiple imputed datasets according to Rubin's rule, [@rubin1987] over which we calculated the performance measures on the estimands.
+
+To calculate the coverage of the parameter coefficients' \95% confidence intervals (CI), we estimate CI with the Wald method.
+Although it is possible to obtain CI through the profile or bootstrap method for the variances of the random effects, we prefer not to estimate them (and hence the coverage) for the random effects parameters due to computational time.
+
 ## Performance measures
 
-We calculated the following evaluation criteria, usually employed to evaluate imputation methods, [@buuren18c] according to the formulas provided in Morris et al.(2019): [@morris2019]
+We calculated the following measures, usually employed to evaluate imputation methods, [@buuren18c] according to the formulas provided in Morris et al.(2019): [@morris2019]
 
@@
 For the simulation study and illustrative examples we used R version 4.0.4 in a linux environment. [@Rpackage]
-<!-- Add a reference to R if you haven't done so yet. Run citation() in R to get a bibtex entry. Mago!!! -->
+
 The Heckman 2L imputation method is available in the mice R package (as \textbf{mice.2l.heckman()}) and also on the github repository \url{https://github.com/johamunoz/Statsmed_Heckman} where you can also find all the codes accompanying this paper and a toy example that explains how to implement the method in mice.
 
-## Results
+## Results from the simulation study
 
@@
 
-### Descriptive results
-We generated data sets of 10 groups of 1000 patients each in each scenario.
+### Descriptive statistics of the simulated datasets
+We generated data sets of 10 clusters of 1000 patients each in each scenario.
 For example, for the scenario in which the error terms followed a normal distribution, out of the 500 datasets generated, we obtained that on average 60.8\% of the Y response was missing, with the lowest missing percentage being 26.84\% and the maximum being 73.67\%.
@@
 #### Continuous incomplete variable
- ```{r rhocont,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable by varing $\\rho$, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
-plot_rho
- ```
 
- <!-- What do you mean with "Even" here? Do you mean to say "Even though" ? Can you add one sentence on what you mean that it is not randomization valid? OK -->
-<!-- What is "Heckman IPD separated" Is that the 2l.Heckman method? Yes jaja-->
-<!-- I would not equate accuracy with CI width -->
 Figure \ref{fig:rhocont} shows the results of simulations where the missing variable was continuous.
@@
 
+ ```{r rhocont,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable by varing $\\rho$, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.7, fig.width = 6}
+plot_rho
+ ```
 Overall, both Heckman-based imputation (1l.
@@
 
-Regarding the coverage, the 2l.Heckman imputation method provided the best coverage values across all the $\rho$ scenarios with a coverage level close to the nominal 95\% interval.
-Even though 2l.Heckman was not properly a randomization-valid method [@buuren18c], as it was not unbiased and had a coverage above 95\% across all the $\rho$ values, it leaded to better results in terms of bias and coverage compared to the evaluated methods.
+Regarding coverage, the 2l.Heckman imputation method provided the best coverage values across all the $\rho$ scenarios with a coverage level close to the nominal 95\% interval.
+Even though 2l.Heckman was not properly a randomization-valid method [@buuren18c], as it was not unbiased and had a coverage above 95\% across all the $\rho$ values, it led to better results in terms of bias and coverage compared to the evaluated methods.
 
-The 2l.Heckman method, by incorporating individual and marginal level information, allowed a bias-variance trade-off.
-Thus, we observed that the 2l.Heckman resulted in better estimates in terms of RMSE than the other methods evaluated.
+The 2l.Heckman method, also resulted in better estimates in terms of RMSE than the other methods evaluated, via a better trade-off between bias and variance.
 
-In particular, the method provided an advantage over the 1l.Heckman method in systematic missingness scenarios, since the latter does not allow the specification of any cluster information implicitly, i.e.
-by adding a group variable in the imputation model.
-This can be seen in the estimates of the random effects parameters of the 1l.Heckman method, which were more biased than those of the other methods in which cluster information was included.
+In particular, our method provided an advantage over the 1l.Heckman method in systematic missingness scenarios, by adding a cluster variable in the imputation model, feature lacking from 1l.Heckman.
+This can be seen in the estimates of the random effects parameters of the 1l.Heckman method, which were more biased than those of the other methods in which cluster information was included, i.e.
+2l.MAR and 2l.Heckman.
 
-Regarding the CI length of the 95th percentile of the estimated parameters, we observed that using the 2l.Heckman model we obtained larger CIs than when using other evaluated methods, which generally increased as $\rho$ moved away from zero.
+Regarding the length of the 95\% CI of the estimated parameters, we observed that using the 2l.Heckman model we obtained larger CIs than when using other evaluated methods, which generally increased as $\rho$ moved away from zero.
 In particular, we observed in $\beta_2$ that the CI length was larger in the MAR scenario.
@@
 
-In case of a bivariate missing variable, we also find (Figure \ref{fig:rhobin}) that the 2l.Heckman method provides unbiased results on coefficient parameters and random effects when the missing variable is binary.
+In case of a bivariate missing variable, we also found (Figure \ref{fig:rhobin}) that the 2l.Heckman method provided unbiased results on coefficient parameters and random effects.
 However, we observe more variability in the estimates of random effects, reflected on longer confidence intervals on the Bias and RMSE criteria.
 
-```{r rhobin,fig.align='center',fig.pos="!htb",fig.cap="Binary incomplete variable by varing $\\rho$, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
+```{r rhobin,fig.align='center',fig.pos="!htb",fig.cap="Binary incomplete variable by varing $\\rho$, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.7, fig.width = 6}
 plot_bin
@@
 
-```{r size,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable under systematical missingness by varing: number of clusters (N); sample size per cluster ($n_i$), where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
+```{r size,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable under systematical missingness by varing: number of clusters (N); sample size per cluster ($n_i$), where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.7, fig.width = 6}
 plot_N
@@
 
-<!-- Again, try to use only one tense: past or present tense. This paragraph mixed both within a single sentence. It is standard to write the results and discussion sections in past tense. -->
-<!-- What do you mean with "the average width of the estimates"?
-Are you referring to the CI width for the individual estimates?
-Or are you referring to the width of the estimates of the bias as resulting from monte carlo error?
--->
-By increasing N, the number of clusters, from 10 to 100, we observed that the bias was not affected but the CI width of the estimates decreased (larger precision) and hence the RMSE decreased.
+By increasing N, the number of clusters, from 10 to 100, we observed that the bias was not affected but the width of the 95\% CI of the estimates decreased (larger precision) and hence the RMSE decreased.
 On the other hand, by reducing the number of units per cluster (from $n_i$=1000 to $n_i$=100) the precision decreased for all coefficients, the bias on coefficient estimates were not drastically affected but the variances of random effects did.
@@
 
-```{r dist,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable with deviations in distribution assumptions, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 4, fig.width = 6}
+```{r dist,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable with deviations in distribution assumptions, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.7, fig.width = 6}
 plot_S
 ```
-<!-- Ideally, the phrase " where the dashed line depicts the true simulated parameter" belongs in a note under the caption. -->
-
-<!-- This next paragraph belongs in the methods section.
-Alternatively, you could make a (sub) section for a sensitivity analysis, where you could describe both the methods and results of this sensitivity analysis (which I did here.
--->
-
-Here, we aimed to investigate how the imputation models behave in settings with departures from the bivariate normal distribution, provide two sensitivity analyses.
-
- 1.
-\textbf{Skewed-t:} We drew error terms from a bivariate skewed student-t distribution using the same location parameter and covariance matrix of the normal distributed settings,  with 4 degrees of freedom and an $\alpha=\{-2,6\}$ parameter which regulates the the slant of the density.
- 2.
-\textbf{Explicit:} In addition we simulated an explicit missingness process, where error terms of the selection and outcome equations were independently normal distributed and the selection of observations depended on the value of the outcome variable, as $ry^*_i=0.3y^*_i+\epsilon^S_{i}$.
-These settings assure that the percentage of missingness on the outcome variable is around 60\% in all the evaluated scenarios.
- 3. \textbf{Normal:} As a reference, we provide the results from the basic scenario of the main simulation study, where the error terms were Bivariate normal distributed.
-
-<!-- Replace "Name 2" with a short name for this scenario, preferably not an equation. Then use the same name in the Figure. Thanks a lot it is awesome :D!!-->
 
@@
 
-Malaria is a mosquito-borne disease and, especially in children and pregnant women, is the leading cause of illness and death in Africa.
+Malaria is a mosquito-borne disease and is the leading cause of illness and death in Africa, especially in children and pregnant women.
 To prevent the spread of the disease, long-lasting nets (LLINs) and indoor residual spraying (IRS) in at-risk households are used as control measures.
@@
 In each sub-district, a sample of households with at least one child aged 2-10 years was surveyed, where information was collected on household conditions and use of preventive measures.
-In addition, finger prick blood samples were taken from each child to determine the prevalence of parasitemia and an etymological study was conducted to estimate mosquito prevalence.
+In addition, finger prick blood samples were taken from each child to determine the prevalence of parasitemia and an entomological study was conducted to estimate mosquito prevalence.
 Details of the project and survey are provided elsewhere.
@@
 
-<!-- I see you wrote both parasitemia and parasitaemia. Choose one -->
 To illustrate our proposed method, following the article by Rugnao et al.
@@
 Under the MAR assumption, the test values of children who were not present during the survey were imputed with the 2l.2stage.bin method of the MICEMD package, where the community was taken as the cluster and the following factors previously associated with parasitemia were used as predictors in the imputation model: sex, two-person mosquito net.
-In addition, we included age as a power 3 spline function, the cluster-level Log10 mean of the number of female anopheline mosquitoes per household estimated from the etymological survey, and the household wealth index from principal components analysis calculated specifically for the surveyed households.
+In addition, we included age as a power 3 spline function, the cluster-level Log10 mean of the number of female anopheline mosquitoes per household estimated from the entomological survey, and the household wealth index from principal components analysis calculated specifically for the surveyed households.
 
@@
 The selection and outcome equation included the same predictor variables as used under the MAR approach.
-In addition, we included a holiday indicator variable as ERV, which was calculated according to school vacation calendars and public holidays in Uganda in 2017.
+In addition, we included a holiday indicator variable as ERV.
+This was calculated according to school vacation calendars and public holidays in Uganda in 2017.
 We examined the association of this ERV with the outcome variable (y) and with the selection indicator (ry), conditioned on the remaining imputation predictors.
-The model results in Table \ref{tab:exclusion} indicate that the holiday indicator could be a plausible ERV variable, as it was significantly associated with ry, but not with y.
+The model results in Table \ref{tab:exclusion} indicate that the holiday indicator could be a plausible ERV variable, as there was strong evidence of an association with ry, but no evidence of an association with y.
 ```{r exclusion, echo = FALSE, warning = FALSE,cache = TRUE}
@@
 According to our imputation approach, non-tested children were estimated to have a higher prevalence of malaria than participants in more than half of the districts analyzed. 
-As can be seen in Figure \ref{fig:region}, for each subregion the prevalence estimates of the approaches do not differ significantly between methods.
+As can be seen in Figure \ref{fig:region}, for each subregion the prevalence estimates of the approaches did not differ significantly between methods.
 However, prevalence estimates under the MNAR assumption (i.e.
-2 level Heckman)  are higher than those estimated under the MAR or MCAR aproaches, except for the East-Central region.
+2 level Heckman) were higher than those estimated under the MAR or MCAR aproaches, except for the East-Central region.
 ```{r region,fig.align='center',fig.width = 5, fig.asp =.67, fig.pos="!htb", fig.cap="Estimates of malaria prevalence by sub-region",echo=FALSE,fig.keep="last",cache=TRUE}
@@
 
-<!-- The age at which children start going to school is not universal, so you need to mention this. done!! thanks -->
-In terms of prevalence by age, there are no significant differences between methods (Figure \ref{fig:region_age}).
-The prevalence estimates for children aged 2 to 6 years are very similar in all regions under the different assumptions.
+In terms of prevalence by age, there were no significant differences between methods (Figure \ref{fig:region_age}).
+The prevalence estimates for children aged 2 to 6 years were very similar in all regions under the different assumptions.
 Assuming that children start going to school after the age of 6, the results could be partly explained by the mobility of 2-6 year olds compared to school-age children, i.e.
 school-age children spend more time outdoors and travel more than younger children.
-<!-- Can you explain in one sentence how mobility of children is related to prevalence here? OK-->
 
@@
 However for school children, prevalences estimated with the Heckman method were found to be higher in the Mid-East and Southwest regions than those obtained with the other methods, whereas in the East-Central region the estimates with the Heckman method are lower.
-A possible reasons for selection bias in surveys of this type is, for example, that daytime visits might favor measurement in sick school children who stay home, leading to overestimated prevalence results as found in the East-Central region.
-[@thedhsprogram2020] Nevertheless, we were unable to find information that suggests or confirms the direction in which malaria prevalence is driven by selection bias in this Uganda study or in other studies similar to this one.
+A possible reason for selection bias in surveys of this type is, for example, that daytime visits might favor measurement in sick school children who stay home, leading to overestimated prevalence results as found in the East-Central region.
+[@thedhsprogram2020] Nevertheless, we were unable to find information confirming the direction in which malaria prevalence is driven by selection bias in this Uganda study or in other studies similar to this one.
 
@@
 For clustered datasets, only imputation methods under the MAR mechanism had previously been proposed.
-Although imputation methods exist to handle MNAR they have only been designed for individual studies.
-This makes them limited in common IPD situations such as systematic missingness or when the proportion of missingness of a variable is very high in one of the included studies.
-In this context, we proposed a new multiple imputation method to handle continuous and binary MNAR covariates specifically for a clustered dataset, which also allows appropriate borrowing of information between the clusters to obtain more reliable imputation results at the individual cluster level.
-<!-- If you say " a new multiple imputation method was proposed", it is unclear who proposed it. It is fine to say " We proposed a new multiple imputation method" (which is also shorter). thanks :D-->
+Although there are imputation methods  that can handle MNAR they can only handle the case of individual studies.
+This puts limit to ctheir use in IPD situations common in IPD-MA such as when there is systematic missingness or when the proportion of missingness of a variable is very high in one of the included studies.
+To address this gap, we proposed a new multiple imputation method for continuous and binary MNAR covariates, that is specifically designed for clustered datasets.
+Oour method, allows borrowing of information between the clusters to obtain more reliable imputation results at the individual cluster level.
 
-From the results of the simulations we can observe that the imputation method we propose can be valid for the imputation of continuous and binary type missing variables that follow a MNAR mechanism according to the Heckman model and that come from multilevel data such as those used in the IPDMA studies.
+In our simulations we observed that the imputation method we proposed id optimal for the imputation of continuous and binary missing variables that follow a MNAR mechanism according to the Heckman model and that come from multilevel data, such data commonly used in the IPDMA.
 
-Overall, the method produced unbiased estimates with convergence close to 95\% for the coefficient parameters and for the random effects parameters resulted in less biased estimates compared to the other methods evaluated.
+Overall, our new method produced unbiased estimates with convergence close to 95\% for the coefficient parameters.
+It also resulted in less biased estimates for the random effects parameters compared to the other methods evaluated.
 
-Empirically, we showed that the proposed method was robust when there are systematic missigness at the cluster level.
+Empirically, we showed that the proposed method was robust when there is systematic missigness at the cluster level.
 This method, in particular, could provide more robust imputation values compared with individual-level imputation methods, as it not only allows imputation of missing values in clusters with systematic missigness, but can also reduce the values of individual clusters toward the overall study mean.
@@
 The advantage of the proposed method over methods that assume MAR is that it allows the imputation of variables from cluster level data following a MAR or MNAR mechanism according to the Heckman model.
-That is to say that under the specification of a valid exclusion variable the method determines by itself which is the most adjustable  correlation parameter between equations ($\rho$), or in general terms the  missingness mechanism  (MAR or MNAR), in each of the clusters evaluated.
+That is to say that under the specification of a valid exclusion variable the method determines which is the most adjustable  correlation parameter between equations ($\rho$), or in general terms the  missingness mechanism  (MAR or MNAR), in each of the clusters evaluated.
 Our implementation of the imputation method was built according to the specifications of the mice R package, which allows, first of all, to be used both on the outcome and on the covariates.
@@
 A major limitation of our method is that it needs a valid restriction variable, which in some contexts is difficult to establish at the individual study level and can be even more challenging if one tries to find a valid exclusion variable across clusters.
+
+In addition to estimate the marginal estimates, the method only uses the clusters with observable information, i.e.
+that are not systematically missing or have sufficient information to estimate the Heckman model.
+The latter could restrict the evaluation of the Heckman model at the cluster level to a certain number of predictors depending on the sample size of the cluster.
+
 Also, the method can be sensitive to both the sample size and the number of studies included in the database.
-On the one hand, a small sample size at the individual study level can affect not only the precision of estimates but also the convergence of the method since it requires a minimum sample size to estimate all the parameters of the Heckman model which can be at least twice the number of parameters required in an imputation model that assumes MAR.
+On the one hand, a small sample size at the individual study level can affect not only the precision of estimates but also the convergence of the method since sample size needed to estimate the parameters of the Heckman model which can be at least twice the sample size required in an imputation model that assumes MAR.
 On the other hand, a high number of studies that may improve the precision of the estimators may also make the estimation of the marginal parameters more difficult and also considerably increase the processing time of our method.
 
-The data were simulated by attributing a constant correlation across all clusters in order to evaluate the performance against M(N)AR assumptions, but in practice this parameter is variable across clusters which can considerably affect the performance of the method.
-Therefore, in future research the effect of this parameter could be more deeply evaluated.
+In our simulation study, data were generated by assuming a constant correlation across all clusters in order to evaluate the performance against M(N)AR assumptions, in practice, however this parameter can be variable across clusters and this can considerably affect the performance of our method.
+Therefore, in future research the effect of this parameter could be further evaluated.
 One might also consider relaxing this assumption of constant correlation to allow for a random effects distribution for the correlation parameter.
@@
 Similarly, less restrictive Heckman based models can be considered in terms of normality distribution of errors and no specification of exclusion variables such as those proposed by Ogundimu & Collins (2019).[@ogundimu_collins19].
-<!-- Add reference to Ogundimu -->
 
 ## Conclusion
-<!-- Always add a conclusion, also in stats papers. I'm thinking it should mention the following. I am not so sure on especially the last sentence, so feel free to rephrase -->
-We have <!--(well actually you) no actually Thomas :D --> proposed an extension to the Heckman model that can account for MNAR, MAR or MCAR of a continuous or binary variable in clustered data sets.
-Our simulations showed that it has favorable statistical properties, when assumptions were met, and provided that the sample size was sufficiently large.
+
+We have proposed an extension to the Heckman model that can account for MNAR, MAR or MCAR of a continuous or binary variable in clustered data sets.
+Our simulations showed that it can have favorable statistical properties, when its assumptions were met, and provided that the sample size is sufficiently large.
 Regarding deviations from distributional assumptions of the error terms, the coefficient parameters were fairly robust in terms of bias, but the intercept was not.
-
-<!-- Can you add this disclaimer for me at the bottom? -->
-<!-- Also, I added recodid. If Harlan or Paul joins as co-author, don't forget to add the Canadian line as well -->
-<!-- If the editors ask you to make a separate Data Availability section, you can add that below here has well -->
 
@@
 ### Acknowledgements {.unnumbered}
-<!-- I usually put recodid as follows in latex: -->
 \begin{minipage}{.08\textwidth}

### 2022-12-04 1.Paper/Paper.Rmd
Update draft

@@
 - name: Matthias Egger 
-  num: d  
+  num: b  
 - name: Orestis Efthimiou
-  num: d
+  num: b
+- name: Vincent Audigier
+  num: c  
 - name: Valentijn M. T. de Jong
-  num: a,b,e
+  num: a,d,f
 - name: Thomas. P. A. Debray
-  num: a,c,e
+  num: a,e,f
 address:
@@
 - num: b
+  org: University of Bern, Switzerland
+- num: c
+  org: CNAM, Laboratoire CEDRIC-MSDMA, 2 rue Conte, 75003 Paris, France  
+- num: d
   org: Data Analytics and Methods Task Force, European Medicines Agency, Amsterdam, The Netherlands
-- num: c
+- num: e
   org: Smart Data Analysis and Statistics, Utrecht, The Netherlands 
-- num: d
-  org: University of Bern, Switzerland
-- num: e
+- num: f
   org: These authors contributed equally
@@
 the data are missing not at random (MNAR).
-Unfortunately, this situation is rather common, in observational studies, registries and other sources of real-world data.
+Unfortunately, this MNAR situation is rather common, in observational studies, registries and other sources of real-world data.
 While several imputation methods have been proposed for addressing individual studies when data are MNAR, their application and validity in large datasets with clustering remains unclear.
@@
    - \usepackage{float}
+   - \usepackage{enumitem}
    - \usepackage{colortbl}
@@
 Traditional imputation methods are designed to address incomplete data sets where variables are MCAR or MAR.
-Their implementation is justified when there is not systematic difference between units with missing or complete data or when the missingness of variable is strongly related to events or variables measured in the study.
+Their implementation is justified when there is not systematic difference between units with missing and  with complete data or when the missingness of a variable is strongly related to variables measured in the study.
 
@@
 Therefore, the MAR mechanism may become more influential than MNAR when more information is recorded.
-For instance, healthcare professionals may decide which vital measures to take at the time of patient admission.
-If low blood pressure (BP) measurements are more likely to be unrecorded, this variable could be assumed to be MNAR, since its missingness depends on its own value, however BP could be considered MAR as more information explaining its missingness is recorded such as the severity of a patient's admission.
+For example, health professionals can decide which vital signs to measure on admission.
+When it is observed that low blood pressure (BP) measurements are not recorded, it could be assumed that this variable is MNAR, as the missigness of BP depends on its own value; however, BP could be considered as MAR variable if additional information is available to explain its missigness, such as the severity of the patient on admission.
 
@@
 
-Selection bias occurs when the inclusion of a subject or an observation into the sample is influenced by unobserved variables  (e.g., the respondent's level of trust toward healthcare entities may cause them to self-select out of the study or refuse to sign consent for a test), which in turn are correlated with unobserved variables related to an outcome of interest (e.g., the result of a blood test).
+Selection bias occurs when the inclusion of an observation into the sample is influenced by unobserved variables  (e.g., the respondent's level of trust toward healthcare entities may cause them to self-select out of the study or refuse to sign consent for a test), which in turn either influence the outcome of interest (e.g., the result of a blood test), or are related to other unobserved variables that influence the outcome of interest.
 [@vella1998]
@@
    \includegraphics[scale=0.35]{DAG_heckman.png}
-   \caption{DAG Heckman selection model: here the nodes (dotted lines = latent variables, continuous lines = observed variables) describe the relationship between $y^*_{ij}$ the latent response and $r^*_{ij}$ the latent selection variables of a $j$th unit in a $i$th cluster,that are dependable on $ \boldsymbol{x_{ij}^O}$ and $ \boldsymbol{x_{ij}^S}$ set of predictors and are correlated through $ \boldsymbol{u_{ij}}$ unobservable variables.}
+   \caption{DAG Heckman selection model: here the nodes (dotted lines = latent variables, continuous lines = observed variables) describe the relationship between $y^*_{ij}$ the latent response and $r^*_{ij}$ the latent selection variables of a $j$th unit in a $i$th cluster, that are dependable on $ \boldsymbol{x_{ij}^O}$ and $ \boldsymbol{x_{ij}^S}$ sets of predictors and are correlated through $ \boldsymbol{u_{ij}}$ unobservable variables.}
    \label{fig:DAG}
@@
 However, to avoid multicollinearity problems, [@puhani1997] it is recommended to include in $\boldsymbol{X^S_{i}}$ at least one variable that is not included in the outcome model.
-[@angrist2001] This variable is commonly known as an exclusion restriction variable (ERV), and should only be associated with the selection in the sample $r^*_{ij}$ (relevance condition) but not with the actual observation $y^*_{ij}$ (exclusion condition).[@gomes_etal20] The ERV should meet the relevance and exclusion conditions in order to provide independent information about the selection process that facilitates the estimation of the Heckman model.
+[@angrist2001] This variable is commonly known as an exclusion restriction variable (ERV), and should only be associated with the selection in the sample $r^*_{ij}$ (relevance condition) but not with the actual observation $y^*_{ij}$ (exclusion condition).[@gomes_etal20] The ERV meets by definition the relevance and exclusion conditions in order to provide independent information about the selection process and to facilitate the estimation of the Heckman model.
 
@@
 Since most statistical analyses assume that the sample units are independent of each other, more complex hierarchical models dealing with nested data are required.
-This hierarchical complexity must not only be taken into account in the observation process,i.e.
-model that connects outcome and predictors, but also in the missing process,i.e related to the distribution and reasons for missingness, thus requiring imputation models that are congenial to the analysis model, i.e., that make the same assumptions about the data.
+This hierarchical complexity must not only be taken into account in the analysis model, but also when dealing with missing data, thus requiring imputation models that are congenial to the analysis model, i.e.
+that make the same assumptions about the data.
 
@@
 
-Briefly, in a first stage  we estimate $\boldsymbol{\theta_i}$ the cluster specific parameters of the Heckman model.
+Briefly, in a first stage  we estimate $\boldsymbol{\theta_i}=\{\boldsymbol{\beta_i^O},\boldsymbol{\beta_i^S},\sigma_i,\rho_i\}$ the cluster specific parameters of the Heckman model.
 That is, we estimate separately for each of the $N$ clusters,the parameters of the two equations in each study, the outcome model and the selection model.
@@
 
-In a random effects model, $\boldsymbol{\theta_i}$ parameters are assumed to be drawn independently and identically from a latent multivariate normal distribution of parameters with a  population mean $\boldsymbol{\theta_m}$ and a population variance $\boldsymbol{\psi}$.
-[@higgins2009] Thus, we can write $\boldsymbol{\theta_i}=\boldsymbol{\theta_m} + \boldsymbol{b_i}$  using $\boldsymbol{b_i} \sim N(0,\boldsymbol{\psi})$ random effects to allow for between-study heterogeneity in observed data relationships, and between-study heterogeneity in missing patterns.
+In a random effects model, $\boldsymbol{\theta_i}$ parameters are assumed to be drawn independently and identically from a latent multivariate normal distribution of parameters with a  population mean $\boldsymbol{\Theta}$ and a population variance $\boldsymbol{\psi}$.
+[@higgins2009] Thus, we can write $\boldsymbol{\theta_i}=\boldsymbol{\Theta} + \boldsymbol{b_i}$  using $\boldsymbol{b_i} \sim N(0,\boldsymbol{\psi})$ random effects to allow for between-study heterogeneity in observed data relationships, and between-study heterogeneity in missing patterns.
 
@@
 We follow a similar approach proposed by Resche-Rigon and White (2018) [@resche-rigon2018] for multilevel imputation of data.
-Briefly, their method was developed to impute variables from a hierarchical structure (i.e., when there are units grouped within a cluster or group).
+Briefly, their method was developed to impute variables from a hierarchical structure (i.e., when there are units grouped within a cluster).
 This method involves estimating an outcome equation (describing the relationships of the observed data to the missing variable) separately in each cluster, after which the parameter estimates of that equation are pooled using random effects meta-analysis.
@@
 With this method, values can be imputed in very common scenarios in IPD, e.g., sporadic and systematic missing patterns.
-In particular, when the response variable is systematically missing within a group, i.e., when $y_{ij}$ are totally missing within a cluster, the imputation values are drawn from a (generalized) linear model conditional on $\boldsymbol{\theta_m}$ the marginal population parameters, i.e., those estimated after pooling the cluster-specific parameters $\boldsymbol{\theta_i}$.
-On the other hand, when the variable is sporadically missing within the cluster, i.e., there are some observed $y_{ij}$ within the cluster, the imputation model is conditional on the shrunk-cluster parameters, i.e., those coming from the shrinkage of $\boldsymbol{\theta_i}$ towards $\boldsymbol{\theta_m}$.
+In particular, when the response variable is systematically missing within a group, i.e., when $y_{ij}$ are totally missing within a cluster, the imputation values are drawn from a (generalized) linear model conditional on $\boldsymbol{\Theta}$ the marginal population parameters, i.e., those estimated after pooling the cluster-specific parameters $\boldsymbol{\theta_i}$.
+On the other hand, when the variable is sporadically missing within the cluster, i.e., there are some observed $y_{ij}$ within the cluster, the imputation model is conditional on the shrunk-cluster parameters, i.e., those coming from the shrinkage of $\boldsymbol{\theta_i}$ towards $\boldsymbol{\Theta}$.
 
-Our imputation approach differs crucially from the previous approach as here we estimate two correlated equations (instead of a single outcome equation) in each cluster, thus obtaining $\boldsymbol{\theta_i}=\{\boldsymbol{\beta_i^O},\boldsymbol{\beta_i^S},\sigma_i,\rho_i\}$ parameters from both equations which are then pooled into a $\boldsymbol{\theta_m}$ parameter set at the marginal or population level.
+Our imputation approach differs crucially from the previous approach as here we estimate two correlated equations (instead of a single outcome equation) in each cluster, thus obtaining $\boldsymbol{\theta_i}=\{\boldsymbol{\beta_i^O},\boldsymbol{\beta_i^S},\sigma_i,\rho_i\}$ parameters from both equations which are then pooled into a $\boldsymbol{\Theta}$ parameter set at the marginal or population level.
 Our imputation method was created for imputing datasets with a single missing outcome or covariate variable, but it can also be used to impute multiple incomplete variables in a data set.
@@
 
-#### Draw $\boldsymbol{\theta^*_m}$ the marginal parameters
+#### Draw the marginal parameters $\boldsymbol{\Theta}$
 
-From the meta-analysis model, we obtain the marginal estimates $\boldsymbol{\widehat{\theta_m}}$ and the between-cluster variance matrix $\boldsymbol{\widehat{\psi}}$,i.e.
-variance-covariance matrix of the random effects, with their corresponding variance-covariance matrices $\boldsymbol{\widehat{S_{\theta_m}}}$ and $\boldsymbol{\widehat{S_{\psi}}}$, which are used to draw the $\boldsymbol{\theta^*_m}$ and $\boldsymbol{\psi^*}$ parameters, from their posterior distribution as follows: [@resche-rigon2013]
+From the meta-analysis model, we obtain the marginal estimates $\boldsymbol{\widehat{\Theta}}$ and the between-cluster variance matrix $\boldsymbol{\widehat{\psi}}$,i.e.
+variance-covariance matrix of the random effects, with their corresponding variance-covariance matrices $\boldsymbol{\widehat{S_{\Theta}}}$ and $\boldsymbol{\widehat{S_{\psi}}}$, which are used to draw the $\boldsymbol{\Theta^*}$ and $\boldsymbol{\psi^*}$ parameters, from their posterior distribution as follows: [@resche-rigon2013]
 
 \begin{align*}
-\boldsymbol{\theta_m^*}& \sim N(\boldsymbol{\widehat{\theta_m}},\boldsymbol{\widehat{S_{\theta_m}}})\\
+\boldsymbol{\Theta^*}& \sim N(\boldsymbol{\widehat{\Theta}},\boldsymbol{\widehat{S_{\Theta}}})\\
 \boldsymbol{\psi^*}& \sim N(\boldsymbol{\widehat{\psi}},\boldsymbol{\widehat{S_{\psi}}})
@@
 
-We draw the shrunked-cluster-parameters $\boldsymbol{\theta^*_i}$ for each cluster $i$ from the following posterior distribution conditional on $\boldsymbol{\theta^*_m}$ and $\boldsymbol{\psi^*}$.
+We draw the shrunked-cluster-parameters $\boldsymbol{\theta^*_i}$ for each $i$ cluster from the following posterior distribution conditional on $\boldsymbol{\Theta^*}$ and $\boldsymbol{\psi^*}$.
 
 \begin{align*}
-\theta^*_i\sim N\left(\left({\boldsymbol{\psi^*}}^{-1}+{\boldsymbol{\widehat{S_{\theta_i}}}}^{-1}\right)^{-1}\left(\boldsymbol{\psi^*}^{-1}\boldsymbol{\theta^*_m}+\boldsymbol{\widehat{S_{\theta_i}}}^{-1}\boldsymbol{\widehat{\theta_i}} \right),\left({\boldsymbol{\psi^*}}^{-1}+{\boldsymbol{\widehat{S_{\theta_i}}}}^{-1}\right)^{-1} \right)
+\theta^*_i\sim N\left(\left({\boldsymbol{\psi^*}}^{-1}+{\boldsymbol{\widehat{S_{\theta_i}}}}^{-1}\right)^{-1}\left(\boldsymbol{\psi^*}^{-1}\boldsymbol{\Theta^*}+\boldsymbol{\widehat{S_{\theta_i}}}^{-1}\boldsymbol{\widehat{\theta_i}} \right),\left({\boldsymbol{\psi^*}}^{-1}+{\boldsymbol{\widehat{S_{\theta_i}}}}^{-1}\right)^{-1} \right)
 \end{align*}
@@
 As can be seen, the mean and variance of the posterior distribution is a combination between the estimated marginal and cluster-specific parameters.
-Here the weights on the cluster-specific parameters $\boldsymbol{\widehat{\theta_i}}$ and the marginal parameters $\boldsymbol{\theta^*_m}$ are inversely proportional to the within cluster variance $\boldsymbol{\widehat{S_{\theta_i}}}$ and between clusters variance $\boldsymbol{\psi^*}$.
+Here the weights on the cluster-specific parameters $\boldsymbol{\widehat{\theta_i}}$ and the marginal parameters $\boldsymbol{\Theta^*}$ are inversely proportional to the within cluster variance $\boldsymbol{\widehat{S_{\theta_i}}}$ and between clusters variance $\boldsymbol{\psi^*}$.
 For example, when $\boldsymbol{\widehat{S_{\theta_i}}} < \boldsymbol{\psi^*}$ the mean of the conditional distribution gives more weight to the estimated cluster-specific parameter.
@@
 
+Briefly, MICE (multiple imputation of chained equations) was built under the fully conditional specification framework, where for each incomplete variable a conditional imputation model is specified based on another variables in the dataset.
+This process is carried out iteratively, so that in each iteration the missing values of an incomplete variable are drawn from the conditional distribution based on the updated variables in the previous iteration.
+
+Our imputation model can then be used in the imputation of any incomplete variable in a dataset following an MNAR mechanism, even if it is an outcome, predictor or auxiliary variable of the main analysis model.
+Furthermore, as implemented according to the MICE framework, the imputation method could be used simultaneously with other imputation methods available in MICE or in add-in packages such as MICEMD [@audigier2021], to impute datasets with multiple incomplete variables that differ in type and missing mechanism.
+
 # Simulation study
@@
 For simplicity we started from "basic scenario" where the database collected information from $N=10$ clusters of $n_i=1000$  individuals.
-However, we altered both $N$ and $n_i$ in sensitivity analyses (see below).
-
+However, we altered both $N$ and $n_i$ in sensitivity analyses (Table \ref{tab:scenarios}).
+```{r scenarios,  echo = FALSE, warning = FALSE,cache = TRUE}
+data2 <- read.table(stringsAsFactors = FALSE, header = TRUE, sep="/", text =
+'Scenario/	Incomplete variable/	Rho/	N; $n_i$/	Missing process
+Base/	Continuous/	0.6/	10;1000/	Heckman(BVN)
+M(N)AR/Continuous, Binary/	MAR:0, MNAR:0.3,0.6,0.9/	10;1000/Heckman(BVN)
+Size and cluster number/	Continuous/	0.6/	10;50,   10;100,   10;1000,  50;1000,  100;1000/	Heckman (BVN)
+Distribution deviations/	Continuous/	0.6/	10;1000	/Heckman(BVN), Heckman(t-skew), Explicit'
+)
+options(kableExtra.latex.load_packages = FALSE) 
+library(kableExtra)
+kbl(data2, booktabs = T, align="l",
+    caption = "Data generation scenarios",
+    col.names = c("Scenario","Incomplete variable", "$\\rho$", "$N;n_i$","Missing process"),
+    row.names=FALSE,escape = FALSE)%>%
+    column_spec(c(1), width = "2,5cm")%>%
+    column_spec(c(2,4), width = "1,5cm")%>%
+    column_spec(c(3), width = "3cm")%>%
+    column_spec(c(5), width = "2.7cm")%>%
+  kable_styling(latex_options = c("striped"))
+```
 For each dataset, we generated $X_{1i}$ a treatment indicator variable from a Bernoulli distribution with a probability of treatment on each cluster equal to 0.6.
@@
 In addition, we included a random effect on the exclusion restriction variable given by $b_3\sim N(0,0.2)$ assuming that the intracluster variation in the exclusion restriction effect is lower than the variation on other coefficient parameters effects.
-The $\rho$ parameter was given different values depending on the simulated missing mechanism (See below additional scenarios) .
+The $\rho$ parameter was given different values depending on the simulated missing mechanism (Table \ref{tab:scenarios}) .
 
@@
 
-- \textbf{Influence of sample size}: we explored different configurations regarding the number of patients per cluster $n_i=\{50,100,1000\}$  and the number of studies $N=\{10,50,100\}$.
+- \textbf{Influence of sample size and cluster number}: we explored different configurations regarding the number of patients per cluster $n_i=\{50,100,1000\}$  and the number of clusters $N=\{10,50,100\}$.
 
 - \textbf{Violation of distributional assumptions:} Here, we aimed to investigate how the imputation models behave in settings with departures from the bivariate normal distribution, performed two sensitivity analyses.
-
- 1.
-\textbf{Skewed-t:} We drew error terms from a bivariate skewed student-t distribution using the same location parameter and covariance matrix of the normal distributed settings,  with 4 degrees of freedom and an $\alpha=\{-2,6\}$ parameter which regulates the the slant of the density.
- 2.
-\textbf{Explicit:} In addition we simulated an explicit missingness process, where error terms of the selection and outcome equations were independently normal distributed and the selection of observations depended on the value of the outcome variable, as $ry^*_i=0.3y^*_i+\epsilon^S_{i}$.
+ \begin{enumerate}[wide=10pt]
+    \item \textbf{Skewed-t:} We drew error terms from a bivariate skewed student-t distribution using the same location parameter and covariance matrix of the normal distributed settings,  with 4 degrees of freedom and an $\alpha=\{-2,6\}$ parameter which regulates the asymmetry of the density.
+    \item  \textbf{Explicit:} In addition we simulated an explicit missingness process, where error terms of the selection and outcome equations were independently normal distributed and the selection of observations depended on the value of the outcome variable, as $ry^*_i=0.3y^*_i+\epsilon^S_{i}$.
 These settings assure that the percentage of missingness on the outcome variable is around 60\% in all the evaluated scenarios.
- 3. \textbf{Normal:} As a reference, we provide the results from the basic scenario of the main simulation study, where the error terms were Bivariate normal distributed.
+    \item \textbf{Normal:} As a reference, we provide the results from the basic scenario of the main simulation study, where the error terms were Bivariate normal distributed. 
+  \end{enumerate}
 
 ## Estimands
-The estimands were the parameter coefficients of the outcome equation $\beta^O={\beta^O_0,\beta^O_1,\beta^O_2}$, with special emphasis on the treatment effect parameter $\beta^0_1$.
+The estimands were the parameter coefficients of the outcome equation $\beta^O=\{\beta^O_0,\beta^O_1,\beta^O_2\}$, with special emphasis on the treatment effect parameter $\beta^0_1$.
 We also report the estimated variance of the random effects and residual errors $\sigma_{b0}^2$,$\sigma_{b1}^2$,$\sigma_{b2}^2$,$\sigma_{e}^2$.
@@
 
-The Heckman 2L imputation method is available in the mice R package (as \textbf{mice.2l.heckman()}) and also on the github repository \url{https://github.com/johamunoz/Statsmed_Heckman} where you can also find all the codes accompanying this paper and a toy example that explains how to implement the method in mice.
+The Heckman 2L imputation method is available in the MICEMD R package [@audigier2021] (as \textbf{mice.2l.heckman()}) and also on the github repository \url{https://github.com/johamunoz/Statsmed_Heckman} where you can also find all the codes accompanying this paper and a toy example that explains how to implement the method in mice.
 
@@
 We generated data sets of 10 clusters of 1000 patients each in each scenario.
-For example, for the scenario in which the error terms followed a normal distribution, out of the 500 datasets generated, we obtained that on average 60.8\% of the Y response was missing, with the lowest missing percentage being 26.84\% and the maximum being 73.67\%.
-At the cluster level, looking only at the sporadic missing clusters, we found an average of 35.25\% missing values, but there were clusters with no missing data at all and up to 98\% missing data in Y.
-Regarding the binary response scenarios, the proportion of treated  patients is around 60% CI[20%,90%].
-Processing times differ in the imputation methods evaluated, for example, for 10 clusters of 1000 units, the 1l.
+For example, for the base scenario (Table \ref{tab:scenarios}) in which the error terms followed a normal distribution, out of the 500 datasets generated, we obtained that on average 60.8\% of the Y response was missing, with the lowest missing percentage being 26.84\% and the maximum being 73.67\%.
+At the cluster level, looking only at the sporadic missing clusters, we found an average of 35.25\% missing values, but there were clusters with no missing data at all and up to 98\% missing data in the outcome.
+Processing times differ in the imputation methods evaluated, the 1l.
 Heckman takes an average of 10 seconds, the 2l.MAR takes 1.5 and the 2l.
@@
 It should be noted that the time of the last two approaches depends strongly on the number of clusters included and the sample size in the data set.
+(See Annex excel file)
 
@@
 when $\rho=0$, all imputation methods provided similar unbiased estimates of the coefficient parameters, but as rho increased, i.e.
-the mechanism becomed MNAR, the estimates for the complete case analysis and the MAR-IPD imputation method became biased.
+the mechanism became MNAR, the estimates for the complete case analysis and the MAR-IPD imputation method became biased.
 
- ```{r rhocont,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable by varing $\\rho$, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.7, fig.width = 6}
+ ```{r rhocont,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable by varing $\\rho$, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.9, fig.width = 6}
 plot_rho
  ```
-Overall, both Heckman-based imputation (1l.
+As it is expected both Heckman-based imputation (1l.
 and 2l.) models provided less biased estimates of the coefficient parameters in MNAR scenarios, but the random effects estimates on the 1l.heckman were far away from the true values as no cluster information was considered at all.
@@
 
-In particular, our method provided an advantage over the 1l.Heckman method in systematic missingness scenarios, by adding a cluster variable in the imputation model, feature lacking from 1l.Heckman.
+In particular, our method provided an advantage over the 1l.Heckman method in systematic missingness scenarios, by adding a cluster variable in the imputation model, a feature lacking from 1l.Heckman.
 This can be seen in the estimates of the random effects parameters of the 1l.Heckman method, which were more biased than those of the other methods in which cluster information was included, i.e.
@@
 
-Regarding the length of the 95\% CI of the estimated parameters, we observed that using the 2l.Heckman model we obtained larger CIs than when using other evaluated methods, which generally increased as $\rho$ moved away from zero.
+Regarding the width of the 95\% CI of the estimated parameters, we observed that using the 2l.Heckman model we obtained larger CIs than when using other evaluated methods, which generally increased as $\rho$ moved away from zero.
 In particular, we observed in $\beta_2$ that the CI length was larger in the MAR scenario.
@@
 
-In case of a bivariate missing variable, we also found (Figure \ref{fig:rhobin}) that the 2l.Heckman method provided unbiased results on coefficient parameters and random effects.
-However, we observe more variability in the estimates of random effects, reflected on longer confidence intervals on the Bias and RMSE criteria.
+In the case of a missing bivariate variable, we also found (Figure \ref{fig:rhobin}) that the 2l.Heckman method provided the least unbiased results on the coefficient parameters.
+However, we observed more unbiased estimates and a larger 95th percentile amplitude in the estimates of the variance of the random effects, especially in the $\sigma_{b2}$ compared to the estimates obtained in the continuous incomplete outcome scenario.
 
-```{r rhobin,fig.align='center',fig.pos="!htb",fig.cap="Binary incomplete variable by varing $\\rho$, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.7, fig.width = 6}
+```{r rhobin,fig.align='center',fig.pos="!htb",fig.cap="Binary incomplete variable by varing $\\rho$, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.9, fig.width = 6}
 plot_bin
@@
 
-```{r size,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable under systematical missingness by varing: number of clusters (N); sample size per cluster ($n_i$), where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.7, fig.width = 6}
+By increasing N, the number of clusters, from 10 to 100, we observed that the bias was not affected but the width of the 95\% CI of the estimates decreased (larger precision) and hence the RMSE decreased.
+On the other hand, by reducing the number of units per cluster (from $n_i$=1000 to $n_i$=100) the precision decreased for all coefficients, the bias on coefficient estimates were not drastically affected but the bias of random effects did.
+
+When we reduced the sample size to 50 patients per study, the bias and RMSE of the $\sigma_b2$ were drastically affected (not shown here but in the Appendix).
+This could be explained in part due to the scarce information on certain clusters, which affects directly the estimation of the Heckman model on those clusters.
+
+```{r size,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable under systematical missingness by varing: number of clusters (N); sample size per cluster ($n_i$), where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.9, fig.width = 6}
 plot_N
@@
 
-By increasing N, the number of clusters, from 10 to 100, we observed that the bias was not affected but the width of the 95\% CI of the estimates decreased (larger precision) and hence the RMSE decreased.
-On the other hand, by reducing the number of units per cluster (from $n_i$=1000 to $n_i$=100) the precision decreased for all coefficients, the bias on coefficient estimates were not drastically affected but the variances of random effects did.
-
-When we reduced the sample size to 50 patients per study (Appendix), the bias and RMSE of the $\sigma_b2$ were drastically affected compared to the previous $N; n_i$ settings.
-This could be explained in part due to the scarce information on certain clusters, which affects directly the estimation of the Heckman model on those clusters.
-
 ### Sensitivity analysis: distributional assumptions
 
-```{r dist,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable with deviations in distribution assumptions, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.7, fig.width = 6}
+When using the Heckman model in an explicit MNAR process, i.e.
+when the probability of loss is associated with the value of the missing variable, we observe that under all imputation methods we obtain biased estimators, with poor coverage.
+Particularly when estimating the intercept we observed more biased estimates compared to those obtained for the other estimated parameters B2 with coverage below 60%.
+
+Our model may not be fully suitable for this type of scenario, especially if the main analysis is focused on estimating absolute prevalence estimates, as it is greatly affected by the bias of the intercept parameter ($\beta_0$) and the width of the 95th percentile CI of all coefficient parameters.
+Also the imputation method was affected in terms of the bias of the random-effects parameters $\sigma_{b2}$.
+
+With respect to the Skewed-t scenario, only the bias of $\beta_0$ was affected for all imputation methods applied.
+But $\beta_0$ estimates for both the 1l.Heckman and 2l.Heckman were drastically affected in terms of bias, with very poor coverage (below 25%).Therefore, the applicability of our method could be questionable in scenarios that do not conform to the BNV assumption.
+
+```{r dist,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable with deviations in distribution assumptions, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.9, fig.width = 6}
 plot_S
 ```
-
-When we used the Heckman model in an explicit MNAR process, we observed that our model might not be completely suitable for this type of scenario, as it highly affected the bias of the intercept parameter ($\beta_0$) and the width of the 95\% CI of all the coefficient parameters \ref{fig:dist}.
-Also the bias of the random effects parameters ($\sigma_0$ and $\sigma_2$) was affected.
-
-With respect to the Skewed-t scenario, only the bias of $\beta_0$ was affected. For both scenarios, it is also seen that the coverage in the $\beta_0$ was highly affected.
-
-<!-- I don't see $\sigma_0$ and $\sigma_2$ in the figure. Are they called $\sigma_{b0}$ and $\sigma_{b2}$ there? -->
-<!-- Can you add some more description of the results of the sensitivity analyses?
-For instance, that in case of the skewed distribution the other methods are also biased and also have poor coverage?
-For the explicit MNAR process, the other methods are also biased.
--->
 
@@
 
-<!-- Can you clarify this? Do you mean that both the campaign and the survey took place in 2013, 2014 and 2017? In that case i suggest rephrasing -->
-<!-- else split it up into two sentences to make it easier to read (see below). -->
 Specifically, in Uganda, under the Uganda LLIN evaluation project, a LLINS distribution campaign was conducted between 2013 and 2014.
@@
 Blood samples were only taken from 8846 children, as 69 were excluded from the study due to lack of consent and 2222 were not present at the time of the survey.
-Although the original data set consists of 164 variables, here we only consider the variables described in Table \ref{tab:descriptive}, which were used as predictors in the imputation model.
+Although the original data set consists of 164 variables, here we only consider the variables described in Table \ref{tab:descriptive}, which were used as predictors in the imputation model and are fully observed in the dataset.
+In this dataset, the parasitaemia test is an incomplete binary result (1=positive test, 0=negative test), which is missing 21\% across the whole dataset.
+
 ```{r descriptive, echo = FALSE, warning = FALSE,cache = TRUE}
@@
 Under the MCAR assumption, prevalence was calculated on the basis of the recorded tests, i.e., we only included patients with a test result.
-Under the MAR assumption, the test values of children who were not present during the survey were imputed with the 2l.2stage.bin method of the MICEMD package, where the community was taken as the cluster and the following factors previously associated with parasitemia were used as predictors in the imputation model: sex, two-person mosquito net.
+Under the MAR assumption, the test values of children who were not present during the survey were imputed with the 2l.2stage.bin method of the MICEMD package, where the community was taken as the cluster and the following factors previously associated with parasitemia were used as predictors in the imputation model: sex, bednet ( indicator of whether only two or fewer persons share a mosquito bed net).
 In addition, we included age as a power 3 spline function, the cluster-level Log10 mean of the number of female anopheline mosquitoes per household estimated from the entomological survey, and the household wealth index from principal components analysis calculated specifically for the surveyed households.
@@
 
-<!-- What are non-participants? Children with missing values? -->
 According to our imputation approach, non-tested children were estimated to have a higher prevalence of malaria than participants in more than half of the districts analyzed. 
@@
 Although there are imputation methods  that can handle MNAR they can only handle the case of individual studies.
-This puts limit to ctheir use in IPD situations common in IPD-MA such as when there is systematic missingness or when the proportion of missingness of a variable is very high in one of the included studies.
+This puts limit to their use in situations common in IPD-MA such as when there is systematic missingness or when the proportion of missingness of a variable is very high in one of the included studies.
 To address this gap, we proposed a new multiple imputation method for continuous and binary MNAR covariates, that is specifically designed for clustered datasets.
-Oour method, allows borrowing of information between the clusters to obtain more reliable imputation results at the individual cluster level.
+Our method, allows borrowing of information between the clusters to obtain more reliable imputation results at the individual cluster level.
 
@@
 
-Empirically, we showed that the proposed method was robust when there is systematic missigness at the cluster level.
-This method, in particular, could provide more robust imputation values compared with individual-level imputation methods, as it not only allows imputation of missing values in clusters with systematic missigness, but can also reduce the values of individual clusters toward the overall study mean.
+Empirically, we showed that the proposed method could be applicable when there is systematic missigness at the cluster level.
+This method, in particular, could provide less biased imputation values compared with individual-level imputation methods, as it not only allows imputation of missing values in clusters with systematic missigness, but can also shrunk the values of individual clusters toward the overall study mean.
 This is especially advantageous in studies with small sample sizes, where an analysis approach that ignores data from other studies may lead to extreme estimates of prevalence.
@@
 
-<!-- But why is it advantageous for these studies?
-If the prevalence truly is large in these studies, they should not be shrunk.
-It may be advantageous (I guess) because these studies are too small to estimate the prevalence reliably.
-I suggest something like this:
-
-"This is particularly advantageous in studies with small sample sizes, where an analysis approach that ignores the data from other studies may lead to extreme estimates of prevalence.
-When using our proposed method, these would be shrunk to the overall mean.
-Sounds perfect!!
-thanks for ur suggestion."
--->
-
 The advantage of the proposed method over methods that assume MAR is that it allows the imputation of variables from cluster level data following a MAR or MNAR mechanism according to the Heckman model.
 That is to say that under the specification of a valid exclusion variable the method determines which is the most adjustable  correlation parameter between equations ($\rho$), or in general terms the  missingness mechanism  (MAR or MNAR), in each of the clusters evaluated.
-Our implementation of the imputation method was built according to the specifications of the mice R package, which allows, first of all, to be used both on the outcome and on the covariates.
-In addition, it offers the option of being used simultaneously with other imputation methods implemented in this package, which is advantageous in databases containing missing variables with different prediction methods and models.
+
+Our implementation of the imputation method was built according to the specifications of the mice R package and is available in the MICEMD package, which allows, first of all, to be used both on the outcome and on the covariates.
+In addition, it offers the option of being used simultaneously with other imputation methods implemented in the MICE package, which is advantageous in databases containing missing variables with different prediction methods and models.
 Finally, the method can be used on systematically and sporadically missing clusters, both for continuous and binary missing variables with heterogeneous effects and error variances.
@@
 
-In addition to estimate the marginal estimates, the method only uses the clusters with observable information, i.e.
+In addition, to estimate the marginal estimates, the method only uses clusters with observable information, i.e.
 that are not systematically missing or have sufficient information to estimate the Heckman model.
-The latter could restrict the evaluation of the Heckman model at the cluster level to a certain number of predictors depending on the sample size of the cluster.
+The latter might restrict the evaluation of the Heckman model at the cluster level to a certain number of predictors depending on the sample size of the cluster.
+This method also requires at least some of the studies to have no systematic missigness in the incomplete variable in question.
+That is, if all studies have the incomplete variable systematically missing, the method is unusable, which can be an important limitation in practice.
 
 Also, the method can be sensitive to both the sample size and the number of studies included in the database.
-On the one hand, a small sample size at the individual study level can affect not only the precision of estimates but also the convergence of the method since sample size needed to estimate the parameters of the Heckman model which can be at least twice the sample size required in an imputation model that assumes MAR.
+On the one hand, a small sample size at the individual study level can affect not only the precision of estimates but also the convergence of the method since sample size needed to estimate the parameters of the Heckman model which can be at least twice the number of parameters required to estimate in an imputation model that assumes MAR.
 On the other hand, a high number of studies that may improve the precision of the estimators may also make the estimation of the marginal parameters more difficult and also considerably increase the processing time of our method.
@@
 One might also consider relaxing this assumption of constant correlation to allow for a random effects distribution for the correlation parameter.
-
-<!-- Does my suggestion make sense? yes!!! thanks a lot-->
 

### 2022-12-04 3.Ilustrative_study/Ilustrative_study.R

@@ -9,65 +9,57 @@ library(ggplot2)
 
 
 #0. Additional required function ----
-# 0.1 Functions to pool estimates first graph ----
-logit <- function(x) {log(x/(1-x))}
-inv.logit <- function(x) {1/(1+exp(-x))}
-inv.logit.SE<- function(logit.se, c) {logit.se * (c*(1-c))}
+# Prevalence functions 
+logit <- function(x){log(x/(1-x))}
+inv_logit <- function(x){exp(x)/(1+exp(x))}
 
-#Function for calculating absolute risk
+#Function for calculating prevalence CI
 f.prevalence <- function(data,outcome_name) {
   a<-summary(data[,get(outcome_name)])
   b<-prop.test(x = a[[2]], n = sum(a), correct = TRUE)
   prop<-b$estimate[[1]]
   ci<-b$conf
-  return(list(prevalence= prop,ci.l=ci[1], ci.u=ci[2]))
+  se<-(ci[2]-ci[1])/(2*1.96)
+  #Replace 0's by 0.000001
+  prevalence<-ifelse(prop==0,1e-10,ifelse(prop==1,1-1e-10,prop))
+  return(list(prev= prevalence,prev_se=se))
 }
-
 
 #Function to calculate absolute risk of an outcome per study and per imputed dataset
 f.abs.perstudy<-function(data, outcome_name,study_name) {
-  
   #Calculate absolute risk of outcome per study
-  inc.outcome<-setDT(data)[, f.prevalence(data=.SD, outcome_name=outcome_name), by = list(get(study_name),.imp)]
-  
-  #Replace 0's by 0.000001
-  inc.outcome$prevalence[inc.outcome$prevalence==0]<-0.000001
-  inc.outcome$ci.l[inc.outcome$ci.l==0]<-0.000001
+  prev_group<-setDT(data)[, f.prevalence(data=.SD, outcome_name=outcome_name), by = list(get(study_name),.imp)]
   
   #Logit transformation
-  inc.outcome[,logit.prevalence:=logit(prevalence)]
-  inc.outcome[,logit.ci.l:=logit(ci.l)]
-  inc.outcome[,logit.ci.u:=logit(ci.u)]
-  inc.outcome[,logit.se:=(logit.ci.u-logit.ci.l)/(2*1.96)]
-  inc.outcome
-}
-
[…]

### 2022-12-04 4.Codes/Functions_HPC.R

@@ -1,4 +1,4 @@
-rm(list = ls())
+#rm(list = ls())
 library(lme4)
 library(data.table)
 library(mvtnorm)
@@ -304,446 +304,8 @@ par_imp1<-list(dm=dm,pred=pred_list,meth=meth_list)
 
 
 # 2.2. 2l.Heckman imputation method ----
-mice.impute.2l.heckman <-function(y,ry,x,wy = NULL, type, pmm = FALSE, meta_method ="reml",...) {
-  
-  
-  # 1. Define variables and dataset----
-  
-  # Rename covariates
-  colnames(x) <- paste0("x_", 1:length(colnames(x))) #change the covariates name for avoiding conflicts when y is covariate
-  bos_name <- colnames(x)[type ==  1] # names of variables present in both outcome and selection model
-  sel_name <- colnames(x)[type == -3] # names of variables in selection model alone
-  out_name <- colnames(x)[type == -4] # names of variables in outcome model alone
-  
-  # # Define y type
-  if (class(y) == "factor" & nlevels(y) == 2){
-    message("the missing variable is assumed to be binomially distributed")
-    family <- "binomial"
-  }else{
-    message("the missing variable is assumed to be normally distributed")
-    family <- "gaussian"
-  }
-  
-  
-  # Check if group variable is defined
-  if (length(colnames(x)[type == -2]) == 0) {
-    message("No group variable has been provided, the Heckman imputation model will be applied globally to the dataset.")
-    Grp_est <- 0 # Group indicator 0: Heckman on full dataset, 1: Heckman at cluster level
-  } else {
-    group_name <- colnames(x)[type == -2]
-    names.clust <- as.character(unique(x[, group_name]))
-    Grp_est <- 1
-  }
-  
-  #Define position of selection and outcome in coefficient vector
-  order <- c(sel_name, bos_name, out_name)
-  send <- length(sel_name) + length(bos_name) + 1 # Index where selection equation ends in coefficient vector
-  oend <- send + length(bos_name) + length(out_name) + 1 # Index where outcome equation ends in coefficient vector
-  
-  # Define outcome and selection equation
-  out <- as.formula(paste0("y", "~", paste(c(bos_name,out_name), collapse = "+")))
[…]

### 2022-12-04 4.Codes/mice.impute.2l.heckman.R

@@ -139,8 +139,8 @@ mice.impute.2l.heckman <-function(y,ry,x,wy = NULL, type, pmm = FALSE, meta_meth
     }
   }
   
-  
-  # 4. Get theta_k and var(theta_k) from full conditional distribution ----
+
+   # 4. Get theta_k and var(theta_k) from full conditional distribution ----
   if (Grp_est == 1) { # Applies imputation at cluster level
     for (i in names.clust) { #Loop across studies
       

### 2022-12-07 4.Codes/Functions_HPC.R
Add ypmm parameter in the function

@@ -305,7 +305,7 @@ par_imp1<-list(dm=dm,pred=pred_list,meth=meth_list)
 
 # 2.2. 2l.Heckman imputation method ----
 source("/Users/jmunozav/Desktop/Statsmed_Heckman/4.Codes/mice.impute.2l.heckman.R")
-#source("/Users/jmunozav/Desktop/Statsmed_Heckman/4.Codes/mice.impute.3l.heckman.R")
+
 #2.3. Get estimates from imputed datasets ----
 get_estimates <- function( data = data, # data with missing values
                            DataID = DataID, #data ID

### 2022-12-07 4.Codes/mice.impute.2l.heckman.R

@@ -18,6 +18,8 @@
 #' -3: Predictor only in the selection model, -4: Predictor only in the outcome 
 #' model}
 #' @param pmm  predictive mean matching can be applied only for for missing continuous variables: "TRUE","FALSE"
+#' @param ypmm vector of possible y values to perform predictive mean matching, in case ypmm is not provided it
+#' uses the observable variables of y.
 #' @param meta_method meta_analysis estimation method for random effects :
 #' "ml" (maximum likelihood), "reml" (restricted maximum likelihood) or "mm"
 #' method of moments.
@@ -59,7 +61,7 @@
 #' @export
 #'
 
-mice.impute.2l.heckman <-function(y,ry,x,wy = NULL, type, pmm = FALSE, meta_method ="reml",...) {
+mice.impute.2l.heckman <-function(y,ry,x,wy = NULL, type, pmm = FALSE, ypmm=NULL, meta_method ="reml",...) {
   
   
   # 1. Define variables and dataset----
@@ -72,10 +74,10 @@ mice.impute.2l.heckman <-function(y,ry,x,wy = NULL, type, pmm = FALSE, meta_meth
   
   # # Define y type
   if (class(y) == "factor" & nlevels(y) == 2){
-    message("the missing variable is assumed to be binomially distributed")
+    message("the missing variable is treated as binary")
     family <- "binomial"
   }else{
-    message("the missing variable is assumed to be normally distributed")
+    message("the missing variable is treated as normal")
     family <- "gaussian"
   }
   
@@ -159,7 +161,7 @@ mice.impute.2l.heckman <-function(y,ry,x,wy = NULL, type, pmm = FALSE, meta_meth
         y.star <- gen_y_star( Xm= Xm, sel_name = sel_name, bos_name = bos_name,
                               out_name =out_name, beta_s_star = star$beta_s_star,
                               beta_o_star = star$beta_o_star, sigma_star = star$sigma_star,
-                              rho_star = star$rho_star, pmm = pmm, y = y, ry = ry)
+                              rho_star = star$rho_star, pmm = pmm, ypmm=ypmm, y = y, ry = ry)
         
         y[!ry & x[, group_name] == as.numeric(i)] <- y.star
       }
[…]

### 2022-12-07 4.Codes/Toy_example.R
Create Toy_example.R

@@ -1,0 +1,90 @@
+library(data.table)
+library(truncnorm)
+library(nlme)
+library(devtools)
+
+#0. Load imputation method and original dataset ----
+# Source imputation function
+source_url("https://github.com/johamunoz/Statsmed_Heckman/blob/main/4.Codes/mice.impute.2l.heckman.R?raw=TRUE")
+# Dataset obtained from "https://www.kaggle.com/datasets/fabinmndez/obesitydata?select=ObesityDataSet_raw_and_data_sinthetic.csv"
+# and it is described here https://www.sciencedirect.com/science/article/pii/S2352340919306985?via%3Dihub
+
+data<-as.data.table(read.csv('/Users/jmunozav/Desktop/ObesityDataSet_raw_and_data_sinthetic.csv', stringsAsFactors=FALSE, fileEncoding="latin1"))
+
+#1. Modify dataset in order to get a variable weight with MNAR missing mechanism ----
+# Generate countries where survey was taken
+data<-data[,c("Weight","Gender","Age","Height","FAVC")] # Select some of the variables of the original dataset
+data$Cluster<-c(rep(1,500),rep(2,500),rep(3,500),rep(4,500),rep(5,111)) # in the imputation method is necessary cluster is included as numeric
+summary(data)
+#  Generation of the exclusion restriction (ERV): The time variable describes the time it takes for a person to answer the weight question.
+data$Time<-rtruncnorm(n=nrow(data), a = 1, b = 10, mean = 5, sd = 2) 
+data[,Gender:=ifelse(Gender=="Female",1,0)]
+data[,FAVC:=ifelse(FAVC=="yes",1,0)] 
+MAR_model<- lme(Weight~Gender+Age+Height+FAVC,random = ~1|Cluster, data = data) 
+# Generate selection and outcome variable (weight)
+data[,XOBO:=predict(MAR_model)]
+data[,XSBS:=15-0.7*Gender-0.06*Age-2*Time] # Simulated selection model
+
+# Simulate bivariate normal correlated errors
+rho=-0.6 # We assume that non-responders are more likely to have a high weight.
+d <- diag(2)
+d[1,1]<-lme$sigma^2
+d[2,1] <- d[1,2] <- rho*lme$sigma
+e_norm <- mvtnorm::rmvnorm(n =nrow(data), mean = rep(0,2), sigma = d)
+
+# Generate latent outcome and selection variables
[…]

### 2022-12-07 4.Codes/Toy_example.R
Update Toy_example.R

@@ -75,16 +75,14 @@ pred["Weight","Time"]  <- -3 # Exclusion restriction
 pred["Weight",c("Height","FAVC")]  <- -4 # Exclusion restriction
 
 # Imputation of weight variable without pmm 
-
-imp <- mice(data = data, meth = meth, pred = pred, maxit = 5, m = 5, seed = 1)
-
+imp <- mice(data = data, meth = meth, pred = pred, maxit = 1, m = 5, seed = 1)
 densityplot(imp)
 
 
 # Imputation of weight variable with (predictive mean matching) pmm: we can also use the ppm approach, by setting pmm= TRUE and providing a vector of donnors ypmm,
 # for instance in this case the vector or donnors are weight from 25 to 200 kilograms and the imputed values are given in this range of values.
 
-imp_pmm <- mice(data = data, meth = meth, pred = pred, maxit = 5, m = 5, seed = 1,pmm=TRUE,ypmm=seq(25,200,0.1))
+imp_pmm <- mice(data = data, meth = meth, pred = pred, maxit = 1, m = 5, seed = 1,pmm=TRUE,ypmm=seq(25,200,0.1))
 
 densityplot(imp_pmm)
 

### 2022-12-09 1.Paper/Paper.Rmd
Update toy example

@@
 abstract: "Missing data is a common problem in medical research, and is commonly addressed using multiple imputation.
-Although traditional imputation methods allow for valid statistical inference when data are missing at random (MAR), their implementation is problematic when the presence of missingness depends on unobserved information, i.e.
+Although traditional imputation methods allow for valid statistical inference when data are missing at random (MAR), their implementation is problematic when the presence of missingness depends on unobserved variables, i.e.
 the data are missing not at random (MNAR).
 Unfortunately, this MNAR situation is rather common, in observational studies, registries and other sources of real-world data.
-While several imputation methods have been proposed for addressing individual studies when data are MNAR, their application and validity in large datasets with clustering remains unclear.
+While several imputation methods have been proposed for addressing individual studies when data are MNAR, their application and validity in large datasets with multilevel structure remains unclear.
 We therefore explored the consequence of MNAR data in hierarchical data in-depth, and proposed a novel multilevel imputation method for common missing patterns in clustered datasets.
@@
 
-keywords: Heckman model; IPDMA; Missing not at random; Selection models;Multiple imputation;
+keywords: Heckman model; IPDMA; Missing not at random; Selection models; Multiple imputation;
 bibliography: bibfile.bib
@@
 For example, the clinical practice research datalink (CRPD) [@herrett2015] is an electronic health record dataset in the UK, which has been used in a variety of medical research, such as the evaluation of health policy and drug efficacy.
-A recent example of an IPD-MA is the emerging risk factor collaboration, [@theemergingriskfactorscollaboration2007] where data were combined from approximately 1.1 million individuals across 104 observational studies to investigate associations of cardiovascular diseases with several predictors.
+A recent example of an IPD-MA is the emerging risk factor collaboration[@theemergingriskfactorscollaboration2007], where data were combined from approximately 1.1 million individuals across 104 observational studies to investigate associations of cardiovascular diseases with several predictors.
 Individuals in these large datasets tend to be clustered in centres, countries or studies, where they have been subject to similar healthcare processes.
 Moreover, clusters may also differ in participant eligibility criteria, follow-up length, predictor and outcome definitions, or in the quality of applied measurement methods.
-Hence, differences (i.e.,heterogeneity) between clusters with respect to baseline covariates and outcomes, while the structure of the correlations between these variables is likely to be different across different clusters.
+Hence, heterogeneity between clusters with respect to baseline covariates and outcomes, while the structure of the correlations between these variables is likely to be different across different clusters.
 
@@
 For example, in registry data it is common that test results are not available for all patients, as the decision to test may be at the discretion of the primary care physician or because the patient refuses to undergo testing.
-It is also possible that variables are systematically missing across clusters.
-For instance, in an IPD-MA, studies may have collected information on different variables.
+It is also possible that variables are systematically missing across clusters.[@resche-rigon2013] For instance, in an IPD-MA, studies may have collected information on different variables.
 Missing values may thus appear for all participants of a study in the combined dataset.
@@
 
-To address the presence of missing data, it is important to consider  the proportion of missingness and the missing mechanism for each incomplete variable.
-Rubin (1976) [@rubin76] identified three missing mechanisms where the probability of missingness: 1) is constant (missing completely  at random; MCAR), 2) depends on observed data only (missing at random; MAR), or 3) depends on unobserved information even after conditioning on all observable variables (missing not at random; MNAR).
+To address the presence of missing data, it is important to consider the missing mechanism for each incomplete variable.
+Rubin (1976) [@rubin76] identified three missing mechanisms where the probability of missingness: 1) is independent from observed or missing values (missing completely  at random; MCAR), 2) depends on observed data only (missing at random; MAR), or 3) depends on unobserved information even after conditioning on all observable variables (missing not at random; MNAR).
 Traditional imputation methods are designed to address incomplete data sets where variables are MCAR or MAR.
@@
 
-Although it is possible to formally rule out MCAR, [@little1988]  it is impossible to assess whether data are MAR or MNAR as the observable data are not enough to test the assumptions of both mechanisms.[@enders2022]  For this reason, researchers often conveniently assume that data are MAR, or present results of a complete case analysis.
+<!-- Although it is possible to formally rule out MCAR[@little1988], it is impossible to assess whether data are MAR or MNAR as the observable data are not enough to test the assumptions of both mechanisms.[@enders2022]  For this reason, researchers often conveniently assume that data are MAR, or present results of a complete case analysis.
 In practice, however, unless missingness is artificially introduced by study design, e.g.
 when a test is only taken on patients with certain characteristics, missingness will often (at least partially) depend on unobserved information.
+-->
 
-The missingness mechanism of a variable may be a mixture of MAR and MNAR, depending on the availability of data that may predict the missingness patterns.
+<!-- The missingness mechanism of a variable may be a mixture of MAR and MNAR, depending on the availability of data that may predict the missingness patterns.
 Therefore, the MAR mechanism may become more influential than MNAR when more information is recorded.
@@
 When it is observed that low blood pressure (BP) measurements are not recorded, it could be assumed that this variable is MNAR, as the missigness of BP depends on its own value; however, BP could be considered as MAR variable if additional information is available to explain its missigness, such as the severity of the patient on admission.
+-->
 
@@
 
-The MNAR mechanism is considered as non-ignorable, in this case it is necessary to specify information about the missingness process in addition to assumptions about the observed data.
-A popular approach to address MNAR is selection models such as the one proposed by Heckman (1976).
+Modelling data under MNAR mechanism implies to specify information about the missingness process in addition to assumptions about the observed data.
+A popular approach to address MNAR mechanism is selection models such as the one proposed by Heckman (1976).
 [@heckman1976] Briefly, the Heckman selection model corrects for selection bias by estimating two linked equations: a outcome equation, where the missing variable is associated with predictors, and a selection equation, that accounts for the inclusion of observations in the sample.
@@
 In section \ref{sec:methods} we define the main steps of the proposed imputation method.
-In section \ref{sec:simulation} we provide the settings and results of a simulation study to evaluate the performance of the proposed imputation method.
+In section \ref{sec:simulation} we provide the settings and results of a simulation study to evaluate the performance of our imputation method.
 In section \ref{sec:illustration} we illustrate the method using the survey information collected in different sub-districts in Uganda to estimate the prevalence of malaria in children.
-Finally in section \ref{sec:discussion} we summarize the results, outline limitations and propose future extension of our method.
+Finally, in section \ref{sec:discussion} we summarize the results, outline limitations and propose future extension of our method.
 
@@
 Generally the same variables can be used on the matrix of predictor variables $\boldsymbol{X^O_{i}}$ and $\boldsymbol{X^S_{i}}$.
-However, to avoid multicollinearity problems, [@puhani1997] it is recommended to include in $\boldsymbol{X^S_{i}}$ at least one variable that is not included in the outcome model.
+However, to avoid multicollinearity problems[@puhani1997], it is recommended to include in $\boldsymbol{X^S_{i}}$ at least one variable that is not included in the outcome model.
 [@angrist2001] This variable is commonly known as an exclusion restriction variable (ERV), and should only be associated with the selection in the sample $r^*_{ij}$ (relevance condition) but not with the actual observation $y^*_{ij}$ (exclusion condition).[@gomes_etal20] The ERV meets by definition the relevance and exclusion conditions in order to provide independent information about the selection process and to facilitate the estimation of the Heckman model.
 
-In the presence of selection bias,the aforementioned outcome equation will yield biased estimates of $\boldsymbol{\beta^O_{i}}$ if no efforts are made to adjust for the non-representativeness of the observed $\boldsymbol{X^O_{i}}$ and $\boldsymbol{y_{i}}$ values.
+In the presence of selection bias, the aforementioned outcome equation will yield biased estimates of $\boldsymbol{\beta^O_{i}}$ if no efforts are made to adjust for the non-representativeness of the observed $\boldsymbol{X^O_{i}}$ and $\boldsymbol{y_{i}}$ values.
 For this reason, the Heckman model aims to jointly estimate the outcome and selection equation by defining a relation between their respective error distributions.
@@
 In our imputation method we estimate the Heckman model using the copula method, as in real life data generating mechanism may deviate a BVN distribution.
-This facilitates the Heckman model estimation and makes it more robust to deviations of distributional assumptions.
+<!-- This facilitates the Heckman model estimation and makes it more robust to deviations of distributional assumptions.
+-->
 
@@
  The Heckman model can be extended to hierarchical settings, i.e., in cases when individuals or sampling units are nested within clusters, as is the case in EHR or IPD.
-In this case, sample units from the same cluster are expected to share some characteristics (e.g., the distribution of variables and correlations between them, relationships between exposure and outcomes, missing processes).
+<!--In this case, sample units from the same cluster are expected to share some characteristics (e.g., the distribution of variables and correlations between them, relationships between exposure and outcomes, missing processes).
 
 Since most statistical analyses assume that the sample units are independent of each other, more complex hierarchical models dealing with nested data are required.
-This hierarchical complexity must not only be taken into account in the analysis model, but also when dealing with missing data, thus requiring imputation models that are congenial to the analysis model, i.e.
+--> This hierarchical complexity must not only be taken into account in the analysis model, but also when dealing with missing data, thus requiring imputation models that are congenial to the analysis model, i.e.
 that make the same assumptions about the data.
@@
 
-Briefly, in a first stage  we estimate $\boldsymbol{\theta_i}=\{\boldsymbol{\beta_i^O},\boldsymbol{\beta_i^S},\sigma_i,\rho_i\}$ the cluster specific parameters of the Heckman model.
-That is, we estimate separately for each of the $N$ clusters,the parameters of the two equations in each study, the outcome model and the selection model.
+Briefly, in a first stage  we estimate $\boldsymbol{\theta_i}=\{\boldsymbol{\beta_i^O},\boldsymbol{\beta_i^S},\sigma_i,\rho_i\}$, for all $i$ between 1 and $N$ clusters, the cluster specific parameters of the Heckman model.
+That is, we estimate separately for each of the clusters, the parameters of the two equations in each study, the outcome model (\ref{eq:1}) and the selection model (\ref{eq:2}).
 In the second stage, we combine all $\boldsymbol{\theta_i}$  by using a random effects multivariate meta-analysis model.
@@
 Initially, we use the copula method to estimate the set of cluster-specific parameters, $\boldsymbol{\widehat{\theta_i}}=\boldsymbol{\{\widehat{\beta^O_i}}, \boldsymbol{\widehat{\beta^S_i}}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, using all $j$ units with observable measurements $y_{ij}^{obs}$ within each cluster $i$.
-The Heckman model is estimated with the \textbf{gjrm} function of the GJRM R package, [@radiceGJRMGeneralisedJoint2021] under the bivariate model with the nonrandom sample selection (BSS) specification, from which we obtain not only the parameters' point estimates $\boldsymbol{\widehat{\theta_i}}$, but also their corresponding $\boldsymbol{\widehat{S(\theta_i)}}$ within-cluster variance-covariance matrix.
+The Heckman model is estimated with the \textbf{gjrm} function of the GJRM R package[@radiceGJRMGeneralisedJoint2021],  under the bivariate model with the nonrandom sample selection (BSS) specification, from which we obtain not only the parameters' point estimates $\boldsymbol{\widehat{\theta_i}}$, but also their corresponding $\boldsymbol{\widehat{S(\theta_i)}}$ within-cluster variance-covariance matrix.
 
@@
 Our imputation model can then be used in the imputation of any incomplete variable in a dataset following an MNAR mechanism, even if it is an outcome, predictor or auxiliary variable of the main analysis model.
-Furthermore, as implemented according to the MICE framework, the imputation method could be used simultaneously with other imputation methods available in MICE or in add-in packages such as MICEMD [@audigier2021], to impute datasets with multiple incomplete variables that differ in type and missing mechanism.
+Furthermore, as implemented according to the MICE framework, the imputation method could be used simultaneously with other imputation methods available in MICE or in add-in packages such as micemd [@audigier2021], to impute datasets with multiple incomplete variables that differ in type and missing mechanism.
 
@@
 We designed a simulation study aimed to compare the performance of alternative methods for imputing a single missing outcome variable in a hierarchical dataset, where the missingness follows a MNAR mechanism.
-In our scenarios we considered systematically missingness,i.e.
-the case some of the studies did not report hte variable at all.
+In our scenarios we considered systematically missingness.
 
@@
 'Scenario/	Incomplete variable/	Rho/	N; $n_i$/	Missing process
-Base/	Continuous/	0.6/	10;1000/	Heckman(BVN)
-M(N)AR/Continuous, Binary/	MAR:0, MNAR:0.3,0.6,0.9/	10;1000/Heckman(BVN)
+Base/	Continuous/	0.6/	10;1000/	Heckman (BVN)
+M(N)AR/Continuous, Binary/ 0 (MAR), 0.3,0.6,0.9 (MNAR)/	10;1000/Heckman (BVN)
 Size and cluster number/	Continuous/	0.6/	10;50,   10;100,   10;1000,  50;1000,  100;1000/	Heckman (BVN)
-Distribution deviations/	Continuous/	0.6/	10;1000	/Heckman(BVN), Heckman(t-skew), Explicit'
+Distribution deviations/	Continuous/	0.6/	10;1000	/Heckman (BVN), Heckman (t-skew), Explicit'
 )
@@
     column_spec(c(2,4), width = "1,5cm")%>%
-    column_spec(c(3), width = "3cm")%>%
+    column_spec(c(3), width = "2,9cm")%>%
     column_spec(c(5), width = "2.7cm")%>%
@@
 
-## Analysis Methods
+## Imputation methods
 
@@
 - \textbf{1l.Heckman:} Multiple imputation based on the Heckman model without no study specification, following the imputation method proposed by Galimard et al.(2016). [@galimard_etal16]
-- \textbf{2l.MAR}: Multiple imputation assuming MAR for hierarchical datasets, we used the multilevel imputation model ( 2l.2stage.norm and 2l.2stage.bin) from the micemd R package, [@audigier2021] which are described in Audigier et al.
+- \textbf{2l.MAR}: Multiple imputation assuming MAR for hierarchical datasets, we used the multilevel imputation model ( 2l.2stage.norm and 2l.2stage.bin) from the micemd R package[@audigier2021], which are described in Audigier et al.
 (2018) paper.
@@
 In case of missing binary variable, we used the same matrix of predictors but on a binary model estimated with the *glmer()* function from the *lme4* R package.
-Then, we pooled the estimates of the $\beta_i^O$ and the variance of the random effect and residual errors of the multiple imputed datasets according to Rubin's rule, [@rubin1987] over which we calculated the performance measures on the estimands.
+Then, we pooled the estimates of the $\beta_i^O$ and the variance of the random effect and residual errors of the multiple imputed datasets according to Rubin's rule[@rubin1987], over which we calculated the performance measures on the estimands.
 
-To calculate the coverage of the parameter coefficients' \95% confidence intervals (CI), we estimate CI with the Wald method.
-Although it is possible to obtain CI through the profile or bootstrap method for the variances of the random effects, we prefer not to estimate them (and hence the coverage) for the random effects parameters due to computational time.
+To calculate the coverage of the parameter coefficients' 95\% confidence intervals (CI), we estimate CI with the Wald method.
+<!-- Although it is possible to obtain CI through the profile or bootstrap method for the variances of the random effects, we prefer not to estimate them (and hence the coverage) for the random effects parameters due to computational time.
+-->
 
@@
 
-We calculated the following measures, usually employed to evaluate imputation methods, [@buuren18c] according to the formulas provided in Morris et al.(2019): [@morris2019]
+We calculated the following measures, usually employed to evaluate imputation methods[@buuren18c], according to the formulas provided in Morris et al.(2019)[@morris2019]:
 
@@
 
-The Heckman 2L imputation method is available in the MICEMD R package [@audigier2021] (as \textbf{mice.2l.heckman()}) and also on the github repository \url{https://github.com/johamunoz/Statsmed_Heckman} where you can also find all the codes accompanying this paper and a toy example that explains how to implement the method in mice.
+The Heckman 2L imputation method is available in the micemd R package [@audigier2021] (as \textbf{mice.2l.heckman()}) and also on the github repository \url{https://github.com/johamunoz/Statsmed_Heckman} where you can also find all the codes accompanying this paper and a toy example that explains how to implement the method in mice.
 
@@
 
-### Descriptive statistics of the simulated datasets
-We generated data sets of 10 clusters of 1000 patients each in each scenario.
+<!-- ### Descriptive statistics of the simulated datasets -->
+<!-- We generated data sets of 10 clusters of 1000 patients each in each scenario.
 For example, for the base scenario (Table \ref{tab:scenarios}) in which the error terms followed a normal distribution, out of the 500 datasets generated, we obtained that on average 60.8\% of the Y response was missing, with the lowest missing percentage being 26.84\% and the maximum being 73.67\%.
@@
 It should be noted that the time of the last two approaches depends strongly on the number of clusters included and the sample size in the data set.
-(See Annex excel file)
+(See Annex excel file) -->
 
@@
 
-For this example, we used data accessed directly from CliniEpiDB, [@staedke] where data were collected from 5195 households with verified consent, inhabited by 11137 residents aged 2-10 years.
+For this example, we used data accessed directly from ClinEpiDB[@staedke], where data were collected from 5195 households with verified consent, inhabited by 11137 residents aged 2-10 years.
 Blood samples were only taken from 8846 children, as 69 were excluded from the study due to lack of consent and 2222 were not present at the time of the survey.
@@
 Under the MCAR assumption, prevalence was calculated on the basis of the recorded tests, i.e., we only included patients with a test result.
-Under the MAR assumption, the test values of children who were not present during the survey were imputed with the 2l.2stage.bin method of the MICEMD package, where the community was taken as the cluster and the following factors previously associated with parasitemia were used as predictors in the imputation model: sex, bednet ( indicator of whether only two or fewer persons share a mosquito bed net).
+Under the MAR assumption, the test values of children who were not present during the survey were imputed with the 2l.2stage.bin method of the micemd package, where the community was taken as the cluster and the following factors previously associated with parasitemia were used as predictors in the imputation model: sex, bednet ( indicator of whether only two or fewer persons share a mosquito bed net).
 In addition, we included age as a power 3 spline function, the cluster-level Log10 mean of the number of female anopheline mosquitoes per household estimated from the entomological survey, and the household wealth index from principal components analysis calculated specifically for the surveyed households.
@@
 
-In our simulations we observed that the imputation method we proposed id optimal for the imputation of continuous and binary missing variables that follow a MNAR mechanism according to the Heckman model and that come from multilevel data, such data commonly used in the IPDMA.
+In our simulations we observed that the imputation method we proposed is optimal for the imputation of continuous and binary missing variables that follow a MNAR mechanism according to the Heckman model and that come from multilevel data, such data commonly used in the IPDMA.
 
@@
 
-Our implementation of the imputation method was built according to the specifications of the mice R package and is available in the MICEMD package, which allows, first of all, to be used both on the outcome and on the covariates.
+Our implementation of the imputation method was built according to the specifications of the mice R package and is available in the micemd package, which allows, first of all, to be used both on the outcome and on the covariates.
 In addition, it offers the option of being used simultaneously with other imputation methods implemented in the MICE package, which is advantageous in databases containing missing variables with different prediction methods and models.
@@
 The latter might restrict the evaluation of the Heckman model at the cluster level to a certain number of predictors depending on the sample size of the cluster.
-This method also requires at least some of the studies to have no systematic missigness in the incomplete variable in question.
+<!-- This method also requires at least some of the studies to have no systematic missigness in the incomplete variable in question.
 That is, if all studies have the incomplete variable systematically missing, the method is unusable, which can be an important limitation in practice.
+-->
 
@@
 
-In our simulation study, data were generated by assuming a constant correlation across all clusters in order to evaluate the performance against M(N)AR assumptions, in practice, however this parameter can be variable across clusters and this can considerably affect the performance of our method.
+In our simulation study, data were generated by assuming a constant correlation across all clusters in order to evaluate the performance against M(N)AR assumptions.
+In practice, however this parameter can be variable across clusters and this can considerably affect the performance of our method.
 Therefore, in future research the effect of this parameter could be further evaluated.

### 2022-12-11 1.Paper/Paper.Rmd
Updated toy example and mice.impute.2l.2stage

@@
 
-The Heckman 2L imputation method is available in the micemd R package [@audigier2021] (as \textbf{mice.2l.heckman()}) and also on the github repository \url{https://github.com/johamunoz/Statsmed_Heckman} where you can also find all the codes accompanying this paper and a toy example that explains how to implement the method in mice.
+The Heckman 2L imputation method is available in the micemd R package [@audigier2021] (as \textbf{mice.2l.2stage.heckman()}) and also on the github repository \url{https://github.com/johamunoz/Statsmed_Heckman} where you can also find all the codes accompanying this paper and a toy example that explains how to implement the method in mice.
 

### 2022-12-11 4.Codes/Toy_example.R

@@ -1,88 +1,205 @@
+rm(list=ls())
 library(data.table)
 library(truncnorm)
 library(nlme)
 library(devtools)
 
-#0. Load imputation method and original dataset ----
-# Source imputation function
-source_url("https://github.com/johamunoz/Statsmed_Heckman/blob/main/4.Codes/mice.impute.2l.heckman.R?raw=TRUE")
-# Dataset obtained from "https://www.kaggle.com/datasets/fabinmndez/obesitydata?select=ObesityDataSet_raw_and_data_sinthetic.csv"
-# and it is described here https://www.sciencedirect.com/science/article/pii/S2352340919306985?via%3Dihub
+#0. Load original dataset ----
 
-data<-as.data.table(read.csv('/Users/jmunozav/Desktop/ObesityDataSet_raw_and_data_sinthetic.csv', stringsAsFactors=FALSE, fileEncoding="latin1"))
+data <- as.data.table(read.csv('/Users/jmunozav/Desktop/Statsmed_Heckman/4.Codes/ObesityDataSet_raw_and_data_sinthetic.csv', stringsAsFactors=FALSE, fileEncoding="latin1"))
 
 #1. Modify dataset in order to get a variable weight with MNAR missing mechanism ----
 # Generate countries where survey was taken
-data<-data[,c("Weight","Gender","Age","Height","FAVC")] # Select some of the variables of the original dataset
-data$Cluster<-c(rep(1,500),rep(2,500),rep(3,500),rep(4,500),rep(5,111)) # in the imputation method is necessary cluster is included as numeric
-summary(data)
+data <- data[,c("Weight","Gender","Age","Height","FAVC")] # Select some of the variables of the original dataset
+data[,Cluster:=c(rep(1,500),rep(2,400),rep(3,450),rep(4,400),rep(5,361))] # in the imputation method is necessary cluster is included as numeric
+N<-length(unique(data$Cluster))
 #  Generation of the exclusion restriction (ERV): The time variable describes the time it takes for a person to answer the weight question.
-data$Time<-rtruncnorm(n=nrow(data), a = 1, b = 10, mean = 5, sd = 2) 
-data[,Gender:=ifelse(Gender=="Female",1,0)]
-data[,FAVC:=ifelse(FAVC=="yes",1,0)] 
[…]

### 2022-12-11 4.Codes/mice.impute.2l.2stage.heckman.R

@@ -3,7 +3,7 @@
 #' Imputes outcome and predictor variables that follow an MNAR mechanism
 #' according to Heckman's model and come from a multilevel database such as
 #' individual participant data.
-#' @aliases mice.impute.2l.heckman 2l.heckman
+#' @aliases mice.impute.2l.2stage.heckman 2l.2stage.heckman
 #' @param y Vector to be imputed
 #' @param ry Logical vector of length \code{length(y)} indicating the
 #' the subset \code{y[ry]} of elements in \code{y} to which the imputation
@@ -17,14 +17,15 @@
 #' 1: Predictor in both the outcome and selection,-2: Cluster id (study id),
 #' -3: Predictor only in the selection model, -4: Predictor only in the outcome 
 #' model}
-#' @param pmm  predictive mean matching can be applied only for for missing continuous variables: "TRUE","FALSE"
-#' @param ypmm vector of possible y values to perform predictive mean matching, in case ypmm is not provided it
-#' uses the observable variables of y.
+#' @param pmm  predictive mean matching can be applied only for for missing 
+#' continuous variables: "FALSE","TRUE"
+#' @param ypmm vector of donor values of y to perform the predictive mean matching, 
+#' in case ypmm is not provided, the observable values of y are used.
 #' @param meta_method meta_analysis estimation method for random effects :
 #' "ml" (maximum likelihood), "reml" (restricted maximum likelihood) or "mm"
 #' method of moments.
 #' @param ... Other named arguments. Not used.
-#' @name mice.impute.2l.heckman
+#' @name mice.impute.2l.2stage.heckman
 #' @return Vector with imputed data, of type binary or continuous
 #' @details Imputes systematically and sporadically missing binary and continuous
 #'  univariate variables that follow a MNAR mechanism according to the Heckman 
@@ -43,25 +44,10 @@
 #' @author Julius Center Methods Group UMC, 2022
 #' @family univariate imputation functions
 #' @keywords datagen
-#' @examples
-#' \dontrun{
-#' library(GJRM)
-#' data(hiv) # load and subset dataset
[…]

### 2022-12-11 4.Codes/Toy_example.R
Update micemd functions

@@ -23,6 +23,7 @@ data[,Ages:=scale(Age)]
 data[,Times:=scale(Time)]
 
 MAR_model <- with(data, lme4::lmer(Weight~Gender+Ages+Heights+FAVC+(1+Heights+Gender+Ages|Cluster)))
+summary(MAR_model)
 # Generate outcome variable
 data[,XOBO:=predict(MAR_model)+15]
 
@@ -135,7 +136,7 @@ geom_histogram(aes(color = as.factor(Cluster),fill= as.factor(Cluster)),
 #3. Imputation model
 # Imputation considering weight is MAR
 
-seed=12345
+seed=123
 
 ini <- mice(dataObs, maxit = 0)
 meth <- ini$method
@@ -168,12 +169,13 @@ meth["Weight"] <- "2l.2stage.heckman"
 #belonging to the selection and outcome as "1", the exclusion restrictions or predictor variables that are only included in the selection equation 
 #as "-3" and those that are only included in the outcome equation as "-4".
 
-pred["Weight","Time"]  <- -3 # Exclusion restriction
+pred["Weight","Time"]  <- -3 # Variable only affects the selection model (Exclusion restriction)
 pred["Weight",c("Height","FAVC")]  <- -4 # Variables only affect the outcome model
 # As Age and Gender are predictors in both selection and outcome model are set as 1
 head(dataObs)
-# Imputation of weight variable without pmm 
 
+
+# Imputation of weight variable without pmm 
 imp <- mice(data = dataObs, meth = meth, pred = pred, seed = seed)
 plot(imp)
 ggmice(imp, ggplot2::aes(x = Weight, group = .imp)) +
@@ -184,13 +186,12 @@ ggmice(imp, ggplot2::aes(x = Weight, group = .imp)) +
   ggplot2::theme(legend.position = "none")+
   facet_wrap(~Cluster)
 
-(summary(complete(imp,"long")$Weight))
 
 # Imputation of weight variable with (predictive mean matching) pmm: we can also use the ppm approach, by setting pmm= TRUE and providing a vector of donnors ypmm,
 # for instance in this case the vector or donnors are weight from 35 to 180 kilograms and the imputed values are given in this range of values.
 
[…]

### 2022-12-11 4.Codes/gen_dataObs.R

@@ -1,0 +1,87 @@
+# Generate data toy example
+rm(list=ls())
+library(data.table)
+library(truncnorm)
+library(nlme)
+library(devtools)
+library(mice)
+
+#0. Load original dataset ----
+
+data <- as.data.table(read.csv('/Users/jmunozav/Desktop/Statsmed_Heckman/4.Codes/ObesityDataSet_raw_and_data_sinthetic.csv', stringsAsFactors=FALSE, fileEncoding="latin1"))
+
+#1. Modify dataset in order to get a variable weight with MNAR missing mechanism ----
+# Generate countries where survey was taken
+data <- data[,c("Weight","Gender","Age","Height","FAVC")] # Select some of the variables of the original dataset
+data[,Cluster:=c(rep(1,500),rep(2,400),rep(3,450),rep(4,400),rep(5,361))] # in the imputation method is necessary cluster is included as numeric
+N<-length(unique(data$Cluster))
+#  Generation of the exclusion restriction (ERV): The time variable describes the time it takes for a person to answer the weight question.
+set.seed(12345)
+data[,Time:=rtruncnorm(n=nrow(data), a = 1, b = 10, mean = 5, sd = 3)] 
+data[,Gender:=(ifelse(Gender=="Female",1,0))]
+data[,FAVC:=(ifelse(FAVC=="yes",1,0))]
+data[,Heights:=scale(Height)]
+data[,Ages:=scale(Age)]
+data[,Times:=scale(Time)]
+
+MAR_model <- with(data, lme4::lmer(Weight~Gender+Ages+Heights+FAVC+(1+Heights+Gender+Ages|Cluster)))
+summary(MAR_model)
+# Generate outcome variable
+data[,XOBO:=predict(MAR_model)+15]
+
+# Generate random slope for Gender Age on selection equation
+alpha_gender <- rnorm(n = N, mean=0, sd = 0.01)
+alpha_age <- rnorm(n = N, mean=0, sd = 0.01)
+nobs<-c(500,400,450,400,361)
+data[,Genderi:=unlist(mapply(rep,alpha_gender,each=nobs))]
+data[,Agei:=unlist(mapply(rep,alpha_age,each=nobs))]
+data[,XSBS:=1.5-(0.7+Genderi)*Gender-(0.5+Agei)*Ages-1.2*Times] # Simulated selection model
+
[…]

### 2022-12-11 5.micemd/mice.impute.2l.2stage.heckman.R

@@ -1,0 +1,117 @@
+mice.impute.2l.2stage.heckman <-function(y,ry,x,wy = NULL, type, pmm = FALSE, ypmm=NULL, meta_method ="reml",...) {
+  
+  # 1. Define variables and dataset----
+  
+  # Rename covariates
+  colnames(x) <- paste0("x_", 1:length(colnames(x))) #change the covariates name for avoiding conflicts when y is covariate
+  bos_name <- colnames(x)[type ==  1] # names of variables present in both outcome and selection model
+  sel_name <- colnames(x)[type == -3] # names of variables in selection model alone
+  out_name <- colnames(x)[type == -4] # names of variables in outcome model alone
+  
+  # # Define y type
+  if (class(y) == "factor" & nlevels(y) == 2){
+    family <- "binomial"
+  }else{
+    family <- "gaussian"
+  }
+  
+  # Check if group variable is defined
+  if (length(colnames(x)[type == -2]) == 0) {
+    message("No group variable has been provided, the Heckman imputation model will be applied globally to the dataset.")
+    Grp_est <- 0 # Group indicator 0: Heckman on full dataset, 1: Heckman at cluster level
+  } else {
+    group_name <- colnames(x)[type == -2]
+    names.clust <- as.character(unique(x[, group_name]))
+    Grp_est <- 1
+  }
+  
+  #Define position of selection and outcome in coefficient vector
+  order <- c(sel_name, bos_name, out_name)
+  send <- length(sel_name) + length(bos_name) + 1 # Index where selection equation ends in coefficient vector
+  oend <- send + length(bos_name) + length(out_name) + 1 # Index where outcome equation ends in coefficient vector
+  
+  # Define outcome and selection equation
+  out <- as.formula(paste0("y", "~", paste(c(bos_name,out_name), collapse = "+")))
+  sel <- as.formula(paste0("ry", "~", paste(c(sel_name, bos_name), collapse = "+")))
+  
+  # Define data & prediction matrix
+  data <- data.frame(ry, y, x[, order])
+  X <- data.frame(cbind(Int=rep(1, nrow(x)), x[, order]))
+  
+  # 2. Step 1: Get the theta estimators for each study ----
+  Syst_nest <- Heck_est <- 1
+  if (Grp_est == 1) { #Heckman at cluster level
+    df_list <- split(data, x[, group_name]) 
+    res_total <- suppressWarnings( lapply( df_list, copulaIPD,
[…]

### 2022-12-11 4.Codes/Toy_example.R
Toy example update

@@ -3,6 +3,7 @@ library(data.table)
 library(truncnorm)
 library(nlme)
 library(devtools)
+library(mice)
 
 #0. Load original dataset ----
 

### 2022-12-16 1.Paper/Paper.Rmd
update manuscript

@@
 Modelling data under MNAR mechanism implies to specify information about the missingness process in addition to assumptions about the observed data.
-A popular approach to address MNAR mechanism is selection models such as the one proposed by Heckman (1976).
+Two major approach have been used to address MNAR mechanism: pattern mixture models[@little1996] and selection models [@vella1998].
+One of the most popular techniques within selection models is the one proposed byHeckman (1976).
 [@heckman1976] Briefly, the Heckman selection model corrects for selection bias by estimating two linked equations: a outcome equation, where the missing variable is associated with predictors, and a selection equation, that accounts for the inclusion of observations in the sample.
@@
 In our imputation method we estimate the Heckman model using the copula method, as in real life data generating mechanism may deviate a BVN distribution.
-<!-- This facilitates the Heckman model estimation and makes it more robust to deviations of distributional assumptions.
--->
 
@@
  The Heckman model can be extended to hierarchical settings, i.e., in cases when individuals or sampling units are nested within clusters, as is the case in EHR or IPD.
-<!--In this case, sample units from the same cluster are expected to share some characteristics (e.g., the distribution of variables and correlations between them, relationships between exposure and outcomes, missing processes).
-
-Since most statistical analyses assume that the sample units are independent of each other, more complex hierarchical models dealing with nested data are required.
---> This hierarchical complexity must not only be taken into account in the analysis model, but also when dealing with missing data, thus requiring imputation models that are congenial to the analysis model, i.e.
+This hierarchical complexity must not only be taken into account in the analysis model, but also when dealing with missing data, thus requiring imputation models that are congenial to the analysis model, i.e.
 that make the same assumptions about the data.
@@
 Different procedures can be adopted to combine information between clusters; however, in our imputation method we opted for the two-stage approach that is often used in meta-analyses.[@simmonds2005] This is because such an approach is computationally less intensive and could potentially generate fewer convergence problems in the estimation of the Heckman hierarchical model compared to other approaches.
+Our method is based on the following conditional imputation model, whose parameters are $\boldsymbol{\theta}=\{\boldsymbol{\beta^O},\boldsymbol{\beta^S},\boldsymbol{\Psi^S},\boldsymbol{\Psi^O},(\sigma_i,\rho_i)_{1<=i<=N}\}$
+\begin{align*}
+\boldsymbol{y^*_{i}}&=\boldsymbol{X^O_{i}}(\boldsymbol{\beta^O}+b^O_{i})+\boldsymbol{\epsilon^O_{i}}\\
+\boldsymbol{r^*_{i}}&=\boldsymbol{X^S_{i}}(\boldsymbol{\beta^S}+b^S_{i})+\boldsymbol{\epsilon^S_{i}}\\
+b^O_{i}&\sim N(0,\boldsymbol{\Psi^O})\\
+b^S_{i}&\sim N(0,\boldsymbol{\Psi^S})
+\end{align*}
+\[\begin{pmatrix}
+\boldsymbol{\epsilon^O_{i}} \\
+\boldsymbol{\epsilon^S_{i}}
+\end{pmatrix}\sim N\left(\begin{pmatrix}
+0 \\
+0
+\end{pmatrix},\begin{pmatrix}
+\sigma^2_i & \rho_i\sigma_i \\
+\rho_i\sigma_i & 1
+\end{pmatrix}\right)
+\]
 
-Briefly, in a first stage  we estimate $\boldsymbol{\theta_i}=\{\boldsymbol{\beta_i^O},\boldsymbol{\beta_i^S},\sigma_i,\rho_i\}$, for all $i$ between 1 and $N$ clusters, the cluster specific parameters of the Heckman model.
-That is, we estimate separately for each of the clusters, the parameters of the two equations in each study, the outcome model (\ref{eq:1}) and the selection model (\ref{eq:2}).
-In the second stage, we combine all $\boldsymbol{\theta_i}$  by using a random effects multivariate meta-analysis model.
+Briefly, in a first stage  we estimate the cluster specific parameters of the Heckman model $\boldsymbol{\theta_i}=\{\boldsymbol{\hat{\beta}_i^O},\boldsymbol{\hat{\beta}_i^S},\hat{\sigma}_i,\hat{\rho}_i\}$, for all $i$ between 1 and $N$ clusters.That is, we estimate separately for each of the clusters, the parameters of the two equations in each study, the outcome model (\ref{eq:1}) and the selection model (\ref{eq:2}) via a copula model.
 
-In a random effects model, $\boldsymbol{\theta_i}$ parameters are assumed to be drawn independently and identically from a latent multivariate normal distribution of parameters with a  population mean $\boldsymbol{\Theta}$ and a population variance $\boldsymbol{\psi}$.
-[@higgins2009] Thus, we can write $\boldsymbol{\theta_i}=\boldsymbol{\Theta} + \boldsymbol{b_i}$  using $\boldsymbol{b_i} \sim N(0,\boldsymbol{\psi})$ random effects to allow for between-study heterogeneity in observed data relationships, and between-study heterogeneity in missing patterns.
+In a second stage, we estimate random effects multivariate meta-analysis model for the coefficient parameters.
+Here $\boldsymbol{\beta_i^O}$ and $\boldsymbol{\beta_i^S}$parameters are assumed to be drawn independently and identically from a latent multivariate normal distribution of parameters with a  population means $\boldsymbol{\beta^O}$ and $\boldsymbol{\beta^S}$ and population variance covariance $\boldsymbol{\psi^O}$ and $\boldsymbol{\psi^S}$ respectively.
+[@higgins2009].Here we assume that $\sigma_i$ and $\rho_i$ are random variables that come from a populational distribution and we estimate random effects univariate meta-analysis models for these parameters.
 
@@
 \label{sec:methods}
-We follow a similar approach proposed by Resche-Rigon and White (2018) [@resche-rigon2018] for multilevel imputation of data.
-Briefly, their method was developed to impute variables from a hierarchical structure (i.e., when there are units grouped within a cluster).
-This method involves estimating an outcome equation (describing the relationships of the observed data to the missing variable) separately in each cluster, after which the parameter estimates of that equation are pooled using random effects meta-analysis.
-
-With this method, values can be imputed in very common scenarios in IPD, e.g., sporadic and systematic missing patterns.
-In particular, when the response variable is systematically missing within a group, i.e., when $y_{ij}$ are totally missing within a cluster, the imputation values are drawn from a (generalized) linear model conditional on $\boldsymbol{\Theta}$ the marginal population parameters, i.e., those estimated after pooling the cluster-specific parameters $\boldsymbol{\theta_i}$.
-On the other hand, when the variable is sporadically missing within the cluster, i.e., there are some observed $y_{ij}$ within the cluster, the imputation model is conditional on the shrunk-cluster parameters, i.e., those coming from the shrinkage of $\boldsymbol{\theta_i}$ towards $\boldsymbol{\Theta}$.
-
-Our imputation approach differs crucially from the previous approach as here we estimate two correlated equations (instead of a single outcome equation) in each cluster, thus obtaining $\boldsymbol{\theta_i}=\{\boldsymbol{\beta_i^O},\boldsymbol{\beta_i^S},\sigma_i,\rho_i\}$ parameters from both equations which are then pooled into a $\boldsymbol{\Theta}$ parameter set at the marginal or population level.
-Our imputation method was created for imputing datasets with a single missing outcome or covariate variable, but it can also be used to impute multiple incomplete variables in a data set.
+We follow a similar approach proposed by Resche-Rigon and White (2018) [@resche-rigon2018] for multilevel data imputation, which allows us to impute values in very common scenarios in IPD, e.g., sporadic and systematic missingness patterns.
+Our imputation method was created to impute datasets with a single missing outcome variable or covariate, but it can also be used to impute multiple incomplete variables in a dataset, as it can be implemented under a multivariate imputation by chain equations (MICE) approach.
 
@@
 Initially, we use the copula method to estimate the set of cluster-specific parameters, $\boldsymbol{\widehat{\theta_i}}=\boldsymbol{\{\widehat{\beta^O_i}}, \boldsymbol{\widehat{\beta^S_i}}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, using all $j$ units with observable measurements $y_{ij}^{obs}$ within each cluster $i$.
-The Heckman model is estimated with the \textbf{gjrm} function of the GJRM R package[@radiceGJRMGeneralisedJoint2021],  under the bivariate model with the nonrandom sample selection (BSS) specification, from which we obtain not only the parameters' point estimates $\boldsymbol{\widehat{\theta_i}}$, but also their corresponding $\boldsymbol{\widehat{S(\theta_i)}}$ within-cluster variance-covariance matrix.
+Here we obtain the estimates not only the parameters' point estimates $\boldsymbol{\widehat{\theta_i}}$, but also their corresponding $\boldsymbol{\widehat{S(\theta_i)}}$ within-cluster variance-covariance matrix.
 
@@
 Similarly we combine all $q$ coefficient parameters of the $\boldsymbol{\beta^S}$ in the selection equation.
+Thus, we can denote the study specific coefficients
+$$\boldsymbol{\hat{\beta^O}_i} = \boldsymbol{\beta^O} +\boldsymbol{ b^O_i} + \boldsymbol{\epsilon'^O_{i}}$$
+$$\boldsymbol{\hat{\beta^S}_i} = \boldsymbol{\beta^S} + b^S_i + \boldsymbol{\epsilon'^S_{i}}$$
+using the random effects 
+ $b^O_i \sim N(0,\boldsymbol{\Psi^O})$ and $b^S_i\sim N(0,\boldsymbol{\Psi^S})$ with $\epsilon'^O_{i}\sim N(0,\sigma^O_{i}(X^O_i{X^O_i}^T))$ and $\epsilon'^S_{i}\sim N(0,\sigma^S_{i}(X^S_i{X^S_i}^T))$.
+From the meta model we obtain the estimates of parameters $\boldsymbol{\beta^S}$,$\boldsymbol{\beta^O}$, $\boldsymbol{\Psi^S}$,$\boldsymbol{\Psi^O}$.
 
-We also perform a univariate random effects meta_analysis on $\sigma'$, the log-transformed parameter of $\sigma$, and another on $\rho'$, the fisher-transformed parameter of $\rho$.
-
-The meta-analysis model is performed with the \textbf{mixmeta} function of the R package mixmeta[@gasparrini2021], which allows the use of maximum likelihood (ML), restricted maximum likelihood (REML), and moments estimation methods.
-For the simulation and illustrative study, we use the restricted REML estimation method, which is recommended as it has a good balance between unbiasedness and efficiency.
-[@viechtbauer2005]
+Assuming that  $\sigma_i$ and $\rho_i$ are random variables comming from a populational distribution, they can express as:
+$$ log(\sigma_i) \sim N( log(\sigma),\psi^{\sigma})$$
+$$ tanh^{-1}(\rho_i)\sim N( tanh^{-1}(\rho),\psi^{\rho})$$ 
+For each of them, we perform a univariate random effects meta_analysis. The model for $\sigma_i$ is given by the model:
+$$log(\hat\sigma_i) = \sigma + b^{\sigma}_i + \epsilon^{\sigma}_{i}$$
+with $b_i^{\sigma} \sim N(0,\psi^{\sigma})$, and $\epsilon_i^{\sigma} \sim N(0,var(log(\hat{\sigma}_i)))$ and  the model for $\rho_i$ is:
+$$tanh^{-1}(\hat\rho_i) = \rho + b^{\rho}_i + \epsilon^{\rho}_{i}$$
+with $b_i^{\rho} \sim N(0,\psi^{\rho})$, and $\epsilon_i^{\rho} \sim N(0,var(tanh^{-1}(\hat{\rho}_i)))$.
 
@@
 
-From the meta-analysis model, we obtain the marginal estimates $\boldsymbol{\widehat{\Theta}}$ and the between-cluster variance matrix $\boldsymbol{\widehat{\psi}}$,i.e.
+From the meta-analysis model, we obtain the marginal estimates $\boldsymbol{\widehat{\Theta}}=\{\widehat{\boldsymbol{\beta^S}},\widehat{\boldsymbol{\beta^O}},\hat{\sigma},\hat{\rho}\}$ and the between-cluster variance matrix $\boldsymbol{\widehat{\psi}}$,i.e.
 variance-covariance matrix of the random effects, with their corresponding variance-covariance matrices $\boldsymbol{\widehat{S_{\Theta}}}$ and $\boldsymbol{\widehat{S_{\psi}}}$, which are used to draw the $\boldsymbol{\Theta^*}$ and $\boldsymbol{\psi^*}$ parameters, from their posterior distribution as follows: [@resche-rigon2013]
@@
 Furthermore, as implemented according to the MICE framework, the imputation method could be used simultaneously with other imputation methods available in MICE or in add-in packages such as micemd [@audigier2021], to impute datasets with multiple incomplete variables that differ in type and missing mechanism.
+
+## Technical details of the implementation 
+The Heckman model is estimated with the \textbf{gjrm} function of the GJRM R package[@radiceGJRMGeneralisedJoint2021],  under the bivariate model with the nonrandom sample selection (BSS) specification.
+The meta-analysis model is performed with the \textbf{mixmeta} function of the R package mixmeta[@gasparrini2021], which allows the use of maximum likelihood (ML), restricted maximum likelihood (REML), and moments estimation methods.
+For the simulation and illustrative study, we use the restricted REML estimation method, which is recommended as it has a good balance between unbiasedness and efficiency.
+[@viechtbauer2005].
 
@@
 
-- \textbf{Bias:} Bias on the coefficient and random effect parameters.
+- \textbf{Bias:} Bias on the coefficient and variance-covariance matrix terms.
 - \textbf{Coverage:} Coverage of the 95\% confidence intervals for the coefficient  parameters.

### 2022-12-16 2.Simulation/simplot.R

@@ -19,32 +19,38 @@ data<- setDT(melt(data, measure.vars = c( paste0("p_Bias_",0:6), paste0("p_Cover
 data[, c("Type","Statistic", "Parameter") := tstrsplit(Statistic, "_", fixed=TRUE)]
 data<- setDT(dcast(data, Scenario+ Variation+ Method +  Parameter+Statistic~Type, value.var = "Value"))
 param_val <- c("0"="beta_0", "1"="beta_1", "2"="beta_2", 
-               "3"="sigma_b0", "4"="sigma_b1", "5"="sigma_b2","6"= "sigma_e")
+               "3"="psi_b0", "4"="psi_b1", "5"="psi_b2","6"= "psi_e")
 data[,Method:=factor(Method, levels = c("CC","1l.Heckman","2l.MAR","2l.Heckman"))]
 
 data[,Parameter := as.factor(param_val[data$Parameter])]
-data[,Parameter := factor(Parameter, levels = c("beta_0","beta_1","beta_2","sigma_b0","sigma_b1","sigma_b2","sigma_e"),
+data[,Parameter := factor(Parameter, levels = c("beta_0","beta_1","beta_2","psi_b0","psi_b1","psi_b2","psi_e"),
                           ordered = TRUE, labels=c(expression(beta[0]),expression(beta[1]),expression(beta[2]),
-                                                   expression(sigma[b*0]),expression(sigma[b*1]),expression(sigma[b*2]),
-                                                   expression(sigma[e])))]
+                                                   expression(sqrt(psi[oo])),expression(sqrt(psi[11])),expression(sqrt(psi[22])),
+                                                   expression(psi[e])))]
 
 
 plotsim<-function(Scenariov,levelsvar,yaxis){
 datagraph<-data[Scenario==Scenariov,]
-datalines <- data.table(Parameter = c("beta_0","beta_1", "beta_2","sigma_b0", "sigma_b1","sigma_b2","sigma_e"),
+datalines <- data.table(Parameter = c("beta_0","beta_1", "beta_2","psi_b0", "psi_b1","psi_b2","psi_e"),
                         Z = c(rep(0,7),rep(0.95,7),rep(0,7),rep(0,7)),
                         Statistic=c(rep("Bias",7),rep("Coverage",7),rep("RMSE",7),rep("Width",7)))
-datalines[,Parameter := factor(Parameter, levels = c("beta_0","beta_1","beta_2","sigma_b0","sigma_b1","sigma_b2","sigma_e"),
+
+
[…]

### 2022-12-16 1.Paper/Paper.Rmd
manuscript update

@@
 Thus, to estimate the parameters of the Heckman method, it is sufficient to specify the marginal distributions of the error terms, and link them with a suitable copula function.
-In our imputation method we estimate the Heckman model using the copula method, as in real life data generating mechanism may deviate a BVN distribution.
+In our imputation method we estimate the Heckman model using the copula method for non-random selection [@wojtys2018] which is a flexible approach that allows the specification of different parametric distributions of the selection and outcome variables and also different types of dependence structure between the two equations.
 
@@
 using the random effects 
- $b^O_i \sim N(0,\boldsymbol{\Psi^O})$ and $b^S_i\sim N(0,\boldsymbol{\Psi^S})$ with $\epsilon'^O_{i}\sim N(0,\sigma^O_{i}(X^O_i{X^O_i}^T))$ and $\epsilon'^S_{i}\sim N(0,\sigma^S_{i}(X^S_i{X^S_i}^T))$.
-From the meta model we obtain the estimates of parameters $\boldsymbol{\beta^S}$,$\boldsymbol{\beta^O}$, $\boldsymbol{\Psi^S}$,$\boldsymbol{\Psi^O}$.
+ $b^O_i \sim N(0,\boldsymbol{\Psi^O})$ and $b^S_i\sim N(0,\boldsymbol{\Psi^S})$ with $\epsilon'^O_{i}\sim N(0,\sigma^O_{i}(X^O_i{X^O_i}^T))$ and $\epsilon'^S_{i}\sim N(0,\sigma^S_{i}(X^S_i{X^S_i}^T)^{-1})$.
 
-Assuming that  $\sigma_i$ and $\rho_i$ are random variables comming from a populational distribution, they can express as:
+We assume that  $\sigma_i$ and $\rho_i$ are random variables coming from a population distribution, so we can express them as:
 $$ log(\sigma_i) \sim N( log(\sigma),\psi^{\sigma})$$
 $$ tanh^{-1}(\rho_i)\sim N( tanh^{-1}(\rho),\psi^{\rho})$$ 
-For each of them, we perform a univariate random effects meta_analysis. The model for $\sigma_i$ is given by the model:
+For each of them, we perform a univariate random effects meta-analysis. The model for $\sigma_i$ is given by the model:
 $$log(\hat\sigma_i) = \sigma + b^{\sigma}_i + \epsilon^{\sigma}_{i}$$
@@
 
+In the case of an incomplete binary variable, the $\sigma$ parameter is not specified, so it is not necessary to perform a random effects meta-analysis model of $\sigma$ or to include it in any of the following steps in the imputation model.
+
 #### Draw the marginal parameters $\boldsymbol{\Theta}$
 
-From the meta-analysis model, we obtain the marginal estimates $\boldsymbol{\widehat{\Theta}}=\{\widehat{\boldsymbol{\beta^S}},\widehat{\boldsymbol{\beta^O}},\hat{\sigma},\hat{\rho}\}$ and the between-cluster variance matrix $\boldsymbol{\widehat{\psi}}$,i.e.
+From the meta-analysis model, we obtain the marginal estimates $\boldsymbol{\widehat{\Theta}}=\{\widehat{\boldsymbol{\beta^S}},\widehat{\boldsymbol{\beta^O}},\hat{\sigma},\hat{\rho}\}$ and the between-cluster variance matrix $\boldsymbol{\widehat{\psi}}=\{\boldsymbol{\Psi^S},\boldsymbol{\Psi^O},\psi^{\sigma},\psi^{\rho}\}$, i.e.
 variance-covariance matrix of the random effects, with their corresponding variance-covariance matrices $\boldsymbol{\widehat{S_{\Theta}}}$ and $\boldsymbol{\widehat{S_{\psi}}}$, which are used to draw the $\boldsymbol{\Theta^*}$ and $\boldsymbol{\psi^*}$ parameters, from their posterior distribution as follows: [@resche-rigon2013]
@@
 For each dataset, we generated $X_{1i}$ a treatment indicator variable from a Bernoulli distribution with a probability of treatment on each cluster equal to 0.6.
-Next, we simulated the mean of two continuous covariates from a multivariate normal distribution $\mu_h\sim N(\begin{psmallmatrix}0\\0\end{psmallmatrix},\begin{psmallmatrix}0.2 & 0.015\\0.015 & 0.2\end{psmallmatrix})$, with $h=\{2,3\}$.
+Next, we simulated the mean of two continuous covariates from a multivariate normal distribution $\begin{psmallmatrix}\mu_2\\\mu_3\end{psmallmatrix}\sim N\left(\begin{psmallmatrix}0\\0\end{psmallmatrix},\begin{psmallmatrix}0.2 & 0.015\\0.015 & 0.2\end{psmallmatrix}\right)$.
 We then simulated for each cluster a baseline covariate $X_{2i}\sim N(\mu_2,1)$ and a exclusion restriction variable  $X_{3i} \sim N(\mu_3,0.5)$.
@@
 \]
-whose $\sigma_i^2$ is variable across clusters and distributed as $log(\sigma_i) \sim N(0,0.05)$.
+whose $\sigma_i^2$ is a random parameter across clusters distributed as $log(\sigma_i) \sim N(0,0.05)$.
 

### 2022-12-16 1.Paper/Paper.Rmd
Update Paper.Rmd

@@
 
-Further, the method can also be extended to other types of variables such as count or ordinal variables.
-Similarly, less restrictive Heckman based models can be considered in terms of normality distribution of errors and no specification of exclusion variables such as those proposed by Ogundimu & Collins (2019).[@ogundimu_collins19].
+Further, the method can also be extended to other copula models for non-random selection, with different distributions of the selection and outcome equations and dependency structure.[@wojtys2018]  Similarly, less restrictive Heckman based models can be considered in terms of normality distribution of errors and no specification of exclusion variables such as those proposed by Ogundimu & Collins (2019).[@ogundimu_collins19].
 

### 2022-12-16 1.Paper/Paper.Rmd
Update Paper.Rmd

@@
 \begin{align*}
- \beta_{hi}^O&= \beta_h^O + b_{hi}^O\\
- \beta_{hi}^S&= \beta_h^S + b_{hi}^S
+ \beta_{i}^O&= \beta^O + b_{i}^O\\
+ \beta_{i}^S&= \beta^S + b_{i}^S
 \end{align*}
-We fixed coefficients $\beta_h^O=\{0.3,1,1\}$  and $\beta_h^S=\{-0.8,1.3,-0.7,1.2\}$ in order to get around 40\% of sporadically missing values on the response $y_{ij}$ in the entire data set.
+We fixed coefficients $\beta^O=\{0.3,1,1\}$  and $\beta^S=\{-0.8,1.3,-0.7,1.2\}$ in order to get around 40\% of sporadically missing values on the response $y_{ij}$ in the entire data set.
 Additionally, we ensured that the $y_{ij}^*$ observations were systematically missing in 20% of the clusters included in the data set, by removing the outcome values in the 20% of the clusters.
 
-We assumed that random effects were independent within equations ($b_{h0}^O\indep b_{h1}^O \indep b_{h2}^O$ and $b_{h0}^S \indep b_{h1}^S \indep b_{h2}^S$), but were linked between both selection and outcome equations through a bivariate normal distributed as:
+We assumed that random effects were independent within equations ($b_{0}^O\indep b_{1}^O \indep b_{2}^O$ and $b_{0}^S \indep b_{1}^S \indep b_{2}^S\indep b_{3}^S$), but were linked between both selection and outcome equations through a bivariate normal distributed as:
 \[\begin{pmatrix}
-b^O_{h} \\
-b^S_{h}
+b^O \\
+b^S
 \end{pmatrix}\sim N\left(\begin{pmatrix}
@@
 0
-\end{pmatrix},\sigma_{bh}^2\begin{pmatrix}
+\end{pmatrix},0.4^2\begin{pmatrix}
 1 & \rho*0.4 \\
@@
 \]
-with the parameters $\sigma_{b0}^2=\sigma_{b1}^2=\sigma_{b2}^2=0.4$.
+for the coefficients of intercept,$X_1$,$X_2$ and $X_3$.
 We considered that the correlation parameter of the random effects between equations is 40% of the value of the assumed correlation parameter between error terms $\rho$.
-In addition, we included a random effect on the exclusion restriction variable given by $b_3\sim N(0,0.2)$ assuming that the intracluster variation in the exclusion restriction effect is lower than the variation on other coefficient parameters effects.
+In addition, we included a random effect on the exclusion restriction variable given by $b_3\sim N(0,0.2^2)$ assuming that the intracluster variation in the exclusion restriction effect is lower than the variation on other coefficient parameters effects.
 The $\rho$ parameter was given different values depending on the simulated missing mechanism (Table \ref{tab:scenarios}) .
@@
 The estimands were the parameter coefficients of the outcome equation $\beta^O=\{\beta^O_0,\beta^O_1,\beta^O_2\}$, with special emphasis on the treatment effect parameter $\beta^0_1$.
-We also report the estimated variance of the random effects and residual errors $\sigma_{b0}^2$,$\sigma_{b1}^2$,$\sigma_{b2}^2$,$\sigma_{e}^2$.
+We also report the estimated standard deviation from the covariance matrix of the random effects, i.e.,  $\sqrt{\psi_{00}}$,$\sqrt{\psi_{11}}$,$\sqrt{\psi_{22}}$.
+
+#### Estimating procedures:
+
+After the imputation procedure, we estimated the following (generalized) mixed linear effect model using the *lmer()* function from the *lme4* R package.[@bates2022]
+$y_{i}=\beta_i^OX_i^O+\epsilon^O_{i}$
+In case of missing binary variable, we used the same matrix of predictors but on a binary model estimated with the *glmer()* function from the *lme4* R package.
+Then, we pooled the estimates of the $\beta_i^O$ and the variance of the random effect and residual errors of the multiple imputed datasets according to Rubin's rule[@rubin1987], over which we calculated the performance measures on the estimands.
+
+To calculate the coverage of the parameter coefficients' 95\% confidence intervals (CI), we estimate CI with the Wald method.
+<!-- Although it is possible to obtain CI through the profile or bootstrap method for the variances of the random effects, we prefer not to estimate them (and hence the coverage) for the random effects parameters due to computational time.
+-->
 
@@
 - \textbf{2l.Heckman}: The proposed imputation method based on the Heckman model for hierarchical datasets.
-
-## Estimating procedures
-
-After the imputation procedure, we estimated the following (generalized) mixed linear effect model using the *lmer()* function from the *lme4* R package.[@bates2022]
-$y_{i}=\beta_i^OX_i^O+\epsilon^O_{i}$
-In case of missing binary variable, we used the same matrix of predictors but on a binary model estimated with the *glmer()* function from the *lme4* R package.
-Then, we pooled the estimates of the $\beta_i^O$ and the variance of the random effect and residual errors of the multiple imputed datasets according to Rubin's rule[@rubin1987], over which we calculated the performance measures on the estimands.
-
-To calculate the coverage of the parameter coefficients' 95\% confidence intervals (CI), we estimate CI with the Wald method.
-<!-- Although it is possible to obtain CI through the profile or bootstrap method for the variances of the random effects, we prefer not to estimate them (and hence the coverage) for the random effects parameters due to computational time.
--->
 
@@
 In the case of a missing bivariate variable, we also found (Figure \ref{fig:rhobin}) that the 2l.Heckman method provided the least unbiased results on the coefficient parameters.
-However, we observed more unbiased estimates and a larger 95th percentile amplitude in the estimates of the variance of the random effects, especially in the $\sigma_{b2}$ compared to the estimates obtained in the continuous incomplete outcome scenario.
+However, we observed more unbiased estimates and a larger 95th percentile amplitude in the estimates of the standard deviations of the random effects, especially in the $\sqrt{\psi_{22}}$ compared to the estimates obtained in the continuous incomplete outcome scenario.
 
@@
 
-When we reduced the sample size to 50 patients per study, the bias and RMSE of the $\sigma_b2$ were drastically affected (not shown here but in the Appendix).
+When we reduced the sample size to 50 patients per study, the bias and RMSE of the $\sqrt{\psi_{22}}$ were drastically affected (not shown here but in the Appendix).
 This could be explained in part due to the scarce information on certain clusters, which affects directly the estimation of the Heckman model on those clusters.
@@
 when the probability of loss is associated with the value of the missing variable, we observe that under all imputation methods we obtain biased estimators, with poor coverage.
-Particularly when estimating the intercept we observed more biased estimates compared to those obtained for the other estimated parameters B2 with coverage below 60%.
+Particularly when estimating the intercept we observed more biased estimates compared to those obtained for the other estimated parameters $\beta_2$ with coverage below 60%.
 
 Our model may not be fully suitable for this type of scenario, especially if the main analysis is focused on estimating absolute prevalence estimates, as it is greatly affected by the bias of the intercept parameter ($\beta_0$) and the width of the 95th percentile CI of all coefficient parameters.
-Also the imputation method was affected in terms of the bias of the random-effects parameters $\sigma_{b2}$.
+Also the imputation method was affected in terms of the bias of the standard deviation of the random-effects parameters $\sqrt{\psi_{22}}$.
 

### 2023-01-24 1.Paper/Paper.Rmd
Update manuscript

@@
 author:
-- name: Johanna Munoz*
+- name: Johanna Muñoz*
   num: a
-- name: Matthias Egger 
-  num: b  
 - name: Orestis Efthimiou
@@
 articletype: Research article
-received: 2022-11-01
-revised: 2022-12-01
-accepted: 2023-01-01
 abstract: "Missing data is a common problem in medical research, and is commonly addressed using multiple imputation.
@@
 Two major approach have been used to address MNAR mechanism: pattern mixture models[@little1996] and selection models [@vella1998].
-One of the most popular techniques within selection models is the one proposed byHeckman (1976).
+One of the most popular techniques within selection models is the one proposed by Heckman (1976).
 [@heckman1976] Briefly, the Heckman selection model corrects for selection bias by estimating two linked equations: a outcome equation, where the missing variable is associated with predictors, and a selection equation, that accounts for the inclusion of observations in the sample.
@@
 However, to avoid multicollinearity problems[@puhani1997], it is recommended to include in $\boldsymbol{X^S_{i}}$ at least one variable that is not included in the outcome model.
-[@angrist2001] This variable is commonly known as an exclusion restriction variable (ERV), and should only be associated with the selection in the sample $r^*_{ij}$ (relevance condition) but not with the actual observation $y^*_{ij}$ (exclusion condition).[@gomes_etal20] The ERV meets by definition the relevance and exclusion conditions in order to provide independent information about the selection process and to facilitate the estimation of the Heckman model.
+[@angrist2001] This variable is commonly known as an exclusion restriction variable (ERV), and should only be associated with the selection in the sample $\boldsymbol{r^*_{i}}$ (relevance condition) but not with the actual observation $\boldsymbol{y^*_{i}}$ (exclusion condition).[@gomes_etal20] The ERV meets by definition the relevance and exclusion conditions in order to provide independent information about the selection process and to facilitate the estimation of the Heckman model.
 
@@
 \[\begin{pmatrix}
-\epsilon^O_{ij} \\
-\epsilon^S_{ij}
-\end{pmatrix}\sim N\left(\begin{pmatrix}
+\boldsymbol{\epsilon^O_{i}} \\
+\boldsymbol{\epsilon^S_{i}}
+\end{pmatrix}\sim N_{n_i}\left(\begin{pmatrix}
 0 \\
@@
 \boldsymbol{r^*_{i}}&=\boldsymbol{X^S_{i}}(\boldsymbol{\beta^S}+b^S_{i})+\boldsymbol{\epsilon^S_{i}}\\
-b^O_{i}&\sim N(0,\boldsymbol{\Psi^O})\\
-b^S_{i}&\sim N(0,\boldsymbol{\Psi^S})
+b^O_{i}&\sim N_p(0,\boldsymbol{\Psi^O})\\
+b^S_{i}&\sim N_q(0,\boldsymbol{\Psi^S})
 \end{align*}
@@
 \boldsymbol{\epsilon^S_{i}}
-\end{pmatrix}\sim N\left(\begin{pmatrix}
+\end{pmatrix}\sim N_{n_i}\left(\begin{pmatrix}
 0 \\
@@
 
-Briefly, in a first stage  we estimate the cluster specific parameters of the Heckman model $\boldsymbol{\theta_i}=\{\boldsymbol{\hat{\beta}_i^O},\boldsymbol{\hat{\beta}_i^S},\hat{\sigma}_i,\hat{\rho}_i\}$, for all $i$ between 1 and $N$ clusters.That is, we estimate separately for each of the clusters, the parameters of the two equations in each study, the outcome model (\ref{eq:1}) and the selection model (\ref{eq:2}) via a copula model.
+Briefly, in a first stage  we estimate the cluster specific parameters of the Heckman model $\widehat{\boldsymbol{\theta_i}}=\boldsymbol{\{\widehat{\beta^O_i}}, \boldsymbol{\widehat{\beta^S_i}}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, for all $i$ between 1 and $N$ clusters.
+That is, we estimate separately for each of the clusters, the parameters of the two equations in each study, the outcome model (\ref{eq:1}) and the selection model (\ref{eq:2}) via a copula model.
 
@@
 Here $\boldsymbol{\beta_i^O}$ and $\boldsymbol{\beta_i^S}$parameters are assumed to be drawn independently and identically from a latent multivariate normal distribution of parameters with a  population means $\boldsymbol{\beta^O}$ and $\boldsymbol{\beta^S}$ and population variance covariance $\boldsymbol{\psi^O}$ and $\boldsymbol{\psi^S}$ respectively.
-[@higgins2009].Here we assume that $\sigma_i$ and $\rho_i$ are random variables that come from a populational distribution and we estimate random effects univariate meta-analysis models for these parameters.
+[@higgins2009] Here we assume that $\sigma_i$ and $\rho_i$ are random variables that come from a populational distribution and we estimate random effects univariate meta-analysis models for these parameters.
 
@@
 
-Initially, we use the copula method to estimate the set of cluster-specific parameters, $\boldsymbol{\widehat{\theta_i}}=\boldsymbol{\{\widehat{\beta^O_i}}, \boldsymbol{\widehat{\beta^S_i}}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, using all $j$ units with observable measurements $y_{ij}^{obs}$ within each cluster $i$.
-Here we obtain the estimates not only the parameters' point estimates $\boldsymbol{\widehat{\theta_i}}$, but also their corresponding $\boldsymbol{\widehat{S(\theta_i)}}$ within-cluster variance-covariance matrix.
+Initially, we use the copula method to estimate the set of cluster-specific parameters, $\widehat{\boldsymbol{\theta_i}}=\boldsymbol{\{\widehat{\beta^O_i}}, \boldsymbol{\widehat{\beta^S_i}}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, using all $j$ units with observable measurements $y_{ij}^{obs}$ within each cluster $i$.
+Here we obtain the estimates not only the parameters' point estimates $\widehat{\boldsymbol{\theta_i}}$, but also their corresponding $\boldsymbol{\widehat{S(\theta_i)}}$ within-cluster variance-covariance matrix.
 
@@
 
-In this step, we pool the parameters $\boldsymbol{\widehat{\theta_i}}$ with a random effects meta-analysis model using only the clusters with observable information, i.e., those with no systematically missing outcome.
+In this step, we pool the parameters $\widehat{\boldsymbol{\theta_i}}$ with a random effects meta-analysis model using only the clusters with observable information, i.e., those with no systematically missing outcome.
 In particular, we pool the $p$ coefficients of the $\boldsymbol{\beta^O}$ in the outcome equation and estimate a multivariate random effects meta-analysis model with them.
@@
 Thus, we can denote the study specific coefficients
-$$\boldsymbol{\hat{\beta^O}_i} = \boldsymbol{\beta^O} +\boldsymbol{ b^O_i} + \boldsymbol{\epsilon'^O_{i}}$$
-$$\boldsymbol{\hat{\beta^S}_i} = \boldsymbol{\beta^S} + b^S_i + \boldsymbol{\epsilon'^S_{i}}$$
+$$\boldsymbol{\widehat{\beta^O_i}} = \boldsymbol{\beta^O} +\boldsymbol{ b^O_i} + \boldsymbol{\epsilon'^O_{i}}$$
+$$\boldsymbol{\widehat{\beta^S_i}} = \boldsymbol{\beta^S} + b^S_i + \boldsymbol{\epsilon'^S_{i}}$$
 using the random effects 
- $b^O_i \sim N(0,\boldsymbol{\Psi^O})$ and $b^S_i\sim N(0,\boldsymbol{\Psi^S})$ with $\epsilon'^O_{i}\sim N(0,\sigma^O_{i}(X^O_i{X^O_i}^T))$ and $\epsilon'^S_{i}\sim N(0,\sigma^S_{i}(X^S_i{X^S_i}^T)^{-1})$.
+ $b^O_i \sim N(0,\boldsymbol{\Psi^O})$ and $b^S_i\sim N(0,\boldsymbol{\Psi^S})$ with sampling errors $\epsilon'^O_{i}$ and $\epsilon'^S_{i}$.
 
@@
 For each of them, we perform a univariate random effects meta-analysis. The model for $\sigma_i$ is given by the model:
-$$log(\hat\sigma_i) = \sigma + b^{\sigma}_i + \epsilon^{\sigma}_{i}$$
-with $b_i^{\sigma} \sim N(0,\psi^{\sigma})$, and $\epsilon_i^{\sigma} \sim N(0,var(log(\hat{\sigma}_i)))$ and  the model for $\rho_i$ is:
-$$tanh^{-1}(\hat\rho_i) = \rho + b^{\rho}_i + \epsilon^{\rho}_{i}$$
-with $b_i^{\rho} \sim N(0,\psi^{\rho})$, and $\epsilon_i^{\rho} \sim N(0,var(tanh^{-1}(\hat{\rho}_i)))$.
 
-In the case of an incomplete binary variable, the $\sigma$ parameter is not specified, so it is not necessary to perform a random effects meta-analysis model of $\sigma$ or to include it in any of the following steps in the imputation model.
+$$log(\widehat{\sigma}_i)= \sigma + b^{\sigma}_i + \epsilon^{\sigma}_{i}$$
+
+with $b_i^{\sigma} \sim N(0,\psi^{\sigma})$, and $\epsilon_i^{\sigma} \sim N(0,var(log(\widehat{\sigma_i})))$.
+In the case of an incomplete binary variable, the $\sigma_i$ parameter is not specified, so it is not necessary to perform a random effects meta-analysis model of $\sigma_i$ or to include it in any of the following steps in the imputation model.
+
+The model for $\rho_i$ is given by:
+$$tanh^{-1}(\widehat{\rho_i}) = \rho + b^{\rho}_i + \epsilon^{\rho}_{i}$$
+with $b_i^{\rho} \sim N(0,\psi^{\rho})$, and $\epsilon_i^{\rho} \sim N(0,var(tanh^{-1}(\widehat{\rho_i})))$.
 
@@
 
-From the meta-analysis model, we obtain the marginal estimates $\boldsymbol{\widehat{\Theta}}=\{\widehat{\boldsymbol{\beta^S}},\widehat{\boldsymbol{\beta^O}},\hat{\sigma},\hat{\rho}\}$ and the between-cluster variance matrix $\boldsymbol{\widehat{\psi}}=\{\boldsymbol{\Psi^S},\boldsymbol{\Psi^O},\psi^{\sigma},\psi^{\rho}\}$, i.e.
+From the meta-analysis model, we obtain the marginal estimates
+$\widehat{\boldsymbol{\Theta}}=\boldsymbol{\{\widehat{\beta^O}}, \boldsymbol{\widehat{\beta^S}}, \widehat{\sigma}, \widehat{\rho}\}$
+and the between-cluster variance matrix
+$\widehat{\boldsymbol{\psi}}=\boldsymbol{\{\widehat{\Psi^O}}, \boldsymbol{\widehat{\Psi^S}}, \widehat{\psi^{\sigma}}, \widehat{\psi^{\rho}}\}$, i.e.
 variance-covariance matrix of the random effects, with their corresponding variance-covariance matrices $\boldsymbol{\widehat{S_{\Theta}}}$ and $\boldsymbol{\widehat{S_{\psi}}}$, which are used to draw the $\boldsymbol{\Theta^*}$ and $\boldsymbol{\psi^*}$ parameters, from their posterior distribution as follows: [@resche-rigon2013]
@@
 \begin{align*}
-\boldsymbol{\Theta^*}& \sim N(\boldsymbol{\widehat{\Theta}},\boldsymbol{\widehat{S_{\Theta}}})\\
-\boldsymbol{\psi^*}& \sim N(\boldsymbol{\widehat{\psi}},\boldsymbol{\widehat{S_{\psi}}})
+\boldsymbol{\Theta^*}& \sim N(\widehat{\boldsymbol{\Theta}},\widehat{\boldsymbol{S_{\Theta}}})\\
+\boldsymbol{\psi^*}& \sim N(\widehat{\boldsymbol{\psi}},\widehat{\boldsymbol{S_{\psi}}})
 \end{align*}
@@
 \begin{align*}
- \boldsymbol{y_{i}^*}&= \boldsymbol{\beta_i^O}\boldsymbol{X_i^O}+\boldsymbol{\epsilon^O_{i}} \\
- \boldsymbol{r_{i}^*}&= \boldsymbol{\beta_i^S}\boldsymbol{X_i^O}+\boldsymbol{\epsilon^S_{i}}
+ \boldsymbol{y_{i}^*}&= \boldsymbol{X_i^O}\boldsymbol{\beta_i^O}+\boldsymbol{\epsilon^O_{i}} \\
+ \boldsymbol{r_{i}^*}&= \boldsymbol{X_i^O}\boldsymbol{\beta_i^S}+\boldsymbol{\epsilon^S_{i}}
 \end{align*}
-Here we assumed that all coefficient parameters varied across studies, by including cluster-specific random effects as:
+where $\boldsymbol{\beta_i^O}=(\beta_{i0}^O,\beta_{i1}^O,\beta_{i2}^O)^T$ and $\boldsymbol{\beta_i^S}=(\beta_{i0}^S,\beta_{i1}^S,\beta_{i2}^S,\beta_{i3}^S)^T$.
+Here we assumed that the coefficient parameter k varied across studies, by including cluster-specific random effects as:
 \begin{align*}
- \beta_{i}^O&= \beta^O + b_{i}^O\\
- \beta_{i}^S&= \beta^S + b_{i}^S
+ \beta_{ik}^O&= \beta_k^O + b_{ik}^O\\
+ \beta_{ik}^S&= \beta_k^S + b_{ik}^S
 \end{align*}
-We fixed coefficients $\beta^O=\{0.3,1,1\}$  and $\beta^S=\{-0.8,1.3,-0.7,1.2\}$ in order to get around 40\% of sporadically missing values on the response $y_{ij}$ in the entire data set.
+Denoting $\boldsymbol{\beta^O}$=$(\beta_{0}^O,\beta_{1}^O,\beta_{2}^O)^T$ and $\boldsymbol{\beta^S}$=$(\beta_{0}^S,\beta_{1}^S,\beta_{2}^S,\beta_{3}^S)^T$, we fixed coefficients $\boldsymbol{\beta^O}$=$(0.3,1,1)$  and $\boldsymbol{\beta^S}$=$(-0.8,1.3,-0.7,1.2)$ in order to get around 40\% of sporadically missing values on the response $y_{ij}$ in the entire data set.
 Additionally, we ensured that the $y_{ij}^*$ observations were systematically missing in 20% of the clusters included in the data set, by removing the outcome values in the 20% of the clusters.
 
-We assumed that random effects were independent within equations ($b_{0}^O\indep b_{1}^O \indep b_{2}^O$ and $b_{0}^S \indep b_{1}^S \indep b_{2}^S\indep b_{3}^S$), but were linked between both selection and outcome equations through a bivariate normal distributed as:
+We assumed that random effects were independent within equations ($b_{i0}^O\indep b_{i1}^O \indep b_{i2}^O$ and $b_{i0}^S \indep b_{i1}^S \indep b_{i2}^S\indep b_{i3}^S$), but were linked between both selection and outcome equations through a bivariate normal distributed as:
 \[\begin{pmatrix}
-b^O \\
-b^S
+b_{ik}^O \\
+b_{ik}^S
 \end{pmatrix}\sim N\left(\begin{pmatrix}
@@
 0
-\end{pmatrix},0.4^2\begin{pmatrix}
-1 & \rho*0.4 \\
-\rho*0.4 &1 
+\end{pmatrix},\begin{pmatrix}
+{\psi^O_{kk}} & \rho_b\sqrt{\psi^O_{kk}}\sqrt{\psi^S_{kk}} \\
+\rho_b \sqrt{\psi^O_{kk}}\sqrt{\psi^S_{kk}} &{\psi^S_{kk}}
 \end{pmatrix}\right)
 \]
-for the coefficients of intercept,$X_1$,$X_2$ and $X_3$.
+with ${\psi^S_{00}}={\psi^O_{00}}={\psi^S_{11}}={\psi^O_{11}}={\psi^S_{22}}={\psi^O_{22}}=0.4$ and $\rho_b=\rho*0.4$.
 We considered that the correlation parameter of the random effects between equations is 40% of the value of the assumed correlation parameter between error terms $\rho$.
-In addition, we included a random effect on the exclusion restriction variable given by $b_3\sim N(0,0.2^2)$ assuming that the intracluster variation in the exclusion restriction effect is lower than the variation on other coefficient parameters effects.
+In addition, we included a random effect on the exclusion restriction variable given by $b_{i3}^S\sim N(0,0.2)$ assuming that the intracluster variation in the exclusion restriction effect is lower than the variation on other coefficient parameters effects.
 The $\rho$ parameter was given different values depending on the simulated missing mechanism (Table \ref{tab:scenarios}) .
@@
 \epsilon^S_{i}
-\end{pmatrix}\sim N\left(\begin{pmatrix}
+\end{pmatrix}\sim N_{n_i}\left(\begin{pmatrix}
 0 \\
@@
 ## Estimands
-The estimands were the parameter coefficients of the outcome equation $\beta^O=\{\beta^O_0,\beta^O_1,\beta^O_2\}$, with special emphasis on the treatment effect parameter $\beta^0_1$.
-We also report the estimated standard deviation from the covariance matrix of the random effects, i.e.,  $\sqrt{\psi_{00}}$,$\sqrt{\psi_{11}}$,$\sqrt{\psi_{22}}$.
+The estimands were the parameter coefficients of the outcome equation $\boldsymbol{\beta^O}=(\beta_{0}^O,\beta_{1}^O,\beta_{2}^O)^T$, with special emphasis on the treatment effect parameter $\beta^0_1$.
+We also report the estimated standard deviation from the covariance matrix of the random effects, i.e.,  $\sqrt{\psi^O_{00}}$,$\sqrt{\psi^O_{11}}$,$\sqrt{\psi^O_{22}}$.
 
@@
 
-After the imputation procedure, we estimated the following (generalized) mixed linear effect model using the *lmer()* function from the *lme4* R package.[@bates2022]
-$y_{i}=\beta_i^OX_i^O+\epsilon^O_{i}$
-In case of missing binary variable, we used the same matrix of predictors but on a binary model estimated with the *glmer()* function from the *lme4* R package.
+After the imputation procedure, for the scenarios with incomple continuos outcome, we estimated the following mixed linear effect model using the *lmer()* function from the *lme4* R package.[@bates2022]
+$$\boldsymbol{y_{i}}=\boldsymbol{X_i^O}\boldsymbol{\beta_i^O}+\boldsymbol{\epsilon^O_{i}}$$
+with $\beta_i^O=\beta^O+b_i^O$, $b_i^O\sim N(0,\Psi^O)$ and $\epsilon^O_{i} \sim N(0,\Sigma_i)$.
+In case of an incomplete binary outcome, we used the same matrix of predictors but on a generalized linear mixed model estimated with the *glmer()* function from the *lme4* R package.
 Then, we pooled the estimates of the $\beta_i^O$ and the variance of the random effect and residual errors of the multiple imputed datasets according to Rubin's rule[@rubin1987], over which we calculated the performance measures on the estimands.
@@
 
-The Heckman 2L imputation method is available in the micemd R package [@audigier2021] (as \textbf{mice.2l.2stage.heckman()}) and also on the github repository \url{https://github.com/johamunoz/Statsmed_Heckman} where you can also find all the codes accompanying this paper and a toy example that explains how to implement the method in mice.
+The 2l.Heckman imputation method is available in the micemd R package [@audigier2021] (as \textbf{mice.2l.2stage.heckman()}) and also on the github repository \url{https://github.com/johamunoz/Statsmed_Heckman} where you can also find all the codes accompanying this paper and a toy example that explains how to implement the method in mice.
 
@@
 
-Figure \ref{fig:rhocont} shows the results of simulations where the missing variable was continuous.
-In the MAR scenario, i.e.
-when $\rho=0$, all imputation methods provided similar unbiased estimates of the coefficient parameters, but as rho increased, i.e.
-the mechanism became MNAR, the estimates for the complete case analysis and the MAR-IPD imputation method became biased.
-
  ```{r rhocont,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable by varing $\\rho$, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.9, fig.width = 6}
@@
  ```
-As it is expected both Heckman-based imputation (1l.
-and 2l.) models provided less biased estimates of the coefficient parameters in MNAR scenarios, but the random effects estimates on the 1l.heckman were far away from the true values as no cluster information was considered at all.
 
-Regarding coverage, the 2l.Heckman imputation method provided the best coverage values across all the $\rho$ scenarios with a coverage level close to the nominal 95\% interval.
-Even though 2l.Heckman was not properly a randomization-valid method [@buuren18c], as it was not unbiased and had a coverage above 95\% across all the $\rho$ values, it led to better results in terms of bias and coverage compared to the evaluated methods.
+Figure \ref{fig:rhocont} shows the results of simulations where the incomplete outcome was continuous.
+In the MAR scenario, i.e.
+when $\rho=0$, all imputation methods provided similar unbiased estimates of the coefficient parameters $\beta^O$ , but as $\rho$ increased, i.e.
+the mechanism became MNAR, the estimates for the complete-case analysis and the 2l.MAR imputation method became further away from the true value.
 
-The 2l.Heckman method, also resulted in better estimates in terms of RMSE than the other methods evaluated, via a better trade-off between bias and variance.
+As expected, both Heckman-based imputation methods (1l.Heckman and 2l.Heckman) gave less biased estimates of the coefficient parameters $\beta^O$ in MNAR scenarios, but the estimates of the random effects parameters $\sqrt{\psi^O_{kk}}$ with $k =\{0,1,2\}$ in the 1l.Heckman method were the most unbiased.
+This could be explained by the fact that 1l.Heckman does not take into account any grouping information.
 
-In particular, our method provided an advantage over the 1l.Heckman method in systematic missingness scenarios, by adding a cluster variable in the imputation model, a feature lacking from 1l.Heckman.
-This can be seen in the estimates of the random effects parameters of the 1l.Heckman method, which were more biased than those of the other methods in which cluster information was included, i.e.
+In terms of coverage, the 2l.Heckman imputation method gave the best coverage values in all $\rho$ scenarios with a coverage level close to the nominal 95\% value.
+Although 2l.Heckman was not properly a randomization-valid method [@buuren18c], as it was not unbiased and had a coverage above 95\% across all $\rho$ values, this method gave better results in terms of bias and coverage when compared to the other methods evaluated.
+The 2l.Heckman method, also resulted in better $\beta^O$ estimates in terms of RMSE than the other methods evaluated, through a better compromise between bias and variance.
+
+In particular, the 2l.Heckman method provides an advantage over the 1l.Heckman method on data sets with systematic omission, by including cluster information in the imputation model.
+This can be seen in the random effects parameter estimates $\sqrt{\psi^O_{kk}}$ of the 1l.Heckman method, which were more biased than those of the other methods where cluster information was included, i.e.
 2l.MAR and 2l.Heckman.
 
-Regarding the width of the 95\% CI of the estimated parameters, we observed that using the 2l.Heckman model we obtained larger CIs than when using other evaluated methods, which generally increased as $\rho$ moved away from zero.
-In particular, we observed in $\beta_2$ that the CI length was larger in the MAR scenario.
-By checking the simulation results further, we noted that the high variability came from one simulation in which the linear model estimation had singularity problems (not shown here).
+Regarding the width of the 95% confidence interval (CI) of the $\beta^O$ parameters, we observed that using the 2l.Heckman method we obtained the widest CIs of $\beta^O$ among the methods evaluated.
+The width of these CI generally increased as $\rho$ moved away from zero.
 
@@
 
-In the case of a missing bivariate variable, we also found (Figure \ref{fig:rhobin}) that the 2l.Heckman method provided the least unbiased results on the coefficient parameters.
-However, we observed more unbiased estimates and a larger 95th percentile amplitude in the estimates of the standard deviations of the random effects, especially in the $\sqrt{\psi_{22}}$ compared to the estimates obtained in the continuous incomplete outcome scenario.
+In the case of an incomplete bivariate outcome, we also found (Figure \ref{fig:rhobin}) that the 2l.Heckman method provided the least unbiased results on the coefficient parameters $\beta^O$ with the best coverage values.
+As for the random effect parameters, we observed unbiased estimates especially $\sqrt{\psi^O_{22}}$ at large values of $\rho$.
 
@@
 ### Sensitivity analysis: number of clusters and sample size of clusters
-We evaluated how robust our method was to variations in the number of clusters and also in the sample size of cluster (Figure \ref{fig:size}).
+We assessed the robustness of our method to variations in the number of clusters and also in the cluster sample size (Figure \ref{fig:size}).
 
-By increasing N, the number of clusters, from 10 to 100, we observed that the bias was not affected but the width of the 95\% CI of the estimates decreased (larger precision) and hence the RMSE decreased.
-On the other hand, by reducing the number of units per cluster (from $n_i$=1000 to $n_i$=100) the precision decreased for all coefficients, the bias on coefficient estimates were not drastically affected but the bias of random effects did.
+By increasing N, the number of clusters, from 10 to 100, we observed that the bias was not affected, but the 95\%CI width decreased (higher precision) and the RMSE decreased as well.
+On the other hand, as we reduced the number of units per cluster (from $n_i$=1000 to $n_i$=100) the precision decreased for all coefficients, the bias of the $\beta^O$ coefficient estimates was not drastically affected, but the bias of the $\sqrt{\psi_{kk}^O}$ random effects parameters was.
 
-When we reduced the sample size to 50 patients per study, the bias and RMSE of the $\sqrt{\psi_{22}}$ were drastically affected (not shown here but in the Appendix).
-This could be explained in part due to the scarce information on certain clusters, which affects directly the estimation of the Heckman model on those clusters.
+When we reduced the sample size to 50 patients per study, the bias and RMSE of the $sqrt{\psi_{22}^O}$ were drastically affected (not shown here, but in the Appendix).
+This could be partly explained by the sparse information on certain clusters, which directly affects the estimation of the Heckman model on these clusters.
 
@@
 When using the Heckman model in an explicit MNAR process, i.e.
-when the probability of loss is associated with the value of the missing variable, we observe that under all imputation methods we obtain biased estimators, with poor coverage.
-Particularly when estimating the intercept we observed more biased estimates compared to those obtained for the other estimated parameters $\beta_2$ with coverage below 60%.
+when the probability that an observation is missing is itself associated with the value of the observation, we observe that under all imputation methods we obtain biased estimators, with poor coverage.
+In particular, when estimating the parameter intercept $\beta^O_0$ we observe more biased estimates compared to the estimates of the other coefficient parameters, with a coverage of less than 60%.
 
-Our model may not be fully suitable for this type of scenario, especially if the main analysis is focused on estimating absolute prevalence estimates, as it is greatly affected by the bias of the intercept parameter ($\beta_0$) and the width of the 95th percentile CI of all coefficient parameters.
-Also the imputation method was affected in terms of the bias of the standard deviation of the random-effects parameters $\sqrt{\psi_{22}}$.
+It is possible that our model is not entirely suitable for this type of scenario, especially if the main analysis is focused on estimating absolute prevalence estimates, as it could be greatly affected by the bias of the ($\beta_0^O$) parameter.
+Also the imputation method was strongly affected in terms of the bias in the $\sqrt{\psi_{22}^O}$ parameter.
 
-With respect to the Skewed-t scenario, only the bias of $\beta_0$ was affected for all imputation methods applied.
-But $\beta_0$ estimates for both the 1l.Heckman and 2l.Heckman were drastically affected in terms of bias, with very poor coverage (below 25%).Therefore, the applicability of our method could be questionable in scenarios that do not conform to the BNV assumption.
+With respect to the Skewed-t scenario, only the bias of $\beta_0^O$ was affected for all imputation methods applied.
+But the estimates of $\beta_0^O$ for both 1l.Heckman and 2l.Heckman were drastically affected in terms of bias, with a very poor coverage (below 25%).Therefore, the applicability of our method might be questionable in scenarios that do not conform to the BNV assumption.
 
@@
 
-For this example, we used data accessed directly from ClinEpiDB[@staedke], where data were collected from 5195 households with verified consent, inhabited by 11137 residents aged 2-10 years.
+For this example, we used data accessed directly from ClinEpiDB[@staedke2021], where data were collected from 5195 households with verified consent, inhabited by 11137 residents aged 2-10 years.
 Blood samples were only taken from 8846 children, as 69 were excluded from the study due to lack of consent and 2222 were not present at the time of the survey.
@@
 \label{sec:discussion}
-We have extended and evaluated methods for multiple imputation of clustered datasets, in situations where some incomplete variables follow a MNAR mechanism.
-For clustered datasets, only imputation methods under the MAR mechanism had previously been proposed.
-Although there are imputation methods  that can handle MNAR they can only handle the case of individual studies.
-This puts limit to their use in situations common in IPD-MA such as when there is systematic missingness or when the proportion of missingness of a variable is very high in one of the included studies.
-To address this gap, we proposed a new multiple imputation method for continuous and binary MNAR covariates, that is specifically designed for clustered datasets.
-Our method, allows borrowing of information between the clusters to obtain more reliable imputation results at the individual cluster level.
+We have extended and evaluated methods for multiple imputation of clustered datasets, in situations where some incomplete variables follow an MNAR mechanism.
+Although there are imputation methods that address incomplete MNAR variables, these can only deal with the case of individual studies.
+This limits their use in common situations in IPD-MA, such as when there is systematic missingness or when the proportion of missingness of a variable is very high in one of the included studies.
+To address this gap, we proposed a new multiple imputation method for incomplete continuous and binary MNAR variables with sporadic as well as systematic missingness, which come from clustered datasets with heterogeneous effects and error variances.
 
-In our simulations we observed that the imputation method we proposed is optimal for the imputation of continuous and binary missing variables that follow a MNAR mechanism according to the Heckman model and that come from multilevel data, such data commonly used in the IPDMA.
+Empirically we observed that our proposed imputation method is well suited for the imputation of both continuous and binary incomplete variables following a MNAR mechanism according to Heckman's model.
+The simulation study showed that the application of the proposed method on clustered datasets with heterogeneous effects and error variances and with sporadically and systemically missing variables resulted in less biased estimates of effects with a convergence close to 95% compared to those obtained using other imputation methods evaluated.
 
-Overall, our new method produced unbiased estimates with convergence close to 95\% for the coefficient parameters.
-It also resulted in less biased estimates for the random effects parameters compared to the other methods evaluated.
+This method, in particular, could yield better results than Heckman-based imputation methods for individual studies, as it not only allows the imputation of missing values in clusters with systematic missigness, but can also reduce the values of individual clusters towards the overall study mean.
+This is especially advantageous in studies with small sample sizes, where an analysis approach that ignores data from other studies may lead to extreme effect estimates.
 
-Empirically, we showed that the proposed method could be applicable when there is systematic missigness at the cluster level.
-This method, in particular, could provide less biased imputation values compared with individual-level imputation methods, as it not only allows imputation of missing values in clusters with systematic missigness, but can also shrunk the values of individual clusters toward the overall study mean.
-This is especially advantageous in studies with small sample sizes, where an analysis approach that ignores data from other studies may lead to extreme estimates of prevalence.
-By using our proposed method, these would be reduced to the overall mean.
+The advantage of the proposed method over methods that assume MAR is that it allows the imputation of variables from cluster-level data following a MAR or MNAR mechanism according to Heckman's model.
+That is, under the specification of a valid exclusion variable the method determines which is the most adjustable inter-equation correlation parameter ($\rho$), or in general terms the missingness mechanism (MAR or MNAR), in each of the clusters evaluated.
 
-The advantage of the proposed method over methods that assume MAR is that it allows the imputation of variables from cluster level data following a MAR or MNAR mechanism according to the Heckman model.
-That is to say that under the specification of a valid exclusion variable the method determines which is the most adjustable  correlation parameter between equations ($\rho$), or in general terms the  missingness mechanism  (MAR or MNAR), in each of the clusters evaluated.
-
-Our implementation of the imputation method was built according to the specifications of the mice R package and is available in the micemd package, which allows, first of all, to be used both on the outcome and on the covariates.
-In addition, it offers the option of being used simultaneously with other imputation methods implemented in the MICE package, which is advantageous in databases containing missing variables with different prediction methods and models.
-Finally, the method can be used on systematically and sporadically missing clusters, both for continuous and binary missing variables with heterogeneous effects and error variances.
+Finally, our imputation method was built according to the specifications of the R mice package and is available in the micemd package.
+This allows the method to be easily and simultaneously used with other imputation methods implemented for the MICE package, which is advantageous in databases containing several incomplete variables that require different imputation methods and imputation models.
 
@@
 The latter might restrict the evaluation of the Heckman model at the cluster level to a certain number of predictors depending on the sample size of the cluster.
-<!-- This method also requires at least some of the studies to have no systematic missigness in the incomplete variable in question.
-That is, if all studies have the incomplete variable systematically missing, the method is unusable, which can be an important limitation in practice.
--->
 
@@
 This project has received funding from the European Union's Horizon 2020 research and innovation programme under ReCoDID grant agreement No 825746.
+In addition, we are grateful for the support of Professor Matthias Egger in conceptualising this article and helping us to find a suitable data set for the illustrative study.
 \end{minipage}

### 2023-01-24 2.Simulation/simplot.R

@@ -24,8 +24,8 @@ data[,Method:=factor(Method, levels = c("CC","1l.Heckman","2l.MAR","2l.Heckman")
 
 data[,Parameter := as.factor(param_val[data$Parameter])]
 data[,Parameter := factor(Parameter, levels = c("beta_0","beta_1","beta_2","psi_b0","psi_b1","psi_b2","psi_e"),
-                          ordered = TRUE, labels=c(expression(beta[0]),expression(beta[1]),expression(beta[2]),
-                                                   expression(sqrt(psi[oo])),expression(sqrt(psi[11])),expression(sqrt(psi[22])),
+                          ordered = TRUE, labels=c(expression(beta[0]**O),expression(beta[1]**O),expression(beta[2]**O),
+                                                   expression(sqrt(psi[oo]**O)),expression(sqrt(psi[11]**O)),expression(sqrt(psi[22]**O)),
                                                    expression(psi[e])))]
 
 
@@ -37,8 +37,8 @@ datalines <- data.table(Parameter = c("beta_0","beta_1", "beta_2","psi_b0", "psi
 
 
 datalines[,Parameter := factor(Parameter, levels = c("beta_0","beta_1","beta_2","psi_b0","psi_b1","psi_b2","psi_e"),
-                               ordered = TRUE, labels=c(expression(beta[0]),expression(beta[1]),expression(beta[2]),
-                                                        expression(sqrt(psi[oo])),expression(sqrt(psi[11])),expression(sqrt(psi[22])),
+                               ordered = TRUE, labels=c(expression(beta[0]**O),expression(beta[1]**O),expression(beta[2]**O),
+                                                        expression(sqrt(psi[oo]**O)),expression(sqrt(psi[11]**O)),expression(sqrt(psi[22]**O)),
                                                         expression(psi[e])))]
 
 

### 2023-11-01 1.Paper/Paper.Rmd
Update 1st revision

@@
 - name: Orestis Efthimiou
-  num: b
+  num: b,c
 - name: Vincent Audigier
-  num: c  
+  num: d  
 - name: Valentijn M. T. de Jong
-  num: a,d,f
+  num: a,e,g
 - name: Thomas. P. A. Debray
-  num: a,e,f
+  num: a,f,g
 address:
@@
 - num: b
-  org: University of Bern, Switzerland
+  org: Institute of Primary Health Care (BIHAM), University of Bern, Bern, Switzerland 
 - num: c
-  org: CNAM, Laboratoire CEDRIC-MSDMA, 2 rue Conte, 75003 Paris, France  
+  org: Institute of Social and Preventive Medicine (ISPM), University of Bern, Bern, Switzerland
 - num: d
+  org: Conservatoire national des arts et métiers (CNAM), Laboratoire CEDRIC-MSDMA, 2 rue Conte, 75003 Paris, France  
+- num: e
   org: Data Analytics and Methods Task Force, European Medicines Agency, Amsterdam, The Netherlands
-- num: e
+- num: f
   org: Smart Data Analysis and Statistics, Utrecht, The Netherlands 
-- num: f
+- num: g
   org: These authors contributed equally
 
-corres: "* Johanna Munoz. \\email{j.munozavila@umcutrecht.nl}"
+corres: "* Johanna Muñoz. \\email{j.munozavila@umcutrecht.nl}"
 presentaddress: Julius Center for Health Sciences and Primary Care, University Medical Center Utrecht, Universiteitsweg 100, 3584 CG Utrecht, the Netherlands
@@
 This method is based on the principles of Heckman selection models and adopts a two-stage meta-analysis approach to impute binary and continuous variables that may be outcomes or predictors and that are systematically or sporadically missing.
-After evaluating the proposed imputation model in simulated scenarios, we illustrate it use in a cross-sectional community survey to estimate the prevalence of malaria parasitemia in children aged 2-10 years in five subregions in Uganda."
+After evaluating the proposed imputation model in simulated scenarios, we illustrate it use in a cross-sectional community survey to estimate the prevalence of malaria parasitemia in children aged 2-10 years in five regions in Uganda."
 
@@
 bibliography: bibfile.bib
+
+graphics: yes
 output: 
@@
      \newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
-
 ---
@@
 # Introduction
-Over the past few years, data sharing efforts have substantially increased, and researchers increasingly often have access to IPD from large combined datasets derived from electronic health records (EHR) or from multiple randomized or observable trials (i.e.
+
+Over the past few years, data sharing efforts have substantially increased, and researchers increasingly often have access to IPD from large combined datasets derived from electronic health records (EHR) or from multiple randomised or observable trials (i.e.
 in IPD meta-analysis, IPD-MA).
-For example, the clinical practice research datalink (CRPD) [@herrett2015] is an electronic health record dataset in the UK, which has been used in a variety of medical research, such as the evaluation of health policy and drug efficacy.
+For example, the clinical practice research datalink (CRPD) [@herrett2015] is an EHR dataset in the UK, which has been used in a variety of medical research, such as the evaluation of health policy and drug efficacy.
 A recent example of an IPD-MA is the emerging risk factor collaboration[@theemergingriskfactorscollaboration2007], where data were combined from approximately 1.1 million individuals across 104 observational studies to investigate associations of cardiovascular diseases with several predictors.
@@
 
-<!-- Although it is possible to formally rule out MCAR[@little1988], it is impossible to assess whether data are MAR or MNAR as the observable data are not enough to test the assumptions of both mechanisms.[@enders2022]  For this reason, researchers often conveniently assume that data are MAR, or present results of a complete case analysis.
-In practice, however, unless missingness is artificially introduced by study design, e.g.
-when a test is only taken on patients with certain characteristics, missingness will often (at least partially) depend on unobserved information.
--->
-
-<!-- The missingness mechanism of a variable may be a mixture of MAR and MNAR, depending on the availability of data that may predict the missingness patterns.
-Therefore, the MAR mechanism may become more influential than MNAR when more information is recorded.
-For example, health professionals can decide which vital signs to measure on admission.
-When it is observed that low blood pressure (BP) measurements are not recorded, it could be assumed that this variable is MNAR, as the missigness of BP depends on its own value; however, BP could be considered as MAR variable if additional information is available to explain its missigness, such as the severity of the patient on admission.
--->
-
 Registries are notoriously prone to incomplete variables that are MNAR, due complex recording processes.
 [@liu2021] For example, laboratory tests are taken only in certain patients based on symptoms that are often incompletely recorded.
-Data from randomized trials may also suffer from MNAR, for example when study participants that experience unfavorable results drop out of the study.
+Data from randomised trials may also suffer from MNAR, for example when study participants that experience unfavourable results drop out of the study.
 Also, heterogeneity of the primary objective or resources of the studies may result in variables relevant to explain the missing process not being recorded at all in some of the studies.
@@
 Modelling data under MNAR mechanism implies to specify information about the missingness process in addition to assumptions about the observed data.
-Two major approach have been used to address MNAR mechanism: pattern mixture models[@little1996] and selection models [@vella1998].
+Two major approaches have been used to address MNAR mechanism: pattern mixture models[@little1996] and selection models [@vella1998].
 One of the most popular techniques within selection models is the one proposed by Heckman (1976).
-[@heckman1976] Briefly, the Heckman selection model corrects for selection bias by estimating two linked equations: a outcome equation, where the missing variable is associated with predictors, and a selection equation, that accounts for the inclusion of observations in the sample.
+[@heckman1976] Briefly, the Heckman selection model corrects for selection bias by estimating two linked equations: an outcome equation, where the missing variable is associated with predictors, and a selection equation, which accounts for the inclusion of observations in the sample.
 An important feature of the Heckman selection model is that it does not assume data to be MNAR, so that it can also be used when data are MCAR or MAR.
@@
 
-Therefore, the aim of this work is to develop a multilevel imputation method for continuous and binary variables that are both sporadically and systematically MNAR, and that this can be applied for an incomplete outcome or multiple incomplete predictors in the data sets.
+Therefore, the aim of this work is to develop a multilevel imputation method for continuous and binary variables that are both sporadically and systematically MNAR, which can be applied for an incomplete outcome or multiple incomplete predictors in the data sets.
 
@@
 In section \ref{sec:illustration} we illustrate the method using the survey information collected in different sub-districts in Uganda to estimate the prevalence of malaria in children.
-Finally, in section \ref{sec:discussion} we summarize the results, outline limitations and propose future extension of our method.
+Finally, in section \ref{sec:discussion} we summarize our results, outline limitations and propose future extensions of our method.
 
 # The Heckman model
-\label{sec:heckman}
-The Heckman selection model was initially proposed as a method to correct for selection bias, in which individuals are not randomly selected from the population, leading to inconsistent estimates and erroneous conclusions.
+
+\label{sec:heckman} The Heckman selection model was initially proposed as a method to correct for selection bias, i.e.
+which individuals are not randomly selected from the population, leading to inconsistent estimates and erroneous conclusions.
 [@heckman1976]
@@
 [@vella1998]
+
+```{=tex}
 \begin{figure}[h!]
@@
    \centering
-   \includegraphics[scale=0.35]{DAG_heckman.png}
-   \caption{DAG Heckman selection model: here the nodes (dotted lines = latent variables, continuous lines = observed variables) describe the relationship between $y^*_{ij}$ the latent response and $r^*_{ij}$ the latent selection variables of a $j$th unit in a $i$th cluster, that are dependable on $ \boldsymbol{x_{ij}^O}$ and $ \boldsymbol{x_{ij}^S}$ sets of predictors and are correlated through $ \boldsymbol{u_{ij}}$ unobservable variables.}
+   \includegraphics[scale=0.35]{Paper_files/figure-latex/DAG_heckman.png}
+   \caption{Grafical representation of the Heckman selection model: here the nodes (dotted lines = latent variables, continuous lines = observed variables) describe the relationship between $y^*_{ij}$ the latent response and $r^*_{ij}$ the latent selection variables of a $j$th unit in a $i$th cluster, that are dependable on $ \boldsymbol{x_{ij}^O}$ and $ \boldsymbol{x_{ij}^S}$ sets of predictors and are correlated through $ \boldsymbol{u_{ij}}$ unobservable variables.}
    \label{fig:DAG}
 \end{figure}
-This is visualized in Figure \ref{fig:DAG}, where for the $j$th individual or unit within the $i$th cluster, there is $y_{ij}^*$, a latent outcome variable, and $r_{ij}^*$, a latent selection variable, which are correlated through $\boldsymbol{u_{ij}}$ an unobserved or unrecorded variable, with $i\in[1,2,...,N]$ and $j\in[1,2,...,n_i]$.
+```
+This is visualised in Figure \ref{fig:DAG}, where for the $j$th individual or unit within the $i$th cluster, there is $y_{ij}^*$, a latent outcome variable, and $r_{ij}^*$, a latent selection variable, which are correlated through $\boldsymbol{u_{ij}}$ an unobserved or unrecorded variable, with $i\in[1,2,...,N]$ and $j\in[1,2,...,n_i]$.
 Here, both latent variables are related to the sets of predictor covariates $\boldsymbol{x_{ij}^O}$ and $\boldsymbol{x_{ij}^S}$.
@@
 
-Here $\boldsymbol{\beta^O_{i}}$ and $\boldsymbol{\beta^S_{i}}$ are $p\times1$ and $q\times1$ coefficient parameter vectors and $\boldsymbol{\epsilon^O_{i}}=(\epsilon^O_{i1},\epsilon^O_{i2},...,\epsilon^O_{in_i})^T$ and $\boldsymbol{\epsilon^S_{i}}=(\epsilon^S_{i1},\epsilon^S_{i2},...,\epsilon^S_{in_i})^T$ are the residual terms vectors for the outcome and selection equations, respectively.
+Here $\boldsymbol{\beta^O_{i}}$ and $\boldsymbol{\beta^S_{i}}$ are $p\times1$ and $q\times1$ coefficient vectors and $\boldsymbol{\epsilon^O_{i}}=(\epsilon^O_{i1},\epsilon^O_{i2},...,\epsilon^O_{in_i})^T$ and $\boldsymbol{\epsilon^S_{i}}=(\epsilon^S_{i1},\epsilon^S_{i2},...,\epsilon^S_{in_i})^T$ are the residual terms vectors for the outcome and selection equations, respectively.
 
@@
 
-In the presence of selection bias, the aforementioned outcome equation will yield biased estimates of $\boldsymbol{\beta^O_{i}}$ if no efforts are made to adjust for the non-representativeness of the observed $\boldsymbol{X^O_{i}}$ and $\boldsymbol{y_{i}}$ values.
+In the presence of selection bias, the aforementioned outcome equation will yield biassed estimates of $\boldsymbol{\beta^O_{i}}$ if no efforts are made to adjust for the non-representativeness of the observed $\boldsymbol{X^O_{i}}$ and $\boldsymbol{y_{i}}$ values.
 For this reason, the Heckman model aims to jointly estimate the outcome and selection equation by defining a relation between their respective error distributions.
-For instance, Heckman’s original model [@heckman1976] assumes that residual terms have a bivariate normal distribution (BVN),
+For instance, Heckman's original model [@heckman1976] assumes that residual terms have a bivariate normal distribution (BVN),
 
-\[\begin{pmatrix}
+$$\begin{pmatrix}
 \boldsymbol{\epsilon^O_{i}} \\
@@
 \end{pmatrix}\right)
-\]
+$$
 
@@
 
-Under the BVN assumption, the parameters of the Heckman model coefficients can be estimated using the two-step Heckman method (H2S) [@heckman1976] or the full information maximum likelihood method (FIML).[@amemiya1984] However, both methods may lead to inconsistent estimators when the true underlying distribution is not the BVN.
-[@gomes_etal20] To overcome this problem, other approaches have been proposed that relax the distribution assumptions, among them copula models.
+Under the BVN distribution, the parameters of the Heckman model can be estimated using the two-step Heckman method [@heckman1976] or the full information maximum likelihood method.[@amemiya1984] However, both methods may lead to inconsistent estimators when the true underlying distribution is not a BVN.
+[@gomes_etal20] To overcome this problem, other approaches have been proposed that relax the distribution assumptions; among them copula models.
 [@smith03] The copula approach uses a function, known as a copula, that joins the marginal distribution of the error terms of the selection and outcome equation which are specified separately.
-The Heckman model can be expressed with the following bivariate normal copula:
-$$ F(\boldsymbol{r^*_{i}},\boldsymbol{y^*_{i}}) =\Phi\left(\boldsymbol{r^*_{i}}-\boldsymbol{X^S_{i}}\boldsymbol{\beta^S_{i}},\frac{\boldsymbol{y^*_{i}}-\boldsymbol{X^O_{i}}\boldsymbol{\beta^O_{i}}}{\sigma_i}, \rho_i\right),$$
-where $\Phi$ is the bivariate standard normal cumulative distribution function.
-Thus, to estimate the parameters of the Heckman method, it is sufficient to specify the marginal distributions of the error terms, and link them with a suitable copula function.
-In our imputation method we estimate the Heckman model using the copula method for non-random selection [@wojtys2018] which is a flexible approach that allows the specification of different parametric distributions of the selection and outcome variables and also different types of dependence structure between the two equations.
+The cumulative joint distribution $F(\epsilon^S_{ij},\epsilon^O_{ij})$ of the error terms is given by: $$ F(\epsilon^S_{ij},\epsilon^O_{ij})=C(F_S(\epsilon^S_{ij}),F_{O}(\epsilon^O_{ij});\rho_i)$$ where $F_S(\epsilon^S_{ij})$ and $F_O(\epsilon^S_{ij})$ are the marginal distributions of the error terms for both equations and C is a copula function with with correlation parameter $\rho_i$.
+Thus, to estimate the parameters of the Heckman method, it is sufficient to specify the marginal distributions of the error terms and 'connect' them with a suitable copula function.
+In our imputation method, we estimate the Heckman model using a gaussian copula and employing the method for non-random selection [@wojtys2018], which is a flexible approach that allows the specification of different parametric distributions of the selection and outcome variables and different types of dependence structure between the two equations.
 
@@
 
- The Heckman model can be extended to hierarchical settings, i.e., in cases when individuals or sampling units are nested within clusters, as is the case in EHR or IPD.
+The Heckman model can be extended to hierarchical settings, i.e., in cases when individuals or sampling units are nested within clusters, as is the case in EHR or IPD meta-analysis.
 This hierarchical complexity must not only be taken into account in the analysis model, but also when dealing with missing data, thus requiring imputation models that are congenial to the analysis model, i.e.
@@
 
-Different procedures can be adopted to combine information between clusters; however, in our imputation method we opted for the two-stage approach that is often used in meta-analyses.[@simmonds2005] This is because such an approach is computationally less intensive and could potentially generate fewer convergence problems in the estimation of the Heckman hierarchical model compared to other approaches.
-Our method is based on the following conditional imputation model, whose parameters are $\boldsymbol{\theta}=\{\boldsymbol{\beta^O},\boldsymbol{\beta^S},\boldsymbol{\Psi^S},\boldsymbol{\Psi^O},(\sigma_i,\rho_i)_{1<=i<=N}\}$
-\begin{align*}
+Different procedures can be adopted to combine information across clusters; however, in our imputation method we opted for the two-stage approach that is often used in meta-analyses.[@simmonds2005] This is because such an approach is computationally less intensive and could potentially generate fewer convergence problems in the estimation of the Heckman hierarchical model compared to other one-stage methods.
+Our method is based on the following conditional imputation model, whose parameters are $\boldsymbol{\theta}=\{\boldsymbol{\beta^O},\boldsymbol{\beta^S},\boldsymbol{\Psi^S},\boldsymbol{\Psi^O},(\sigma_i,\rho_i)_{1<=i<=N}\}$ \begin{align*}
 \boldsymbol{y^*_{i}}&=\boldsymbol{X^O_{i}}(\boldsymbol{\beta^O}+b^O_{i})+\boldsymbol{\epsilon^O_{i}}\\
@@
 b^S_{i}&\sim N_q(0,\boldsymbol{\Psi^S})
-\end{align*}
-\[\begin{pmatrix}
+\end{align*} $$\begin{pmatrix}
 \boldsymbol{\epsilon^O_{i}} \\
@@
 \end{pmatrix}\right)
-\]
+$$
 
-Briefly, in a first stage  we estimate the cluster specific parameters of the Heckman model $\widehat{\boldsymbol{\theta_i}}=\boldsymbol{\{\widehat{\beta^O_i}}, \boldsymbol{\widehat{\beta^S_i}}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, for all $i$ between 1 and $N$ clusters.
-That is, we estimate separately for each of the clusters, the parameters of the two equations in each study, the outcome model (\ref{eq:1}) and the selection model (\ref{eq:2}) via a copula model.
+Briefly, at the first stage we estimate the cluster-specific parameters of the Heckman model $\widehat{\boldsymbol{\theta_i}}=\boldsymbol{\{\widehat{\beta^O_i}}, \boldsymbol{\widehat{\beta^S_i}}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, for clusters, i.e.
+for all $i$ in 1,2,...,$N$.
+That is, for each cluster, we estimate separately the parameters of the two equations, the outcome model (\ref{eq:1}) and the selection model (\ref{eq:2}) via a copula model.
 
-In a second stage, we estimate random effects multivariate meta-analysis model for the coefficient parameters.
-Here $\boldsymbol{\beta_i^O}$ and $\boldsymbol{\beta_i^S}$parameters are assumed to be drawn independently and identically from a latent multivariate normal distribution of parameters with a  population means $\boldsymbol{\beta^O}$ and $\boldsymbol{\beta^S}$ and population variance covariance $\boldsymbol{\psi^O}$ and $\boldsymbol{\psi^S}$ respectively.
-[@higgins2009] Here we assume that $\sigma_i$ and $\rho_i$ are random variables that come from a populational distribution and we estimate random effects univariate meta-analysis models for these parameters.
+At the second stage, we fit a random effects multivariate meta-analysis model for the coefficients.
+Here $\boldsymbol{\beta_i^O}$ and $\boldsymbol{\beta_i^S}$parameters are assumed to be drawn independently and identically from a latent multivariate normal distribution of parameters with across-cluster means $\boldsymbol{\beta^O}$ and $\boldsymbol{\beta^S}$ and across-cluster variance covariance $\boldsymbol{\psi^O}$ and $\boldsymbol{\psi^S}$ respectively.
+[@higgins2009] Here we assume that $\sigma_i$ and $\rho_i$ are random variables that follow a distribution across clusters, and we estimate random effects univariate meta-analysis models for these parameters.
 
 # Using the Heckman model to impute missing data
+
 \label{sec:methods}
-We follow a similar approach proposed by Resche-Rigon and White (2018) [@resche-rigon2018] for multilevel data imputation, which allows us to impute values in very common scenarios in IPD, e.g., sporadic and systematic missingness patterns.
-Our imputation method was created to impute datasets with a single missing outcome variable or covariate, but it can also be used to impute multiple incomplete variables in a dataset, as it can be implemented under a multivariate imputation by chain equations (MICE) approach.
+
+When dealing with MNAR data, it is necessary to make assumptions about the distribution of the missing data.
+Unfortunately, these assumptions cannot directly be verified from the observed data.
+Generally, MNAR imputation methods can be categorized into two distinct approaches: pattern mixture models and selection models, which rely on specific assumptions (please refer to the Appendix for details).
+Our imputation method is based in the Heckman selection model, which is estimated through a Gaussian copula model, and inherently embodies certain assumptions (as outlined in Table \ref{tab:assumptions}).
+
+```{r assumptions,  echo = FALSE, warning = FALSE,cache = TRUE}
+library(dplyr)
+library(kableExtra)
+assumptions <- data.frame(
+  c(" - The MNAR mechanism can be characterized as indirectly non-ignorable, i.e., the likelihood of a missing value in the incomplete variable is not determined by the variable's own value (i.e., self-masking), but instead depends on other unobserved variables that are also correlated with the incomplete variable.",
+     " - The error terms of both selection and outcome equations are connected by a Gaussian distribution and they are linearly related through a correlation parameter $\\rho$.",   
+    " -  An exclusion restriction variable, i.e. a predictor variable that appears in the selection equation but is not included in the outcome equation.",
+    " - Hierarchical Level-2 structure: cluster-specific parameters also follow a Gaussian underlying distribution and the coefficients in the outcome equation are uncorrelated with those in the selection equation."))
+
+assumptions %>%
+  kable( booktabs = T, escape = F,
+         caption = "Assumptions of the 2l.2stage.heckman imputation method",
+         col.names = linebreak(c("Imputation method assumptions"), align = "l"))%>%
+  column_spec(c(1), width = "15cm")%>%
+   kable_styling(latex_options = c("striped"))
+
+```
+
+We follow a similar approach proposed by Resche-Rigon and White (2018) [@resche-rigon2018] for multilevel data imputation, which allows us to impute values in very common scenarios in IPD meta-analysis, e.g., sporadic and systematic missingness patterns.
+Our imputation method was created to impute datasets with a single missing outcome variable or covariate, but it can also be used to impute multiple incomplete variables in a dataset, as it can be implemented under a multivariate imputation by chained equations (MICE) approach.
 
@@
 In the latter case, in the outcome equation (\ref{eq:1}), the incomplete predictor variable must be specified as the dependent variable and $\boldsymbol{X^O_{i}}$ as the set of covariates associated with it.
-On the other hand, in the selection model(\ref{eq:2}), $\boldsymbol{r^*_{i}}$ corresponds to the latent variable of selection of the incomplete predictor variable together with its $\boldsymbol{X^S_{i}}$ associated set of predictors.
+On the other hand, in the selection equation(\ref{eq:2}), $\boldsymbol{r^*_{i}}$ corresponds to the latent variable of selection of the incomplete predictor variable together with its $\boldsymbol{X^S_{i}}$ associated set of predictors.
 
 ## Univariate imputation
-Given an outcome variable $\boldsymbol{y}=(y_1,y_2,...,y_N)^T$, that consists of $y^{miss}_{ij}$ missing and $y^{obs}_{ij}$ observable values, we generate independent draws from the posterior predictive distribution for the missing data, $y_{ij}^{miss}$, given the observable data information $y_{ij}^{obs}$.
-$$p(y_{ij}^{miss}|y_{ij}^{obs})=\int_{\boldsymbol{\theta}} p(y_{ij}^{miss}|\boldsymbol{\theta},y_{ij}^{obs})p(\boldsymbol{\theta}|y_{ij}^{obs})\mathrm{d}\boldsymbol{\theta}$$
-Here we implicitly assume vague prior distributions for each of the parameters included in the parameter vector $\boldsymbol{\theta}$.
+
+Given an outcome variable $\boldsymbol{y}=(y_1,y_2,...,y_N)^T$, which consists of $y^{miss}_{ij}$ missing and $y^{obs}_{ij}$ observable values, we generate independent draws from the posterior predictive distribution for the missing data, $y_{ij}^{miss}$, given the observable data information $y_{ij}^{obs}$.
+$$p(y_{ij}^{miss}|y_{ij}^{obs})=\int_{\boldsymbol{\theta}} p(y_{ij}^{miss}|\boldsymbol{\theta},y_{ij}^{obs})p(\boldsymbol{\theta}|y_{ij}^{obs})\mathrm{d}\boldsymbol{\theta}$$ Here we implicitly assume vague prior distributions for each of the parameters included in the parameter vector $\boldsymbol{\theta}$.
 Because the integration can be performed computationally by sampling from the posterior predictive distribution $p(\boldsymbol{\theta}|y_{ij}^{obs})$, our imputation method can be carried out in the following two steps:
@@
 
-#### Fit $p(y_{ij}^{obs}|\boldsymbol{\theta_i})$, the heckman selection model at cluster level
+#### Fit $p(y_{ij}^{obs}|\boldsymbol{\theta_i})$, the heckman selection model at the cluster level
 
-Initially, we use the copula method to estimate the set of cluster-specific parameters, $\widehat{\boldsymbol{\theta_i}}=\boldsymbol{\{\widehat{\beta^O_i}}, \boldsymbol{\widehat{\beta^S_i}}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, using all $j$ units with observable measurements $y_{ij}^{obs}$ within each cluster $i$.
+Initially, we use the copula method to estimate the set of cluster-specific parameters, $\widehat{\boldsymbol{\theta_i}}=\boldsymbol{\{\widehat{\beta^O_i}}, \boldsymbol{\widehat{\beta^S_i}}, \widehat{\sigma_i}, \widehat{\rho_i}\}$, using all $j$ units with observable measurements $y_{ij}^{obs}$ within cluster $i$.
 Here we obtain the estimates not only the parameters' point estimates $\widehat{\boldsymbol{\theta_i}}$, but also their corresponding $\boldsymbol{\widehat{S(\theta_i)}}$ within-cluster variance-covariance matrix.
@@
 In this step, we pool the parameters $\widehat{\boldsymbol{\theta_i}}$ with a random effects meta-analysis model using only the clusters with observable information, i.e., those with no systematically missing outcome.
-In particular, we pool the $p$ coefficients of the $\boldsymbol{\beta^O}$ in the outcome equation and estimate a multivariate random effects meta-analysis model with them.
-Similarly we combine all $q$ coefficient parameters of the $\boldsymbol{\beta^S}$ in the selection equation.
-Thus, we can denote the study specific coefficients
-$$\boldsymbol{\widehat{\beta^O_i}} = \boldsymbol{\beta^O} +\boldsymbol{ b^O_i} + \boldsymbol{\epsilon'^O_{i}}$$
-$$\boldsymbol{\widehat{\beta^S_i}} = \boldsymbol{\beta^S} + b^S_i + \boldsymbol{\epsilon'^S_{i}}$$
-using the random effects 
- $b^O_i \sim N(0,\boldsymbol{\Psi^O})$ and $b^S_i\sim N(0,\boldsymbol{\Psi^S})$ with sampling errors $\epsilon'^O_{i}$ and $\epsilon'^S_{i}$.
+In particular, we pool the $p$ coefficients of $\boldsymbol{\beta^O}$ in the outcome equation and fit a multivariate random effects meta-analysis model with them.
+Similarly we combine all $q$ coefficients of $\boldsymbol{\beta^S}$ in the selection equation.
+Thus, we can denote the study specific coefficients $$\boldsymbol{\widehat{\beta^O_i}} = \boldsymbol{\beta^O} +\boldsymbol{ b^O_i} + \boldsymbol{\epsilon'^O_{i}}$$ $$\boldsymbol{\widehat{\beta^S_i}} = \boldsymbol{\beta^S} + b^S_i + \boldsymbol{\epsilon'^S_{i}}$$ using the random effects $b^O_i \sim N(0,\boldsymbol{\Psi^O})$ and $b^S_i\sim N(0,\boldsymbol{\Psi^S})$ with sampling errors $\epsilon'^O_{i}$ and $\epsilon'^S_{i}$.
 
-We assume that  $\sigma_i$ and $\rho_i$ are random variables coming from a population distribution, so we can express them as:
-$$ log(\sigma_i) \sim N( log(\sigma),\psi^{\sigma})$$
-$$ tanh^{-1}(\rho_i)\sim N( tanh^{-1}(\rho),\psi^{\rho})$$ 
-For each of them, we perform a univariate random effects meta-analysis. The model for $\sigma_i$ is given by the model:
+We assume that $\sigma_i$ and $\rho_i$ are random variables coming from an across-cluster distribution, so we can express them as: $$ log(\sigma_i) \sim N( log(\sigma),\psi^{\sigma})$$ $$ tanh^{-1}(\rho_i)\sim N( tanh^{-1}(\rho),\psi^{\rho})$$ For each of them, we perform a univariate random effects meta-analysis.
+The model for $\sigma_i$ is given by the model:
 
@@
 
-The model for $\rho_i$ is given by:
-$$tanh^{-1}(\widehat{\rho_i}) = \rho + b^{\rho}_i + \epsilon^{\rho}_{i}$$
-with $b_i^{\rho} \sim N(0,\psi^{\rho})$, and $\epsilon_i^{\rho} \sim N(0,var(tanh^{-1}(\widehat{\rho_i})))$.
+The model for $\rho_i$ is given by: $$tanh^{-1}(\widehat{\rho_i}) = \rho + b^{\rho}_i + \epsilon^{\rho}_{i}$$ with $b_i^{\rho} \sim N(0,\psi^{\rho})$, and $\epsilon_i^{\rho} \sim N(0,var(tanh^{-1}(\widehat{\rho_i})))$.
 
@@
 
-From the meta-analysis model, we obtain the marginal estimates
-$\widehat{\boldsymbol{\Theta}}=\boldsymbol{\{\widehat{\beta^O}}, \boldsymbol{\widehat{\beta^S}}, \widehat{\sigma}, \widehat{\rho}\}$
-and the between-cluster variance matrix
-$\widehat{\boldsymbol{\psi}}=\boldsymbol{\{\widehat{\Psi^O}}, \boldsymbol{\widehat{\Psi^S}}, \widehat{\psi^{\sigma}}, \widehat{\psi^{\rho}}\}$, i.e.
+From the meta-analysis model, we obtain the marginal estimates $\widehat{\boldsymbol{\Theta}}=\boldsymbol{\{\widehat{\beta^O}}, \boldsymbol{\widehat{\beta^S}}, \widehat{\sigma}, \widehat{\rho}\}$ and the between-cluster variance matrix $\widehat{\boldsymbol{\psi}}=\boldsymbol{\{\widehat{\Psi^O}}, \boldsymbol{\widehat{\Psi^S}}, \widehat{\psi^{\sigma}}, \widehat{\psi^{\rho}}\}$, i.e.
 variance-covariance matrix of the random effects, with their corresponding variance-covariance matrices $\boldsymbol{\widehat{S_{\Theta}}}$ and $\boldsymbol{\widehat{S_{\psi}}}$, which are used to draw the $\boldsymbol{\Theta^*}$ and $\boldsymbol{\psi^*}$ parameters, from their posterior distribution as follows: [@resche-rigon2013]
 
+```{=tex}
 \begin{align*}
@@
 \end{align*}
-
+```
 #### Draw the cluster parameters $\boldsymbol{\theta^*_i}$
@@
 
+```{=tex}
 \begin{align*}
@@
 \end{align*}
-
-As can be seen, the mean and variance of the posterior distribution is a combination between the estimated marginal and cluster-specific parameters.
+```
+As can be seen, the mean and variance of the posterior distribution is a combination of the estimated marginal and cluster-specific parameters.
 Here the weights on the cluster-specific parameters $\boldsymbol{\widehat{\theta_i}}$ and the marginal parameters $\boldsymbol{\Theta^*}$ are inversely proportional to the within cluster variance $\boldsymbol{\widehat{S_{\theta_i}}}$ and between clusters variance $\boldsymbol{\psi^*}$.
@@
 Conversely, when $\boldsymbol{\widehat{S_{\theta_i}}}>\boldsymbol{\psi^*}$, more weight is given to the estimated marginal parameters.
-Therefore, in case of a cluster with systematic missingness, it is as if the within-cluster variance is infinite ($\boldsymbol{\widehat{S_{\theta_i}}}\to\infty$), so all the weight is assigned to the parameters estimated at the marginal level.
+Therefore, in case of a cluster with systematic missingness, the within-cluster variance is effectively infinite ($\boldsymbol{\widehat{S_{\theta_i}}}\to\infty$), so that all the weight is assigned to the parameters estimated at the marginal level.
 That is, when there is no information in a cluster, we rely entirely on the marginal information.
@@
 
-Having $\boldsymbol{\theta^*_i}$, the shrunk-cluster parameters vector for each cluster, we back-transform $\sigma^*$ and $\rho^*$ to the original scale.
+Having estimated $\boldsymbol{\theta^*_i}$, the shrunk-cluster parameters vector for each cluster, we back-transform $\sigma^*$ and $\rho^*$ to the original scale.
 Then $y^{miss}_{ij}$, the missing values, can be drawn from $p(y^{miss}_{ij}|\boldsymbol{\theta^*_i})$, their predictive distribution given $\boldsymbol{\theta^*_i}$, as follows:
@@
 
-The imputed value of $y_{ij}^{miss}$ can be drawn from the conditional expectation of $y_{ij}$ on unobserved measurements: [@greene2018]
-\begin{align*}
+The imputed value of $y_{ij}^{miss}$ can be drawn from the conditional expectation of $y_{ij}$ on unobserved measurements: [@greene2018] \begin{align*}
 \mu&= E[y_{ij}|r_{ij}=0,{\boldsymbol{\beta^O_{i}}^*},\boldsymbol{{\beta^{S}_{i}}^*},{\rho_{i}}^*,{\sigma_i}^*]\\
@@
 When $y_{ij}^{miss}$ is a binary variable, the imputed value is drawn from a Bernoulli distribution with a proportion parameter $p_{ij}^*$ given by $P[y_{ij}=1|r_{ij}=0]$, that is the conditional probability that $y_{ij}=1$ given that the measure is unobservable ($r_{ij}=0$).
-The $p_{ij}^*$ is obtained from a bivariate probit model, as follows: [@greene2018]
-\begin{align*}
+The $p_{ij}^*$ is obtained from a bivariate probit model, as follows: [@greene2018] \begin{align*}
 p^*_{ij} &= P[y_{ij}=1|r_{ij}=0,{\boldsymbol{\beta^O_{i}}^*},{\boldsymbol{\beta^{S}_{i}}^*},{\rho_{i}}^*]\\
@@
 y_{ij}^{miss}&\sim Ber(p^*_{ij})
-\end{align*}
-where  $\Phi_2(.)$ corresponds to the bivariate normal cumulative distribution function.
+\end{align*} where $\Phi_2(.)$ corresponds to the bivariate normal cumulative distribution function.
 
@@
 
-Briefly, MICE (multiple imputation of chained equations) was built under the fully conditional specification framework, where for each incomplete variable a conditional imputation model is specified based on another variables in the dataset.
+Briefly, MICE (multiple imputation of chained equations) was built under the fully conditional specification framework, where for each incomplete variable a conditional imputation model is specified based on other variables in the dataset.
 This process is carried out iteratively, so that in each iteration the missing values of an incomplete variable are drawn from the conditional distribution based on the updated variables in the previous iteration.
 
-Our imputation model can then be used in the imputation of any incomplete variable in a dataset following an MNAR mechanism, even if it is an outcome, predictor or auxiliary variable of the main analysis model.
+Our imputation model can then be used in the imputation of any incomplete variable in a dataset following an MNAR mechanism, even if it is an outcome or predictor of the main analysis model, of for auxiliary variables.
 Furthermore, as implemented according to the MICE framework, the imputation method could be used simultaneously with other imputation methods available in MICE or in add-in packages such as micemd [@audigier2021], to impute datasets with multiple incomplete variables that differ in type and missing mechanism.
@@
 ## Technical details of the implementation
-The Heckman model is estimated with the \textbf{gjrm} function of the GJRM R package[@radiceGJRMGeneralisedJoint2021],  under the bivariate model with the nonrandom sample selection (BSS) specification.
-The meta-analysis model is performed with the \textbf{mixmeta} function of the R package mixmeta[@gasparrini2021], which allows the use of maximum likelihood (ML), restricted maximum likelihood (REML), and moments estimation methods.
-For the simulation and illustrative study, we use the restricted REML estimation method, which is recommended as it has a good balance between unbiasedness and efficiency.
+
+The Heckman model is estimated with the \textbf{gjrm} function of the GJRM R package[@radiceGJRMGeneralisedJoint2021], under the bivariate sample selection model (BSS) specification with a bivariate normal error distribution (BivD="N").
+The meta-analysis model is estimated with the \textbf{mixmeta} function of the R package mixmeta[@gasparrini2021], which allows the use of maximum likelihood (ML), restricted maximum likelihood (REML), and moments estimation methods.
+For the simulation and illustrative study, we use the restricted REML estimation method, which is recommended as it has a good balance between unbiassedness and efficiency.
 [@viechtbauer2005].
@@
 # Simulation study
+
 \label{sec:simulation}
@@
 ## Aim
+
 We designed a simulation study aimed to compare the performance of alternative methods for imputing a single missing outcome variable in a hierarchical dataset, where the missingness follows a MNAR mechanism.
@@
 We generated the data from a Heckman selection model with bivariate normal distribution error terms.
-For simplicity we started from "basic scenario" where the database collected information from $N=10$ clusters of $n_i=1000$  individuals.
-However, we altered both $N$ and $n_i$ in sensitivity analyses (Table \ref{tab:scenarios}).
+For simplicity we started from a "basic scenario",i.e.
+where the database collected information from $N=10$ clusters of $n_i=1000$ individuals.
+Subsequently, we altered both $N$ and $n_i$ in additional analyses (Table \ref{tab:scenarios}).
+
 ```{r scenarios,  echo = FALSE, warning = FALSE,cache = TRUE}
@@
 Size and cluster number/	Continuous/	0.6/	10;50,   10;100,   10;1000,  50;1000,  100;1000/	Heckman (BVN)
-Distribution deviations/	Continuous/	0.6/	10;1000	/Heckman (BVN), Heckman (t-skew), Explicit'
+Distribution deviations/	Continuous/	0.6/	10;1000	/Heckman (BVN), Heckman (t-skew), Self-masking'
 )
 options(kableExtra.latex.load_packages = FALSE) 
+library(gamm4)
 library(kableExtra)
@@
 ```
-For each dataset, we generated $X_{1i}$ a treatment indicator variable from a Bernoulli distribution with a probability of treatment on each cluster equal to 0.6.
+
+For each dataset, we generated $X_{1i}$, a treatment indicator variable, from a Bernoulli distribution with a probability of treatment on each cluster equal to 0.6.
 Next, we simulated the mean of two continuous covariates from a multivariate normal distribution $\begin{psmallmatrix}\mu_2\\\mu_3\end{psmallmatrix}\sim N\left(\begin{psmallmatrix}0\\0\end{psmallmatrix},\begin{psmallmatrix}0.2 & 0.015\\0.015 & 0.2\end{psmallmatrix}\right)$.
@@
 For the selection equation we included both variables and the $X_{3i}$ exclusion restriction variable, $\boldsymbol{X_i^S}=[1,\boldsymbol{X_{1i}},\boldsymbol{X_{2i}},\boldsymbol{X_{3i}}]$.
-Then in case of a missing continuous variable, we calculate the latent variables  $\boldsymbol{y_{i}^*}$ and $\boldsymbol{r^*_{i}}$ as follows:
-\begin{align*}
+Then in case of a missing continuous variable, we calculate the latent variables $\boldsymbol{y_{i}^*}$ and $\boldsymbol{r^*_{i}}$ as follows: \begin{align*}
  \boldsymbol{y_{i}^*}&= \boldsymbol{X_i^O}\boldsymbol{\beta_i^O}+\boldsymbol{\epsilon^O_{i}} \\
  \boldsymbol{r_{i}^*}&= \boldsymbol{X_i^O}\boldsymbol{\beta_i^S}+\boldsymbol{\epsilon^S_{i}}
-\end{align*}
-where $\boldsymbol{\beta_i^O}=(\beta_{i0}^O,\beta_{i1}^O,\beta_{i2}^O)^T$ and $\boldsymbol{\beta_i^S}=(\beta_{i0}^S,\beta_{i1}^S,\beta_{i2}^S,\beta_{i3}^S)^T$.
-Here we assumed that the coefficient parameter k varied across studies, by including cluster-specific random effects as:
-\begin{align*}
- \beta_{ik}^O&= \beta_k^O + b_{ik}^O\\
- \beta_{ik}^S&= \beta_k^S + b_{ik}^S
-\end{align*}
-Denoting $\boldsymbol{\beta^O}$=$(\beta_{0}^O,\beta_{1}^O,\beta_{2}^O)^T$ and $\boldsymbol{\beta^S}$=$(\beta_{0}^S,\beta_{1}^S,\beta_{2}^S,\beta_{3}^S)^T$, we fixed coefficients $\boldsymbol{\beta^O}$=$(0.3,1,1)$  and $\boldsymbol{\beta^S}$=$(-0.8,1.3,-0.7,1.2)$ in order to get around 40\% of sporadically missing values on the response $y_{ij}$ in the entire data set.
+\end{align*} where $\boldsymbol{\beta_i^O}=(\beta_{i0}^O,\beta_{i1}^O,\beta_{i2}^O)^T$ and $\boldsymbol{\beta_i^S}=(\beta_{i0}^S,\beta_{i1}^S,\beta_{i2}^S,\beta_{i3}^S)^T$.
+Here we assumed that the coefficient k varied across studies, by including cluster-specific random effects as: \begin{align*}
+ \beta_{ik}^O&= \beta_k^O + b_{ik}^O, k=0,1,2\\
+ \beta_{ik}^S&= \beta_k^S + b_{ik}^S, k=0,1,2,3
+\end{align*} Denoting $\boldsymbol{\beta^O}$=$(\beta_{0}^O,\beta_{1}^O,\beta_{2}^O)^T$ and $\boldsymbol{\beta^S}$=$(\beta_{0}^S,\beta_{1}^S,\beta_{2}^S,\beta_{3}^S)^T$, we fixed coefficients $\boldsymbol{\beta^O}$=$(0.3,1,1)$ and $\boldsymbol{\beta^S}$=$(-0.8,1.3,-0.7,1.2)$ in order to get around 40% of sporadically missing values on the response $y_{ij}$ in the entire data set.
 Additionally, we ensured that the $y_{ij}^*$ observations were systematically missing in 20% of the clusters included in the data set, by removing the outcome values in the 20% of the clusters.
 
-We assumed that random effects were independent within equations ($b_{i0}^O\indep b_{i1}^O \indep b_{i2}^O$ and $b_{i0}^S \indep b_{i1}^S \indep b_{i2}^S\indep b_{i3}^S$), but were linked between both selection and outcome equations through a bivariate normal distributed as:
-\[\begin{pmatrix}
+We assumed that random effects were independent within equations ($b_{i0}^O\indep b_{i1}^O \indep b_{i2}^O$ and $b_{i0}^S \indep b_{i1}^S \indep b_{i2}^S\indep b_{i3}^S$), but were linked between both selection and outcome equations through a bivariate normal distributed as: $$\begin{pmatrix}
 b_{ik}^O \\
@@
 \end{pmatrix}\right)
-\]
-with ${\psi^S_{00}}={\psi^O_{00}}={\psi^S_{11}}={\psi^O_{11}}={\psi^S_{22}}={\psi^O_{22}}=0.4$ and $\rho_b=\rho*0.4$.
+$$ with $k=0,1,2$, ${\psi^S_{00}}={\psi^O_{00}}={\psi^S_{11}}={\psi^O_{11}}={\psi^S_{22}}={\psi^O_{22}}=0.4$ and $\rho_b=\rho*0.4$.
 We considered that the correlation parameter of the random effects between equations is 40% of the value of the assumed correlation parameter between error terms $\rho$.
@@
 
-As regards the error terms, they were bivariate normal distributed as:
-\[\begin{pmatrix}
+As regards the error terms, they were bivariate normal distributed as: $$\begin{pmatrix}
 \epsilon^O_{i} \\
@@
 \end{pmatrix}\right)
-\]
-whose $\sigma_i^2$ is a random parameter across clusters distributed as $log(\sigma_i) \sim N(0,0.05)$.
+$$ whose $\sigma_i^2$ is a random parameter across clusters distributed as $log(\sigma_i) \sim N(0,0.05)$.
 
-## Adittional scenarios
+## Additional scenarios
 
@@
 -   \textbf{M(N)AR scenarios:} We explored a missing MAR mechanism ($\rho=0$), a scenario when data followed a MNAR mechanism with a low ($\rho=0.3$), intermediate ($\rho=0.6$) and strong correlation ($\rho=0.9$) between $y^*$ and $r^*$.
-We also explored a scenario when the missing variable is binary.
-Therefore we simulated a $y_{i}$ binary incomplete variable, by keeping similar parameters to the ones used in the simulation of a missing continuous variable, but here we defined the observable binary variable as:
-\begin{align*}
+We also explored a scenario when the missing variable was binary.
+Therefore we simulated $y_{i}$ as a binary incomplete variable, by keeping similar parameters to the ones used in the simulation of a missing continuous variable, but here after defining the observed binary variable as: \begin{align*}
       r_{i}&=I(r*_{i}>0)\\
@@
 
-- \textbf{Violation of distributional assumptions:} Here, we aimed to investigate how the imputation models behave in settings with departures from the bivariate normal distribution, performed two sensitivity analyses.
+-   \textbf{Violation of distributional assumptions:} Here, we aimed to investigate how the imputation models behave in settings with departures from the bivariate normal distribution, performing two sensitivity analyses.
+
+    ```{=tex}
     \begin{enumerate}[wide=10pt]
       \item \textbf{Skewed-t:} We drew error terms from a bivariate skewed student-t distribution using the same location parameter and covariance matrix of the normal distributed settings,  with 4 degrees of freedom and an $\alpha=\{-2,6\}$ parameter which regulates the asymmetry of the density.
-    \item  \textbf{Explicit:} In addition we simulated an explicit missingness process, where error terms of the selection and outcome equations were independently normal distributed and the selection of observations depended on the value of the outcome variable, as $ry^*_i=0.3y^*_i+\epsilon^S_{i}$.
-These settings assure that the percentage of missingness on the outcome variable is around 60\% in all the evaluated scenarios.
+      \item  \textbf{Self-masking:} In addition we simulated a self-masking missingness process, where error terms of the selection and outcome equations were independently normal distributed and the selection of observations depended on the value of the outcome variable, as $r^*_i=0.3y^*_i+\epsilon^S_{i}$.
+These settings led to around 60\% missingness of the outcome variable across all the evaluated scenarios.
       \item \textbf{Normal:} As a reference, we provide the results from the basic scenario of the main simulation study, where the error terms were Bivariate normal distributed. 
     \end{enumerate}
+    ```
 
 ## Estimands
+
 The estimands were the parameter coefficients of the outcome equation $\boldsymbol{\beta^O}=(\beta_{0}^O,\beta_{1}^O,\beta_{2}^O)^T$, with special emphasis on the treatment effect parameter $\beta^0_1$.
@@
 
-After the imputation procedure, for the scenarios with incomple continuos outcome, we estimated the following mixed linear effect model using the *lmer()* function from the *lme4* R package.[@bates2022]
-$$\boldsymbol{y_{i}}=\boldsymbol{X_i^O}\boldsymbol{\beta_i^O}+\boldsymbol{\epsilon^O_{i}}$$
-with $\beta_i^O=\beta^O+b_i^O$, $b_i^O\sim N(0,\Psi^O)$ and $\epsilon^O_{i} \sim N(0,\Sigma_i)$.
-In case of an incomplete binary outcome, we used the same matrix of predictors but on a generalized linear mixed model estimated with the *glmer()* function from the *lme4* R package.
+After the imputation procedure with incomplete continuous outcome, we estimated the following mixed linear effect model using the *lmer()* function from the *lme4* R package.[@bates2022] $$\boldsymbol{y_{i}}=\boldsymbol{X_i^O}\boldsymbol{\beta_i^O}+\boldsymbol{\epsilon^O_{i}}$$ with $\beta_i^O=\beta^O+b_i^O$, $b_i^O\sim N(0,\Psi^O)$ and $\epsilon^O_{i} \sim N(0,\Sigma_i)$.
+In case of an incomplete binary outcome, we used the same matrix of predictors but we fit on a generalized linear mixed model with the *glmer()* function from the *lme4* R package.
 Then, we pooled the estimates of the $\beta_i^O$ and the variance of the random effect and residual errors of the multiple imputed datasets according to Rubin's rule[@rubin1987], over which we calculated the performance measures on the estimands.
 
-To calculate the coverage of the parameter coefficients' 95\% confidence intervals (CI), we estimate CI with the Wald method.
-<!-- Although it is possible to obtain CI through the profile or bootstrap method for the variances of the random effects, we prefer not to estimate them (and hence the coverage) for the random effects parameters due to computational time.
--->
+To calculate the coverage of the parameter coefficients' 95% confidence intervals (CI), we estimate CI with the Wald method.
 
@@
 -   \textbf{Complete case analysis (CCA):} We removed all patients with missing observations.
-- \textbf{1l.Heckman:} Multiple imputation based on the Heckman model without no study specification, following the imputation method proposed by Galimard et al.(2016). [@galimard_etal16]
-- \textbf{2l.MAR}: Multiple imputation assuming MAR for hierarchical datasets, we used the multilevel imputation model ( 2l.2stage.norm and 2l.2stage.bin) from the micemd R package[@audigier2021], which are described in Audigier et al.
-(2018) paper.
-[@audigier_etal18]
-
+-   \textbf{1l.Heckman:} Multiple imputation based on the Heckman model for 1-level incomplete variable,i.e.
+with no study specification, following the imputation method proposed by Galimard et al.(2016).
+[@galimard_etal16]
+-   \textbf{2l.MAR}: Multiple imputation assuming MAR with 2-level hierarchical datasets.
+We used the multilevel imputation model ( 2l.2stage.norm and 2l.2stage.bin) from the micemd R package[@audigier2021], which are described by Audigier et al.
+(2018) [@audigier_etal18]
 -   \textbf{2l.Heckman}: The proposed imputation method based on the Heckman model for hierarchical datasets.
@@
 -   \textbf{Bias:} Bias on the coefficient and variance-covariance matrix terms.
-- \textbf{Coverage:} Coverage of the 95\% confidence intervals for the coefficient  parameters.
-- \textbf{Width}: Average width of the confidence interval on the coefficient parameters.
+-   \textbf{Coverage:} Coverage of the 95% confidence intervals for the coefficients.
+-   \textbf{Width}: Average width of the confidence interval on the coefficients.
 -   \textbf{RMSE:} Root mean squared error of the coefficient and random effect parameters.
 
-In addition, in the appendix table we reported the empirical standard errors (EmpSE), Monte Carlo standard errors (ModSE) on the coefficient  parameters, average processing time (time in seconds) and the percentage of datasets where the imputation method converged (run), i.e., the imputation method generated an output.
+In addition, in the appendix table we reported the empirical standard errors (EmpSE), Monte Carlo standard errors (ModSE) on the coefficients, average processing time (time in seconds) and the percentage of datasets where the imputation method converged (run), i.e., the imputation method generated an output.
 
 ## Software
+
 For the simulation study and illustrative examples we used R version 4.0.4 in a linux environment. [@Rpackage]
 
-The 2l.Heckman imputation method is available in the micemd R package [@audigier2021] (as \textbf{mice.2l.2stage.heckman()}) and also on the github repository \url{https://github.com/johamunoz/Statsmed_Heckman} where you can also find all the codes accompanying this paper and a toy example that explains how to implement the method in mice.
+The 2l.2stage.Heckman imputation method is available in the micemd R package [@audigier2021] (as \textbf{mice.2l.2stage.heckman()}) and also on the github repository \url{https://github.com/johamunoz/Statsmed_Heckman} where we also provide all codes accompanying this paper as well as a toy example that explains how to implement the method in mice.
 
@@
 
-<!-- ### Descriptive statistics of the simulated datasets -->
-<!-- We generated data sets of 10 clusters of 1000 patients each in each scenario.
-For example, for the base scenario (Table \ref{tab:scenarios}) in which the error terms followed a normal distribution, out of the 500 datasets generated, we obtained that on average 60.8\% of the Y response was missing, with the lowest missing percentage being 26.84\% and the maximum being 73.67\%.
-At the cluster level, looking only at the sporadic missing clusters, we found an average of 35.25\% missing values, but there were clusters with no missing data at all and up to 98\% missing data in the outcome.
-Processing times differ in the imputation methods evaluated, the 1l.
-Heckman takes an average of 10 seconds, the 2l.MAR takes 1.5 and the 2l.
-Heckman takes about 18 seconds.
-It should be noted that the time of the last two approaches depends strongly on the number of clusters included and the sample size in the data set.
-(See Annex excel file) -->
-
 ### Results M(N)AR scenarios
@@
 
- ```{r rhocont,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable by varing $\\rho$, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.9, fig.width = 6}
+```{r rhocont,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable by varying $\\rho$, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.9, fig.width = 6}
 plot_rho
@@
 In the MAR scenario, i.e.
-when $\rho=0$, all imputation methods provided similar unbiased estimates of the coefficient parameters $\beta^O$ , but as $\rho$ increased, i.e.
+when $\rho=0$, all imputation methods provided similar unbiassed estimates of the coefficients $\beta^O$ , but as $\rho$ increased, i.e.
 the mechanism became MNAR, the estimates for the complete-case analysis and the 2l.MAR imputation method became further away from the true value.
 
-As expected, both Heckman-based imputation methods (1l.Heckman and 2l.Heckman) gave less biased estimates of the coefficient parameters $\beta^O$ in MNAR scenarios, but the estimates of the random effects parameters $\sqrt{\psi^O_{kk}}$ with $k =\{0,1,2\}$ in the 1l.Heckman method were the most unbiased.
+As expected, both Heckman-based imputation methods (1l.Heckman and 2l.Heckman) gave less biassed estimates of the coefficients $\beta^O$ in MNAR scenarios, but the estimates of the random effects parameters $\sqrt{\psi^O_{kk}}$ with $k =\{0,1,2\}$ in the 1l.Heckman method were the most unbiassed.
 This could be explained by the fact that 1l.Heckman does not take into account any grouping information.
 
-In terms of coverage, the 2l.Heckman imputation method gave the best coverage values in all $\rho$ scenarios with a coverage level close to the nominal 95\% value.
-Although 2l.Heckman was not properly a randomization-valid method [@buuren18c], as it was not unbiased and had a coverage above 95\% across all $\rho$ values, this method gave better results in terms of bias and coverage when compared to the other methods evaluated.
+In terms of coverage, the 2l.Heckman imputation method gave the best coverage values in all $\rho$ scenarios with a coverage level close to the nominal 95% value.
+Although 2l.Heckman was not properly a randomisation-valid method [@buuren18c], as it was not unbiassed and had a coverage above 95% across all $\rho$ values, this method gave better results in terms of bias and coverage when compared to the other methods evaluated.
 The 2l.Heckman method, also resulted in better $\beta^O$ estimates in terms of RMSE than the other methods evaluated, through a better compromise between bias and variance.
 
-In particular, the 2l.Heckman method provides an advantage over the 1l.Heckman method on data sets with systematic omission, by including cluster information in the imputation model.
-This can be seen in the random effects parameter estimates $\sqrt{\psi^O_{kk}}$ of the 1l.Heckman method, which were more biased than those of the other methods where cluster information was included, i.e.
+In particular, the 2l.Heckman method provides an advantage over the 1l.Heckman method on data sets with systematically missing variables, by including cluster information in the imputation model.
+This can be seen in the random effects parameter estimates $\sqrt{\psi^O_{kk}}$ of the 1l.Heckman method, which were more biassed than those of the other methods where cluster information was included, i.e.
 2l.MAR and 2l.Heckman.
@@
 
-In the case of an incomplete bivariate outcome, we also found (Figure \ref{fig:rhobin}) that the 2l.Heckman method provided the least unbiased results on the coefficient parameters $\beta^O$ with the best coverage values.
-As for the random effect parameters, we observed unbiased estimates especially $\sqrt{\psi^O_{22}}$ at large values of $\rho$.
+In the case of an incomplete bivariate outcome, we also found (Figure \ref{fig:rhobin}) that the 2l.Heckman method provided the least unbiassed results on the coefficients $\beta^O$ with the best coverage values.
+As for the random effect parameters, we observed unbiassed estimates especially $\sqrt{\psi^O_{22}}$ at large values of $\rho$.
 
-```{r rhobin,fig.align='center',fig.pos="!htb",fig.cap="Binary incomplete variable by varing $\\rho$, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.9, fig.width = 6}
+```{r rhobin,fig.align='center',fig.pos="!htb",fig.cap="Binary incomplete variable by varying $\\rho$, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.9, fig.width = 6}
 plot_bin
@@
 ### Sensitivity analysis: number of clusters and sample size of clusters
-We assessed the robustness of our method to variations in the number of clusters and also in the cluster sample size (Figure \ref{fig:size}).
 
-By increasing N, the number of clusters, from 10 to 100, we observed that the bias was not affected, but the 95\%CI width decreased (higher precision) and the RMSE decreased as well.
-On the other hand, as we reduced the number of units per cluster (from $n_i$=1000 to $n_i$=100) the precision decreased for all coefficients, the bias of the $\beta^O$ coefficient estimates was not drastically affected, but the bias of the $\sqrt{\psi_{kk}^O}$ random effects parameters was.
-
-When we reduced the sample size to 50 patients per study, the bias and RMSE of the $sqrt{\psi_{22}^O}$ were drastically affected (not shown here, but in the Appendix).
-This could be partly explained by the sparse information on certain clusters, which directly affects the estimation of the Heckman model on these clusters.
-
-```{r size,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable under systematical missingness by varing: number of clusters (N); sample size per cluster ($n_i$), where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.9, fig.width = 6}
+```{r size,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable under systematic missingness by varying: number of clusters (N); sample size per cluster ($n_i$), where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.9, fig.width = 6}
 plot_N
@@
 
+We assessed the robustness of our method to variations in the number of clusters and also in the cluster sample size (Figure \ref{fig:size}).
+
+By increasing N, the number of clusters, from 10 to 100, we observed that the bias was not affected, but the 95%CI width decreased (higher precision) and the RMSE decreased as well.
+On the other hand, as we reduced the number of units per cluster (from $n_i$=1000 to $n_i$=100) precision decreased for all coefficients, the bias of the $\beta^O$ coefficient estimates was not drastically affected, but the bias of the $\sqrt{\psi_{kk}^O}$ random effects parameters was.
+
+When we reduced the sample size to 50 patients per study, bias and RMSE of the $sqrt{\psi_{22}^O}$ were drastically affected (not shown here, but in the Appendix).
+This could be partly explained by the sparse information on certain clusters, which directly affects the estimation of the Heckman model on these clusters.
+
 ### Sensitivity analysis: distributional assumptions
 
-When using the Heckman model in an explicit MNAR process, i.e.
-when the probability that an observation is missing is itself associated with the value of the observation, we observe that under all imputation methods we obtain biased estimators, with poor coverage.
-In particular, when estimating the parameter intercept $\beta^O_0$ we observe more biased estimates compared to the estimates of the other coefficient parameters, with a coverage of less than 60%.
+```{r dist,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable with deviations in distribution assumptions, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.9, fig.width = 6}
+plot_S
+```
+
+When using the Heckman model in an self-masking MNAR process, i.e.
+when the probability that an observation is missing is associated with the value of the observation, we observe that under all imputation methods we obtain biassed estimators, with poor coverage.
+In particular, when estimating the parameter intercept $\beta^O_0$ we observe more biassed estimates compared to the estimates of the other coefficients, with a coverage of less than 60%.
 
@@
 With respect to the Skewed-t scenario, only the bias of $\beta_0^O$ was affected for all imputation methods applied.
-But the estimates of $\beta_0^O$ for both 1l.Heckman and 2l.Heckman were drastically affected in terms of bias, with a very poor coverage (below 25%).Therefore, the applicability of our method might be questionable in scenarios that do not conform to the BNV assumption.
+But the estimates of $\beta_0^O$ for both 1l.Heckman and 2l.Heckman were drastically affected in terms of bias, with a very poor coverage (below 25%).
+Therefore, the applicability of our method might be questionable in scenarios that do not conform to the BNV assumption.
 
-```{r dist,fig.align='center',fig.pos="!htb",fig.cap="Continuous incomplete variable with deviations in distribution assumptions, where dashed lines depict the target performance criteria value",echo=FALSE,fig.keep="last",fig.height = 3.9, fig.width = 6}
-plot_S
-```
+Please note that our approach relies on specific assumptions, and any deviation from these assumptions may have an impact on its performance.
+Therefore, it is advisable for users to carefully assess these assumptions and determine whether they align with a plausible missing mechanism in their data.
+
+As we employ a Gaussian copula function to estimate the Heckman model, it is important to recognize that deviations from normality assumptions could lead to biased and less accurate estimates [@ogundimu_hutton16].
+In certain situations, it is recommended to compare the results with those obtained using alternative imputation methods, such as predicted mean matching (PMM), which can be especially beneficial to prevent the generation of imputation values that fall outside the range of observable values.
+Alternatively, one may explore the possibility of refining our method by utilizing a copula function that better fits the error dependence distribution of the data.
 
 # An illustrative study
+
 \label{sec:illustration}
-```{r, include=FALSE,cache = TRUE}
+
+```{r, include=FALSE}
 library(here)
-
  source(here('3.Ilustrative_study','Ilustrative_study.R'),local = knitr::knit_global())
@@
 Although the original data set consists of 164 variables, here we only consider the variables described in Table \ref{tab:descriptive}, which were used as predictors in the imputation model and are fully observed in the dataset.
-In this dataset, the parasitaemia test is an incomplete binary result (1=positive test, 0=negative test), which is missing 21\% across the whole dataset.
+In this dataset, the parasitaemia test is an incomplete binary result (1=positive test, 0=negative test), which is missing 21% across the whole dataset.
 
-```{r descriptive, echo = FALSE, warning = FALSE,cache = TRUE}
+To illustrate our proposed method, following the article by Rugnao et al.
+(2019),[@rugnao2019] we estimated the prevalence of parasitemia by region and by age after approximately 3 years of LLIN campaigns started.
+We estimated parasitemia prevalence using 3 approaches that made different assumptions on the missingness mechanism: MCAR, MAR and MNAR.
+
+```{r descriptive, echo = FALSE, warning = FALSE,cache = TRUE,fig.pos='H'}
 options(kableExtra.latex.load_packages = FALSE) 
@@
 
-To illustrate our proposed method, following the article by Rugnao et al.
-(2019),[@rugnao2019] we estimated the prevalence of parasitemia by subregion and by age after approximately 3 years of LLIN campaigns started.
-We estimated parasitemia prevalence using 3 approaches that made different assumptions on the missingness mechanism: MCAR, MAR and MNAR.
-
 Under the MCAR assumption, prevalence was calculated on the basis of the recorded tests, i.e., we only included patients with a test result.
 Under the MAR assumption, the test values of children who were not present during the survey were imputed with the 2l.2stage.bin method of the micemd package, where the community was taken as the cluster and the following factors previously associated with parasitemia were used as predictors in the imputation model: sex, bednet ( indicator of whether only two or fewer persons share a mosquito bed net).
-In addition, we included age as a power 3 spline function, the cluster-level Log10 mean of the number of female anopheline mosquitoes per household estimated from the entomological survey, and the household wealth index from principal components analysis calculated specifically for the surveyed households.
+In addition, we included age as a cubic spline function, the cluster-level Log10 mean of the number of female anopheline mosquitoes per household estimated from the entomological survey, and the household wealth index from principal components analysis calculated specifically for the surveyed households.
 
@@
 This was calculated according to school vacation calendars and public holidays in Uganda in 2017.
-We examined the association of this ERV with the outcome variable (y) and with the selection indicator (ry), conditioned on the remaining imputation predictors.
+We examined the association of this ERV with the outcome variable (y) and with the selection indicator (r), conditioned on the remaining imputation predictors.
 The model results in Table \ref{tab:exclusion} indicate that the holiday indicator could be a plausible ERV variable, as there was strong evidence of an association with ry, but no evidence of an association with y.
+
 ```{r exclusion, echo = FALSE, warning = FALSE,cache = TRUE}
@@
     caption = "Evaluation of Holidays as exclusion restriction variable",
-    col.names = c("Predictors/response","Test result (y)","Test taken (ry)"),
+    col.names = c("Predictors/response","Test result (y)","Test taken (r)"),
     row.names=FALSE)%>%
@@
 
-According to our imputation approach, non-tested children were estimated to have a higher prevalence of malaria than participants in more than half of the districts analyzed. 
-As can be seen in Figure \ref{fig:region}, for each subregion the prevalence estimates of the approaches did not differ significantly between methods.
+According to our imputation approach, non-tested children were estimated to have a higher prevalence of malaria than participants in more than half of the districts analysed.
+As can be seen in Figure \ref{fig:region}, for each region the prevalence estimates of the approaches did not differ significantly between methods.
 However, prevalence estimates under the MNAR assumption (i.e.
-2 level Heckman) were higher than those estimated under the MAR or MCAR aproaches, except for the East-Central region.
-```{r region,fig.align='center',fig.width = 5, fig.asp =.67, fig.pos="!htb", fig.cap="Estimates of malaria prevalence by sub-region",echo=FALSE,fig.keep="last",cache=TRUE}
+2 level Heckman) were higher than those estimated under the MAR or MCAR approaches, except for the East-Central region.
+
+```{r region,fig.align='center',fig.width = 5, fig.asp =.67, fig.pos="!htb", fig.cap="Estimates of malaria prevalence by sub-region.The ranges of plausible prevalence values are represented by dashed lines behind the CC method.",echo=FALSE,fig.keep="last",cache = TRUE}
 plot_dist
@@
 However for school children, prevalences estimated with the Heckman method were found to be higher in the Mid-East and Southwest regions than those obtained with the other methods, whereas in the East-Central region the estimates with the Heckman method are lower.
-A possible reason for selection bias in surveys of this type is, for example, that daytime visits might favor measurement in sick school children who stay home, leading to overestimated prevalence results as found in the East-Central region.
-[@thedhsprogram2020] Nevertheless, we were unable to find information confirming the direction in which malaria prevalence is driven by selection bias in this Uganda study or in other studies similar to this one.
+A possible reason for selection bias in surveys of this type is, for example, that daytime visits might favour measurement in sick school children who stay home, leading to overestimated prevalence results as found in the East-Central region.\cite{thedhsprogram2020} Nevertheless, we were unable to find information confirming the direction in which malaria prevalence is driven by selection bias in this Uganda study or in other studies similar to this one.
 
 # Discussion
-\label{sec:discussion}
-We have extended and evaluated methods for multiple imputation of clustered datasets, in situations where some incomplete variables follow an MNAR mechanism.
+
+\label{sec:discussion} We have extended and evaluated methods for multiple imputation of clustered datasets, in situations where some incomplete variables follow a MNAR mechanism.
 Although there are imputation methods that address incomplete MNAR variables, these can only deal with the case of individual studies.
 This limits their use in common situations in IPD-MA, such as when there is systematic missingness or when the proportion of missingness of a variable is very high in one of the included studies.
-To address this gap, we proposed a new multiple imputation method for incomplete continuous and binary MNAR variables with sporadic as well as systematic missingness, which come from clustered datasets with heterogeneous effects and error variances.
+To address this gap, we proposed a new multiple imputation method for incomplete continuous and binary MNAR variables with sporadic as well as systematic missingness, applicable for the case of clustered datasets with heterogeneous effects and error variances.
 
-Empirically we observed that our proposed imputation method is well suited for the imputation of both continuous and binary incomplete variables following a MNAR mechanism according to Heckman's model.
-The simulation study showed that the application of the proposed method on clustered datasets with heterogeneous effects and error variances and with sporadically and systemically missing variables resulted in less biased estimates of effects with a convergence close to 95% compared to those obtained using other imputation methods evaluated.
+In a simulation study, we found that our proposed imputation method is well suited for the imputation of both continuous and binary incomplete variables following a MNAR mechanism indirectly non-ignorable.
+The simulation study showed that the application of the proposed method on clustered datasets with heterogeneous effects and error variances and with sporadically and systemically missing variables resulted in less biassed estimates of effects with a convergence close to 95% compared to those obtained using other imputation methods evaluated.
 
-This method, in particular, could yield better results than Heckman-based imputation methods for individual studies, as it not only allows the imputation of missing values in clusters with systematic missigness, but can also reduce the values of individual clusters towards the overall study mean.
-This is especially advantageous in studies with small sample sizes, where an analysis approach that ignores data from other studies may lead to extreme effect estimates.
+Our method may yield better results than Heckman-based imputation methods for individual studies, as it not only enables the imputation of missing values within clusters exhibiting systematic missingness but also has the capacity to adjust values within individual clusters, aligning them more closely with the overall study mean.
+This can be especially advantageous in studies with small sample sizes, where an analysis approach that ignores data from other studies may lead to extreme effect estimates.
 
@@
 
-In addition, to estimate the marginal estimates, the method only uses clusters with observable information, i.e.
+In addition, to estimate marginal estimates, the method only uses clusters with observable information, i.e.
 that are not systematically missing or have sufficient information to estimate the Heckman model.
@@
 
-Also, the method can be sensitive to both the sample size and the number of studies included in the database.
-On the one hand, a small sample size at the individual study level can affect not only the precision of estimates but also the convergence of the method since sample size needed to estimate the parameters of the Heckman model which can be at least twice the number of parameters required to estimate in an imputation model that assumes MAR.
-On the other hand, a high number of studies that may improve the precision of the estimators may also make the estimation of the marginal parameters more difficult and also considerably increase the processing time of our method.
+Also, the method can be sensitive to both the sample size of the individual studies, as well as the number of studies included in the database.
+On the one hand, a small sample size at the individual study level can affect not only the precision of estimates, but also the convergence of the method since the sample size required to estimate the parameters of the Heckman model can be at least twice the number of parameters required to estimate in an imputation model that assumes MAR.
+On the other hand, a large number of studies, which may improve precision of the estimations may also make the estimation of the marginal parameters more difficult and may also considerably increase the processing time of our method.
 
 In our simulation study, data were generated by assuming a constant correlation across all clusters in order to evaluate the performance against M(N)AR assumptions.
-In practice, however this parameter can be variable across clusters and this can considerably affect the performance of our method.
-Therefore, in future research the effect of this parameter could be further evaluated.
-One might also consider relaxing this assumption of constant correlation to allow for a random effects distribution for the correlation parameter.
+In practice, however, this parameter can be variable across clusters which can considerably affect the performance of our method.
+Therefore, the effect of this parameter could be further evaluated in future research.
+One might also consider relaxing the assumption of constant correlation to allow for a random effects distribution of the correlation parameter.
 
@@
 We have proposed an extension to the Heckman model that can account for MNAR, MAR or MCAR of a continuous or binary variable in clustered data sets.
-Our simulations showed that it can have favorable statistical properties, when its assumptions were met, and provided that the sample size is sufficiently large.
-Regarding deviations from distributional assumptions of the error terms, the coefficient parameters were fairly robust in terms of bias, but the intercept was not.
+Our simulations showed that it can have favourable statistical properties, when its assumptions were met and provided that the sample size is sufficiently large.
+Regarding deviations from distributional assumptions of the error terms, the estimated parameters were fairly robust in terms of bias, but the intercept was not.
 
 # Footnotes {.unnumbered}
+
 ### Disclaimer {.unnumbered}
-The views expressed in this paper are the personal views of the authors and may not be understood or quoted as being made on behalf of or reflecting the position of the regulatory agency/agencies or organizations with which the authors are employed/affiliated.
+
+The views expressed in this paper are the personal views of the authors and may not be understood or quoted as being made on behalf of or reflecting the position of the regulatory agency/agencies or organisations with which the authors are employed/affiliated.
 
 ### Acknowledgements {.unnumbered}
+
+```{=tex}
 \begin{minipage}{.08\textwidth}
@@
 In addition, we are grateful for the support of Professor Matthias Egger in conceptualising this article and helping us to find a suitable data set for the illustrative study.
+We would like to thank the Associate Editor, and the reviewer for their helpful comments that have substantially improved the manuscript.
 \end{minipage}
+```

### 2023-11-01 2.Simulation/simplot.R

@@ -21,7 +21,7 @@ data<- setDT(dcast(data, Scenario+ Variation+ Method +  Parameter+Statistic~Type
 param_val <- c("0"="beta_0", "1"="beta_1", "2"="beta_2", 
                "3"="psi_b0", "4"="psi_b1", "5"="psi_b2","6"= "psi_e")
 data[,Method:=factor(Method, levels = c("CC","1l.Heckman","2l.MAR","2l.Heckman"))]
-
+levels(data$Method)<-c("CC","1l.heckman","2l.MAR","2l.2stage.heckman")
 data[,Parameter := as.factor(param_val[data$Parameter])]
 data[,Parameter := factor(Parameter, levels = c("beta_0","beta_1","beta_2","psi_b0","psi_b1","psi_b2","psi_e"),
                           ordered = TRUE, labels=c(expression(beta[0]**O),expression(beta[1]**O),expression(beta[2]**O),
@@ -72,8 +72,19 @@ plot<-ggplot(datagraph[!Parameter%in%c("psi[e]")],
 return(plot)
 }
 
+width=5
+height=width/1.1
+width/height
 
 plot_rho<-plotsim(Scenariov="Rho",levelsvar=c("0","0.3","0.6","0.9"),yaxis=expression(rho))
+plot_rho
+ggsave(filename = "Figure2.png", width = width, height = height, dpi=380)
 plot_bin<-plotsim(Scenariov="Bin",levelsvar=c("0","0.3","0.6","0.9"),yaxis=expression(rho))
+plot_bin
+ggsave(filename = "Figure3.png", width = width, height = height, dpi=380)
 plot_N<-plotsim(Scenariov="N",levelsvar=c("10;50","10;100","10;1000","50;1000","100;1000"),yaxis="Number of clusters; sample size per cluster")
-plot_S<-plotsim(Scenariov="S",levelsvar=c("Skewed-t","Normal","Explicit"),yaxis="Missing process")
+plot_N
+ggsave(filename = "Figure4.png", width = width, height = height, dpi=380)
+plot_S<-plotsim(Scenariov="S",levelsvar=c("Skewed-t","Normal","Self-masking"),yaxis="Missing process")
+plot_S
+ggsave(filename = "Figure5.png", width = width, height = height, dpi=380)

### 2023-11-01 3.Ilustrative_study/Additional_functions.R

@@ -43,16 +43,26 @@ f.abs.perstudy<-function(data, outcome_name,study_name) {
 #Results in absolute risk per study
 
 f.pool<-function(data,m){
-  logit.abs <- mean(data$logit.prevalence)
-  within <- mean(data$logit.se^2)
-  between <- (1 + (1/m)) * var(data$logit.prevalence)
-  logit.var <- within + between
-  logit.se <- sqrt(logit.var)
-  logit.ci.lb <- logit.abs + qnorm(0.05/2)     * logit.se
-  logit.ci.ub <- logit.abs + qnorm(1 - 0.05/2) * logit.se
-  prevalence<-inv.logit(logit.abs)*100
-  ci.lb<-inv.logit(logit.ci.lb)*100
-  ci.ub<-inv.logit(logit.ci.ub)*100
+
+  alpha <- 0.05
+  tr_est <- data$logit.prevalence
+  tr_se  <- data$logit.se
+  n <- length(tr_est)  #number observations
+  mu <- mean(tr_est, na.rm = T) # pool estimate
+  w_var <- mean(tr_se^2, na.rm = T) # within variance
+  b_var <- var(tr_est, na.rm = T) # between variance
+  t_var <- sum(w_var,b_var,b_var/n,na.rm=T) # total variance
+  t_se <- sqrt(t_var) # total standard error
+  r <-  sum(b_var,(b_var / n),na.rm=T)/ w_var # relative increase variance due to missing values
+  v <- (n - 1) * (1 + r^-1)^2 # degrees of freedom
+  t <- qt(1-alpha/2, v) #t criteria
+  if (is.infinite(v)|is.na(v)){ # t can not be calculated e.g 1 only observation it is aprox to normal 
+    t <- qnorm(1-alpha/2)
+  }
+  
+  prevalence <- inv.logit(mu)*100 # mean(est, na.rm = T)
+  ci.lb <- inv.logit(mu - t_se*t)*100
+  ci.ub <- inv.logit(mu + t_se*t)*100
   return(list(prevalence=prevalence,ci.lb=ci.lb,ci.ub=ci.ub))
 }
 
@@ -62,14 +72,6 @@ f.abs.poolrubin <-function(data,study_name) {
   setnames(abs.outcome, "get", study_name)
   return(abs.outcome)
 }
-
-
-f.abs.poolrubinall <-function(data,study_name) {
-  abs.outcome<-setDT(data)[, f.pool(data=.SD, m<-max(data$.imp)), by = list(get)]
[…]

### 2023-11-01 3.Ilustrative_study/Ilustrative_study.R

@@ -7,11 +7,7 @@ library(tidymv)
 library(ggplot2)
 
 
-
 #0. Additional required function ----
-# Prevalence functions 
-logit <- function(x){log(x/(1-x))}
-inv_logit <- function(x){exp(x)/(1+exp(x))}
 
 #Function for calculating prevalence CI
 f.prevalence <- function(data,outcome_name) {
@@ -31,7 +27,7 @@ f.abs.perstudy<-function(data, outcome_name,study_name) {
   prev_group<-setDT(data)[, f.prevalence(data=.SD, outcome_name=outcome_name), by = list(get(study_name),.imp)]
   
   #Logit transformation
-  prev_group[,prev_logit:=logit(prev)] 
+  prev_group[,prev_logit:=qlogis(prev)] 
   prev_group[,prev_se_logit:=prev_se/(prev*(1-prev))]
   return(prev_group)
 }  
@@ -47,9 +43,9 @@ f.pool<-function(data,m){
   pool_LC_logit=pool_est_logit-pool_se_logit*t
   pool_UC_logit=pool_est_logit+pool_se_logit*t
   
-  prevalence<-inv_logit(pool_est_logit)*100
-  ci.lb<-inv_logit(pool_LC_logit)*100
-  ci.ub<-inv_logit(pool_UC_logit)*100
+  prevalence<-plogis(pool_est_logit)*100
+  ci.lb<-plogis(pool_LC_logit)*100
+  ci.ub<-plogis(pool_UC_logit)*100
   return(data.frame(cbind(prevalence=prevalence,ci.lb=ci.lb,ci.ub=ci.ub)))
 }
 
@@ -80,9 +76,9 @@ pred_model<-function(model,source){
   mod[,age := as.numeric(age)]
   mod[,fit := mm%*%coef]
   mod[,se.fit := sqrt(diag(mm %*% pool_variance(model) %*% t(mm)))]
-  mod[,pred := fn(fit)*100]
-  mod[,lb:=fn(fit - 1.96 * se.fit)*100]
-  mod[,ub:=fn(fit + 1.96 * se.fit)*100]
+  mod[,pred := plogis(fit)*100]
+  mod[,lb:=plogis(fit - 1.96 * se.fit)*100]
+  mod[,ub:=plogis(fit + 1.96 * se.fit)*100]
   mod[,source:=source]
 }
 
@@ -92,7 +88,7 @@ pred_model<-function(model,source){
[…]

### 2023-11-01 4.Codes/Toy_example.R

@@ -1,217 +1,218 @@
 rm(list=ls())
-library(data.table)
-library(truncnorm)
+require(mice)
+require(micemd)
+library(tidyverse)
+library(broom.mixed)
 library(nlme)
-library(devtools)
-library(mice)
+library(ggplot2)
+library(ggmice)
+library(patchwork)
 
-#0. Load original dataset ----
+## Load the dataset ----
+# data("Obesity",pkg="micemd")
 
-data <- as.data.table(read.csv('/Users/jmunozav/Desktop/Statsmed_Heckman/4.Codes/ObesityDataSet_raw_and_data_sinthetic.csv', stringsAsFactors=FALSE, fileEncoding="latin1"))
+## Explore the dataset ----
 
-#1. Modify dataset in order to get a variable weight with MNAR missing mechanism ----
-# Generate countries where survey was taken
-data <- data[,c("Weight","Gender","Age","Height","FAVC")] # Select some of the variables of the original dataset
-data[,Cluster:=c(rep(1,500),rep(2,400),rep(3,450),rep(4,400),rep(5,361))] # in the imputation method is necessary cluster is included as numeric
-N<-length(unique(data$Cluster))
-#  Generation of the exclusion restriction (ERV): The time variable describes the time it takes for a person to answer the weight question.
-set.seed(12345)
-data[,Time:=rtruncnorm(n=nrow(data), a = 1, b = 10, mean = 5, sd = 3)] 
-data[,Gender:=(ifelse(Gender=="Female",1,0))]
-data[,FAVC:=(ifelse(FAVC=="yes",1,0))]
-data[,Heights:=scale(Height)]
-data[,Ages:=scale(Age)]
-data[,Times:=scale(Time)]
+### Missing patterns ----
+md.pattern(Obesity)
+by(Obesity,INDICES = Obesity$Cluster,summary)
 
-MAR_model <- with(data, lme4::lmer(Weight~Gender+Ages+Heights+FAVC+(1+Heights+Gender+Ages|Cluster)))
-summary(MAR_model)
-# Generate outcome variable
-data[,XOBO:=predict(MAR_model)+15]
+myplots <- lapply(1:5, function(i) {
+  ggmice::plot_pattern(setDT(Obesity)[Cluster==i])+ 
[…]

### 2023-11-01 4.Codes/gen_dataObs.R

@@ -1,87 +1,0 @@
-# Generate data toy example
-rm(list=ls())
-library(data.table)
-library(truncnorm)
-library(nlme)
-library(devtools)
-library(mice)
-
-#0. Load original dataset ----
-
-data <- as.data.table(read.csv('/Users/jmunozav/Desktop/Statsmed_Heckman/4.Codes/ObesityDataSet_raw_and_data_sinthetic.csv', stringsAsFactors=FALSE, fileEncoding="latin1"))
-
-#1. Modify dataset in order to get a variable weight with MNAR missing mechanism ----
-# Generate countries where survey was taken
-data <- data[,c("Weight","Gender","Age","Height","FAVC")] # Select some of the variables of the original dataset
-data[,Cluster:=c(rep(1,500),rep(2,400),rep(3,450),rep(4,400),rep(5,361))] # in the imputation method is necessary cluster is included as numeric
-N<-length(unique(data$Cluster))
-#  Generation of the exclusion restriction (ERV): The time variable describes the time it takes for a person to answer the weight question.
-set.seed(12345)
-data[,Time:=rtruncnorm(n=nrow(data), a = 1, b = 10, mean = 5, sd = 3)] 
-data[,Gender:=(ifelse(Gender=="Female",1,0))]
-data[,FAVC:=(ifelse(FAVC=="yes",1,0))]
-data[,Heights:=scale(Height)]
-data[,Ages:=scale(Age)]
-data[,Times:=scale(Time)]
-
-MAR_model <- with(data, lme4::lmer(Weight~Gender+Ages+Heights+FAVC+(1+Heights+Gender+Ages|Cluster)))
-summary(MAR_model)
-# Generate outcome variable
-data[,XOBO:=predict(MAR_model)+15]
-
-# Generate random slope for Gender Age on selection equation
-alpha_gender <- rnorm(n = N, mean=0, sd = 0.01)
-alpha_age <- rnorm(n = N, mean=0, sd = 0.01)
-nobs<-c(500,400,450,400,361)
-data[,Genderi:=unlist(mapply(rep,alpha_gender,each=nobs))]
-data[,Agei:=unlist(mapply(rep,alpha_age,each=nobs))]
-data[,XSBS:=1.5-(0.7+Genderi)*Gender-(0.5+Agei)*Ages-1.2*Times] # Simulated selection model
-
[…]

### 2023-11-01 4.Codes/gendata_Obesity.R

@@ -1,0 +1,133 @@
+# Synthetic data based on a obesity survey retrieved from
+# https://www.kaggle.com/datasets/fabinmndez/obesitydata/
+
+rm(list=ls())
+
+library(data.table)
+library(truncnorm)
+library(nlme)
+library(devtools)
+library(mice)
+library(mvtnorm)
+
+
+generate_data <- function(betao,betas){
+  
+  set.seed(1234)
+  N <- 5
+  Ns <- c(500,400,450,400,361)
+  Gender <- Age <- Height <- BMI <- FamOb <- Time <- Cluster <- bo1 <-bo2<- bs1 <-bs2 <- NULL
+  sigmae <- exp( rnorm( n= N, mean=log(1.8), sd = 0.3))
+  es <- eo <- NULL
+  # Effects
+  
+  # Random effects
+  # Generate random intercept and slope outcome
+  d <-diag(2)
+  rho= -0.4
+  d[1,1] <- 15^2
+  d[2,2] <- 6^2
+  d[2,1] <- d[1,2] <- rho*sqrt(d[1,1]*d[2,2])
+  bo <- rmvnorm(n = N, mean = rep(0,2), sigma = d)
+  # Generate random intercept and slope selection
+  d <-diag(2)
+  rho= 0.4
+  d[1,1] <- 0.8^2
+  d[2,2] <- 0.05^2
+  d[2,1] <- d[1,2] <- rho*sqrt(d[1,1]*d[2,2])
+  bs <- rmvnorm(n = N, mean = rep(0,2), sigma = d)
+  
+  for (i in 1:N){
+    
+    n <- Ns[i]
+    Gender <- c( Gender, rbinom(n = n, size = 1, prob = 0.49))
+    Age <- c( Age, round(rtruncnorm(n, a = 14, b = 61, mean = 24, sd = 6.34), 0)) # rounded to integers
+    Height <- c( Height, rtruncnorm(n, a = 1.45, b = 1.98, mean = 1.7, sd = 0.09))
+    BMI <- c(BMI,rtruncnorm(n, a = 13, b = 50, mean = 29, sd = 8))
+    FamOb <- c(FamOb,rbinom(n = n, size = 1, prob = 0.85))
+    Time<- c(Time,rtruncnorm(n=n, a = 1, b = 10, mean = 5, sd = 2.5))
+    Cluster <- c(Cluster,rep(i,n))
[…]

### 2023-11-01 4.Codes/mice.impute.2l.2stage.heckman.R

@@ -1,29 +1,34 @@
+library(lme4)
+library(data.table)
+library(mvtnorm)
+library(fMultivar)
+library(data.table)
+library(mgcv)
+library(GJRM)
+library(Matrix)
+library(mixmeta)
+library(mvtnorm)
+library(pbivnorm)
+library(mice)
+library(micemd)
+library(glmmTMB)
+
+
 #' Imputation based on Heckman model for multilevel data.
-#'
 #' Imputes outcome and predictor variables that follow an MNAR mechanism
 #' according to Heckman's model and come from a multilevel database such as
 #' individual participant data.
 #' @aliases mice.impute.2l.2stage.heckman 2l.2stage.heckman
+#'
 #' @param y Vector to be imputed
-#' @param ry Logical vector of length \code{length(y)} indicating the
-#' the subset \code{y[ry]} of elements in \code{y} to which the imputation
-#' model is fitted. The \code{ry} generally distinguishes the observed
-#' (\code{TRUE}) and missing values (\code{FALSE}) in \code{y}.
-#' @param x Numeric design matrix with \code{length(y)} rows with predictors for
-#' \code{y}. Matrix \code{x} may have no missing values.
-#' @param wy Logical vector of length \code{length(y)}. A \code{TRUE} value
-#' indicates locations in \code{y} for which imputations are created.
-#' @param type type of the variable in the prediction model {0: No predictor,
-#' 1: Predictor in both the outcome and selection,-2: Cluster id (study id),
-#' -3: Predictor only in the selection model, -4: Predictor only in the outcome 
-#' model}
-#' @param pmm  predictive mean matching can be applied only for for missing 
-#' continuous variables: "FALSE","TRUE"
-#' @param ypmm vector of donor values of y to perform the predictive mean matching, 
-#' in case ypmm is not provided, the observable values of y are used.
-#' @param meta_method meta_analysis estimation method for random effects :
-#' "ml" (maximum likelihood), "reml" (restricted maximum likelihood) or "mm"
-#' method of moments.
[…]

### 2023-11-01 5.micemd/copulaIPD.R

@@ -36,7 +36,8 @@ copulaIPD <- function(data, sel, out, family, send) {
     gam2 <- try(mgcv::gam(formula=out,data = data,family = family, method ="REML"))
     fit_ind <- 0
     if(!any(inherits(gam1, "try-error"))&!any(inherits(gam2, "try-error"))){
-      vp<-c(diag(gam1$Vp),diag(gam2$Vp))
+      coefficients <- c(gam1$coefficients,gam2$coefficients)
+      vp <- c(diag(gam1$Vp),diag(gam2$Vp))
       if(all(c(!is.na(coefficients),vp!=0))){
         s     <- ifelse(family != "binomial",1,0) 
         ncol1 <- ncol(gam1$Vp)

### 2023-11-01 5.micemd/mice.impute.2l.2stage.heckman.R

@@ -1,10 +1,14 @@
-mice.impute.2l.2stage.heckman <-function(y,ry,x,wy = NULL, type, pmm = FALSE, ypmm=NULL, meta_method ="reml",...) {
+mice.impute.2l.2stage.heckman <- function(y, ry, x, wy=NULL, type, pmm=FALSE, ypmm=NULL, meta_method="reml", pred_std=FALSE,...){
   
+   if (pred_std == TRUE){
+     x <- apply(x, MARGIN = 2, FUN = function(X) (if(length(unique(X))<=2&sum(unique(X))==1){X} else{scale(X)}))
+  }
+
   # 1. Define variables and dataset----
   
   # Rename covariates
   colnames(x) <- paste0("x_", 1:length(colnames(x))) #change the covariates name for avoiding conflicts when y is covariate
-  bos_name <- colnames(x)[type ==  1] # names of variables present in both outcome and selection model
+  bos_name <- colnames(x)[type ==  2] # names of variables present in both outcome and selection model
   sel_name <- colnames(x)[type == -3] # names of variables in selection model alone
   out_name <- colnames(x)[type == -4] # names of variables in outcome model alone
   

### 2023-11-01 3.Ilustrative_study/Ilustrative_study.R
Update 1st revision

@@ -215,7 +215,7 @@ pred["test","sage2"]<- 2
 
 #3.3. Multiple imputation with heckman model----
 meth<-ini$method
-meth[c("test")]<-"3l.2stage.heckman"
+meth[c("test")]<-"2l.2stage.heckman"
 quickpred(hdataF0f)
 #Full correlation of parameters
 data_heck <- mice(hdataF0f, # dataset with missing values
